### 2014-12-12 RIP-MISDP-Paper/Paper_v1.tex
added subfolder for Paper about exact RIP constants via MISDP

\documentclass{elsarticle}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\newtheorem{thm}{Theorem}
\newdefinition{rmk}{Definition}

\begin{document}
\title{Computing exact restricted isometry constants via mixed integer semidefinite programming}
\author[tud]{T. Gally}
\ead{gally@mathematik.tu-darmstadt.de}
\author[tud]{M. E. Pfetsch}
\ead{pfetsch@mathematik.tu-darmstadt.de}
\address[tud]{Technische Universit\"at Darmstadt, Dolivostra\ss e 15, 64293 Darmstadt, Germany}
\begin{abstract}
An important task in compressed sensing is finding the sparsest solution to an underdetermined system of linear equations. An important result is that, under certain conditions, this is equivalent to minimizing the 1-Norm and thus
solving a linear program. One such condition is the restricted isometry property for which no efficient algorithms are known to compute the optimal constants, the so called restricted isometry constants. In this paper we propose a
mixed integer semidefinite program for computing these constants. Numerical results are presented comparing this to ealier SDP relaxations for the same task.
\end{abstract}
\maketitle
\section{Introduction}
Finding the sparsest solution of an underdetermined system of linear equations is a problem that has been the subject of much recent research in compressed sensing. Given a matrix $A \in \mathbb{R}^{m \times n}$ with $m < n$ and 
right-hand side $b \in \mathbb{R}^m$, we are interested in a solution of 
\begin{align}\label{l0min}
 \text{min} \quad & \|x\|_0 \\
 \text{s.t.} \quad & Ax = b, \nonumber
\end{align}
where $\|x\|_0$ stands for the cardinality and thus the number of nonzero components of a vector $x$. As it is $\mathcal{NP}$-hard to solve (\ref{l0min}), see for example \cite{GareyJohnson}, we would like to solve the problem
\begin{align}\label{l1min}
 \text{min} \quad & \|x\|_0 \\
 \text{s.t.} \quad & Ax = b, \nonumber
\end{align}
instead, as this can be cast as an LP. Results in compressed sensing show, that there are many conditions which imply that the problems (\ref{l0min}) and (\ref{l1min}) are indeed equivalent. First results about this have been 
established in \cite{DH01}, and have later been followed up for example by \cite{CRT05}, \cite{Zha08} or \cite{BCT11}. One such condition is the restricted isometry property
\begin{equation}\label{RIPintro}
 (1-\delta_k)\|x\|_2^2 \leq \|Ax\|_2^2 \leq (1+\delta_k)\|x\|_2^2,
\end{equation}
that was first introduced in \cite{CT05}, if it holds for all $k$-sparse vectors $x$ for a sufficiently small restricted isometry constant $\delta_k$, then problems (\ref{l0min}) and (\ref{l1min}) have the same optimal solution. 
Even tough it was shown in \cite{BDDW08}, that this condition holds for multiple classes of random matrices with probability exponentially going to one, in \cite{PT14} it was shown, that computing the optimal restricted isometry 
constants in (\ref{RIPintro}) for a given matrix $A$ is again an $\mathcal{NP}$-hard task. \\
In the past SDP-Relaxations were proposed in \cite{Asp07} and \cite{Asp08} to compute upper bounds on the values of the restricted isometry 
constant, which could then be used to verify the equivalance between (\ref{l0min}) and (\ref{l1min}), but there could also be cases, where the results of those relaxations weren't sufficient, but the real constants would have been. 
In this paper we will construct a mixed integer semidefinite program to compute the exact constants in (\ref{RIPintro}), to be able to verify in all cases if those sufficient conditions for the equivalence hold for a given matrix. 
The main contribution of this paper is theorem \ref{Rk1thm}, which shows that the optimal solution of that MISDP indeed allows us to compute the exact restricted isometry constant.\\
In the following we will start by introducing the restricted isometry property and corresponding results about the equivalence of (\ref{l0min}) and (\ref{l1min}). In the next section we will present the SDP-relaxations in \cite{Asp07} 
and \cite{Asp08}.Afterwards we will construct a mixed integer semidefinite programme and show that it produces exact restricted isometry constants before presenting numerical results and comparisons with the SDP-Relaxations in 
\cite{Asp07} and \cite{Asp08}.

\section{The restriced isometry property}

In this section we will discuss some general results about the restricted isometry property and it's relationship with the optimization problems (\ref{l0min}) and (\ref{l1min}). To start, we will properly define the restricted
isometry property as in \cite{CT05}:

\begin{rmk} \label{ripdef}
The \emph{restricted isometry property (RIP)} of order $k$ holds for a matrix $A \in \mathbb{R}^{m \times n}$, if there exists a minimal \emph{restricted isometry constant} $\delta_k$, such that for all $x \in \Sigma_k := 
\{x \in \mathbb{R}^n : \|x\|_0 \leq k \}$ we have
 \begin{equation}\label{rip}
  (1-\delta_k)\|x\|_2^2 \leq \|Ax\|_2^2 \leq (1+\delta_k)\|x\|_2^2.
 \end{equation}
 \end{rmk}

 If the restricted isometry property holds for a sufficiently small restricted isometry constant, then the optimal solutions of (\ref{l0min}) and (\ref{l1min}) coincide. The best known result of this type is probably the following from
 \cite{Can08}:

\begin{thm} \label{RIPsqrt2}
 If $A$ satisfies $\delta_{2k} < \sqrt{2}-1$, then for every $x \in \mathbb{R}^n$
 \begin{equation*}\label{l1ErrorRIP}
  \|x^* - x \|_{\ell_1} \leq C \|x - x_k \|_{\ell_1},
 \end{equation*}
where $x^*$ is an optimal solution for (\ref{l1min}) with right-hand side $b = Ax$, $x_k$ is the best approximation of $x$ with $k$-sparse vectors and $C$ is a constant depending only on $k$ and $\delta_{2k}$. If $x$ is $k$-sparse, then the right-hand side of
(\ref{l1ErrorRIP}) is zero, so $x$ is the only $k$-sparse vector with $A\hat{x} = Ax$ and recovered by (\ref{l1min}) and therefore the optimal solutions of (\ref{l0min}) and (\ref{l1min}) coincide.\end{thm}

There are also many other results like Theorem \ref{RIPsqrt2}, for example in \cite{fou10} the condition was relaxed to $\delta_{2k} < 0.4652$ and in \cite{CWX10} it was shown for $\delta_k < 0.307$. Also all of these results can be
generalized to the case where $\|Ax - b\|_{\ell_2} \leq \epsilon$. The restricted isometry property can also be formulated in an asymetric way like in \cite{FL09}:

\begin{rmk} \label{asymRIPdef}
The \emph{restricted isometry ratio} $\gamma_k$ is defined as $\gamma_k := \beta_k^2 / \alpha_k^2$, where $\alpha_k$ and $\beta_k$ are the largest and smallest constant satisfying
\begin{equation}\label{asymRIP}
   \alpha_k^2\|x\|_2^2 \leq \|Ax\|_2^2 \leq \beta_k^2\|x\|_2^2
\end{equation}
for all $x \in \Sigma_k$.
\end{rmk}

This is a generalization of Definition \ref{ripdef} in the form that if $A$ satisfies the restricted isometry property of order $k$ with restricted isometry constant $\delta_k$, than it has a restricted isometry constant of at most 
$(1 - \delta_k) / (1 + \delta_k)$. For this definition a condition like Theorem \ref{RIPsqrt2} can again be shown like in \cite{FL09}:

\begin{thm} \label{asymRIPtheorem}
If $A$ has restricted isometry ratio $\gamma_{2k} \leq 4 \sqrt{2} - 3 \approx 2.6569$, then every $k$-sparse vector is exactly recovered by (\ref{l1min}). 
\end{thm}

Again this result can be generalized to the case of imperfect data. One advantage of Defintion \ref{asymRIPdef} and Theorem \ref{asymRIPtheorem} is, that they are indifferent under scaling of the system of linear equations. If a system
$Ax=b$ is scaled to $\lambda Ax = \lambda b$, then the condition of Theorem \ref{RIPsqrt2} might no longer be valid, while in Definition \ref{asymRIPdef} we can use $\tilde{\alpha_k} = \lambda \alpha_k, \tilde{\beta_k} = \lambda \beta_k$
and therefore $\tilde{\gamma_k} = \lambda^2 \alpha_k^2 / \lambda^2 \beta_k^2 = \gamma_k$. \\
In the following we will try to compute the optimal constants $\alpha_k$ and $\beta_k$ in (\ref{asymRIP}). These can then be used to compute either the restricted isometry ratio $\gamma_k := \beta_k^2 / \alpha_k^2$ or the restricted
isometry constant $\delta_k = \text{max} \{ 1 - \alpha_k^2, \beta_k^2 - 1 \}$.

\section{Previous SDP relaxations}

In this section we will discuss two known SDP relaxations to compute an upper bound on the restricted isometry constant $\delta_k$ in (\ref{rip}), as they have some ideas in common with the mixed integer semidefinite program we want to 
propose and will also be checked against in the section on numerical results. The idea of these two relaxations as well as the MISDP in the next section is that the optimal constant $\alpha_k$ in (\ref{asymRIP}) can be computed
with the non-convex quadratic optimization problem
\begin{align}\label{QP}
 \text{min} \quad & \|Ax\|_2^2 \nonumber \\
 \text{s.t.} \quad & \|x\|_2^2 = 1 \tag{QP} \\
 & \|x\|_0 \leq k. \nonumber
\end{align}
The corresponding maximization problem can be solved to compute $\beta_k$. In \cite{Asp07} the non-convex quadratic equality constraint $\|x\|_2^2 = 1$ is tackled by the technique of semidefinite lifting first introduced in \cite{GW95} 
for the max-cut problem. The idea is to use a new matrix variable $X=xx^T$ for the quadratic terms, where the condition $X=xx^T$ is then enforced by the equivalent constraints $X \succeq 0$ and $\textbf{Rk}(X) = 1$, where 
$X \succeq 0$ means that $X$ is symmetric and positive semidefinite. In the next step the non-convex rank constraint is then relaxed and the inequality $\|x\|_0 \leq k$ is substituted by the coarser constraint $1^T|X|1 \leq k$ to form
the relaxation
\begin{align}\label{Asp07}
 \text{min} \quad & \textbf{Tr}(A^TAX) \nonumber \\
 \text{s.t.} \quad & \textbf{Tr}(X) = 1 \nonumber \\
 & 1^T|X|1 \leq k \tag{Asp07} \\
 & X \succeq 0. \nonumber
\end{align}

In \cite{Asp08} instead the problem

\begin{equation}\label{phi}
 \phi(\rho) = \max_{\|x\|_2 \leq 1} x^T \Sigma x - \rho \|x\|_0
\end{equation}
is tackled for a symmetric positive semidefinite matrix $\Sigma$, which for $\Sigma = A^TA$, by duality, gives an upper bound on $\beta_k$ via

\begin{equation*}
 \beta_k^2 \leq \inf_{\rho \geq 0} \phi(\rho) + \rho k.
\end{equation*}
To solve (\ref{phi}) again the semidefinite lifting technique is used to compute an upper bound on $\phi$ via
\begin{align}\label{Asp08}
 \text{max} \quad & \sum_{i=1}^n\textbf{Tr}(P_iB_i) \nonumber \\
 \text{s.t.} \quad & \textbf{Tr}(X) = 1 \nonumber \\
 & X \succeq 0 \tag{Asp08} \\
 & X \succeq P_i \succeq 0 \quad \forall i \leq n, \nonumber
\end{align}
for $B_i = b_ib_i^T - \rho I,$ where $b_i$ is the $i$-th column of a squareroot of $\Sigma$. To compute the parameter $\rho$, the authors of \cite{Asp07} propose to minimize over the interval
\begin{equation*}
 \max_{i \notin I} (b_i^Tx)^2 \leq \rho \leq  \min_{i \in I} (b_i^Tx)^2
\end{equation*}
the convex function
\begin{equation*}
 \text{gap}(\rho) = \lambda_{\max} \left( \sum_{i=1}^n Y_i \right) - \sum_{i \in I} ((b_i^Tx)^2 - \rho),
\end{equation*}
where $I$ is a given sparsity pattern, found by some heuristic for example, $x$ the largest eigenvector of $\sum_{i \in I} b_i b_i^T$ and
\begin{equation*}
 Y_i = \begin{cases}\frac{B_ixx^TB_i}{x^TB_ix}, & \text{for } i \in I \\ \max\{0, \rho \frac{b_i^Tb_i - \rho}{\rho - (b_i^Tx)^2} \} \frac{(I-xx^T)b_ib_i^T(I-xx^T)}{\|(I-xx^T)b_i\|_2^2}, & \text{for } i \in I^c\end{cases}. 
 %x in the second case sill as above?
\end{equation*}

\section{A MISDP for the restricted isometry property}

In this section we want to compute the optimal constants $\alpha_k$ and $\beta_k$ in (\ref{asymRIP}) using mixed integer semidefinite programming. We will again start with the non-convex quadratic problem
\begin{align}\label{QP2}
 \text{min} \quad & \|Ax\|_2^2 \nonumber \\
 \text{s.t.} \quad & \|x\|_2^2 = 1 \tag{QP} \\
 & \|x\|_0 \leq k. \nonumber
\end{align}

For the non-convex quadratic equality constraint $\|x\|_2^2 = 1$ we will again use semidefinite lifting as $\|Ax\|_2^2 = x^TA^TAx = \textbf{Tr}(A^TAX)$ and $\textbf{Tr}(X) = \sum_{i=1}^n x_i^2$. 
The cardinality-constraint in (\ref{QP}) can equivalently be written using binary variables $z_i$ as $-z_i \leq x_i \leq z_i$ and $\sum_{i=1}^n z_i \leq k$. We can thus rewrite (\ref{QP}) equivalently to
% In this section we want to compute the optimal constants $\alpha_k$ and $\beta_k$ in (\ref{asymRIP}) using mixed integer semidefinite programming. A non-convex quadratic optimization problem to compute $\alpha_k$ can be formulated as
% 
% \begin{align}\label{QP}
%  \text{min} \quad & \|Ax\|_2^2 \nonumber \\
%  \text{s.t.} \quad & \|x\|_2^2 = 1 \tag{QP} \\
%  & \|x\|_0 \leq k. \nonumber
% \end{align}
% 
% By switching $\text{min}$ for $\text{max}$ in (\ref{QP}) we also get a corresponding optimization problem to compute $\beta_k$ and therefore also $\gamma_k$ and $\delta_k$.
% The cardinality-constraint in (\ref{QP}) can equivalently be written using binary variables $z_i$ as $-z_i \leq x_i \leq z_i$ and $\sum_{i=1}^n z_i \leq k$. For the non-convex quadratic equality constraint $\|x\|_2^2 = 1$ we will
% use the technique of semidefinite lifting first introduced in \cite{GW95} for the max-cut problem. The idea is to use a new matrix variable $X=xx^T$ for the quadratic terms, where the condition $X=xx^T$ is then enforced by the 
% equivalent constraints $X \succeq 0$ and $\textbf{Rk}(X) = 1$, where $X \succeq 0$ means that $X$ is symmetric and positive semidefinite. Using that $\|Ax\|_2^2 = x^TA^TAx = \textbf{Tr}(A^TAX)$ and 
% $\textbf{Tr}(X) = \sum_{i=1}^n x_i^2$, we can thus rewrite (\ref{QP}) equivalently to
\begin{align}\label{Rk1MISDP}
 \text{min} \quad & \textbf{Tr}(A^TAX) \nonumber \\
 \text{s.t.} \quad & \textbf{Tr}(X) = 1 \nonumber \\
 & -z_j \leq X_{ij} \leq z_j \quad \forall j \leq n \nonumber \\
 & \sum_{i=1}^n z_i \leq k \tag{Rk1MISDP} \\
 & \textbf{Rk}(X) = 1 \nonumber \\
 & X \succeq 0  \nonumber \\
 & z \in \{0,1\}^n \nonumber
\end{align}

The next step would then be to relax the non-convex rank constraint again to arrive at the following relaxation which computes a lower bound on the optimal constant $\alpha_k$ in (\ref{asymRIP}):
\begin{align}\label{MISDP}
 \text{min} \quad & \textbf{Tr}(A^TAX) \nonumber \\
 \text{s.t.} \quad & \textbf{Tr}(X) = 1 \nonumber \\
 & -z_j \leq X_{ij} \leq z_j \quad \forall j \leq n \nonumber \\
 & \sum_{i=1}^n z_i \leq k \tag{MISDP} \\
 & X \succeq 0  \nonumber \\
 & z \in \{0,1\}^n \nonumber
\end{align}

In the following we want to show that (\ref{MISDP}) not only computes a lower bound but the exact value of $\alpha_k$. For this we will use results about the geometry of semidefinite programs to show that there always exists an 
optimal solution to (\ref{MISDP}) with rank 1 so that the optimal objective values of (\ref{Rk1MISDP}) and (\ref{MISDP}) have to agree. We will use the following result from \cite{pat98} about the relationship between the rank of
solutions of semidefinite programs and the dimension of corresponding faces, meaning a subset $F$ of the convex set $P$ such that $x \in F, y, z \in P$ and $x = 0.5y + 0.5z$ imply $y,z \in F$:

\begin{thm}\label{patakiLemma}
 Let $X \in F$, where $F$ is a face of
 \begin{equation*}
 P := \{X \in S_n : X \succeq 0, \textbf{Tr}(A_i X) = b_i \ \forall i \leq M\}
 \end{equation*}
 for symmetric matrices $A_i \in S_n, i \leq m$. Then
 \begin{equation*}
  \frac{1}{2}\textbf{Rk}(X)\cdot(\textbf{Rk}(X)+1) \leq m + \textbf{dim}(F).
 \end{equation*}
\end{thm}

Now we will use this theorem on the projection of the feasible set of (\ref{MISDP}) onto the space of all indices $i$, such that $z_i = 1$ for a given optimal solution of (\ref{MISDP}), to show that there always exists an optimal 
solution to (\ref{MISDP}) with rank one.

\begin{thm}\label{Rk1thm}
 For every $A \in \mathbb{R}^{m \times n}$ and $k \geq 1$ there exists an optimal solution $X^*, z^*$ to (\ref{MISDP}) such that $\textbf{Rk}(X^*) = 1$. 
\end{thm}
\begin{proof}
 Let $\hat{X}, \hat{z}$ be an optimal solution to (\ref{MISDP}) (because of $\textbf{Tr}(X) = 1$ the problem is bounded and $X_{11} = 1, X_{ij} = 0 \ \forall i,j \neq 1, z_1 = 1, z_i = 0 \ \forall i \neq 0$ is always feasible)
 with $\sum_{i=1}^n z_i = \ell \leq k$. Define
 \begin{align}\label{Proj}
 \text{min} \quad & \textbf{Tr}(\tilde{A}^T\tilde{A}X) \nonumber \\
 \text{s.t.} \quad & \textbf{Tr}(X) = 1 \tag{Proj} \\
 & X \succeq 0 , X \in S_\ell, \nonumber
 \end{align}
where $\tilde{A} = (A_{i_1}, A_{i_2}, ..., A_{i_\ell})$ consists of those columns $\{{i_1}, {i_2}, ..., {i_\ell}\} =: T$ of $A$ for which $\hat{z}_{i_k} = 1$. \\
Now let $\tilde{X}$ be the matrix consisting of those rows and columns 
$i_k$ of $\hat{X}$ such that $\hat{z}_{i_k} = 1$. Then $\hat{X}$ is feasible for (\ref{Proj}) with the same objective value as the optimal solution $\hat{X}$ of (\ref{MISDP}), as we only removed zero rows and columns of $\hat{X}$ 
and $(A^TA)\hat{X}$, which do not influence the trace or the positive semidefiniteness. \\
On the other hand we can lift any optimal solution $\tilde{X}$ of (\ref{Proj}) to a feasible point for (\ref{MISDP}) with identical objective value by setting $\bar{z}_i$ to one if and only if $i \in T$
and $\bar{X}_{i_j i_k} = \tilde{X}_{jk}$ if and only if $i_j, i_k \in T$ and $\bar{X}_{ij}=0$ otherwise. \\
So the optimal objective values of (\ref{Proj}) and (\ref{MISDP}) agree and with this construction we also get that 
$\textbf{Rk}(\bar{X}) = \textbf{Rk}(\tilde{X})$, so it suffices to show that (\ref{Proj}) always allows for a solution of rank one.
Let $\breve{X}$ be an extreme point of (\ref{Proj}) (as the set of optimal points of (\ref{Proj}) is a non-empty, convex, compact subset of $\mathbb{R}^{\ell(\ell+1)/2}$ such a point exists as a consequence of corollary 18.5.1 in
\cite{Roc70}) with corresponding face $F=\{\breve{X}\}$. Then using theorem \ref{patakiLemma} on (\ref{Proj}) with $\tilde{m} = 1$ LP-constraint gives
\begin{equation*}
 \frac{1}{2}\textbf{Rk}(\breve{X})\cdot(\textbf{Rk}(\breve{X})+1) \leq \tilde{m} + \textbf{dim}(F) = 1 + 0 = 1.
\end{equation*}
As $\breve{X}=0$ is infeasible for $\textbf{Tr}(X)=1$, we must have $\textbf{Rk}(\breve{X}) = 1$, which can then be lifted to an optimal solution of (\ref{MISDP}) with rank one.
\end{proof}

As Theorem \ref{Rk1thm} tells us that (\ref{MISDP}) always has a solution of rank one and the rank-constraint was the only constraint we relaxed to get (\ref{MISDP}) from (\ref{QP}), this means that their optimal objective values 
have to agree and we can compute the optimal constant $\alpha_k$ for the RIP by solving (\ref{MISDP}). As the same argumentation also holds for the maximization problem for $\beta_k$, this allows us to compute the restricted isometry
constant and the restrcited isometry ratio by mixed integer semidefinite programming.

\section{Numerical Results}

For solving \ref{MISDP} we used the MISDP-Solver SCIP\_SDP originally developed in \cite{MS12} and \cite{Mar13}, which combines the branch-and-bound-framework of SCIP \cite{SCIP} with interior-point-SDP-solvers, in this case 
with DSDP \cite{DSDP}.

\section*{Acknowledgments}
The authors would like to thank the German Research Foundation (DFG) for funding as parts of this research have been carried out within the Collaborative Research Center 805.

\section*{References}
\bibliographystyle{abbrv}
\bibliography{Paperbib}
\end{document}

### 2014-12-18 RIP-MISDP-Paper/Paper_v1.tex
added macros, replaces \mathbb{R} by \R (from dsfonts)

@@
 \usepackage{amsthm}
-\usepackage{amsfonts}
+\usepackage{dsfont}
 \newtheorem{thm}{Theorem}
 \newdefinition{rmk}{Definition}
+
+\DeclareMathOperator{\argmin}{argmin}
+\DeclareMathOperator{\argmax}{argmax}
+\DeclareMathOperator{\supp}{supp}
+
+\newcommand{\suchthat}{\,:\,}
+\newcommand{\abs}[1]{\lvert{#1}\rvert}
+\newcommand{\card}[1]{\lvert{#1}\rvert}
+\newcommand{\define}{\coloneqq}
+\newcommand{\norm}[1]{\lVert{#1}\rVert}
+\newcommand{\Norm}[2]{\lVert{#1}\rVert_{#2}}
+\newcommand{\skal}[2]{\langle{#1},{#2}\rangle}
+\newcommand{\order}[1]{O\left({#1}\right)}
+\newcommand{\T}{^{\top}}
+
+% sets
+\newcommand{\R}{\mathds{R}}
+\newcommand{\Z}{\mathds{Z}}
+\newcommand{\ones}{\mathds{1}}
 
@@
 Finding the sparsest solution of an underdetermined system of linear equations is a problem that has been the subject of much recent research in compressed sensing.
-Given a matrix $A \in \mathbb{R}^{m \times n}$ with $m < n$ and
-right-hand side $b \in \mathbb{R}^m$, we are interested in a solution of 
+Given a matrix $A \in \R^{m \times n}$ with $m < n$ and
+right-hand side $b \in \R^m$, we are interested in a solution of 
 \begin{align}\label{l0min}
@@
 \begin{rmk} \label{ripdef}
-The \emph{restricted isometry property (RIP)} of order $k$ holds for a matrix $A \in \mathbb{R}^{m \times n}$, if there exists a minimal \emph{restricted isometry constant} $\delta_k$, such that for all $x \in \Sigma_k :=
-\{x \in \mathbb{R}^n : \|x\|_0 \leq k \}$ we have
+The \emph{restricted isometry property (RIP)} of order $k$ holds for a matrix $A \in \R^{m \times n}$, if there exists a minimal \emph{restricted isometry constant} $\delta_k$, such that for all $x \in \Sigma_k :=
+\{x \in \R^n : \|x\|_0 \leq k \}$ we have
  \begin{equation}\label{rip}
@@
 \begin{thm} \label{RIPsqrt2}
- If $A$ satisfies $\delta_{2k} < \sqrt{2}-1$, then for every $x \in \mathbb{R}^n$
+ If $A$ satisfies $\delta_{2k} < \sqrt{2}-1$, then for every $x \in \R^n$
  \begin{equation*}\label{l1ErrorRIP}
@@
 \begin{thm}\label{Rk1thm}
- For every $A \in \mathbb{R}^{m \times n}$ and $k \geq 1$ there exists an optimal solution $X^*, z^*$ to (\ref{MISDP}) such that $\textbf{Rk}(X^*) = 1$. 
+ For every $A \in \R^{m \times n}$ and $k \geq 1$ there exists an optimal solution $X^*, z^*$ to (\ref{MISDP}) such that $\textbf{Rk}(X^*) = 1$. 
 \end{thm}
@@
 $\textbf{Rk}(\bar{X}) = \textbf{Rk}(\tilde{X})$, so it suffices to show that (\ref{Proj}) always allows for a solution of rank one.
-Let $\breve{X}$ be an extreme point of (\ref{Proj}) (as the set of optimal points of (\ref{Proj}) is a non-empty, convex, compact subset of $\mathbb{R}^{\ell(\ell+1)/2}$ such a point exists as a consequence of corollary 18.5.1 in
+Let $\breve{X}$ be an extreme point of (\ref{Proj}) (as the set of optimal points of (\ref{Proj}) is a non-empty, convex, compact subset of $\R^{\ell(\ell+1)/2}$ such a point exists as a consequence of corollary 18.5.1 in
 \cite{Roc70}) with corresponding face $F=\{\breve{X}\}$. Then using theorem \ref{patakiLemma} on (\ref{Proj}) with $\tilde{m} = 1$ LP-constraint gives

### 2014-12-18 RIP-MISDP-Paper/Paper_v1.tex
try \ones and replace ^T by \T

@@
 for the max-cut problem.
-The idea is to use a new matrix variable $X=xx^T$ for the quadratic terms, where the condition $X=xx^T$ is then enforced by the equivalent constraints $X \succeq 0$ and $\textbf{Rk}(X) = 1$, where
+The idea is to use a new matrix variable $X=xx\T$ for the quadratic terms, where the condition $X=xx\T$ is then enforced by the equivalent constraints $X \succeq 0$ and $\textbf{Rk}(X) = 1$, where
 $X \succeq 0$ means that $X$ is symmetric and positive semidefinite.
-In the next step the non-convex rank constraint is then relaxed and the inequality $\|x\|_0 \leq k$ is substituted by the coarser constraint $1^T|X|1 \leq k$ to form
+In the next step the non-convex rank constraint is then relaxed and the inequality $\|x\|_0 \leq k$ is substituted by the coarser constraint $1\T|X|1 \leq k$ to form
 the relaxation
 \begin{align}\label{Asp07}
- \text{min} \quad & \textbf{Tr}(A^TAX) \nonumber \\
+ \text{min} \quad & \textbf{Tr}(A\T A X) \nonumber \\
  \text{s.t.} \quad & \textbf{Tr}(X) = 1 \nonumber \\
- & 1^T|X|1 \leq k \tag{Asp07} \\
+ & \ones\T|X|\ones \leq k \tag{Asp07} \\
  & X \succeq 0. \nonumber
@@
 \begin{equation}\label{phi}
- \phi(\rho) = \max_{\|x\|_2 \leq 1} x^T \Sigma x - \rho \|x\|_0
+ \phi(\rho) = \max_{\|x\|_2 \leq 1} x\T \Sigma x - \rho \|x\|_0
 \end{equation}
-is tackled for a symmetric positive semidefinite matrix $\Sigma$, which for $\Sigma = A^TA$, by duality, gives an upper bound on $\beta_k$ via
+is tackled for a symmetric positive semidefinite matrix $\Sigma$, which for
+$\Sigma = A\T A$, by duality, gives an upper bound on $\beta_k$ via
 
@@
 \end{align}
-for $B_i = b_ib_i^T - \rho I,$ where $b_i$ is the $i$-th column of a squareroot of $\Sigma$. To compute the parameter $\rho$, the authors of \cite{Asp07} propose to minimize over the interval
+for $B_i = b_ib_i\T - \rho I,$ where $b_i$ is the $i$-th column of a squareroot of $\Sigma$. To compute the parameter $\rho$, the authors of \cite{Asp07} propose to minimize over the interval
 \begin{equation*}
- \max_{i \notin I} (b_i^Tx)^2 \leq \rho \leq  \min_{i \in I} (b_i^Tx)^2
+ \max_{i \notin I} (b_i\T x)^2 \leq \rho \leq  \min_{i \in I} (b_i\T x)^2
 \end{equation*}
@@
 \begin{equation*}
- \text{gap}(\rho) = \lambda_{\max} \left( \sum_{i=1}^n Y_i \right) - \sum_{i \in I} ((b_i^Tx)^2 - \rho),
+ \text{gap}(\rho) = \lambda_{\max} \left( \sum_{i=1}^n Y_i \right) -
+ \sum_{i \in I} ((b_i\T x)^2 - \rho),
 \end{equation*}
-where $I$ is a given sparsity pattern, found by some heuristic for example, $x$ the largest eigenvector of $\sum_{i \in I} b_i b_i^T$ and
+where $I$ is a given sparsity pattern, found by some heuristic for example, $x$ the largest eigenvector of $\sum_{i \in I} b_i b_i\T$ and
 \begin{equation*}
- Y_i = \begin{cases}\frac{B_ixx^TB_i}{x^TB_ix}, & \text{for } i \in I \\ \max\{0, \rho \frac{b_i^Tb_i - \rho}{\rho - (b_i^Tx)^2} \} \frac{(I-xx^T)b_ib_i^T(I-xx^T)}{\|(I-xx^T)b_i\|_2^2}, & \text{for } i \in I^c\end{cases}.
+ Y_i = \begin{cases}\frac{B_ixx\T B_i}{x\T B_ix}, & \text{for } i \in I \\
+   \max\{0, \rho \frac{b_i\T b_i - \rho}{\rho - (b_i\T x)^2} \} \frac{(I-xx\T)b_ib_i\T(I-xx\T)}{\|(I-xx\T)b_i\|_2^2}, & \text{for } i \in I^c\end{cases}. 
  %x in the second case sill as above?
@@
 
-For the non-convex quadratic equality constraint $\|x\|_2^2 = 1$ we will again use semidefinite lifting as $\|Ax\|_2^2 = x^TA^TAx = \textbf{Tr}(A^TAX)$ and $\textbf{Tr}(X) = \sum_{i=1}^n x_i^2$. 
+For the non-convex quadratic equality constraint $\|x\|_2^2 = 1$ we will
+again use semidefinite lifting as $\|Ax\|_2^2 = x\T A\T Ax =
+\textbf{Tr}(A\T A X)$ and $\textbf{Tr}(X) = \sum_{i=1}^n x_i^2$. 
 The cardinality-constraint in (\ref{QP}) can equivalently be written using binary variables $z_i$ as $-z_i \leq x_i \leq z_i$ and $\sum_{i=1}^n z_i \leq k$.
@@
 % use the technique of semidefinite lifting first introduced in \cite{GW95} for the max-cut problem.
-The idea is to use a new matrix variable $X=xx^T$ for the quadratic terms, where the condition $X=xx^T$ is then enforced by the
-% equivalent constraints $X \succeq 0$ and $\textbf{Rk}(X) = 1$, where $X \succeq 0$ means that $X$ is symmetric and positive semidefinite. Using that $\|Ax\|_2^2 = x^TA^TAx = \textbf{Tr}(A^TAX)$ and 
+The idea is to use a new matrix variable $X=xx\T$ for the quadratic terms, where the condition $X=xx\T$ is then enforced by the
+% equivalent constraints $X \succeq 0$ and $\textbf{Rk}(X) = 1$, where $X
+% \succeq 0$ means that $X$ is symmetric and positive semidefinite. Using
+% that $\|Ax\|_2^2 = x\T A\T Ax = \textbf{Tr}(A\T A X)$ and 
 % $\textbf{Tr}(X) = \sum_{i=1}^n x_i^2$, we can thus rewrite (\ref{QP}) equivalently to
 \begin{align}\label{Rk1MISDP}
- \text{min} \quad & \textbf{Tr}(A^TAX) \nonumber \\
+ \text{min} \quad & \textbf{Tr}(A\T A X) \nonumber \\
  \text{s.t.} \quad & \textbf{Tr}(X) = 1 \nonumber \\
@@
 \begin{align}\label{MISDP}
- \text{min} \quad & \textbf{Tr}(A^TAX) \nonumber \\
+ \text{min} \quad & \textbf{Tr}(A\T A X) \nonumber \\
  \text{s.t.} \quad & \textbf{Tr}(X) = 1 \nonumber \\
@@
  \begin{align}\label{Proj}
- \text{min} \quad & \textbf{Tr}(\tilde{A}^T\tilde{A}X) \nonumber \\
+ \text{min} \quad & \textbf{Tr}(\tilde{A}\T\tilde{A}X) \nonumber \\
  \text{s.t.} \quad & \textbf{Tr}(X) = 1 \tag{Proj} \\
@@
 Then $\hat{X}$ is feasible for (\ref{Proj}) with the same objective value as the optimal solution $\hat{X}$ of (\ref{MISDP}), as we only removed zero rows and columns of $\hat{X}$
-and $(A^TA)\hat{X}$, which do not influence the trace or the positive semidefiniteness. \\
+and $(A\T A)\hat{X}$, which do not influence the trace or the positive semidefiniteness. \\
 On the other hand we can lift any optimal solution $\tilde{X}$ of (\ref{Proj}) to a feasible point for (\ref{MISDP}) with identical objective value by setting $\bar{z}_i$ to one if and only if $i \in T$

### 2014-12-18 RIP-MISDP-Paper/Paper_v1.tex
minor changes

@@
 \documentclass{elsarticle}
+
 \usepackage{amsmath}
@@
 \usepackage{dsfont}
+
 \newtheorem{thm}{Theorem}
@@
 \begin{document}
+
 \title{Computing exact restricted isometry constants via mixed integer semidefinite programming}
@@
 \address[tud]{Technische Universit\"at Darmstadt, Dolivostra\ss e 15, 64293 Darmstadt, Germany}
+
 \begin{abstract}
@@
 \end{abstract}
+
 \maketitle
+
 \section{Introduction}

### 2014-12-18 RIP-MISDP-Paper/Paper_v1.tex
remove \\

@@
 Even tough it was shown in \cite{BDDW08}, that this condition holds for multiple classes of random matrices with probability exponentially going to one, in \cite{PT14} it was shown, that computing the optimal restricted isometry
-constants in (\ref{RIPintro}) for a given matrix $A$ is again an $\mathcal{NP}$-hard task. \\
+constants in (\ref{RIPintro}) for a given matrix $A$ is again an
+$\mathcal{NP}$-hard task.
+
 In the past SDP-Relaxations were proposed in \cite{Asp07} and \cite{Asp08} to compute upper bounds on the values of the restricted isometry 
@@
 In this paper we will construct a mixed integer semidefinite program to compute the exact constants in (\ref{RIPintro}), to be able to verify in all cases if those sufficient conditions for the equivalence hold for a given matrix.
-The main contribution of this paper is theorem \ref{Rk1thm}, which shows that the optimal solution of that MISDP indeed allows us to compute the exact restricted isometry constant.\\
+The main contribution of this paper is theorem \ref{Rk1thm}, which shows
+that the optimal solution of that MISDP indeed allows us to compute the
+exact restricted isometry constant.
+
 In the following we will start by introducing the restricted isometry property and corresponding results about the equivalence of (\ref{l0min}) and (\ref{l1min}).
@@
 $Ax=b$ is scaled to $\lambda Ax = \lambda b$, then the condition of Theorem \ref{RIPsqrt2} might no longer be valid, while in Definition \ref{asymRIPdef} we can use $\tilde{\alpha_k} = \lambda \alpha_k, \tilde{\beta_k} = \lambda \beta_k$
-and therefore $\tilde{\gamma_k} = \lambda^2 \alpha_k^2 / \lambda^2 \beta_k^2 = \gamma_k$. \\
+and therefore $\tilde{\gamma_k} = \lambda^2 \alpha_k^2 / \lambda^2
+\beta_k^2 = \gamma_k$.
+
 In the following we will try to compute the optimal constants $\alpha_k$ and $\beta_k$ in (\ref{asymRIP}).
@@
  \end{align}
-where $\tilde{A} = (A_{i_1}, A_{i_2}, ..., A_{i_\ell})$ consists of those columns $\{{i_1}, {i_2}, ..., {i_\ell}\} =: T$ of $A$ for which $\hat{z}_{i_k} = 1$. \\
+where $\tilde{A} = (A_{i_1}, A_{i_2}, ..., A_{i_\ell})$ consists of those
+columns $\{{i_1}, {i_2}, ..., {i_\ell}\} =: T$ of $A$ for which
+$\hat{z}_{i_k} = 1$.
+
 Now let $\tilde{X}$ be the matrix consisting of those rows and columns 
@@
 Then $\hat{X}$ is feasible for (\ref{Proj}) with the same objective value as the optimal solution $\hat{X}$ of (\ref{MISDP}), as we only removed zero rows and columns of $\hat{X}$
-and $(A\T A)\hat{X}$, which do not influence the trace or the positive semidefiniteness. \\
+and $(A\T A)\hat{X}$, which do not influence the trace or the positive
+semidefiniteness.
+
 On the other hand we can lift any optimal solution $\tilde{X}$ of (\ref{Proj}) to a feasible point for (\ref{MISDP}) with identical objective value by setting $\bar{z}_i$ to one if and only if $i \in T$
-and $\bar{X}_{i_j i_k} = \tilde{X}_{jk}$ if and only if $i_j, i_k \in T$ and $\bar{X}_{ij}=0$ otherwise. \\
+and $\bar{X}_{i_j i_k} = \tilde{X}_{jk}$ if and only if $i_j, i_k \in T$
+and $\bar{X}_{ij}=0$ otherwise.
+
 So the optimal objective values of (\ref{Proj}) and (\ref{MISDP}) agree and with this construction we also get that 

### 2014-12-22 RIP-MISDP-Paper/Paper_v1.tex
try some documentclass options
added the section on numerical results to the paper and made some general improvements

@@
-\documentclass{elsarticle}
+\documentclass[a4paper,11pt,1p]{elsarticle}
 
@@
 \usepackage{dsfont}
+\usepackage{booktabs}
+\usepackage{multirow}
 
@@
 \newcommand{\T}{^{\top}}
+\newcommand{\Tr}{\textbf{Tr}}
+\newcommand{\Rk}{\textbf{Rk}}
 
@@
 \ead{pfetsch@mathematik.tu-darmstadt.de}
-\address[tud]{Technische Universit\"at Darmstadt, Dolivostra\ss e 15, 64293 Darmstadt, Germany}
+\address[tud]{Technische Universit\"at Darmstadt, Department of Mathematics, Dolivostra\ss e 15, 64293 Darmstadt, Germany}
 
@@
 \begin{align}\label{l0min}
- \text{min} \quad & \|x\|_0 \\
+ \text{min} \quad & \|x\|_0 \tag{$\ell_0$-min}\\
  \text{s.t.} \quad & Ax = b, \nonumber
@@
 \begin{align}\label{l1min}
- \text{min} \quad & \|x\|_0 \\
- \text{s.t.} \quad & Ax = b, \nonumber
+ \text{min} \quad & \|x\|_1 \tag{$\ell_1$-min}\\
+ \text{s.t.} \quad & Ax = b \nonumber
 \end{align}
@@
 \end{equation}
-that was first introduced in \cite{CT05}, if it holds for all $k$-sparse vectors $x$ for a sufficiently small restricted isometry constant $\delta_k$, then problems (\ref{l0min}) and (\ref{l1min}) have the same optimal solution.
-Even tough it was shown in \cite{BDDW08}, that this condition holds for multiple classes of random matrices with probability exponentially going to one, in \cite{PT14} it was shown, that computing the optimal restricted isometry
-constants in (\ref{RIPintro}) for a given matrix $A$ is again an
-$\mathcal{NP}$-hard task.
+that was first introduced in \cite{CT05}.
+If it holds for all $k$-sparse vectors $x$ for a sufficiently small restricted isometry constant $\delta_k$, then problems (\ref{l0min}) and (\ref{l1min}) have the same optimal solution.
+Even though it was shown in \cite{BDDW08}, that this condition holds for multiple classes of random matrices with probability exponentially going to one, in \cite{PT14} it was shown, that computing the optimal restricted isometry
+constants in (\ref{RIPintro}) for a given matrix $A$ is again an $\mathcal{NP}$-hard task.
 
@@
 The main contribution of this paper is theorem \ref{Rk1thm}, which shows
-that the optimal solution of that MISDP indeed allows us to compute the
+that the optimal solution of this MISDP indeed allows us to compute the
 exact restricted isometry constant.
@@
 
-In this section we will discuss some general results about the restricted isometry property and it's relationship with the optimization problems (\ref{l0min}) and (\ref{l1min}).
+In this section we will discuss some general results about the restricted isometry property and its relationship with the optimization problems (\ref{l0min}) and (\ref{l1min}).
 To start, we will properly define the restricted
@@
  If the restricted isometry property holds for a sufficiently small restricted isometry constant, then the optimal solutions of (\ref{l0min}) and (\ref{l1min}) coincide.
-The best known result of this type is probably the following from
- \cite{Can08}:
+To show this, one needs estimates of the form
+  \begin{equation}\label{l1ErrorRIP}
+  \|x^* - x \| \leq C \|x - x_k \|,
+ \end{equation}
+ where $x^*$ is an optimal solution of (\ref{l1min}) with $b = Ax$ and $x_k$ is the best approximation of some vector $x \in \R^n$ with $k$-sparse vectors.
+If we get a result like (\ref{l1ErrorRIP}), and we know that the vector $x$, which in the
+ applications would be the sparse signal that needs to be recovered, is already $k$-sparse itself, then the best approximation of $x$ with $k$-sparse vectors is clearly $x$ itself.
+This means that the right-hand-side in
+ (\ref{l1ErrorRIP}) equals zero, and therefore the unique optimal solution of (\ref{l1min}) has to be $x$ again.
+Now (\ref{l1ErrorRIP}) not only holds for this $x$ exclusively, but for any $k$-sparse vector $y$ with $Ay=Ax=b$.
+And
+ because of $Ay=Ax=b$, we get the same problem in (\ref{l1min}) and therefore also the same solution $x^*$ which needs to equal $y$ because of (\ref{l1ErrorRIP}), so this means that there can only be a single vector $x$ with sparsity
+ at most $k$ for which $Ax=b$ holds, which then also has to solve (\ref{l0min}), which gives us the equivalance of these two problems.
+
+ The best-known result of this type is probably the following from \cite{Can08}:
 
@@
  If $A$ satisfies $\delta_{2k} < \sqrt{2}-1$, then for every $x \in \R^n$
- \begin{equation*}\label{l1ErrorRIP}
+ \begin{equation*}
   \|x^* - x \|_{\ell_1} \leq C \|x - x_k \|_{\ell_1},
@@
 where $x^*$ is an optimal solution for (\ref{l1min}) with right-hand side $b = Ax$, $x_k$ is the best approximation of $x$ with $k$-sparse vectors and $C$ is a constant depending only on $k$ and $\delta_{2k}$.
-If $x$ is $k$-sparse, then the right-hand side of
-(\ref{l1ErrorRIP}) is zero, so $x$ is the only $k$-sparse vector with $A\hat{x} = Ax$ and recovered by (\ref{l1min}) and therefore the optimal solutions of (\ref{l0min}) and (\ref{l1min}) coincide.\end{thm}
+If $x$ is $k$-sparse,
+then the optimal solutions of (\ref{l0min}) and (\ref{l1min}) coincide.\end{thm}
 
@@
 Also all of these results can be
-generalized to the case where $\|Ax - b\|_{\ell_2} \leq \epsilon$. The restricted isometry property can also be formulated in an asymetric way like in \cite{FL09}:
+generalized to the case where $\|Ax - b\|_{\ell_2} \leq \epsilon$. The restricted isometry property can also be formulated in an asymmetric way like in \cite{FL09}:
 
@@
 
-This is a generalization of Definition \ref{ripdef} in the form that if $A$ satisfies the restricted isometry property of order $k$ with restricted isometry constant $\delta_k$, than it has a restricted isometry constant of at most
-$(1 - \delta_k) / (1 + \delta_k)$. For this definition a condition like Theorem \ref{RIPsqrt2} can again be shown like in \cite{FL09}:
+This is a generalization of Definition \ref{ripdef} in the form that if $A$ satisfies the restricted isometry property of order $k$ with restricted isometry constant $\delta_k$, then it has a restricted isometry constant of at most
+$(1 - \delta_k) / (1 + \delta_k)$. Also in the other direction, if $A$ satisfies (\ref{asymRIP}) with constants $\alpha_k$ and $\beta_k$, then it will also satisfy (\ref{rip}) for 
+$\delta_k = \text{max} \{ 1 - \alpha_k^2, \beta_k^2 - 1 \}$.
+So $A$ satisfies definition \ref{asymRIPdef} if and only if it satisfies definition \ref{ripdef}, but as in (\ref{rip}) usually only one of the two inequalities will
+be sharp in contrast to (\ref{asymRIP}), the asymmetric version will be more precise. For this definition a condition like Theorem \ref{RIPsqrt2} can again be shown like in \cite{FL09}:
 
@@
 
-\section{Previous SDP relaxations}
+\section{SDP relaxations}
 
@@
 propose and will also be checked against in the section on numerical results.
-The idea of these two relaxations as well as the MISDP in the next section is that the optimal constant $\alpha_k$ in (\ref{asymRIP}) can be computed
+The idea of these two relaxations as well as the MISDP in the next section is that the optimal constant $\alpha_k^2$ in (\ref{asymRIP}) can be computed
 with the non-convex quadratic optimization problem
@@
 \end{align}
-The corresponding maximization problem can be solved to compute $\beta_k$.
+The corresponding maximization problem can be solved to compute $\beta_k^2$.
 In \cite{Asp07} the non-convex quadratic equality constraint $\|x\|_2^2 = 1$ is tackled by the technique of semidefinite lifting first introduced in \cite{GW95}
 for the max-cut problem.
-The idea is to use a new matrix variable $X=xx\T$ for the quadratic terms, where the condition $X=xx\T$ is then enforced by the equivalent constraints $X \succeq 0$ and $\textbf{Rk}(X) = 1$, where
+The idea is to use a new matrix variable $X=xx\T$ for the quadratic terms, where the condition $X=xx\T$ is then enforced by the equivalent constraints $X \succeq 0$ and $\Rk(X) = 1$, where
 $X \succeq 0$ means that $X$ is symmetric and positive semidefinite.
@@
 \begin{align}\label{Asp07}
- \text{min} \quad & \textbf{Tr}(A\T A X) \nonumber \\
- \text{s.t.} \quad & \textbf{Tr}(X) = 1 \nonumber \\
- & \ones\T|X|\ones \leq k \tag{Asp07} \\
+ \text{min} \quad & \Tr(A\T A X) \nonumber \\
+ \text{s.t.} \quad & \Tr(X) = 1 \nonumber \\
+ & \ones\T|X|\ones \leq k \tag{A1} \\
  & X \succeq 0. \nonumber
@@
 \begin{equation}\label{phi}
- \phi(\rho) = \max_{\|x\|_2 \leq 1} x\T \Sigma x - \rho \|x\|_0
+ \phi(\rho) = \max_{\|x\|_2 \leq 1} x\T \Sigma x - \rho \|x\|_0,
 \end{equation}
-is tackled for a symmetric positive semidefinite matrix $\Sigma$, which for
+originally stemming from sparse principal component analysis, is tackled for a symmetric positive semidefinite matrix $\Sigma$, which for
 $\Sigma = A\T A$, by duality, gives an upper bound on $\beta_k$ via
@@
 \end{equation*}
-To solve (\ref{phi}) again the semidefinite lifting technique is used to compute an upper bound on $\phi$ via
+
+As one way to solve (\ref{phi}) again the semidefinite lifting technique is used to compute an upper bound on $\phi$ via
 \begin{align}\label{Asp08}
- \text{max} \quad & \sum_{i=1}^n\textbf{Tr}(P_iB_i) \nonumber \\
- \text{s.t.} \quad & \textbf{Tr}(X) = 1 \nonumber \\
- & X \succeq 0 \tag{Asp08} \\
+ \text{max} \quad & \sum_{i=1}^n\Tr(P_iB_i) \nonumber \\
+ \text{s.t.} \quad & \Tr(X) = 1 \nonumber \\
+ & X \succeq 0 \tag{A2-Primal} \\
  & X \succeq P_i \succeq 0 \quad \forall i \leq n, \nonumber
 \end{align}
-for $B_i = b_ib_i\T - \rho I,$ where $b_i$ is the $i$-th column of a squareroot of $\Sigma$. To compute the parameter $\rho$, the authors of \cite{Asp07} propose to minimize over the interval
+ for $B_i = b_ib_i\T - \rho I,$ where $b_i$ is the $i$-th column of a squareroot of $\Sigma$. The corresponding dual problem can then be written as
+
+ \begin{align}\label{Asp08Dual}
+ \text{min} \quad & \lambda_\text{max}\left(\sum_{i=1}^n Y_i \right) \nonumber \\
+ \text{s.t.} \quad & Y_i \succeq B_i \quad \forall i \leq n \tag{A2-Dual} \\
+ & Y_i \succeq 0 \quad \ \ \forall i \leq n, \nonumber
+\end{align}
+which only has half as many variables as (\ref{Asp08}) and one less linear and semidefinite constraint.
+Then (\ref{Asp08Dual}) can be solved for non-negative $\rho$, which should be smaller than $\Sigma_{11}$, because otherwise the
+optimal solution for (\ref{phi}) will always be $x=0$.
+
+A lower bound for $\alpha_k$ can be computed via
+
 \begin{equation*}
- \max_{i \notin I} (b_i\T x)^2 \leq \rho \leq  \min_{i \in I} (b_i\T x)^2
+ \alpha_k^2 \geq \sup_{\rho \geq 0} \psi(\rho) - \rho k,
 \end{equation*}
-the convex function
+with 
 \begin{equation*}
- \text{gap}(\rho) = \lambda_{\max} \left( \sum_{i=1}^n Y_i \right) -
- \sum_{i \in I} ((b_i\T x)^2 - \rho),
+ \psi(\rho) = \min_{\|x\|_2 \leq 1} x\T \Sigma x + \rho \|x\|_0.
 \end{equation*}
-where $I$ is a given sparsity pattern, found by some heuristic for example, $x$ the largest eigenvector of $\sum_{i \in I} b_i b_i\T$ and
+In this case problem (\ref{Asp08}) has to be solved for $\Sigma = \alpha I - A\T A$, where $\alpha$ has to be big enough to make $\Sigma$ positive semidefinite again.
+If $\bar{\Psi}$ is the corresponding optimal solution, this implies
+a lower bound on $\alpha_k$ as
+
 \begin{equation*}
- Y_i = \begin{cases}\frac{B_ixx\T B_i}{x\T B_ix}, & \text{for } i \in I \\
-   \max\{0, \rho \frac{b_i\T b_i - \rho}{\rho - (b_i\T x)^2} \} \frac{(I-xx\T)b_ib_i\T(I-xx\T)}{\|(I-xx\T)b_i\|_2^2}, & \text{for } i \in I^c\end{cases}. 
- %x in the second case sill as above?
+ \alpha_k^2 \geq \sup_{\rho \geq 0} (M - \bar{\Psi}) - \rho k.
 \end{equation*}
+
+%To compute the parameter $\rho$, the authors of \cite{Asp07} propose to minimize over the interval
+% \begin{equation*}
+%  \max_{i \notin I} (b_i\T x)^2 \leq \rho \leq  \min_{i \in I} (b_i\T x)^2
+% \end{equation*}
+% the convex function
+% \begin{equation*}
+%  \text{gap}(\rho) = \lambda_{\max} \left( \sum_{i=1}^n Y_i \right) -
+%  \sum_{i \in I} ((b_i\T x)^2 - \rho),
+% \end{equation*}
+% where $I$ is a given sparsity pattern, found by some heuristic for example, $x$ the largest eigenvector of $\sum_{i \in I} b_i b_i\T$ and
+% \begin{equation*}
+%  Y_i = \begin{cases}\frac{B_ixx\T B_i}{x\T B_ix}, & \text{for } i \in I \\
+%    \max\{0, \rho \frac{b_i\T b_i - \rho}{\rho - (b_i\T x)^2} \} \frac{(I-xx\T)b_ib_i\T(I-xx\T)}{\|(I-xx\T)b_i\|_2^2}, & \text{for } i \in I^c\end{cases}. 
+%  %x in the second case sill as above?
+% \end{equation*}
 
@@
 again use semidefinite lifting as $\|Ax\|_2^2 = x\T A\T Ax =
-\textbf{Tr}(A\T A X)$ and $\textbf{Tr}(X) = \sum_{i=1}^n x_i^2$. 
+\Tr(A\T A X)$ and $\Tr(X) = \sum_{i=1}^n x_i^2$. 
 The cardinality-constraint in (\ref{QP}) can equivalently be written using binary variables $z_i$ as $-z_i \leq x_i \leq z_i$ and $\sum_{i=1}^n z_i \leq k$.
@@
 The idea is to use a new matrix variable $X=xx\T$ for the quadratic terms, where the condition $X=xx\T$ is then enforced by the
-% equivalent constraints $X \succeq 0$ and $\textbf{Rk}(X) = 1$, where $X
+% equivalent constraints $X \succeq 0$ and $\Rk(X) = 1$, where $X
 % \succeq 0$ means that $X$ is symmetric and positive semidefinite. Using
-% that $\|Ax\|_2^2 = x\T A\T Ax = \textbf{Tr}(A\T A X)$ and 
-% $\textbf{Tr}(X) = \sum_{i=1}^n x_i^2$, we can thus rewrite (\ref{QP}) equivalently to
+% that $\|Ax\|_2^2 = x\T A\T Ax = \Tr(A\T A X)$ and 
+% $\Tr(X) = \sum_{i=1}^n x_i^2$, we can thus rewrite (\ref{QP}) equivalently to
 \begin{align}\label{Rk1MISDP}
- \text{min} \quad & \textbf{Tr}(A\T A X) \nonumber \\
- \text{s.t.} \quad & \textbf{Tr}(X) = 1 \nonumber \\
+ \text{min} \quad & \Tr(A\T A X) \nonumber \\
+ \text{s.t.} \quad & \Tr(X) = 1 \nonumber \\
  & -z_j \leq X_{ij} \leq z_j \quad \forall j \leq n \nonumber \\
  & \sum_{i=1}^n z_i \leq k \tag{Rk1MISDP} \\
- & \textbf{Rk}(X) = 1 \nonumber \\
+ & \Rk(X) = 1 \nonumber \\
  & X \succeq 0  \nonumber \\
@@
 \begin{align}\label{MISDP}
- \text{min} \quad & \textbf{Tr}(A\T A X) \nonumber \\
- \text{s.t.} \quad & \textbf{Tr}(X) = 1 \nonumber \\
+ \text{min} \quad & \Tr(A\T A X) \nonumber \\
+ \text{s.t.} \quad & \Tr(X) = 1 \nonumber \\
  & -z_j \leq X_{ij} \leq z_j \quad \forall j \leq n \nonumber \\
@@
  \begin{equation*}
- P := \{X \in S_n : X \succeq 0, \textbf{Tr}(A_i X) = b_i \ \forall i \leq M\}
+ P := \{X \in S_n : X \succeq 0, \Tr(A_i X) = b_i \ \forall i \leq m\}
  \end{equation*}
@@
  \begin{equation*}
-  \frac{1}{2}\textbf{Rk}(X)\cdot(\textbf{Rk}(X)+1) \leq m + \textbf{dim}(F).
+  \frac{1}{2}\Rk(X)\cdot(\Rk(X)+1) \leq m + \textbf{dim}(F).
  \end{equation*}
@@
 \begin{thm}\label{Rk1thm}
- For every $A \in \R^{m \times n}$ and $k \geq 1$ there exists an optimal solution $X^*, z^*$ to (\ref{MISDP}) such that $\textbf{Rk}(X^*) = 1$. 
+ For every $A \in \R^{m \times n}$ and $k \geq 1$ there exists an optimal solution $X^*, z^*$ to (\ref{MISDP}) such that $\Rk(X^*) = 1$. 
 \end{thm}
 \begin{proof}
- Let $\hat{X}, \hat{z}$ be an optimal solution to (\ref{MISDP}) (because of $\textbf{Tr}(X) = 1$ the problem is bounded and $X_{11} = 1, X_{ij} = 0 \ \forall i,j \neq 1, z_1 = 1, z_i = 0 \ \forall i \neq 0$ is always feasible)
+ Let $\hat{X}, \hat{z}$ be an optimal solution to (\ref{MISDP}) (because of $\Tr(X) = 1$ the problem is bounded and $X_{11} = 1, X_{ij} = 0 \ \forall i,j \neq 1, z_1 = 1, z_i = 0 \ \forall i \neq 0$ is always feasible for $k > 0$,
+ so an optimal solution exists)
  with $\sum_{i=1}^n z_i = \ell \leq k$. Define
  \begin{align}\label{Proj}
- \text{min} \quad & \textbf{Tr}(\tilde{A}\T\tilde{A}X) \nonumber \\
- \text{s.t.} \quad & \textbf{Tr}(X) = 1 \tag{Proj} \\
+ \text{min} \quad & \Tr(\tilde{A}\T\tilde{A}X) \nonumber \\
+ \text{s.t.} \quad & \Tr(X) = 1 \tag{Proj} \\
  & X \succeq 0 , X \in S_\ell, \nonumber
@@
 $i_k$ of $\hat{X}$ such that $\hat{z}_{i_k} = 1$.
-Then $\hat{X}$ is feasible for (\ref{Proj}) with the same objective value as the optimal solution $\hat{X}$ of (\ref{MISDP}), as we only removed zero rows and columns of $\hat{X}$
+Then $\tilde{X}$ is feasible for (\ref{Proj}) with the same objective value as the optimal solution $\hat{X}$ of (\ref{MISDP}), as we only removed zero rows and columns of $\hat{X}$
 and $(A\T A)\hat{X}$, which do not influence the trace or the positive
@@
 So the optimal objective values of (\ref{Proj}) and (\ref{MISDP}) agree and with this construction we also get that 
-$\textbf{Rk}(\bar{X}) = \textbf{Rk}(\tilde{X})$, so it suffices to show that (\ref{Proj}) always allows for a solution of rank one.
-Let $\breve{X}$ be an extreme point of (\ref{Proj}) (as the set of optimal points of (\ref{Proj}) is a non-empty, convex, compact subset of $\R^{\ell(\ell+1)/2}$ such a point exists as a consequence of corollary 18.5.1 in
+$\Rk(\bar{X}) = \Rk(\tilde{X})$, so it suffices to show that (\ref{Proj}) always allows for a solution of rank one.
+Let $\breve{X}$ be an extreme point of (\ref{Proj}) (as the set of optimal points of (\ref{Proj}) is a non-empty, convex, compact subset of $\R^{\ell(\ell+1)/2}$, such a point exists as a consequence of corollary 18.5.1 in
 \cite{Roc70}) with corresponding face $F=\{\breve{X}\}$. Then using theorem \ref{patakiLemma} on (\ref{Proj}) with $\tilde{m} = 1$ LP-constraint gives
 \begin{equation*}
- \frac{1}{2}\textbf{Rk}(\breve{X})\cdot(\textbf{Rk}(\breve{X})+1) \leq \tilde{m} + \textbf{dim}(F) = 1 + 0 = 1.
+ \frac{1}{2}\Rk(\breve{X})\cdot(\Rk(\breve{X})+1) \leq \tilde{m} + \textbf{dim}(F) = 1 + 0 = 1.
 \end{equation*}
-As $\breve{X}=0$ is infeasible for $\textbf{Tr}(X)=1$, we must have $\textbf{Rk}(\breve{X}) = 1$, which can then be lifted to an optimal solution of (\ref{MISDP}) with rank one.
+As $\breve{X}=0$ is infeasible for $\Tr(X)=1$, we must have $\Rk(\breve{X}) = 1$, which can then be lifted to an optimal solution of (\ref{MISDP}) with rank one.
 \end{proof}
@@
 As the same argumentation also holds for the maximization problem for $\beta_k$, this allows us to compute the restricted isometry
-constant and the restrcited isometry ratio by mixed integer semidefinite programming.
+constant and the restricted isometry ratio by mixed integer semidefinite programming.
 
@@
 
-For solving \ref{MISDP} we used the MISDP-Solver SCIP\_SDP originally developed in \cite{MS12} and \cite{Mar13}, which combines the branch-and-bound-framework of SCIP \cite{SCIP} with interior-point-SDP-solvers, in this case
-with DSDP \cite{DSDP}.
+For solving \ref{MISDP} we used the MISDP-Solver SCIP\_SDP originally developed in \cite{MS12} and \cite{Mar13}, which combines the branch-and-bound-framework of SCIP 3.1 \cite{SCIP} with interior-point-SDP-solvers, in this case
+with the dual-scaling interior-point-solver DSDP 5.8 \cite{DSDP}.
+The code was overhauled and ported to C in contrast to the cited original version.
+The continuous SDPs were solved using the command-line version of DSDP.
+The testset
+consisted of 14 matrices of seven different types, and for each matrix we computed
+the left- and right-hand restricted isometry constant $\alpha$ and $\beta$ using the problems (\ref{MISDP}), (\ref{Asp07}) and (\ref{Asp08Dual}).
+In the latter case we solved 15 SDPs for $\rho$ between $0$ and the maximum diagonal
+entry of $A\T A$.
+The matrices were band matrices with band size three or five with entries within the band chosen uniformly in $\{0,1\}$, binary matrices with all entries chosen uniformly in $\{0,1\}$ as well as matrices with
+entries chosen with distribution $\mathcal{N}(0,1)$ and rank one matrices $A = aa\T$, with $\mathcal{N}(0,1)$-distributed entries of $a$.
+In addition to these four kinds of matrices, we also used three more with theoretical results
+about their restricted isometry constants, namely matrices with $N(0, 1/m)$-distributed entries, with entries chosen uniformly in $\pm 1/\sqrt{m}$ and finaly with distribution
+\begin{equation}\nonumber
+A_{ij} = \begin{cases} + \sqrt{3/m} & \text{with probability } \frac{1}{6} \\ 0 & \text{with probability } \frac{2}{3} \\ - \sqrt{3/m} & \text{with probability } \frac{1}{6}. \end{cases}
+\end{equation}
+For the latter three kinds of matrices it was shown in \cite{BDDW08}, that for $n \rightarrow \infty$, sufficiently small $k$ and given $\delta$ they satisfy the restricted isometry property (\ref{rip}) of order $k$ for this $\delta$
+with probability converging exponentially to one.
+But the needed proportion between $n$ and $k$ is so small that we can't expect the generated matrices in small dimensions to satisfy the RIP.
+For each type we randomly generated two
+matrices using MATLAB 8.3.0 \cite{MAT} with rows ranging from 15 to 30, columns from 25 to 40 and order $k$ between three and five.
+
+The tests were performed on an Intel i7-4770 with 3.4 GHz.
+For each instance we set a time limit of one hour, if a problem didnot finish within this time, we counted 3600 seconds and used the dual bound as the result.
+In our test this
+only happened once for the right-hand-side of a $15 \times 30$ binary matrix with order $k=5$ in (\ref{MISDP}).
+
+The average gap of the relaxations (\ref{Asp07}) and (\ref{Asp08Dual}) in contrast to the exact solution computed
+with (\ref{MISDP}) is shown in table \ref{results}. Results for the left hand sides of the band and rank one matrices where omitted, as they equalled zero within numerical exactness for all problems.
+
+\begin{table}[ht]
+   \centering
+        \caption{Average gap of relaxations for restricted isometry constants $\alpha_k$ and $\beta_k$}
+   \label{results}
+   \begin{tabular*}{\textwidth}{l @{\extracolsep{\fill}} c r r c r r}
+   \toprule
+  & & \multicolumn{2}{c}{$\alpha_k^2$} & \phantom{abc} & \multicolumn{2}{c}{$\beta_k^2$} \\
+ \cmidrule{3-4} \cmidrule{6-7} 
+type of matrix & & (\ref{Asp07}) & (\ref{Asp08Dual}) & & (\ref{Asp07}) & (\ref{Asp08Dual}) \\
+\midrule
+$\mathcal{N}(0,1)$ & & $85.45\%$ & todo &  & $23.84\%$ & todo\\
+binary & & $85.28\%$ & todo &  & $47.12\%$ & todo\\
+band matrix & & - & todo &  & $12.97\%$ & todo\\
+rank 1 & & - & todo &  & $50.94\%$ & todo\\
+$\mathcal{N}(0,1/m)$ & & $88.36\%$ & todo &  & $26.05\%$ & todo\\
+$\pm 1/\sqrt{m}$ & & $85.35\%$ & todo &  & $29.53\%$ & todo\\
+$0, \pm \sqrt{3/m}$ & & $87.45\%$ & todo &  & $23.05\%$ & todo\\
+\midrule
+average & & $86.38\%$ & todo &  & $29.92\%$ & todo \\
+\bottomrule
+\end{tabular*}
+\end{table}
+
+As we can see from the results, for the right- and especially for the left-hand-side for all of our instances there was quite a big gap for the relaxation in (\ref{Asp07}) which we were able to close with our exact optimization
+problem (\ref{MISDP}).
+For the right-hand-side the relaxation from \cite{Asp07} performed quite well for band matrices with a gap of $13 \%$ and satisfactorly for most other classes of matrices, except for binary matrices, where in
+our testset the average gap reached almost $50\%$.
+On the left-hand-side however, the gap of the relaxation was between $85\%$ and $90\%$ for all classes, which would make it extremely hard to verify the requirements of theorem
+\ref{RIPsqrt2} or theorem \ref{asymRIPtheorem}.
+
+The average solving times of (\ref{MISDP}), (\ref{Asp07}) and (\ref{Asp08Dual}) in seconds are shown in table \ref{times}.
+
+\begin{table}[ht]
+   \centering
+        \caption{Average solving times for left- and right-hand-sides (seconds)}
+   \label{times}
+   \begin{tabular*}{\textwidth}{l @{\extracolsep{\fill}} c r r c r r c r r}
+   \toprule
+  & & \multicolumn{2}{c}{(\ref{MISDP})} & \phantom{abc} & \multicolumn{2}{c}{(\ref{Asp07})} & \phantom{abc} & \multicolumn{2}{c}{(\ref{Asp08Dual})} \\
+ \cmidrule{3-4} \cmidrule{6-7} \cmidrule{9-10} 
+type of matrix & & l & r &  & l & r &  & l & r \\
+\midrule
+$\mathcal{N}(0,1)$ & & 2145.35 & 1489.05 & & 5.35 & 5.6 & & todo & todo\\
+binary & & 2009.65 & 2586.0 & & 5.05 & 3.85 & & todo & todo\\
+band matrix & & 75.65 & 732.75 & & 3.1 & 5.55 & & todo & todo\\
+rank 1 & & 222.85 & 1658.1 & & 4.4 & 5.05 & & todo & todo\\
+$\mathcal{N}(0,1/m)$ & & 2035.05 & 1488.05& & 7.4 & 5.8 & & todo & todo\\
+$\pm 1/\sqrt{m}$ & & 2106.3 & 1537.45 & & 3.7 & 4.3 & & todo & todo\\
+$0, \pm \sqrt{3/m}$ & & 1731.45 & 1309.8 & & 5.6 & 5.35 & & todo & todo\\
+\midrule
+average & & 1475.19 & 1543.03 & & 4.94 & 5.07 & & todo & todo \\
+\bottomrule
+\end{tabular*}
+\end{table}
+
+Here we can see the advantage of the relaxation proposed in \cite{Asp07}.
+While it won't produce exact results, it can solve instances, where the mixed integer semidefinite problem takes 30 or more minutes, within a few seconds.
+With the mixed integer semidefinite program we could only solve instances up to size $40$ and order $5$ within our time limit of 60 minutes, which is far from the amount needed for the instances used in practical applications.
+But this still gives us a way to compute restricted isometry constants apart from total enumeration, if they need to be known exactly.
+Also mixed integer semidefinite programming is still a relatively new subject of research, and like
+in the mixed integer linear case, there should be hope for improvements in the algorithms in the future, which would make this approach much more feasible.
+
+In table \ref{nodes} we compare the number of nodes in the branch-and-bound-tree for (\ref{MISDP}) with the number of eigenvalue problems that would have to be solved to compute the restricted isometry constants via total enumeration.
+For this we computed the ratio of branch-and-bound nodes to submatrices of order $k$ of $A$, of which there exist $\binom{n}{k}$.
+
+\begin{table}[ht]
+   \centering
+        \caption{Branch-and-bound-nodes as percentage of total enumeration}
+   \label{nodes}
+   \begin{tabular*}{\textwidth}{l @{\extracolsep{\fill}} r r}
+   \toprule
+type of matrix & left & right \\
+\midrule
+$\mathcal{N}(0,1)$ & $76,94\%$ & $59.52\%$\\
+binary & $89.40\%$ & $138.87\%$\\
+band matrix & $0.77\%$ & $0.17\%$ \\
+rank 1 & $1.01\%$ & $30.64\%$\\
+$\mathcal{N}(0,1/m)$ & $59.85\%$ & $58.85\%$\\
+$\pm 1/\sqrt{m}$ & $90.74\%$ & $69.49\%$\\
+$0, \pm \sqrt{3/m}$ & $75.96\%$ & $60.68\%$\\
+\midrule
+average & $56.38\%$ & $61.87\%$ \\
+\bottomrule
+\end{tabular*}
+\end{table}
+
+While the mixed integer semidefinite program performed quite well for band and rank one matrices, it didn't perform too well for most other clases, sometimes even needing more branch-and-bound nodes than there are submatrices of $A$
+(which can happen, as the latter only appear as leaves in the branch-and-bound-tree).
+One reason for this is that sparse eigenvalues will usually be lower than non-sparse eigenvalues, leading to fractional solution as long as there
+are still unfixed binary variables, while integer solutions only appear in the leaves.
+But for these computations we only used general heuristics and branching rules as implemented in SCIP, so there might be chances for improvements with
+specially tailored heuristics and branching rules for this kind of problem.
+Also our tests showed, that the ratio gets smaller with larger $k$, so for bigger instances the performance of the algorithm in contrast to total enumeration
+should increase.
+
+\section{Conclusion}
+
+We have seen that exact restricted isometry constants can be computed not only via total enumeration but also by solving a mixed integer semidefinite program.
+Because of this exactness result, this new problem could drastically
+improve the bounds computed with the semidefinite relaxations proposed in \cite{Asp07} and \cite{Asp08}, but this came at the cost of a much higher complexity of computation.
+We could only solve these problems for relatively small
+instances, but this could change with sufficient advances in the area of mixed integer semidefinite programming, especially with the addition of special cutting planes, which proved to be the main enhancement needed to make mixed
+integer linear solvers as fast as they are today.
+Also special heuristics and branching rules could improve the performance of the branch-and-bound-algorithm and allow us to solve bigger instances of the proposed MISDP.
 

### 2015-01-02 RIP-MISDP-Paper/Paper_v1.tex
some style changes for title

@@
 
-\title{Computing exact restricted isometry constants via mixed integer semidefinite programming}
-\author[tud]{T. Gally}
+\title{Computing exact restricted isometry constants via\\ mixed integer semidefinite programming}
+\author[tud]{Tristan Gally}
 \ead{gally@mathematik.tu-darmstadt.de}
-\author[tud]{M. E. Pfetsch}
+\author[tud]{Marc E. Pfetsch}
 \ead{pfetsch@mathematik.tu-darmstadt.de}
-\address[tud]{Technische Universit\"at Darmstadt, Department of Mathematics, Dolivostra\ss e 15, 64293 Darmstadt, Germany}
+\address[tud]{Technische Universit\"at Darmstadt, Department of Mathematics,\\ Dolivostra\ss e 15, 64293 Darmstadt, Germany}
 

### 2015-01-02 RIP-MISDP-Paper/Paper_v1.tex
rework introduction

@@
 
+% other commands
+\newcommand{\NP}{\text{NP}}
+\newcommand{\comment}[1]{\marginpar{\raggedright\tiny {#1}}}
+
 \begin{document}
@@
 \section{Introduction}
-Finding the sparsest solution of an underdetermined system of linear equations is a problem that has been the subject of much recent research in compressed sensing.
-Given a matrix $A \in \R^{m \times n}$ with $m < n$ and
-right-hand side $b \in \R^m$, we are interested in a solution of 
+
+Finding a sparsest solution of an underdetermined system of linear
+equations is one key problem in compressed sensing and has been
+thoroughly investigated in recent years. Given a matrix $A \in \R^{m \times n}$ with $m < n$ and 
+right-hand side $b \in \R^m$, the problem is to compute
 \begin{align}\label{l0min}
- \text{min} \quad & \|x\|_0 \tag{$\ell_0$-min}\\
+ \text{min} \quad & \|x\|_0 \tag{$P_0$}\\
  \text{s.t.} \quad & Ax = b, \nonumber
 \end{align}
-where $\|x\|_0$ stands for the cardinality and thus the number of nonzero components of a vector $x$.
-As it is $\mathcal{NP}$-hard to solve (\ref{l0min}), see for example \cite{GareyJohnson}, we would like to solve the problem
+where $\|x\|_0$ denotes the number of nonzero components of~$x$. This
+problem is \NP-hard, see for example the book by Garey and
+Johnson~\cite{GareyJohnson}. One key observation in compressed sensing is
+that a solution of~\eqref{l0min} can be found by solving the following
+convex optimization problem if certain conditions are fulfilled:
 \begin{align}\label{l1min}
- \text{min} \quad & \|x\|_1 \tag{$\ell_1$-min}\\
- \text{s.t.} \quad & Ax = b \nonumber
+ \text{min} \quad & \|x\|_1 \tag{$P_1$}\\
+ \text{s.t.} \quad & Ax = b. \nonumber
 \end{align}
-instead, as this can be cast as an LP.
-Results in compressed sensing show, that there are many conditions which imply that the problems (\ref{l0min}) and (\ref{l1min}) are indeed equivalent.
-First results about this have been
-established in \cite{DH01}, and have later been followed up for example by \cite{CRT05}, \cite{Zha08} or \cite{BCT11}. One such condition is the restricted isometry property
-\begin{equation}\label{RIPintro}
- (1-\delta_k)\|x\|_2^2 \leq \|Ax\|_2^2 \leq (1+\delta_k)\|x\|_2^2,
-\end{equation}
-that was first introduced in \cite{CT05}.
-If it holds for all $k$-sparse vectors $x$ for a sufficiently small restricted isometry constant $\delta_k$, then problems (\ref{l0min}) and (\ref{l1min}) have the same optimal solution.
-Even though it was shown in \cite{BDDW08}, that this condition holds for multiple classes of random matrices with probability exponentially going to one, in \cite{PT14} it was shown, that computing the optimal restricted isometry
-constants in (\ref{RIPintro}) for a given matrix $A$ is again an $\mathcal{NP}$-hard task.
+In fact, this so-called \emph{basis pursuit} problem can be reformulated as
+a linear program (LP) and efficient specialized solution algorithms are
+available both in theory and practice.\comment{Add references:
+  $\ell_1$-homotopy, SPGL1, Chen, Donoho, and Saunders, \dots}
 
-In the past SDP-Relaxations were proposed in \cite{Asp07} and \cite{Asp08} to compute upper bounds on the values of the restricted isometry 
-constant, which could then be used to verify the equivalance between (\ref{l0min}) and (\ref{l1min}), but there could also be cases, where the results of those relaxations weren't sufficient, but the real constants would have been.
-In this paper we will construct a mixed integer semidefinite program to compute the exact constants in (\ref{RIPintro}), to be able to verify in all cases if those sufficient conditions for the equivalence hold for a given matrix.
-The main contribution of this paper is theorem \ref{Rk1thm}, which shows
-that the optimal solution of this MISDP indeed allows us to compute the
-exact restricted isometry constant.
+Several conditions in compressed sensing imply that a solution
+of~\eqref{l0min} can be \emph{recovered} by solving \eqref{l1min}, in which
+case one speaks of $\ell_0$-$\ell_1$-\emph{equivalence}. First results in
+this direction were established by Chen, Donoho, and
+Saunders~\cite{CheDS99} and Donoho and Huo~\cite{DH01}, with many articles
+following. We therefore refer the reader to the excellent book by Foucart
+and Rauhut~\cite{FouR13} for further information.
 
-In the following we will start by introducing the restricted isometry property and corresponding results about the equivalence of (\ref{l0min}) and (\ref{l1min}).
-In the next section we will present the SDP-relaxations in \cite{Asp07}
-and \cite{Asp08}.Afterwards we will construct a mixed integer semidefinite programme and show that it produces exact restricted isometry constants before presenting numerical results and comparisons with the SDP-Relaxations in
+One important condition is the \emph{restricted isometry property} (RIP),
+which was first introduced by Cand\`es and Tao~\cite{CT05}. For each $k \in
+\{1, \dots, n\}$, the RIP involves a \emph{restricted isometry constant}
+$\delta_k$ depending on~$k$ and~$A$. If $\delta_k$ is sufficiently small
+and there exists $\tilde{x} \in \R^n$ with $A \tilde{x} = b$ and
+$\Norm{\tilde{x}}{0} \leq k$, then problems~\eqref{l0min} and~\eqref{l1min}
+are equivalent; see Section~\ref{sec:RIP} for more details. There are
+several types of matrices~$A$ for which the RIP holds, for instance random
+matrices, see, e.g., Baraniuk et al.~\cite{BDDW08}. Although the RIP does
+not provide necessary conditions, it is still one of the major tools for
+establishing for $\ell_0$-$\ell_1$-equivalence. However, in~\cite{PT14} it
+was shown that computing the restricted isometry constants for a given
+matrix~$A$ and~$k$ is NP-hard.
+
+One the other hand, d'Aspremont et al.~\cite{Asp07,Asp08} proposed several
+semidefinite relaxations to compute upper bounds on the restricted isometry
+constants. Note that small upper bounds suffice to guarantee
+$\ell_0$-$\ell_1$-equivalence between~\eqref{l0min} and~\eqref{l1min}, but
+there exist cases in which the bounds are too weak to give a guarantee.
+
+The goal of this paper is to compute optimal restricted isometry constants
+via a mixed integer semidefinite program (SDP). To this end, we will
+introduce an SDP with binary variables and a rank-1 constraint. In order to
+utilize this SDP, we will show that the rank constraint can be relaxed
+without changing the optimal value. Moreover, we discuss a solution
+methodology to solve the resulting mixed integer SDP. It will turn out that
+the resulting problems are quite hard to solve, even for matrices of
+smaller size. However, the results allow an evaluation of the SDP
+relaxations of d'Aspremont et al. Depending on the input matrices, these
+bounds can be quite tight or weak.
+
+In the following, we will start by introducing the restricted isometry
+property and corresponding results about the equivalence of~\eqref{l0min}
+and~\eqref{l1min} in Section~\ref{sec:RIP}. In Section~\ref{sec:SDPrelax},
+we will then review the SDP-relaxations of~\cite{Asp07}
+and~\cite{Asp08}. Afterwards we will construct a mixed integer semidefinite
+programme and show that it produces exact restricted isometry constants in
+Section~\ref{sec:MISDP}. Finally, in Section~\ref{sec:numerical_results},
+we present numerical results and comparisons with the SDP relaxations in
 \cite{Asp07} and \cite{Asp08}.
@@
 \section{The restriced isometry property}
+\label{sec:RIP}
 
@@
 \section{SDP relaxations}
+\label{sec:SDPrelax}
 
@@
 \section{A MISDP for the restricted isometry property}
+\label{sec:MISDP}
 
@@
 \section{Numerical Results}
+\label{sec:numerical_results}
 

### 2015-01-02 RIP-MISDP-Paper/Paper_v1.tex
\text{min|max} -> \min|\max

@@
 \begin{align}\label{l0min}
- \text{min} \quad & \|x\|_0 \tag{$P_0$}\\
+ \min \quad & \|x\|_0 \tag{$P_0$}\\
  \text{s.t.} \quad & Ax = b, \nonumber
@@
 \begin{align}\label{l1min}
- \text{min} \quad & \|x\|_1 \tag{$P_1$}\\
+ \min \quad & \|x\|_1 \tag{$P_1$}\\
  \text{s.t.} \quad & Ax = b. \nonumber
@@
 $(1 - \delta_k) / (1 + \delta_k)$. Also in the other direction, if $A$ satisfies (\ref{asymRIP}) with constants $\alpha_k$ and $\beta_k$, then it will also satisfy (\ref{rip}) for 
-$\delta_k = \text{max} \{ 1 - \alpha_k^2, \beta_k^2 - 1 \}$.
+$\delta_k = \max \{ 1 - \alpha_k^2, \beta_k^2 - 1 \}$.
 So $A$ satisfies definition \ref{asymRIPdef} if and only if it satisfies definition \ref{ripdef}, but as in (\ref{rip}) usually only one of the two inequalities will
@@
 These can then be used to compute either the restricted isometry ratio $\gamma_k := \beta_k^2 / \alpha_k^2$ or the restricted
-isometry constant $\delta_k = \text{max} \{ 1 - \alpha_k^2, \beta_k^2 - 1 \}$.
+isometry constant $\delta_k = \max \{ 1 - \alpha_k^2, \beta_k^2 - 1 \}$.
 
@@
 \begin{align}\label{QP}
- \text{min} \quad & \|Ax\|_2^2 \nonumber \\
+ \min \quad & \|Ax\|_2^2 \nonumber \\
  \text{s.t.} \quad & \|x\|_2^2 = 1 \tag{QP} \\
@@
 \begin{align}\label{Asp07}
- \text{min} \quad & \Tr(A\T A X) \nonumber \\
+ \min \quad & \Tr(A\T A X) \nonumber \\
  \text{s.t.} \quad & \Tr(X) = 1 \nonumber \\
@@
 \begin{align}\label{Asp08}
- \text{max} \quad & \sum_{i=1}^n\Tr(P_iB_i) \nonumber \\
+ \max \quad & \sum_{i=1}^n\Tr(P_iB_i) \nonumber \\
  \text{s.t.} \quad & \Tr(X) = 1 \nonumber \\
@@
 \begin{align}\label{Asp08Dual}
- \text{min} \quad & \lambda_\text{max}\left(\sum_{i=1}^n Y_i \right) \nonumber \\
+  \min \quad & \lambda_{\max}\left(\sum_{i=1}^n Y_i \right) \nonumber \\
   \text{s.t.} \quad & Y_i \succeq B_i \quad \forall i \leq n \tag{A2-Dual} \\
@@
 \begin{align}\label{QP2}
- \text{min} \quad & \|Ax\|_2^2 \nonumber \\
+ \min \quad & \|Ax\|_2^2 \nonumber \\
  \text{s.t.} \quad & \|x\|_2^2 = 1 \tag{QP} \\
@@
 % \begin{align}\label{QP}
-%  \text{min} \quad & \|Ax\|_2^2 \nonumber \\
+%  \min \quad & \|Ax\|_2^2 \nonumber \\
 %  \text{s.t.} \quad & \|x\|_2^2 = 1 \tag{QP} \\
@@
 % 
-% By switching $\text{min}$ for $\text{max}$ in (\ref{QP}) we also get a corresponding optimization problem to compute $\beta_k$ and therefore also $\gamma_k$ and $\delta_k$.
+% By switching $\min$ for $\max$ in (\ref{QP}) we also get a corresponding optimization problem to compute $\beta_k$ and therefore also $\gamma_k$ and $\delta_k$.
 % The cardinality-constraint in (\ref{QP}) can equivalently be written using binary variables $z_i$ as $-z_i \leq x_i \leq z_i$ and $\sum_{i=1}^n z_i \leq k$.
@@
 \begin{align}\label{Rk1MISDP}
- \text{min} \quad & \Tr(A\T A X) \nonumber \\
+ \min \quad & \Tr(A\T A X) \nonumber \\
  \text{s.t.} \quad & \Tr(X) = 1 \nonumber \\
@@
 \begin{align}\label{MISDP}
- \text{min} \quad & \Tr(A\T A X) \nonumber \\
+ \min \quad & \Tr(A\T A X) \nonumber \\
  \text{s.t.} \quad & \Tr(X) = 1 \nonumber \\
@@
  \begin{align}\label{Proj}
- \text{min} \quad & \Tr(\tilde{A}\T\tilde{A}X) \nonumber \\
+ \min \quad & \Tr(\tilde{A}\T\tilde{A}X) \nonumber \\
  \text{s.t.} \quad & \Tr(X) = 1 \tag{Proj} \\

### 2015-01-02 RIP-MISDP-Paper/Paper_v1.tex
\|...\| -> \Norm{...} everywhere

@@
 \begin{align}\label{l0min}
- \min \quad & \|x\|_0 \tag{$P_0$}\\
+ \min \quad & \Norm{x}{0} \tag{$P_0$}\\
  \text{s.t.} \quad & Ax = b, \nonumber
 \end{align}
-where $\|x\|_0$ denotes the number of nonzero components of~$x$. This
+where $\Norm{x}{0}$ denotes the number of nonzero components of~$x$. This
 problem is \NP-hard, see for example the book by Garey and
@@
 \begin{align}\label{l1min}
- \min \quad & \|x\|_1 \tag{$P_1$}\\
+ \min \quad & \Norm{x}{1} \tag{$P_1$}\\
  \text{s.t.} \quad & Ax = b. \nonumber
@@
 The \emph{restricted isometry property (RIP)} of order $k$ holds for a matrix $A \in \R^{m \times n}$, if there exists a minimal \emph{restricted isometry constant} $\delta_k$, such that for all $x \in \Sigma_k :=
-\{x \in \R^n : \|x\|_0 \leq k \}$ we have
+\{x \in \R^n : \Norm{x}{0} \leq k \}$ we have
  \begin{equation}\label{rip}
-  (1-\delta_k)\|x\|_2^2 \leq \|Ax\|_2^2 \leq (1+\delta_k)\|x\|_2^2.
+  (1-\delta_k)\Norm{x}{2}^2 \leq \Norm{Ax}{2}^2 \leq (1+\delta_k) \Norm{x}{2}^2.
  \end{equation}
@@
   \begin{equation}\label{l1ErrorRIP}
-  \|x^* - x \| \leq C \|x - x_k \|,
+  \Norm{x^* - x} \leq C \Norm{x - x_k},
  \end{equation}
@@
  \begin{equation*}
-  \|x^* - x \|_{\ell_1} \leq C \|x - x_k \|_{\ell_1},
+  \Norm{x^* - x}{1} \leq C \Norm{x - x_k}{1},
  \end{equation*}
@@
 Also all of these results can be
-generalized to the case where $\|Ax - b\|_{\ell_2} \leq \epsilon$. The restricted isometry property can also be formulated in an asymmetric way like in \cite{FL09}:
+generalized to the case where $\Norm{Ax - b}{2} \leq \epsilon$. The restricted isometry property can also be formulated in an asymmetric way like in \cite{FL09}:
 
@@
 \begin{equation}\label{asymRIP}
-   \alpha_k^2\|x\|_2^2 \leq \|Ax\|_2^2 \leq \beta_k^2\|x\|_2^2
+   \alpha_k^2 \Norm{x}{2}^2 \leq \Norm{Ax}{2}^2 \leq \beta_k^2 \Norm{x}{2}^2
 \end{equation}
@@
 \begin{align}\label{QP}
- \min \quad & \|Ax\|_2^2 \nonumber \\
- \text{s.t.} \quad & \|x\|_2^2 = 1 \tag{QP} \\
- & \|x\|_0 \leq k. \nonumber
+ \min \quad & \Norm{Ax}{2}^2 \nonumber \\
+ \text{s.t.} \quad & \Norm{x}{2}^2 = 1 \tag{QP} \\
+ & \Norm{x}{0} \leq k. \nonumber
 \end{align}
 The corresponding maximization problem can be solved to compute $\beta_k^2$.
-In \cite{Asp07} the non-convex quadratic equality constraint $\|x\|_2^2 = 1$ is tackled by the technique of semidefinite lifting first introduced in \cite{GW95}
+In \cite{Asp07} the non-convex quadratic equality constraint $\Norm{x}{2}^2 = 1$ is tackled by the technique of semidefinite lifting first introduced in \cite{GW95}
 for the max-cut problem.
@@
 $X \succeq 0$ means that $X$ is symmetric and positive semidefinite.
-In the next step the non-convex rank constraint is then relaxed and the inequality $\|x\|_0 \leq k$ is substituted by the coarser constraint $1\T|X|1 \leq k$ to form
+In the next step the non-convex rank constraint is then relaxed and the inequality $\Norm{x}{0} \leq k$ is substituted by the coarser constraint $1\T|X|1 \leq k$ to form
 the relaxation
@@
 \begin{equation}\label{phi}
- \phi(\rho) = \max_{\|x\|_2 \leq 1} x\T \Sigma x - \rho \|x\|_0,
+ \phi(\rho) = \max_{\Norm{x}{2} \leq 1} x\T \Sigma x - \rho \Norm{x}{0},
 \end{equation}
@@
 \begin{equation*}
- \psi(\rho) = \min_{\|x\|_2 \leq 1} x\T \Sigma x + \rho \|x\|_0.
+ \psi(\rho) = \min_{\Norm{x}{2} \leq 1} x\T \Sigma x + \rho \Norm{x}{0}.
 \end{equation*}
@@
 %  Y_i = \begin{cases}\frac{B_ixx\T B_i}{x\T B_ix}, & \text{for } i \in I \\
-%    \max\{0, \rho \frac{b_i\T b_i - \rho}{\rho - (b_i\T x)^2} \} \frac{(I-xx\T)b_ib_i\T(I-xx\T)}{\|(I-xx\T)b_i\|_2^2}, & \text{for } i \in I^c\end{cases}. 
+%    \max\{0, \rho \frac{b_i\T b_i - \rho}{\rho - (b_i\T x)^2} \} \frac{(I-xx\T)b_ib_i\T(I-xx\T)}{\Norm{(I-xx\T)b_i}{2}^2}, & \text{for } i \in I^c\end{cases}. 
 %  %x in the second case sill as above?
@@
 \begin{align}\label{QP2}
- \min \quad & \|Ax\|_2^2 \nonumber \\
- \text{s.t.} \quad & \|x\|_2^2 = 1 \tag{QP} \\
- & \|x\|_0 \leq k. \nonumber
+ \min \quad & \Norm{Ax}{2}^2 \nonumber \\
+ \text{s.t.} \quad & \Norm{x}{2}^2 = 1 \tag{QP} \\
+ & \Norm{x}{0} \leq k. \nonumber
 \end{align}
 
-For the non-convex quadratic equality constraint $\|x\|_2^2 = 1$ we will
-again use semidefinite lifting as $\|Ax\|_2^2 = x\T A\T Ax =
+For the non-convex quadratic equality constraint $\Norm{x}{2}^2 = 1$ we will
+again use semidefinite lifting as $\Norm{Ax}{2}^2 = x\T A\T Ax =
 \Tr(A\T A X)$ and $\Tr(X) = \sum_{i=1}^n x_i^2$. 
@@
 % \begin{align}\label{QP}
-%  \min \quad & \|Ax\|_2^2 \nonumber \\
-%  \text{s.t.} \quad & \|x\|_2^2 = 1 \tag{QP} \\
-%  & \|x\|_0 \leq k. \nonumber
+%  \min \quad & \Norm{Ax}{2}^2 \nonumber \\
+%  \text{s.t.} \quad & \Norm{x}{2}^2 = 1 \tag{QP} \\
+%  & \Norm{x}{0} \leq k. \nonumber
 % \end{align}
@@
 % The cardinality-constraint in (\ref{QP}) can equivalently be written using binary variables $z_i$ as $-z_i \leq x_i \leq z_i$ and $\sum_{i=1}^n z_i \leq k$.
-For the non-convex quadratic equality constraint $\|x\|_2^2 = 1$ we will
+For the non-convex quadratic equality constraint $\Norm{x}{2}^2 = 1$ we will
 % use the technique of semidefinite lifting first introduced in \cite{GW95} for the max-cut problem.
@@
 % \succeq 0$ means that $X$ is symmetric and positive semidefinite. Using
-% that $\|Ax\|_2^2 = x\T A\T Ax = \Tr(A\T A X)$ and 
+% that $\Norm{Ax}{2}^2 = x\T A\T Ax = \Tr(A\T A X)$ and 
 % $\Tr(X) = \sum_{i=1}^n x_i^2$, we can thus rewrite (\ref{QP}) equivalently to

### 2015-01-02 RIP-MISDP-Paper/Paper_v1.tex
add remark on denoising case

@@
 not provide necessary conditions, it is still one of the major tools for
-establishing for $\ell_0$-$\ell_1$-equivalence. However, in~\cite{PT14} it
-was shown that computing the restricted isometry constants for a given
-matrix~$A$ and~$k$ is NP-hard.
+establishing for $\ell_0$-$\ell_1$-equivalence.
 
-One the other hand, d'Aspremont et al.~\cite{Asp07,Asp08} proposed several
+We note that recovery results can also be derived for the \emph{denoising}
+case, in which instead of~\eqref{l0min} one considers
+\begin{align}\label{l0minDenoising}
+  \min \quad & \Norm{x}{0} \tag{$P_0^\delta$}\\
+  \text{s.t.} \quad & \Norm{Ax - b}{2} \leq \delta, \nonumber
+\end{align}
+for some $\delta \geq 0$; the $\ell_1$-problem~\eqref{l1min} is changed
+similarly. For more details we again refer the reader to~\cite{FouR13}.
+
+However, one drawback of the RIP is that computing the restricted isometry
+constants for a given matrix~$A$ and~$k$ is NP-hard, see~\cite{PT14}. One
+the other hand, d'Aspremont et al.~\cite{Asp07,Asp08} proposed several
 semidefinite relaxations to compute upper bounds on the restricted isometry

### 2015-01-02 RIP-MISDP-Paper/Paper_v1.tex
reworked Section 2

@@
 \usepackage{multirow}
+\usepackage{mathtools}
 
@@
 
-\DeclareMathOperator{\argmin}{argmin}
-\DeclareMathOperator{\argmax}{argmax}
+\DeclareMathOperator*{\argmin}{argmin}
+\DeclareMathOperator*{\argmax}{argmax}
 \DeclareMathOperator{\supp}{supp}
@@
 
-In this section we will discuss some general results about the restricted isometry property and its relationship with the optimization problems (\ref{l0min}) and (\ref{l1min}).
-To start, we will properly define the restricted
-isometry property as in \cite{CT05}:
+In this section we will review some general results about the restricted
+isometry property and its relationship with the optimization problems
+(\ref{l0min}) and (\ref{l1min}). We start with the definition of the RIP as
+in \cite{CT05}:
 
-\begin{rmk} \label{ripdef}
-The \emph{restricted isometry property (RIP)} of order $k$ holds for a matrix $A \in \R^{m \times n}$, if there exists a minimal \emph{restricted isometry constant} $\delta_k$, such that for all $x \in \Sigma_k :=
-\{x \in \R^n : \Norm{x}{0} \leq k \}$ we have
+\begin{rmk}\label{ripdef}
+  The \emph{restricted isometry property} (RIP) of order~$k$ holds for a
+  matrix $A \in \R^{m \times n}$ with constant~$\delta \geq 0$, if
   \begin{equation}\label{rip}
-  (1-\delta_k)\Norm{x}{2}^2 \leq \Norm{Ax}{2}^2 \leq (1+\delta_k) \Norm{x}{2}^2.
+    (1 - \delta) \Norm{x}{2}^2 \leq \Norm{Ax}{2}^2 \leq (1+\delta) \Norm{x}{2}^2
   \end{equation}
+  for all $x \in \Sigma_k \define \{x \in \R^n \suchthat \Norm{x}{0} \leq k\}$. The
+  smallest constant~$\delta$ such that~\eqref{rip} holds is called
+  \emph{restricted isometry constant} (RIC)~$\delta_k$.
 \end{rmk}
 
- If the restricted isometry property holds for a sufficiently small restricted isometry constant, then the optimal solutions of (\ref{l0min}) and (\ref{l1min}) coincide.
+For this paper, it is useful to take an asymmetric viewpoint and
+distinguish between the lower and upper bounds in~\eqref{rip}, as was done
+in~\cite{FL09}:
+
+\begin{rmk}\label{asymRIPdef}
+ The \emph{lower} and \emph{upper RIC} are defined as
+ \begin{align}
+   \alpha_k & \define \argmax_{\alpha \geq 0}\; \{ \alpha^2 \Norm{x}{2}^2
+   \leq \Norm{Ax}{2}^2,\; \text{ for all }x \in \Sigma_k\}\text{ and }\label{lowerRIP}\\
+   \beta_k & \define \argmin_{\beta \geq 1}\; \{ \Norm{Ax}{2}^2 \leq \beta^2
+   \Norm{x}{2}^2,\; \text{ for all }x \in \Sigma_k\},\label{upperRIP}
+ \end{align}
+ respectively. The \emph{restricted isometry ratio} (RIR) $\gamma_k$ is
+ defined as $\gamma_k \define \beta_k^2 / \alpha_k^2$.
+\end{rmk}
+
+\comment{Der ganze Absatz ist \"uberarbeitet, z.B.\ die RIR korrigiert
+  (statt $(1 - \delta_k)/(1 + \delta_k)$). Korrekt?}
+%
+Definition~\ref{asymRIPdef} is a generalization of Definition~\ref{ripdef}
+in the following sense: if the matrix~$A$ satisfies the RIP of order $k$
+with RIC $\delta_k$, then it has a RIR of at most $(1 + \delta_k) / (1 -
+\delta_k)$. Conversely, if $A$ has lower/upper RICs $\alpha_k$ and
+$\beta_k$, respectively, then it will also satisfy~\eqref{rip} for
+$\delta_k = \max \{1 - \alpha_k^2,\; \beta_k^2 - 1\}$. So $A$ has a small
+RIC if and only if it has a large lower and a small upper RIC. But since
+in~\eqref{rip} usually only one of the two inequalities will be sharp, the
+asymmetric version yields more information about~$A$.
+
+ % If the restricted isometry property holds for a sufficiently small restricted isometry constant, then the optimal solutions of (\ref{l0min}) and (\ref{l1min}) coincide.
 To show this, one needs estimates of the form
-  \begin{equation}\label{l1ErrorRIP}
-  \Norm{x^* - x} \leq C \Norm{x - x_k},
- \end{equation}
- where $x^*$ is an optimal solution of (\ref{l1min}) with $b = Ax$ and $x_k$ is the best approximation of some vector $x \in \R^n$ with $k$-sparse vectors.
+ %  \begin{equation}\label{l1ErrorRIP}
+ %  \Norm{x^* - x} \leq C \Norm{x - x_k},
+ % \end{equation}
+ % where $x^*$ is an optimal solution of (\ref{l1min}) with $b = Ax$ and $x_k$ is the best approximation of some vector $x \in \R^n$ with $k$-sparse vectors.
 If we get a result like (\ref{l1ErrorRIP}), and we know that the vector $x$, which in the
- applications would be the sparse signal that needs to be recovered, is already $k$-sparse itself, then the best approximation of $x$ with $k$-sparse vectors is clearly $x$ itself.
+ % applications would be the sparse signal that needs to be recovered, is already $k$-sparse itself, then the best approximation of $x$ with $k$-sparse vectors is clearly $x$ itself.
 This means that the right-hand-side in
- (\ref{l1ErrorRIP}) equals zero, and therefore the unique optimal solution of (\ref{l1min}) has to be $x$ again.
+ % (\ref{l1ErrorRIP}) equals zero, and therefore the unique optimal solution of (\ref{l1min}) has to be $x$ again.
 Now (\ref{l1ErrorRIP}) not only holds for this $x$ exclusively, but for any $k$-sparse vector $y$ with $Ay=Ax=b$.
 And
- because of $Ay=Ax=b$, we get the same problem in (\ref{l1min}) and therefore also the same solution $x^*$ which needs to equal $y$ because of (\ref{l1ErrorRIP}), so this means that there can only be a single vector $x$ with sparsity
- at most $k$ for which $Ax=b$ holds, which then also has to solve (\ref{l0min}), which gives us the equivalance of these two problems.
+ % because of $Ay=Ax=b$, we get the same problem in (\ref{l1min}) and therefore also the same solution $x^*$ which needs to equal $y$ because of (\ref{l1ErrorRIP}), so this means that there can only be a single vector $x$ with sparsity
+ % at most $k$ for which $Ax=b$ holds, which then also has to solve (\ref{l0min}), which gives us the equivalance of these two problems.
 
- The best-known result of this type is probably the following from \cite{Can08}:
+If the restricted isometry property holds for a sufficiently small
+restricted isometry constant, then the optimal solutions of~\eqref{l0min}
+and~\eqref{l1min} coincide.  The best-known result of this type is probably
+the following:
 
-\begin{thm} \label{RIPsqrt2}
- If $A$ satisfies $\delta_{2k} < \sqrt{2}-1$, then for every $x \in \R^n$
- \begin{equation*}
-  \Norm{x^* - x}{1} \leq C \Norm{x - x_k}{1},
- \end{equation*}
-where $x^*$ is an optimal solution for (\ref{l1min}) with right-hand side $b = Ax$, $x_k$ is the best approximation of $x$ with $k$-sparse vectors and $C$ is a constant depending only on $k$ and $\delta_{2k}$.
-If $x$ is $k$-sparse,
-then the optimal solutions of (\ref{l0min}) and (\ref{l1min}) coincide.\end{thm}
-
-There are also many other results like Theorem \ref{RIPsqrt2}, for example in \cite{fou10} the condition was relaxed to $\delta_{2k} < 0.4652$ and in \cite{CWX10} it was shown for $\delta_k < 0.307$.
-Also all of these results can be
-generalized to the case where $\Norm{Ax - b}{2} \leq \epsilon$. The restricted isometry property can also be formulated in an asymmetric way like in \cite{FL09}:
-
-\begin{rmk} \label{asymRIPdef}
-The \emph{restricted isometry ratio} $\gamma_k$ is defined as $\gamma_k := \beta_k^2 / \alpha_k^2$, where $\alpha_k$ and $\beta_k$ are the largest and smallest constant satisfying
-\begin{equation}\label{asymRIP}
-   \alpha_k^2 \Norm{x}{2}^2 \leq \Norm{Ax}{2}^2 \leq \beta_k^2 \Norm{x}{2}^2
-\end{equation}
-for all $x \in \Sigma_k$.
-\end{rmk}
-
-This is a generalization of Definition \ref{ripdef} in the form that if $A$ satisfies the restricted isometry property of order $k$ with restricted isometry constant $\delta_k$, then it has a restricted isometry constant of at most
-$(1 - \delta_k) / (1 + \delta_k)$. Also in the other direction, if $A$ satisfies (\ref{asymRIP}) with constants $\alpha_k$ and $\beta_k$, then it will also satisfy (\ref{rip}) for 
-$\delta_k = \max \{ 1 - \alpha_k^2, \beta_k^2 - 1 \}$.
-So $A$ satisfies definition \ref{asymRIPdef} if and only if it satisfies definition \ref{ripdef}, but as in (\ref{rip}) usually only one of the two inequalities will
-be sharp in contrast to (\ref{asymRIP}), the asymmetric version will be more precise. For this definition a condition like Theorem \ref{RIPsqrt2} can again be shown like in \cite{FL09}:
-
-\begin{thm} \label{asymRIPtheorem}
-If $A$ has restricted isometry ratio $\gamma_{2k} \leq 4 \sqrt{2} - 3 \approx 2.6569$, then every $k$-sparse vector is exactly recovered by (\ref{l1min}). 
+\begin{thm}[Cand\'es~\cite{Can08}]\label{RIPsqrt2}
+  Let $A$ satisfy $\delta_{2k} < \sqrt{2}-1 \approx 0.4142$. If there
+  exists $\tilde{x} \in \R^n$ with $\Norm{\tilde{x}}{0} \leq k$ such that
+  $A\tilde{x} = b$, then the optimal solutions of~\eqref{l0min}
+  and~\eqref{l1min} coincide.
 \end{thm}
 
-Again this result can be generalized to the case of imperfect data.
-One advantage of Defintion \ref{asymRIPdef} and Theorem \ref{asymRIPtheorem} is, that they are indifferent under scaling of the system of linear equations.
-If a system
-$Ax=b$ is scaled to $\lambda Ax = \lambda b$, then the condition of Theorem \ref{RIPsqrt2} might no longer be valid, while in Definition \ref{asymRIPdef} we can use $\tilde{\alpha_k} = \lambda \alpha_k, \tilde{\beta_k} = \lambda \beta_k$
-and therefore $\tilde{\gamma_k} = \lambda^2 \alpha_k^2 / \lambda^2
-\beta_k^2 = \gamma_k$.
+There are also many other results similar to Theorem~\ref{RIPsqrt2}: For
+example, Foucart~\cite{fou10} relaxed the condition to $\delta_{2k} <
+0.4652$. Moreover, Cai et al.~\cite{CWX10} gave the condition $\delta_k <
+0.307$ for Theorem~\ref{RIPsqrt2} to hold. Note that all of these results
+can be generalized to the denoising case~\eqref{l0minDenoising}.
 
-In the following we will try to compute the optimal constants $\alpha_k$ and $\beta_k$ in (\ref{asymRIP}).
-These can then be used to compute either the restricted isometry ratio $\gamma_k := \beta_k^2 / \alpha_k^2$ or the restricted
-isometry constant $\delta_k = \max \{ 1 - \alpha_k^2, \beta_k^2 - 1 \}$.
+Similar results hold for the asymmetric case. For instance, we have:
+
+\begin{thm}[Foucart and Lai~\cite{FL09}]\label{asymRIPtheorem}
+  Let $A$ satisfy $\gamma_{2k} \leq 4 \sqrt{2} - 3 \approx 2.6569$. If
+  there exists $\tilde{x} \in \R^n$ with $\Norm{\tilde{x}}{0} \leq k$ such
+  that $A\tilde{x} = b$, then the optimal solutions of~\eqref{l0min}
+  and~\eqref{l1min} coincide.
+\end{thm}
+
+Again this result can be generalized to the denoising case. One advantage
+of Defintion~\ref{asymRIPdef} and Theorem~\ref{asymRIPtheorem} is that they
+are invariant under scaling of $Ax = b$. The condition of
+Theorem~\ref{RIPsqrt2} holding for~$A x = b$ might not hold for the scaled
+$\lambda Ax = \lambda b$. However, in Definition \ref{asymRIPdef}, we can
+use $\tilde{\alpha}_k = \lambda \alpha_k$, $\tilde{\beta}_k = \lambda
+\beta_k$, and therefore $\tilde{\gamma}_k = \lambda^2 \alpha_k^2 /
+\lambda^2 \beta_k^2 = \gamma_k$.
+
+In the following we will design an approach to compute the optimal
+constants $\alpha_k$ and $\beta_k$. These can then be used to compute
+either the restricted isometry ratio $\gamma_k := \beta_k^2 / \alpha_k^2$
+or the restricted isometry constant $\delta_k = \max \{ 1 - \alpha_k^2,\;
+\beta_k^2 - 1 \}$.
 

### 2015-01-02 RIP-MISDP-Paper/Paper_v1.tex
use \define and \enifed

@@
 \newcommand{\define}{\coloneqq}
+\newcommand{\enifed}{\eqqcolon}
 \newcommand{\norm}[1]{\lVert{#1}\rVert}
@@
 constants $\alpha_k$ and $\beta_k$. These can then be used to compute
-either the restricted isometry ratio $\gamma_k := \beta_k^2 / \alpha_k^2$
+either the restricted isometry ratio $\gamma_k \define \beta_k^2 / \alpha_k^2$
 or the restricted isometry constant $\delta_k = \max \{ 1 - \alpha_k^2,\;
@@
  \begin{equation*}
- P := \{X \in S_n : X \succeq 0, \Tr(A_i X) = b_i \ \forall i \leq m\}
+ P \define \{X \in S_n : X \succeq 0, \Tr(A_i X) = b_i \ \forall i \leq m\}
  \end{equation*}
@@
 where $\tilde{A} = (A_{i_1}, A_{i_2}, ..., A_{i_\ell})$ consists of those
-columns $\{{i_1}, {i_2}, ..., {i_\ell}\} =: T$ of $A$ for which
+columns $\{{i_1}, {i_2}, ..., {i_\ell}\} \enifed T$ of $A$ for which
 $\hat{z}_{i_k} = 1$.

### 2015-01-02 RIP-MISDP-Paper/Paper_v1.tex
rename environment for definitions

@@
 \newtheorem{thm}{Theorem}
-\newdefinition{rmk}{Definition}
+\newdefinition{definition}{Definition}
 
@@
 
-\begin{rmk}\label{ripdef}
+\begin{definition}\label{ripdef}
   The \emph{restricted isometry property} (RIP) of order~$k$ holds for a
@@
   \emph{restricted isometry constant} (RIC)~$\delta_k$.
-\end{rmk}
+\end{definition}
 
@@
 
-\begin{rmk}\label{asymRIPdef}
+\begin{definition}\label{asymRIPdef}
  The \emph{lower} and \emph{upper RIC} are defined as
@@
  defined as $\gamma_k \define \beta_k^2 / \alpha_k^2$.
-\end{rmk}
+\end{definition}
 

### 2015-01-02 RIP-MISDP-Paper/Paper_v1.tex
trace and rank in normal font
use \abs and \ones

@@
 \newcommand{\T}{^{\top}}
-\newcommand{\Tr}{\textbf{Tr}}
-\newcommand{\Rk}{\textbf{Rk}}
+\newcommand{\Tr}{\text{Tr}}
+\newcommand{\Rk}{\text{Rk}}
 
@@
 The idea is to use a new matrix variable $X=xx\T$ for the quadratic terms, where the condition $X=xx\T$ is then enforced by the equivalent constraints $X \succeq 0$ and $\Rk(X) = 1$, where
-$X \succeq 0$ means that $X$ is symmetric and positive semidefinite.
-In the next step the non-convex rank constraint is then relaxed and the inequality $\Norm{x}{0} \leq k$ is substituted by the coarser constraint $1\T|X|1 \leq k$ to form
+$X \succeq 0$ means that $X$ is symmetric and positive semidefinite. In the
+next step the non-convex rank constraint is then relaxed and the inequality
+$\Norm{x}{0} \leq k$ is substituted by the coarser constraint $\ones\T
+\abs{X} \ones \leq k$ to form
 the relaxation
@@
  \text{s.t.} \quad & \Tr(X) = 1 \nonumber \\
- & \ones\T|X|\ones \leq k \tag{A1} \\
+ & \ones\T \abs{X} \ones \leq k \tag{A1} \\
  & X \succeq 0. \nonumber

### 2015-01-08 RIP-MISDP-Paper/Paper_v1.tex
changed the alpha to Big M for left hand side in Asp08

@@
 \comment{Der ganze Absatz ist \"uberarbeitet, z.B.\ die RIR korrigiert
-  (statt $(1 - \delta_k)/(1 + \delta_k)$). Korrekt?}
+  (statt $(1 - \delta_k)/(1 + \delta_k)$). Korrekt? Korrekt}
 %
@@
 \end{equation*}
-In this case problem (\ref{Asp08}) has to be solved for $\Sigma = \alpha I - A\T A$, where $\alpha$ has to be big enough to make $\Sigma$ positive semidefinite again.
+In this case problem (\ref{Asp08}) or equivalently (\ref{Asp08Dual}) has to be solved for $\Sigma = M I - A\T A$, where $M$ has to be big enough to make $\Sigma$ positive semidefinite again.
 If $\bar{\Psi}$ is the corresponding optimal solution, this implies

### 2015-01-11 RIP-MISDP-Paper/Paper_v1.tex
add list of l1 algorithms

@@
 a linear program (LP) and efficient specialized solution algorithms are
-available both in theory and practice.\comment{Add references:
-  $\ell_1$-homotopy, SPGL1, Chen, Donoho, and Saunders, \dots}
+available both in theory and practice: as an extremely selected list
+consider the articles~\cite{BecBC11,CheDS99,OsbPT00,vdBF08}. For more
+information we refer the reader to~\cite{FouR13} and the practical
+comparison in~\cite{LorPT15}.
+% Note: do not write out author names to keep text short at this place.
+
+% long version:
+%
+% consider the articles by Chen et al.~\cite{CheDS99} (interior point
+% algorithm), Osbourne et al.~\cite{OsbPT00} and Malioutov et
+% al.~\cite{MalCW05} ($\ell_1$-homotopy), van den Berg et al.~\cite{vdBF08}
+% (SPGL1), and Becker et al.~\cite{BecBC11} (NESTA). More information can be
+% found in Foucart and Rauhut~\cite{FouR13} and the practical comparison
+% in~\cite{LorPT15}.
 
@@
 matrices, see, e.g., Baraniuk et al.~\cite{BDDW08}. Although the RIP does
-not provide necessary conditions, it is still one of the major tools for
+not provide necessary conditions, it is one of the major tools for
 establishing for $\ell_0$-$\ell_1$-equivalence.

### 2015-02-09 RIP-MISDP-Paper/Paper_v1.tex
add reference to theorectical comparision of d'Aspremont et al.
made some changes to the definitions and the presentation of the d'Asp07-Relaxation

@@
 \newcommand{\Rk}{\text{Rk}}
+\newcommand{\vect}{\text{vec}}
 
 % sets
+\newcommand{\C}{\mathds{C}}
 \newcommand{\R}{\mathds{R}}
@@
 relaxations of d'Aspremont et al. Depending on the input matrices, these
-bounds can be quite tight or weak.
+bounds can be quite tight or weak. This complements the theoretical
+approximation guarantees developed by d'Aspremont et
+al. in~\cite{AspBG14}.\comment{Also add reference to~\cite{AspBG14} in
+  Section~3.}
 
@@
   \end{equation}
-  for all $x \in \Sigma_k \define \{x \in \R^n \suchthat \Norm{x}{0} \leq k\}$. The
-  smallest constant~$\delta$ such that~\eqref{rip} holds is called
-  \emph{restricted isometry constant} (RIC)~$\delta_k$.
+  for all $x \in \Sigma_k \define \{x \in \R^n \suchthat \Norm{x}{0} \leq k\}$.
+\end{definition}
+
+As the following results depend on the constant $\delta$ in (\ref{rip}) being
+small enough, we are generally interested in the smallest constant satisfying
+(\ref{rip}).
+
+\begin{definition}\label{ricdef}
+For a matrix $A \in \R^{m \times n}$ and order~$k$, the smallest constant in 
+(\ref{rip}), meaning
+\begin{equation}\label{ric}
+ \delta_k \define \argmin_{\delta \geq 0} \{(1 - \delta) \Norm{x}{2}^2 \leq \Norm{Ax}{2}^2 \leq (1+\delta) \Norm{x}{2}^2 ,\; \text{ for all }x \in \Sigma_k\},
+\end{equation}
+is called \emph{restricted isometry constant} (RIC).
 \end{definition}
@@
 \delta_k)$. Conversely, if $A$ has lower/upper RICs $\alpha_k$ and
-$\beta_k$, respectively, then it will also satisfy~\eqref{rip} for
-$\delta_k = \max \{1 - \alpha_k^2,\; \beta_k^2 - 1\}$. So $A$ has a small
+$\beta_k$, respectively, then this implies a restricted isometry constant
+of $\delta_k = \max \{1 - \alpha_k^2,\; \beta_k^2 - 1\}$ in~\eqref{rip}.
+\comment{$\max \{1 - \alpha_k^2,\; \beta_k^2 - 1\}$ erf\"ullt RIP nicht nur,
+sondern ist auch das kleinste $\delta$, das es tut, weshalb die asym. wirklich
+st\"arker ist.} So $A$ has a small
 RIC if and only if it has a large lower and a small upper RIC. But since
@@
 example, Foucart~\cite{fou10} relaxed the condition to $\delta_{2k} <
-0.4652$. Moreover, Cai et al.~\cite{CWX10} gave the condition $\delta_k <
+0.4652$ and Foucart and Rauhut~\cite{FouR13} to $\delta_{2k} < 0.6246$.
+\comment{Ergebnis ist f\"ur $\C$, sollte aber auch f\"ur $\R$
+gelten, da man RIC \"uber $\C$ durch Aufspaltung in Real- und 
+Imagin\"arteil durch reelle RIC absch\"atzen k\"onnen sollte.}
+Moreover, Cai et al.~\cite{CWX10} gave the condition $\delta_k <
 0.307$ for Theorem~\ref{RIPsqrt2} to hold. Note that all of these results
@@
 propose and will also be checked against in the section on numerical results.
-The idea of these two relaxations as well as the MISDP in the next section is that the optimal constant $\alpha_k^2$ in (\ref{asymRIP}) can be computed
+The idea of these two relaxations as well as the MISDP in the next section is that the optimal constant $\alpha_k^2$ in (\ref{lowerRIP}) can be computed
 with the non-convex quadratic optimization problem
@@
 The idea is to use a new matrix variable $X=xx\T$ for the quadratic terms, where the condition $X=xx\T$ is then enforced by the equivalent constraints $X \succeq 0$ and $\Rk(X) = 1$, where
-$X \succeq 0$ means that $X$ is symmetric and positive semidefinite. In the
-next step the non-convex rank constraint is then relaxed and the inequality
-$\Norm{x}{0} \leq k$ is substituted by the coarser constraint $\ones\T
-\abs{X} \ones \leq k$ to form
-the relaxation
+$X \succeq 0$ means that $X$ is symmetric and positive semidefinite. The objective and the normalization-constraint can then be expressed in $X$ as  $\Norm{Ax}{2}^2 = x\T A\T Ax =
+\Tr(A\T A X)$ and $\Norm{x}{2}^2 = \sum_{i=1}^n x_i^2 = \Tr(X)$. In the
+next step the inequality $\Norm{x}{0} \leq k$ is then substituted by the coarser constraint $\ones\T
+\abs{X} \ones \leq k$, which follows from the fact that $\ones\T\abs{X} \ones = \Norm{\vect(X)}{1}\leq \sqrt{\Norm{\vect(X)}{0}} \Norm{\vect(X)}{2}$ and $\Norm{\vect(X)}{0}=\Norm{x}{0}^2$ in the rank 1 case, where $\vect(X)$ is a vector
+in $\R^{n^2}$ consisting of all entries of $X$. 
+Relaxing the non-convex Rank-constraint finally leads to the following relaxation:
 \begin{align}\label{Asp07}
@@
 
-In \cite{Asp08} instead the problem
+Another SDP relaxation was proposed in \cite{Asp08}. There the problem
 
@@
 
-In this section we want to compute the optimal constants $\alpha_k$ and $\beta_k$ in (\ref{asymRIP}) using mixed integer semidefinite programming.
+In this section we want to compute the optimal constants $\alpha_k$ and $\beta_k$ in (\ref{lowerRIP}) and (\ref{upperRIP}) using mixed integer semidefinite programming.
 We will again start with the non-convex quadratic problem
@@
 For the non-convex quadratic equality constraint $\Norm{x}{2}^2 = 1$ we will
-again use semidefinite lifting as $\Norm{Ax}{2}^2 = x\T A\T Ax =
-\Tr(A\T A X)$ and $\Tr(X) = \sum_{i=1}^n x_i^2$. 
+again use semidefinite lifting, introducing the new variable $X=xx\T$ like in the last section. 
 The cardinality-constraint in (\ref{QP}) can equivalently be written using binary variables $z_i$ as $-z_i \leq x_i \leq z_i$ and $\sum_{i=1}^n z_i \leq k$.
@@
 
-The next step would then be to relax the non-convex rank constraint again to arrive at the following relaxation which computes a lower bound on the optimal constant $\alpha_k$ in (\ref{asymRIP}):
+The next step would then be to relax the non-convex rank constraint again to arrive at the following relaxation which computes a lower bound on the optimal constant $\alpha_k$ in (\ref{lowerRIP}):
 \begin{align}\label{MISDP}

### 2015-02-09 RIP-MISDP-Paper/Paper_v1.tex
added a paragraph about SPCA to the introduction
added two papers about SPCA to the bib-file, the rest of the changes come from the 
formating of JabRef

@@
 \{1, \dots, n\}$, the RIP involves a \emph{restricted isometry constant}
-$\delta_k$ depending on~$k$ and~$A$. If $\delta_k$ is sufficiently small
+$\delta_k$ bounding the $k$-sparse eigenvalues of $A$. If $\delta_k$ is sufficiently small
 and there exists $\tilde{x} \in \R^n$ with $A \tilde{x} = b$ and
@@
 However, one drawback of the RIP is that computing the restricted isometry
-constants for a given matrix~$A$ and~$k$ is NP-hard, see~\cite{PT14}. One
+constants for a given matrix~$A$ and~$k$ is NP-hard, see~\cite{PT14}. On
 the other hand, d'Aspremont et al.~\cite{Asp07,Asp08} proposed several
@@
 there exist cases in which the bounds are too weak to give a guarantee.
+
+The relaxations in \cite{Asp07,Asp08} were originally derived for the deeply
+connected problem of sparse principal component analysis (SPCA). Principal
+component analysis (PCA) is about finding orthogonal directions maximizing the 
+variance given in matrix $A$. SPCA tries to fix the problem that the components
+of PCA usually consist of small contributions of all axes, which make the results
+hard to interpret, by enforcing sparsity on the components. SPCA was
+originally proposed by Zuo et al. in \cite{SPCA}, where a penalty term was added 
+to the objective. Other authors like Moghaddam et al. in \cite{sbSPCA} also proposed 
+computing a sparse maximal eigenvalue for SPCA, like it is done for the RIP.
 

### 2015-02-09 RIP-MISDP-Paper/Paper_v1.tex
made some changes to the presentation of the relaxation in d'Asp 08

@@
 approximation guarantees developed by d'Aspremont et
-al. in~\cite{AspBG14}.\comment{Also add reference to~\cite{AspBG14} in
-  Section~3.}
+al. in~\cite{AspBG14}.
 
@@
 \end{equation}
-originally stemming from sparse principal component analysis, is tackled for a symmetric positive semidefinite matrix $\Sigma$, which for
-$\Sigma = A\T A$, by duality, gives an upper bound on $\beta_k$ via
-
+originally stemming from sparse principal component analysis, is tackled for a symmetric positive semidefinite matrix $\Sigma$. For
+$\Sigma = A\T A$ we can use this to compute an upper bound on $\beta_k$ via Lagrange-relaxation as
 \begin{equation*}
- \beta_k^2 \leq \inf_{\rho \geq 0} \phi(\rho) + \rho k.
+\beta_k^2 = \max_{\|x\|_2 \leq 1, \|x\|_0 \leq k} x^TA^TAx \leq \inf_{\rho \geq 0} \max_{\|x\|_2\leq 1} x^TA^TAx - \rho (\|x\|_0-k) = \inf_{\rho \geq 0} \phi(\rho) + \rho k.
 \end{equation*}
 
-As one way to solve (\ref{phi}) again the semidefinite lifting technique is used to compute an upper bound on $\phi$ via
+As one way to solve (\ref{phi}) again the semidefinite lifting technique is used in \cite{Asp08} to compute an upper bound on $\phi$ via
 \begin{align}\label{Asp08}
@@
 
-A lower bound for $\alpha_k$ can be computed via
-
+As mentioned in \cite{Asp08}, a lower bound for $\alpha_k$ can be computed via
 \begin{equation*}
@@
 a lower bound on $\alpha_k$ as
-
 \begin{equation*}
@@
 
-%To compute the parameter $\rho$, the authors of \cite{Asp07} propose to minimize over the interval
+In \cite{AspBG14} explicit bounds on the quality of the semidefinite relaxation where derived using randomization arguments.
+
+%To compute the parameter $\rho$, the authors of \cite{Asp08} propose to minimize over the interval
 % \begin{equation*}

### 2015-02-10 RIP-MISDP-Paper/Paper_v1.tex
made some changes to the chapter on the MISDP
changed forall quantor to for i=1,...,n
switched some accents in the rank one proof
added a remark that we also have rank one equivalence in all B&B-nodes
added a paragraph about finding rank one solutions via Pataki algorithm for SPCA etc.

@@
 \newdefinition{definition}{Definition}
+\newdefinition{rmk}{Remark}
 
@@
 The idea is to use a new matrix variable $X=xx\T$ for the quadratic terms, where the condition $X=xx\T$ is then enforced by the equivalent constraints $X \succeq 0$ and $\Rk(X) = 1$, where
-$X \succeq 0$ means that $X$ is symmetric and positive semidefinite. The objective and the normalization-constraint can then be expressed in $X$ as  $\Norm{Ax}{2}^2 = x\T A\T Ax =
+$X \succeq 0$ means that $X$ is symmetric (shortly written \mbox{$X \in S_n$}) and positive semidefinite.
+The objective and the normalization-constraint can then be expressed in $X$ as  $\Norm{Ax}{2}^2 = x\T A\T Ax =
 \Tr(A\T A X)$ and $\Norm{x}{2}^2 = \sum_{i=1}^n x_i^2 = \Tr(X)$. In the
@@
  & X \succeq 0 \tag{A2-Primal} \\
- & X \succeq P_i \succeq 0 \quad \forall i \leq n, \nonumber
+ & X \succeq P_i \succeq 0 \quad \text{for } i = 1, ..., n, \nonumber
 \end{align}
@@
   \text{s.t.} \quad & Y_i \succeq B_i \quad \forall i \leq n \tag{A2-Dual} \\
-  & Y_i \succeq 0 \quad \ \ \forall i \leq n, \nonumber
+  & Y_i \succeq 0 \quad \ \ \text{for } i = 1, ..., n, \nonumber
 \end{align}
@@
  \text{s.t.} \quad & \Tr(X) = 1 \nonumber \\
- & -z_j \leq X_{ij} \leq z_j \quad \forall j \leq n \nonumber \\
- & \sum_{i=1}^n z_i \leq k \tag{Rk1MISDP} \\
+ & -z_j \leq X_{ij} \leq z_j \quad \text{for } j = 1, ..., n \nonumber \\
+ & \sum_{i=1}^n z_i \leq k \tag{M1} \\
  & \Rk(X) = 1 \nonumber \\
  & X \succeq 0  \nonumber \\
- & z \in \{0,1\}^n \nonumber
+ & z \in \{0,1\}^n. \nonumber
 \end{align}
+\comment{test whether to use only diagonal inequalities $-z_j \leq X_{jj} \leq z_j$, or only those for $j$ (which are redundant in the SDP because of diagonal dominance) or all for $i$ and $j$ (which are even more redundant
+because of symmetrie)}
 
 The next step would then be to relax the non-convex rank constraint again to arrive at the following relaxation which computes a lower bound on the optimal constant $\alpha_k$ in (\ref{lowerRIP}):
+
 \begin{align}\label{MISDP}
@@
  \text{s.t.} \quad & \Tr(X) = 1 \nonumber \\
- & -z_j \leq X_{ij} \leq z_j \quad \forall j \leq n \nonumber \\
+ & -z_j \leq X_{ij} \leq z_j \quad \text{for } j = 1, ..., n \nonumber \\
  & \sum_{i=1}^n z_i \leq k \tag{MISDP} \\
@@
 
-In the following we want to show that (\ref{MISDP}) not only computes a lower bound but the exact value of $\alpha_k$.
+In the following we will to show that (\ref{MISDP}) not only computes a lower bound but the exact value of $\alpha_k$.
 For this we will use results about the geometry of semidefinite programs to show that there always exists an
 optimal solution to (\ref{MISDP}) with rank 1 so that the optimal objective values of (\ref{Rk1MISDP}) and (\ref{MISDP}) have to agree.
-We will use the following result from \cite{pat98} about the relationship between the rank of
-solutions of semidefinite programs and the dimension of corresponding faces, meaning a subset $F$ of the convex set $P$ such that $x \in F, y, z \in P$ and $x = 0.5y + 0.5z$ imply $y,z \in F$:
+We will use the following result by Pataki \cite{pat98} about the relationship between the rank of
+solutions of semidefinite programs and the dimension of corresponding faces, meaning a subset $F$ of the convex set $P$ such that \mbox{$x \in F, y, z \in P$} and $x = 0.5y + 0.5z$ imply $y,z \in F$:
 
@@
  \begin{equation*}
- P \define \{X \in S_n : X \succeq 0, \Tr(A_i X) = b_i \ \forall i \leq m\}
+ P \define \{X \in S_n : X \succeq 0, \Tr(A_i X) = b_i \ \text{for } i = 1, ..., m\}
  \end{equation*}
@@
 \begin{proof}
- Let $\hat{X}, \hat{z}$ be an optimal solution to (\ref{MISDP}) (because of $\Tr(X) = 1$ the problem is bounded and $X_{11} = 1, X_{ij} = 0 \ \forall i,j \neq 1, z_1 = 1, z_i = 0 \ \forall i \neq 0$ is always feasible for $k > 0$,
- so an optimal solution exists)
- with $\sum_{i=1}^n z_i = \ell \leq k$. Define
+ Let $\hat{X}, \hat{z}$ be an optimal solution to (\ref{MISDP}) with $\sum_{i=1}^n z_i = \ell \leq k$ 
+ (because of $\Tr(X) = 1$ the problem is bounded and $X_{11} = 1, X_{ij} = 0 \ \forall i,j \neq 1,$ \mbox{$z_1 = 1, z_i = 0 \ \forall i \neq 1$} is always feasible for $k > 0$, so an optimal solution exists).
+ Define
  \begin{align}\label{Proj}
- \min \quad & \Tr(\tilde{A}\T\tilde{A}X) \nonumber \\
- \text{s.t.} \quad & \Tr(X) = 1 \tag{Proj} \\
- & X \succeq 0 , X \in S_\ell, \nonumber
- \end{align}
+ \min \quad & \Tr(\tilde{A}\T\tilde{A}\tilde{X}) \nonumber \\
+ \text{s.t.} \quad & \Tr(\tilde{X}) = 1 \tag{Proj} \\
+ & \tilde{X} \succeq 0 , \tilde{X} \in S_\ell, \nonumber 
+ \end{align}\comment{$\tilde{X} \in S_\ell$ eigentlich in $\tilde{X} \succeq 0$ enthalten, aber betont halt Dimension, so drinnlassen oder \"andern?}
 where $\tilde{A} = (A_{i_1}, A_{i_2}, ..., A_{i_\ell})$ consists of those
@@
 
-Now let $\tilde{X}$ be the matrix consisting of those rows and columns 
+Now let $\bar{X}$ be the matrix consisting of those rows and columns 
 $i_k$ of $\hat{X}$ such that $\hat{z}_{i_k} = 1$.
-Then $\tilde{X}$ is feasible for (\ref{Proj}) with the same objective value as the optimal solution $\hat{X}$ of (\ref{MISDP}), as we only removed zero rows and columns of $\hat{X}$
+Then $\bar{X}$ is feasible for (\ref{Proj}) with the same objective value as the optimal solution $\hat{X}$ of (\ref{MISDP}), as we only removed zero rows and columns of $\hat{X}$
 and $(A\T A)\hat{X}$, which do not influence the trace or the positive
@@
 
-On the other hand we can lift any optimal solution $\tilde{X}$ of (\ref{Proj}) to a feasible point for (\ref{MISDP}) with identical objective value by setting $\bar{z}_i$ to one if and only if $i \in T$
-and $\bar{X}_{i_j i_k} = \tilde{X}_{jk}$ if and only if $i_j, i_k \in T$
-and $\bar{X}_{ij}=0$ otherwise.
+On the other hand we can lift any optimal solution $\check{X}$ of (\ref{Proj}) to a feasible point for (\ref{MISDP}) with identical objective value by setting $\bar{z}_i$ to one if and only if $i \in T$
+and $X^{\prime}_{i_j i_k} = \check{X}_{jk}$ if and only if $i_j, i_k \in T$
+and $X^{\prime}_{ij}=0$ otherwise.
+This solution will then be feasible for (\ref{MISDP}), as $\Tr(\check{X}) = 1$ and $\check{X}\in S_\ell$ imply $0 \leq \check{X}_{jj} = X^{\prime}_{i_j i_j} \leq 1$ and therefore
+$X^{\prime}_{ij} \leq z^{\prime}_j$ by diagonal dominance of semidefinite matrices.
 
 So the optimal objective values of (\ref{Proj}) and (\ref{MISDP}) agree and with this construction we also get that 
-$\Rk(\bar{X}) = \Rk(\tilde{X})$, so it suffices to show that (\ref{Proj}) always allows for a solution of rank one.
+$\Rk(X^{\prime}) = \Rk(\check{X})$, so it suffices to show that (\ref{Proj}) always allows for a solution of rank one.
 Let $\breve{X}$ be an extreme point of (\ref{Proj}) (as the set of optimal points of (\ref{Proj}) is a non-empty, convex, compact subset of $\R^{\ell(\ell+1)/2}$, such a point exists as a consequence of corollary 18.5.1 in
@@
 constant and the restricted isometry ratio by mixed integer semidefinite programming.
+
+\begin{rmk}\label{Rk1rmk}
+Theorem \ref{Rk1thm} still holds if additional constraints are added fixing some of the binary variables $z_j$ to $0$ or $1$.
+In this case there will always exist an optimal solution to problem (\ref{MISDP}) with these added
+constraints with rank $1$, such that the optimal objective value of this problem will also equal that of (\ref{Rk1MISDP}) with the added constraints.
+This means in particular, that solving problem (\ref{MISDP}) using a
+branch-and-bound-approach is the same as solving the original problem (\ref{Rk1MISDP}) with a branch-and-bound-code.
+\end{rmk}
+
+Theorem \ref{Rk1thm} and Remark \ref{Rk1rmk} do not guarantee that any found solution will have rank one, they only guarantee that there exists a rank one solution.
+But this isn't a problem for applications in the context of the RIP,
+as we are here only interested in the optimal objective value, as this gives us the constant $\alpha_k$ or $\beta_k$ for the RIP.
+For other applications like the SPCA such an extreme solution can also be generated with an algorithm
+proposed by Pataki in \cite{coneLP}.
 

### 2016-03-09 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
added version of paper using IEEE style (tables do not yet fit)

--- 2015-02-10 RIP-MISDP-Paper/Paper_v1.tex
@@
-\documentclass[a4paper,11pt,1p]{elsarticle}
+\documentclass[journal]{IEEEtran}
 
@@
 
-\newtheorem{thm}{Theorem}
-\newdefinition{definition}{Definition}
-\newdefinition{rmk}{Remark}
+\newtheorem{theorem}{Theorem}
+\newtheorem{definition}{Definition}
+\newtheorem{remark}{Remark}
 
@@
 \title{Computing exact restricted isometry constants via\\ mixed integer semidefinite programming}
-\author[tud]{Tristan Gally}
-\ead{gally@mathematik.tu-darmstadt.de}
-\author[tud]{Marc E. Pfetsch}
-\ead{pfetsch@mathematik.tu-darmstadt.de}
-\address[tud]{Technische Universit\"at Darmstadt, Department of Mathematics,\\ Dolivostra\ss e 15, 64293 Darmstadt, Germany}
+\author{Tristan~Gally and Marc~E.~Pfetsch
+  \thanks{This work was supported by the German Research Foundation (DFG) through Collaborative Research Center 805 and Priority Program 1798.}
+  \thanks{The  authors  are  with  the  Research  Group  Optimization,  TU  Darmstadt,
+  Darmstadt  64293,  Germany  (e-mail:  gally@mathematik.tu-darmstadt.de;
+  pfetsch@mathematik.tu-darmstadt.de).}}
+
+\maketitle
 
@@
 
-\maketitle
-
 \section{Introduction}
 
-Finding a sparsest solution of an underdetermined system of linear
+\IEEEPARstart{F}{inding} a sparsest solution of an underdetermined system of linear
 equations is one key problem in compressed sensing and has been
@@
 
-\begin{thm}[Cand\'es~\cite{Can08}]\label{RIPsqrt2}
+\begin{theorem}[Cand\'es~\cite{Can08}]\label{RIPsqrt2}
   Let $A$ satisfy $\delta_{2k} < \sqrt{2}-1 \approx 0.4142$. If there
@@
   and~\eqref{l1min} coincide.
-\end{thm}
+\end{theorem}
 
@@
 
-\begin{thm}[Foucart and Lai~\cite{FL09}]\label{asymRIPtheorem}
+\begin{theorem}[Foucart and Lai~\cite{FL09}]\label{asymRIPtheorem}
   Let $A$ satisfy $\gamma_{2k} \leq 4 \sqrt{2} - 3 \approx 2.6569$. If
@@
 
-\begin{thm}\label{patakiLemma}
+\begin{theorem}\label{patakiLemma}
  Let $X \in F$, where $F$ is a face of
@@
  \end{equation*}
-\end{thm}
+\end{theorem}
 
@@
 
-\begin{thm}\label{Rk1thm}
+\begin{theorem}\label{Rk1thm}
  For every $A \in \R^{m \times n}$ and $k \geq 1$ there exists an optimal solution $X^*, z^*$ to (\ref{MISDP}) such that $\Rk(X^*) = 1$. 
-\end{thm}
+\end{theorem}
 \begin{proof}
@@
 
-Now let $\bar{X}$ be the matrix consisting of those rows and columns 
-$i_k$ of $\hat{X}$ such that $\hat{z}_{i_k} = 1$.
-Then $\bar{X}$ is feasible for (\ref{Proj}) with the same objective value as the optimal solution $\hat{X}$ of (\ref{MISDP}), as we only removed zero rows and columns of $\hat{X}$
-and $(A\T A)\hat{X}$, which do not influence the trace or the positive
-semidefiniteness.
+Now let $\bar{X}$ be the $\ell \times \ell$ matrix with all rows and columns outside of $T$ removed.
+Then $\bar{X}$ is feasible for (\ref{Proj}) with the same objective value as the optimal solution $\hat{X}$ of (\ref{MISDP}),
+as we only removed zero rows and columns of $\hat{X}$ and $(A\T A)\hat{X}$, which do not influence the trace or the positive semidefiniteness.
 
-On the other hand we can lift any optimal solution $\check{X}$ of (\ref{Proj}) to a feasible point for (\ref{MISDP}) with identical objective value by setting $\bar{z}_i$ to one if and only if $i \in T$
-and $X^{\prime}_{i_j i_k} = \check{X}_{jk}$ if and only if $i_j, i_k \in T$
+On the other hand we can lift any optimal solution $\check{X}$ of (\ref{Proj}) to a feasible point for (\ref{MISDP}) with identical objective value by setting $z^{\prime}_i$ to one if and only if $i \in T$
+and $X^{\prime}_{i_{j_1} i_{j_2}} = \check{X}_{j_1j_2}$ if and only if $i_{j_1}, i_{j_2} \in T$
 and $X^{\prime}_{ij}=0$ otherwise.
@@
 
-\begin{rmk}\label{Rk1rmk}
+\begin{remark}\label{Rk1rmk}
 Theorem \ref{Rk1thm} still holds if additional constraints are added fixing some of the binary variables $z_j$ to $0$ or $1$.
@@
 branch-and-bound-approach is the same as solving the original problem (\ref{Rk1MISDP}) with a branch-and-bound-code.
-\end{rmk}
+\end{remark}
 
@@
 
+\section{Solving MISDPs}
+\label{sec:solving_misdps}
+
+For solving the mixed-integer semidefinite programs we used the MISDP-Solver SCIP\_SDP originally developed in \cite{MS12} and \cite{Mar13}, which combines the branch-and-bound-framework of SCIP 3.1 \cite{SCIP} with
+interior-point SDP-solvers.
+An important property for semidefinite programs is the slater condition, that states that there is a strictly feasible point, in this case a positive definite matrix, for the primal and dual problem,
+which is needed for strict duality \cite{alizadeh95, helmberghabil} and therefore an imposed precondition for most SDP-solvers.
+This will however usually be violated in a branch-and-bound-context, when variables get fixed and the
+variable bounds are added as linear constraints, which will then be transformed to diagonal entries of the semidefinite program.
+\comment{will man das ausf\"uhrlicher erkl\"aren oder eine Quelle daf\"ur zitieren (wenn man eine
+findet) ?
+} To keep the slater condition intact, the code will remove all fixed variables from the SDP formulation, as well as removing zero-rows from the SDP-matrices, which might appear after fixing variables to zero.
+The
+SCIP\_SDP-code also extracts some easy LP-conditions from the SDP-constraints, to use the presolving capabilities of SCIP for linear constraints.
+
+The semidefinite problems appearing as relaxations are then solved using interior-point-methods, for details on 
+those see for example the books by Ye and Nesterov and Nemirovski \cite{yeip, neneip}. In this case the dual-scaling interior-point-solver DSDP 5.8 \cite{DSDP} was used. 
+Branching was always conducted on the binary variable with value closest to one, as a node-selection rule best first search was used, as it needs the smallest number of iteration to find the optimal solution \cite{achterbergphd},
+and unlike linear solvers warmstart-capabilities cannot be used to profit from the small changes of the problems in the breadth-first-search.
+
+%The code was overhauled and ported to C in contrast to the cited original version.
+
 \section{Numerical Results}
@@
 
-For solving \ref{MISDP} we used the MISDP-Solver SCIP\_SDP originally developed in \cite{MS12} and \cite{Mar13}, which combines the branch-and-bound-framework of SCIP 3.1 \cite{SCIP} with interior-point-SDP-solvers, in this case
-with the dual-scaling interior-point-solver DSDP 5.8 \cite{DSDP}.
-The code was overhauled and ported to C in contrast to the cited original version.
+For solving (\ref{MISDP}) we used the MISDP-Solver SCIP\_SDP presented in chapter \ref{sec:solving_misdps}.
 The continuous SDPs were solved using the command-line version of DSDP.
-The testset
-consisted of 14 matrices of seven different types, and for each matrix we computed
-the left- and right-hand restricted isometry constant $\alpha$ and $\beta$ using the problems (\ref{MISDP}), (\ref{Asp07}) and (\ref{Asp08Dual}).
-In the latter case we solved 15 SDPs for $\rho$ between $0$ and the maximum diagonal
-entry of $A\T A$.
-The matrices were band matrices with band size three or five with entries within the band chosen uniformly in $\{0,1\}$, binary matrices with all entries chosen uniformly in $\{0,1\}$ as well as matrices with
-entries chosen with distribution $\mathcal{N}(0,1)$ and rank one matrices $A = aa\T$, with $\mathcal{N}(0,1)$-distributed entries of $a$.
-In addition to these four kinds of matrices, we also used three more with theoretical results
-about their restricted isometry constants, namely matrices with $N(0, 1/m)$-distributed entries, with entries chosen uniformly in $\pm 1/\sqrt{m}$ and finaly with distribution
+The testset consisted of 14 matrices of seven
+different types, and for each matrix we computed the left- and right-hand restricted isometry constant $\alpha$ and $\beta$ using the problems (\ref{MISDP}), (\ref{Asp07}) and (\ref{Asp08Dual}).
+In the latter case we solved 15
+SDPs for $\rho$ between $0$ and the maximum diagonal entry of $A\T A$.
+The matrices were band matrices with band size three or five with entries within the band chosen uniformly in $\{0,1\}$, binary matrices with all entries chosen
+uniformly in $\{0,1\}$ as well as matrices with entries chosen with distribution $\mathcal{N}(0,1)$ and rank one matrices $A = aa\T$, with $\mathcal{N}(0,1)$-distributed entries of $a$.
+In addition to these four kinds of matrices,
+we also used three more with theoretical results about their restricted isometry constants, namely matrices with $N(0, 1/m)$-distributed entries, with entries chosen uniformly in $\pm 1/\sqrt{m}$ and finaly with distribution
 \begin{equation}\nonumber
@@
 But the needed proportion between $n$ and $k$ is so small that we can't expect the generated matrices in small dimensions to satisfy the RIP.
-For each type we randomly generated two
-matrices using MATLAB 8.3.0 \cite{MAT} with rows ranging from 15 to 30, columns from 25 to 40 and order $k$ between three and five.
+For each type we used \mbox{MATLAB 8.3.0
+\cite{MAT}} to randomly generated nine matrices of three different sizes with rows ranging from 15 to 30, columns from 25 to 40 and order $k$ between three and five.
 
@@
 Because of this exactness result, this new problem could drastically
-improve the bounds computed with the semidefinite relaxations proposed in \cite{Asp07} and \cite{Asp08}, but this came at the cost of a much higher complexity of computation.
-We could only solve these problems for relatively small
-instances, but this could change with sufficient advances in the area of mixed integer semidefinite programming, especially with the addition of special cutting planes, which proved to be the main enhancement needed to make mixed
-integer linear solvers as fast as they are today.
-Also special heuristics and branching rules could improve the performance of the branch-and-bound-algorithm and allow us to solve bigger instances of the proposed MISDP.
+improve the bounds computed with the semidefinite relaxations proposed in \cite{Asp07} and \cite{Asp08}, but this came at the cost of a much higher complexity of computation, at least in comparison to \cite{Asp07}.
+We could only
+solve these problems for relatively small instances, but this could change with sufficient advances in the area of mixed integer semidefinite programming, especially with the addition of special cutting planes, which proved to be
+the main enhancement needed to make mixed integer linear solvers as fast as they are today.
+Also special heuristics and branching rules could improve the performance of the branch-and-bound-algorithm and allow us to solve bigger
+instances of the proposed MISDP.
 
-\section*{Acknowledgments}
-The authors would like to thank the German Research Foundation (DFG) for funding as parts of this research have been carried out within the Collaborative Research Center 805.
-
-\section*{References}
 \bibliographystyle{abbrv}

### 2016-03-09 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
reworked all parts of the paper except for the numerical results

@@
 \usepackage{mathtools}
+\usepackage{url}
 
@@
 not provide necessary conditions, it is one of the major tools for
-establishing for $\ell_0$-$\ell_1$-equivalence.
+establishing the $\ell_0$-$\ell_1$-equivalence.
 
@@
 The goal of this paper is to compute optimal restricted isometry constants
-via a mixed integer semidefinite program (SDP). To this end, we will
-introduce an SDP with binary variables and a rank-1 constraint. In order to
-utilize this SDP, we will show that the rank constraint can be relaxed
+via a mixed integer semidefinite program (MISDP). To this end, we will
+introduce an MISDP with a rank-1 constraint. In order to
+utilize this MISDP, we will show that the rank constraint can be relaxed
 without changing the optimal value. Moreover, we discuss a solution
-methodology to solve the resulting mixed integer SDP. It will turn out that
+methodology to solve the resulting MISDP. It will turn out that
 the resulting problems are quite hard to solve, even for matrices of
@@
 and~\cite{Asp08}. Afterwards we will construct a mixed integer semidefinite
-programme and show that it produces exact restricted isometry constants in
+programm and show that it produces exact restricted isometry constants in
 Section~\ref{sec:MISDP}. Finally, in Section~\ref{sec:numerical_results},
@@
 \begin{equation}\label{ric}
- \delta_k \define \argmin_{\delta \geq 0} \{(1 - \delta) \Norm{x}{2}^2 \leq \Norm{Ax}{2}^2 \leq (1+\delta) \Norm{x}{2}^2 ,\; \text{ for all }x \in \Sigma_k\},
+ \begin{aligned}
+  \delta_k \define \argmin_{\delta \geq 0} \{&(1 - \delta) \Norm{x}{2}^2 \leq \Norm{Ax}{2}^2 \leq (1+\delta) \Norm{x}{2}^2 ,\\
+  & \text{ for all }x \in \Sigma_k\},
+ \end{aligned}
 \end{equation}
@@
    \alpha_k & \define \argmax_{\alpha \geq 0}\; \{ \alpha^2 \Norm{x}{2}^2
-   \leq \Norm{Ax}{2}^2,\; \text{ for all }x \in \Sigma_k\}\text{ and }\label{lowerRIP}\\
+   \leq \Norm{Ax}{2}^2,\; \text{ for all }x \in \Sigma_k\}, \label{lowerRIP}\\
    \beta_k & \define \argmin_{\beta \geq 1}\; \{ \Norm{Ax}{2}^2 \leq \beta^2
@@
 
-\comment{Der ganze Absatz ist \"uberarbeitet, z.B.\ die RIR korrigiert
-  (statt $(1 - \delta_k)/(1 + \delta_k)$). Korrekt? Korrekt}
 %
@@
 of $\delta_k = \max \{1 - \alpha_k^2,\; \beta_k^2 - 1\}$ in~\eqref{rip}.
-\comment{$\max \{1 - \alpha_k^2,\; \beta_k^2 - 1\}$ erf\"ullt RIP nicht nur,
-sondern ist auch das kleinste $\delta$, das es tut, weshalb die asym. wirklich
-st\"arker ist.} So $A$ has a small
-RIC if and only if it has a large lower and a small upper RIC. But since
-in~\eqref{rip} usually only one of the two inequalities will be sharp, the
-asymmetric version yields more information about~$A$.
+So $A$ has a small RIC if and only if it has a large lower and a small upper 
+RIC. But since in~\eqref{rip} usually only one of the two inequalities will 
+be sharp, the asymmetric version yields more information about~$A$.
 
@@
 Again this result can be generalized to the denoising case. One advantage
-of Defintion~\ref{asymRIPdef} and Theorem~\ref{asymRIPtheorem} is that they
+of Definition~\ref{asymRIPdef} and Theorem~\ref{asymRIPtheorem} is that they
 are invariant under scaling of $Ax = b$. The condition of
-Theorem~\ref{RIPsqrt2} holding for~$A x = b$ might not hold for the scaled
+Theorem~\ref{RIPsqrt2} holding for~$A x = b$ does not imply it for the scaled
 $\lambda Ax = \lambda b$. However, in Definition \ref{asymRIPdef}, we can
@@
 The idea of these two relaxations as well as the MISDP in the next section is that the optimal constant $\alpha_k^2$ in (\ref{lowerRIP}) can be computed
-with the non-convex quadratic optimization problem
+by the non-convex quadratic optimization problem
 \begin{align}\label{QP}
@@
 Another SDP relaxation was proposed in \cite{Asp08}. There the problem
-
 \begin{equation}\label{phi}
@@
 originally stemming from sparse principal component analysis, is tackled for a symmetric positive semidefinite matrix $\Sigma$. For
-$\Sigma = A\T A$ we can use this to compute an upper bound on $\beta_k$ via Lagrange-relaxation as
+$\Sigma = A\T A$ we can use this to compute an upper bound on $\beta_k$ via Lagrangian Relaxation as
 \begin{equation*}
-\beta_k^2 = \max_{\|x\|_2 \leq 1, \|x\|_0 \leq k} x^TA^TAx \leq \inf_{\rho \geq 0} \max_{\|x\|_2\leq 1} x^TA^TAx - \rho (\|x\|_0-k) = \inf_{\rho \geq 0} \phi(\rho) + \rho k.
+ \begin{aligned}
+  \beta_k^2 & = \max_{\|x\|_2 \leq 1, \|x\|_0 \leq k} x^TA^TAx \\
+	    & \leq \inf_{\rho \geq 0} \max_{\|x\|_2\leq 1} x^TA^TAx - \rho (\|x\|_0-k) \\
+	    & = \inf_{\rho \geq 0} \phi(\rho) + \rho k.
+ \end{aligned}
 \end{equation*}
@@
 for $B_i = b_ib_i\T - \rho I,$ where $b_i$ is the $i$-th column of a squareroot of $\Sigma$. The corresponding dual problem can then be written as 
-
 \begin{align}\label{Asp08Dual}
@@
 If $\bar{\Psi}$ is the corresponding optimal solution, this implies
-a lower bound on $\alpha_k$ as
+a lower bound on $\alpha_k$ in
 \begin{equation*}
@@
 \end{align}
-
 For the non-convex quadratic equality constraint $\Norm{x}{2}^2 = 1$ we will
@@
 The cardinality-constraint in (\ref{QP}) can equivalently be written using binary variables $z_i$ as $-z_i \leq x_i \leq z_i$ and $\sum_{i=1}^n z_i \leq k$.
-We can thus rewrite (\ref{QP}) equivalently to
+We can thus rewrite (\ref{QP}) equivalently as
 % In this section we want to compute the optimal constants $\alpha_k$ and $\beta_k$ in (\ref{asymRIP}) using mixed integer semidefinite programming.
@@
 \end{align}
-\comment{test whether to use only diagonal inequalities $-z_j \leq X_{jj} \leq z_j$, or only those for $j$ (which are redundant in the SDP because of diagonal dominance) or all for $i$ and $j$ (which are even more redundant
-because of symmetrie)}
-
 The next step would then be to relax the non-convex rank constraint again to arrive at the following relaxation which computes a lower bound on the optimal constant $\alpha_k$ in (\ref{lowerRIP}):
-
 \begin{align}\label{MISDP}
@@
 
-Now let $\bar{X}$ be the $\ell \times \ell$ matrix with all rows and columns outside of $T$ removed.
+Now let $\bar{X}$ be the $\ell \times \ell$ submatrix of $\hat{X}$ with all rows and columns outside of $T$ removed.
 Then $\bar{X}$ is feasible for (\ref{Proj}) with the same objective value as the optimal solution $\hat{X}$ of (\ref{MISDP}),
@@
 and $X^{\prime}_{ij}=0$ otherwise.
-This solution will then be feasible for (\ref{MISDP}), as $\Tr(\check{X}) = 1$ and $\check{X}\in S_\ell$ imply $0 \leq \check{X}_{jj} = X^{\prime}_{i_j i_j} \leq 1$ and therefore
+This solution will then be feasible for (\ref{MISDP}), since $\Tr(\check{X}) = 1$ and $\check{X}\in S_\ell$ imply $0 \leq \check{X}_{jj} = X^{\prime}_{i_j i_j} \leq 1$ and therefore
 $X^{\prime}_{ij} \leq z^{\prime}_j$ by diagonal dominance of semidefinite matrices.
@@
 $\Rk(X^{\prime}) = \Rk(\check{X})$, so it suffices to show that (\ref{Proj}) always allows for a solution of rank one.
-Let $\breve{X}$ be an extreme point of (\ref{Proj}) (as the set of optimal points of (\ref{Proj}) is a non-empty, convex, compact subset of $\R^{\ell(\ell+1)/2}$, such a point exists as a consequence of corollary 18.5.1 in
-\cite{Roc70}) with corresponding face $F=\{\breve{X}\}$. Then using theorem \ref{patakiLemma} on (\ref{Proj}) with $\tilde{m} = 1$ LP-constraint gives
+Let $\breve{X}$ be an extreme point of (\ref{Proj}) (as the set of optimal points of (\ref{Proj}) is a non-empty, convex, compact subset of $\R^{\ell(\ell+1)/2}$, such a point exists
+\cite[Corollary 18.5.1]{Roc70}) with corresponding face $F=\{\breve{X}\}$. Then using Theorem~\ref{patakiLemma} on (\ref{Proj}) with $\tilde{m} = 1$ LP-constraint gives
 \begin{equation*}
@@
 \end{equation*}
-As $\breve{X}=0$ is infeasible for $\Tr(X)=1$, we must have $\Rk(\breve{X}) = 1$, which can then be lifted to an optimal solution of (\ref{MISDP}) with rank one.
+Since $\breve{X}=0$ is infeasible for $\Tr(X)=1$, we must have $\Rk(\breve{X}) = 1$, which can then be lifted to an optimal solution of (\ref{MISDP}) with rank one.
 \end{proof}
 
-As Theorem \ref{Rk1thm} tells us that (\ref{MISDP}) always has a solution of rank one and the rank-constraint was the only constraint we relaxed to get (\ref{MISDP}) from (\ref{QP}), this means that their optimal objective values
+Since Theorem~\ref{Rk1thm} guarantees that (\ref{MISDP}) always has a solution of rank one and the rank-constraint was the only constraint we relaxed to get (\ref{MISDP}) from (\ref{QP}), this means that their optimal objective values
 have to agree and we can compute the optimal constant $\alpha_k$ for the RIP by solving (\ref{MISDP}).
-As the same argumentation also holds for the maximization problem for $\beta_k$, this allows us to compute the restricted isometry
+Since the same argumentation also holds for the maximization problem for $\beta_k$, this allows us to compute the restricted isometry
 constant and the restricted isometry ratio by mixed integer semidefinite programming.
@@
 
+To solve \eqref{MISDP} successfully using a branch-and-bound-approach, it is important to have strong semidefinite relaxations for fractional $z$ variables. 
+Therefore we will further strengthen the constraint coupling the continuous and binary variables to force the nonzero entries of $z$ towards one in the
+semidefinite relaxations.
+
+\begin{remark}\label{HalfConstraint}
+We can substitute the constraint $-z_j \leq X_{ij} \leq z_j$ for all non-diagonal entries of $X$ by the stronger constraint 
+\begin{equation}\label{eq:HalfConstraint}
+  -\frac{1}{2}z_j \leq X_{ij} \leq \frac{1}{2}z_j.
+\end{equation}
+Let $X$ satisfy $\Tr(X)=1$ and $X\succeq 0$, which immediately implies $X_{ii}\geq 0$.
+Since all minors of $X$ need to be nonnegative, we get that
+\begin{equation*}
+  X_{ij}^2 \leq  X_{ii} X_{jj} \leq \frac{1}{4}
+\end{equation*}
+and therefore $-1/2 \leq X_{ij} \leq 1/2$, which implies the equivalence between
+$-z_j \leq X_{ij} \leq z_j$ and~\eqref{eq:HalfConstraint} for all integer values 
+of $z$, while~\eqref{eq:HalfConstraint} might cut off some solutions for fractional $z$.
+\end{remark}
+
 Theorem \ref{Rk1thm} and Remark \ref{Rk1rmk} do not guarantee that any found solution will have rank one, they only guarantee that there exists a rank one solution.
-But this isn't a problem for applications in the context of the RIP,
-as we are here only interested in the optimal objective value, as this gives us the constant $\alpha_k$ or $\beta_k$ for the RIP.
+But this is not a problem for applications in the context of the RIP,
+since we are only interested in the optimal objective value, since it implies the constant $\alpha_k$ or $\beta_k$ for the RIP.
 For other applications like the SPCA such an extreme solution can also be generated with an algorithm
 proposed by Pataki in \cite{coneLP}.
-
-\section{Solving MISDPs}
-\label{sec:solving_misdps}
-
-For solving the mixed-integer semidefinite programs we used the MISDP-Solver SCIP\_SDP originally developed in \cite{MS12} and \cite{Mar13}, which combines the branch-and-bound-framework of SCIP 3.1 \cite{SCIP} with
-interior-point SDP-solvers.
-An important property for semidefinite programs is the slater condition, that states that there is a strictly feasible point, in this case a positive definite matrix, for the primal and dual problem,
-which is needed for strict duality \cite{alizadeh95, helmberghabil} and therefore an imposed precondition for most SDP-solvers.
-This will however usually be violated in a branch-and-bound-context, when variables get fixed and the
-variable bounds are added as linear constraints, which will then be transformed to diagonal entries of the semidefinite program.
-\comment{will man das ausf\"uhrlicher erkl\"aren oder eine Quelle daf\"ur zitieren (wenn man eine
-findet) ?
-} To keep the slater condition intact, the code will remove all fixed variables from the SDP formulation, as well as removing zero-rows from the SDP-matrices, which might appear after fixing variables to zero.
-The
-SCIP\_SDP-code also extracts some easy LP-conditions from the SDP-constraints, to use the presolving capabilities of SCIP for linear constraints.
-
-The semidefinite problems appearing as relaxations are then solved using interior-point-methods, for details on 
-those see for example the books by Ye and Nesterov and Nemirovski \cite{yeip, neneip}. In this case the dual-scaling interior-point-solver DSDP 5.8 \cite{DSDP} was used. 
-Branching was always conducted on the binary variable with value closest to one, as a node-selection rule best first search was used, as it needs the smallest number of iteration to find the optimal solution \cite{achterbergphd},
-and unlike linear solvers warmstart-capabilities cannot be used to profit from the small changes of the problems in the breadth-first-search.
-
-%The code was overhauled and ported to C in contrast to the cited original version.
 
@@
 
-For solving (\ref{MISDP}) we used the MISDP-Solver SCIP\_SDP presented in chapter \ref{sec:solving_misdps}.
-The continuous SDPs were solved using the command-line version of DSDP.
-The testset consisted of 14 matrices of seven
+For solving the mixed-integer semidefinite programs we used the MISDP-Solver SCIP-SDP \cite{SCIP-SDP} originally developed in \cite{MS12} and \cite{Mar13}, which combines the branch-and-bound-framework of SCIP 3.1 \cite{SCIP} with
+interior-point SDP-solvers. The continuous SDPs were solved using the command-line version of SDPA~\cite{SDPA6,SDPA7}. The testset consisted of 63 matrices of seven 
 different types, and for each matrix we computed the left- and right-hand restricted isometry constant $\alpha$ and $\beta$ using the problems (\ref{MISDP}), (\ref{Asp07}) and (\ref{Asp08Dual}).
@@
 The tests were performed on an Intel i7-4770 with 3.4 GHz.
-For each instance we set a time limit of one hour, if a problem didnot finish within this time, we counted 3600 seconds and used the dual bound as the result.
+For each instance we set a time limit of one hour, if a problem did not finish within this time, we counted 3600 seconds and used the dual bound as the result.
 In our test this

### 2016-03-10 www/doc-2.0.0/html/html/search/all_2.js
removed generated documentation from www-folder

@@ -1,34 +1,0 @@
-var searchData=
-[
-  ['calc_5frelax',['calc_relax',['../relax__sdp_8c.html#ad457f6e9ed719ae515e8fb5c61d792ee',1,'relax_sdp.c']]],
-  ['check_5fif_5fsolved',['CHECK_IF_SOLVED',['../sdpi_8c.html#a9e12d019967fa73efce3798eb1cffb33',1,'CHECK_IF_SOLVED():&#160;sdpi.c'],['../sdpisolver__dsdp_8c.html#ad35156183cbcc3eff18e8678b90cf7bf',1,'CHECK_IF_SOLVED():&#160;sdpisolver_dsdp.c'],['../sdpisolver__sdpa_8cpp.html#ad35156183cbcc3eff18e8678b90cf7bf',1,'CHECK_IF_SOLVED():&#160;sdpisolver_sdpa.cpp']]],
-  ['check_5fif_5fsolved_5fbool',['CHECK_IF_SOLVED_BOOL',['../sdpi_8c.html#afacc609ddc5d9dbb3dcdedc72e3bd45b',1,'CHECK_IF_SOLVED_BOOL():&#160;sdpi.c'],['../sdpisolver__dsdp_8c.html#ab0f77ac0eb633c7971fc162028e51732',1,'CHECK_IF_SOLVED_BOOL():&#160;sdpisolver_dsdp.c'],['../sdpisolver__sdpa_8cpp.html#ab0f77ac0eb633c7971fc162028e51732',1,'CHECK_IF_SOLVED_BOOL():&#160;sdpisolver_sdpa.cpp']]],
-  ['close_5fon_5fexit_5f',['close_on_exit_',['../classScipStreamBuffer.html#a413412c76a68951cfce4cd42a3c62704',1,'ScipStreamBuffer']]],
-  ['columns',['columns',['../classscip_1_1SDPBlock.html#a1f80a8b6e6c70b1f401448c50545e42e',1,'scip::SDPBlock']]],
-  ['compconstmatafterfixings',['compConstMatAfterFixings',['../sdpi_8c.html#a1d1b0de601ae6a5211422465d6580a82',1,'sdpi.c']]],
[…]

### 2016-03-10 www/doc-2.0.0/html/html/search/all_9.js

@@ -1,9 +1,0 @@
-var searchData=
-[
-  ['main',['main',['../main_8cpp.html#a3c04138a5bfe5d72780bb7e82a18e627',1,'main.cpp']]],
-  ['main_2ecpp',['main.cpp',['../main_8cpp.html',1,'']]],
-  ['max_5fsize',['max_size',['../classBlockMemoryAllocator.html#aa2909192b7f799f2b115ed99aff4028c',1,'BlockMemoryAllocator']]],
-  ['move_5f1x1_5fblocks_5fto_5flp',['move_1x1_blocks_to_lp',['../cons__sdp_8c.html#a0a6052e778b5ebbe7770a80935098325',1,'cons_sdp.c']]],
-  ['multiaggrvar',['multiaggrVar',['../cons__sdp_8c.html#abc88ebef3100f908bf1822bc71fd28ff',1,'cons_sdp.c']]],
-  ['multiplyconstraintmatrix',['multiplyConstraintMatrix',['../cons__sdp_8c.html#afe2bd14c1755e635d23858cc54ef147b',1,'cons_sdp.c']]]
-];

### 2016-03-10 www/doc-2.0.0/html/html/search/defines_6.js

@@ -1,7 +1,0 @@
-var searchData=
-[
-  ['relax_5fdesc',['RELAX_DESC',['../relax__sdp_8c.html#a07d360b6abb7d664459afac79a17c4aa',1,'relax_sdp.c']]],
-  ['relax_5ffreq',['RELAX_FREQ',['../relax__sdp_8c.html#a41e344358aacae1ad7568b32033f115b',1,'relax_sdp.c']]],
-  ['relax_5fname',['RELAX_NAME',['../relax__sdp_8c.html#a399d3a11c8956582fa1d90f32edf3b0e',1,'relax_sdp.c']]],
-  ['relax_5fpriority',['RELAX_PRIORITY',['../relax__sdp_8c.html#a1980c30a6c657567fd76ccb6dd029feb',1,'relax_sdp.c']]]
-];

### 2016-03-10 www/doc-2.0.0/html/html/search/files_6.js

@@ -1,5 +1,0 @@
-var searchData=
-[
-  ['prop_5fsdpredcost_2ec',['prop_sdpredcost.c',['../prop__sdpredcost_8c.html',1,'']]],
-  ['prop_5fsdpredcost_2eh',['prop_sdpredcost.h',['../prop__sdpredcost_8h.html',1,'']]]
-];

### 2016-03-10 www/doc-2.0.0/html/html/search/functions_1.js

@@ -1,4 +1,0 @@
-var searchData=
-[
-  ['blockmemoryallocator',['BlockMemoryAllocator',['../classBlockMemoryAllocator.html#a7c44120711310beea29f0b313a050ee5',1,'BlockMemoryAllocator::BlockMemoryAllocator(SCIP *scip)'],['../classBlockMemoryAllocator.html#ac00acf02a9701db424e118bd9ac986ed',1,'BlockMemoryAllocator::BlockMemoryAllocator(const BlockMemoryAllocator &amp;other)'],['../classBlockMemoryAllocator.html#a4637cd0cc8cb202f806b095557c8837d',1,'BlockMemoryAllocator::BlockMemoryAllocator(const BlockMemoryAllocator&lt; U &gt; &amp;other)']]]
-];

### 2016-03-10 www/doc-2.0.0/html/html/search/functions_3.js

@@ -1,7 +1,0 @@
-var searchData=
-[
-  ['deallocate',['deallocate',['../classBlockMemoryAllocator.html#a99f0c929fba2010aaa10cd22f2009ddc',1,'BlockMemoryAllocator']]],
-  ['destroy',['destroy',['../classBlockMemoryAllocator.html#abea8f0b9ec21fb92276f0f796c5d2014',1,'BlockMemoryAllocator']]],
-  ['diaggezero',['diagGEzero',['../cons__sdp_8c.html#af4034c85b03cb731ef54872e2fb64459',1,'cons_sdp.c']]],
-  ['dropcomments',['dropComments',['../namespacescip.html#a12da53d5a9f21751e885856b7f7168f7',1,'scip']]]
-];

### 2016-03-10 www/doc-2.0.0/html/html/search/functions_7.js

@@ -1,4 +1,0 @@
-var searchData=
-[
-  ['lpblock',['LPBlock',['../classscip_1_1LPBlock.html#ad45ca560126346bc3e212bb30c61282f',1,'scip::LPBlock']]]
-];

### 2016-03-10 www/doc-2.0.0/html/html/search/functions_d.js

@@ -1,4 +1,0 @@
-var searchData=
-[
-  ['underflow',['underflow',['../classScipStreamBuffer.html#a00b190ec1f66d68d58b5ef1b377b1716',1,'ScipStreamBuffer']]]
-];

### 2016-03-10 www/doc-2.0.0/html/html/search/search.js

@@ -1,815 +1,0 @@
-// Search script generated by doxygen
-// Copyright (C) 2009 by Dimitri van Heesch.
-
-// The code in this file is loosly based on main.js, part of Natural Docs,
-// which is Copyright (C) 2003-2008 Greg Valure
-// Natural Docs is licensed under the GPL.
-
-var indexSectionsWithContent =
-{
-  0: "abcdefgilmnoprstuvx~",
-  1: "blors",
-  2: "s",
-  3: "bcdlmoprstx",
-  4: "abcdefilmoprsux~",
-  5: "bcdfgnrsv",
-  6: "cdloprsv",
-  7: "s",
-  8: "s",
-  9: "o",
-  10: "bcdefprs",
-  11: "ao"
-};
-
-var indexSectionNames =
-{
-  0: "all",
-  1: "classes",
-  2: "namespaces",
-  3: "files",
-  4: "functions",
-  5: "variables",
-  6: "typedefs",
-  7: "enums",
-  8: "enumvalues",
-  9: "related",
-  10: "defines",
-  11: "pages"
-};
-
-function convertToId(search)
-{
-  var result = '';
-  for (i=0;i<search.length;i++)
-  {
-    var c = search.charAt(i);
-    var cn = c.charCodeAt(0);
-    if (c.match(/[a-z0-9\u0080-\uFFFF]/))
-    {
-      result+=c;
-    }
-    else if (cn<16) 
-    {
-      result+="_0"+cn.toString(16);
-    }
-    else 
-    {
-      result+="_"+cn.toString(16);
-    }
-  }
-  return result;
-}
-
-function getXPos(item)
-{
-  var x = 0;
-  if (item.offsetWidth)
-  {
-    while (item && item!=document.body)
-    {
-      x   += item.offsetLeft;
-      item = item.offsetParent;
-    }
-  }
-  return x;
-}
-
-function getYPos(item)
-{
-  var y = 0;
-  if (item.offsetWidth)
-  {
-     while (item && item!=document.body)
-     {
-       y   += item.offsetTop;
-       item = item.offsetParent;
-     }
-  }
-  return y;
-}
-
-/* A class handling everything associated with the search panel.
-
-   Parameters:
-   name - The name of the global variable that will be 
-          storing this instance.  Is needed to be able to set timeouts.
-   resultPath - path to use for external files
-*/
-function SearchBox(name, resultsPath, inFrame, label)
-{
[…]

### 2016-03-10 www/doc-2.0.0/html/html/search/typedefs_4.js

@@ -1,4 +1,0 @@
-var searchData=
-[
-  ['pointer',['pointer',['../classBlockMemoryAllocator.html#a6145be2b5d1802635717410b4dca6a7a',1,'BlockMemoryAllocator::pointer()'],['../classBlockMemoryAllocator_3_01void_01_4.html#a5b14ada9c4dd4efc3a2ae67ee12d6025',1,'BlockMemoryAllocator&lt; void &gt;::pointer()']]]
-];

### 2016-03-10 www/doc-2.0.0/html/search/all_0.js

@@ -1,7 +1,0 @@
-var searchData=
-[
-  ['address',['address',['../classBlockMemoryAllocator.html#aab61d00366dbb389a729f646df9877c5',1,'BlockMemoryAllocator::address(reference x) const '],['../classBlockMemoryAllocator.html#ad7f6e9aaef497a3afdeb0750a56bbce6',1,'BlockMemoryAllocator::address(const_reference x) const ']]],
-  ['allocate',['allocate',['../classBlockMemoryAllocator.html#aa179675af57d9b44295a579f1ae03d37',1,'BlockMemoryAllocator']]],
-  ['allvarsfixed',['allVarsFixed',['../relax__sdp_8c.html#a53c3e9ea74f8ee83caa5252a1fd923b3',1,'relax_sdp.c']]],
-  ['additional_20parameters',['Additional Parameters',['../PARAMETERS.html',1,'']]]
-];

### 2016-03-10 www/doc-2.0.0/html/search/all_12.js

@@ -1,5 +1,0 @@
-var searchData=
-[
-  ['xsgetn',['xsgetn',['../classScipStreamBuffer.html#a824e7103bcf1fefac3d12218f1b43fca',1,'ScipStreamBuffer']]],
-  ['xternal_2ec',['xternal.c',['../xternal_8c.html',1,'']]]
-];

### 2016-03-10 www/doc-2.0.0/html/search/all_3.js

@@ -1,33 +1,0 @@
-var searchData=
-[
-  ['data',['data',['../structscip_1_1LProw.html#a8dfd171ae4cc637cbaa29d0d46c1a3d7',1,'scip::LProw']]],
-  ['deallocate',['deallocate',['../classBlockMemoryAllocator.html#a99f0c929fba2010aaa10cd22f2009ddc',1,'BlockMemoryAllocator']]],
-  ['default_5fcoupledvars',['DEFAULT_COUPLEDVARS',['../branch__sdpinfobjective_8c.html#a1696bb6df771765803369402b5cae3ff',1,'DEFAULT_COUPLEDVARS():&#160;branch_sdpinfobjective.c'],['../branch__sdpobjective_8c.html#a1696bb6df771765803369402b5cae3ff',1,'DEFAULT_COUPLEDVARS():&#160;branch_sdpobjective.c']]],
-  ['default_5fobjlimit',['DEFAULT_OBJLIMIT',['../relax__sdp_8c.html#a02d1082fdc72cac0e806d1d323265851',1,'relax_sdp.c']]],
-  ['default_5fsdprcbin',['DEFAULT_SDPRCBIN',['../prop__sdpredcost_8c.html#aa7672baad9af21f4c0eb7323fb6c17e5',1,'prop_sdpredcost.c']]],
-  ['default_5fsdprcintcont',['DEFAULT_SDPRCINTCONT',['../prop__sdpredcost_8c.html#aa77a1ed2879c88ca21b46d5243b6baac',1,'prop_sdpredcost.c']]],
-  ['default_5fsdpsolverepsilon',['DEFAULT_SDPSOLVEREPSILON',['../relax__sdp_8c.html#afd8c705668d5b5953cafed1ba00dfe0b',1,'relax_sdp.c']]],
-  ['default_5fsdpsolverfeastol',['DEFAULT_SDPSOLVERFEASTOL',['../relax__sdp_8c.html#a55b0b22621ca445ab5c865322c6e3a54',1,'relax_sdp.c']]],
[…]

### 2016-03-10 www/doc-2.0.0/html/search/all_b.js

@@ -1,12 +1,0 @@
-var searchData=
-[
-  ['overview',['Overview',['../index.html',1,'']]],
-  ['objreader_5fsdpa_2ecpp',['objreader_sdpa.cpp',['../objreader__sdpa_8cpp.html',1,'']]],
-  ['objreader_5fsdpa_2eh',['objreader_sdpa.h',['../objreader__sdpa_8h.html',1,'']]],
-  ['objreadersdpa',['ObjReaderSDPA',['../classscip_1_1ObjReaderSDPA.html#a6f27dda2c99d31b9c244a667006f8f5a',1,'scip::ObjReaderSDPA']]],
-  ['objreadersdpa',['ObjReaderSDPA',['../classscip_1_1ObjReaderSDPA.html',1,'scip']]],
-  ['operator_21_3d',['operator!=',['../classBlockMemoryAllocator.html#a8c0d42fa69862be8db6ca6222b52fc0e',1,'BlockMemoryAllocator::operator!=()'],['../BlockMemoryAllocator_8h.html#a51728588824fd98124289aa1c8ce46a9',1,'operator!=():&#160;BlockMemoryAllocator.h']]],
-  ['operator_3d',['operator=',['../classBlockMemoryAllocator.html#ac71648c0f079527d9d10b3799addb34d',1,'BlockMemoryAllocator']]],
-  ['operator_3d_3d',['operator==',['../classBlockMemoryAllocator.html#ad303f8862f2722ce9cbe4df5b53122db',1,'BlockMemoryAllocator::operator==()'],['../BlockMemoryAllocator_8h.html#ad0d94434c480a2e2d21619aa44af267b',1,'operator==():&#160;BlockMemoryAllocator.h']]],
-  ['other',['other',['../structBlockMemoryAllocator_1_1rebind.html#a286c6f3eb304795c559c9de0e0e3ebd6',1,'BlockMemoryAllocator::rebind::other()'],['../structBlockMemoryAllocator_3_01void_01_4_1_1rebind.html#af1768fdb56a58afc1e19bb25d4ecd964',1,'BlockMemoryAllocator&lt; void &gt;::rebind::other()']]]
-];

### 2016-03-10 www/doc-2.0.0/html/search/classes_1.js

@@ -1,5 +1,0 @@
-var searchData=
-[
-  ['lpblock',['LPBlock',['../classscip_1_1LPBlock.html',1,'scip']]],
-  ['lprow',['LProw',['../structscip_1_1LProw.html',1,'scip']]]
-];

### 2016-03-10 www/doc-2.0.0/html/search/defines_1.js

@@ -1,16 +1,0 @@
-var searchData=
-[
-  ['check_5fif_5fsolved',['CHECK_IF_SOLVED',['../sdpi_8c.html#a9e12d019967fa73efce3798eb1cffb33',1,'CHECK_IF_SOLVED():&#160;sdpi.c'],['../sdpisolver__dsdp_8c.html#ad35156183cbcc3eff18e8678b90cf7bf',1,'CHECK_IF_SOLVED():&#160;sdpisolver_dsdp.c'],['../sdpisolver__sdpa_8cpp.html#ad35156183cbcc3eff18e8678b90cf7bf',1,'CHECK_IF_SOLVED():&#160;sdpisolver_sdpa.cpp']]],
-  ['check_5fif_5fsolved_5fbool',['CHECK_IF_SOLVED_BOOL',['../sdpi_8c.html#afacc609ddc5d9dbb3dcdedc72e3bd45b',1,'CHECK_IF_SOLVED_BOOL():&#160;sdpi.c'],['../sdpisolver__dsdp_8c.html#ab0f77ac0eb633c7971fc162028e51732',1,'CHECK_IF_SOLVED_BOOL():&#160;sdpisolver_dsdp.c'],['../sdpisolver__sdpa_8cpp.html#ab0f77ac0eb633c7971fc162028e51732',1,'CHECK_IF_SOLVED_BOOL():&#160;sdpisolver_sdpa.cpp']]],
-  ['conshdlr_5fcheckpriority',['CONSHDLR_CHECKPRIORITY',['../cons__sdp_8c.html#a1bccb692871b1e48ddeefe66a7dbd76b',1,'cons_sdp.c']]],
-  ['conshdlr_5fdelaysepa',['CONSHDLR_DELAYSEPA',['../cons__sdp_8c.html#adf171f22d972da1f289d5b9ac05adab5',1,'cons_sdp.c']]],
-  ['conshdlr_5fdesc',['CONSHDLR_DESC',['../cons__sdp_8c.html#a1fa323744fa4e37db6300fbd6be84b7f',1,'cons_sdp.c']]],
-  ['conshdlr_5feagerfreq',['CONSHDLR_EAGERFREQ',['../cons__sdp_8c.html#ac939edbf15ddf8e569bf0285a8a385e1',1,'cons_sdp.c']]],
[…]

### 2016-03-10 www/doc-2.0.0/html/search/defines_5.js

@@ -1,9 +1,0 @@
-var searchData=
-[
-  ['prop_5fdelay',['PROP_DELAY',['../prop__sdpredcost_8c.html#a7004d49c05d1b2473ff254d4c4d575fd',1,'prop_sdpredcost.c']]],
-  ['prop_5fdesc',['PROP_DESC',['../prop__sdpredcost_8c.html#a01f6c19c723e33bb044d0e466828ac80',1,'prop_sdpredcost.c']]],
-  ['prop_5ffreq',['PROP_FREQ',['../prop__sdpredcost_8c.html#aef556cb12e7377f21103a6ffe5764c30',1,'prop_sdpredcost.c']]],
-  ['prop_5fname',['PROP_NAME',['../prop__sdpredcost_8c.html#a2062ba7b5292b2cb9456209022cbdac7',1,'prop_sdpredcost.c']]],
-  ['prop_5fpriority',['PROP_PRIORITY',['../prop__sdpredcost_8c.html#af54d22d68dc88e0702c6387381260bdd',1,'prop_sdpredcost.c']]],
-  ['prop_5ftiming',['PROP_TIMING',['../prop__sdpredcost_8c.html#a95706b8e33186c2eea7c0bc857974b46',1,'prop_sdpredcost.c']]]
-];

### 2016-03-10 www/doc-2.0.0/html/search/enums_0.js

@@ -1,4 +1,0 @@
-var searchData=
-[
-  ['scip_5fsdpparam',['SCIP_SDPParam',['../type__sdpi_8h.html#a4da74acdf76b107865a14e6235736ecb',1,'type_sdpi.h']]]
-];

### 2016-03-10 www/doc-2.0.0/html/search/enumvalues_0.js

@@ -1,8 +1,0 @@
-var searchData=
-[
-  ['scip_5fsdppar_5fepsilon',['SCIP_SDPPAR_EPSILON',['../type__sdpi_8h.html#a4da74acdf76b107865a14e6235736ecba6a8eeb8e29dd92a4d3b4c7d34aea44df',1,'type_sdpi.h']]],
-  ['scip_5fsdppar_5ffeastol',['SCIP_SDPPAR_FEASTOL',['../type__sdpi_8h.html#a4da74acdf76b107865a14e6235736ecbaf70f6673bba11efd01eb563751557eab',1,'type_sdpi.h']]],
-  ['scip_5fsdppar_5fobjlimit',['SCIP_SDPPAR_OBJLIMIT',['../type__sdpi_8h.html#a4da74acdf76b107865a14e6235736ecbaf2d7cff478ef4a21433ebdc0300cb884',1,'type_sdpi.h']]],
-  ['scip_5fsdppar_5fsdpinfo',['SCIP_SDPPAR_SDPINFO',['../type__sdpi_8h.html#a4da74acdf76b107865a14e6235736ecba750cbe3fec9a21f97cec62eb4ec5cd14',1,'type_sdpi.h']]],
-  ['scip_5fsdppar_5fslatercheck',['SCIP_SDPPAR_SLATERCHECK',['../type__sdpi_8h.html#a4da74acdf76b107865a14e6235736ecba4ca6515ba8300692a64c746c507d4de4',1,'type_sdpi.h']]]
-];

### 2016-03-10 www/doc-2.0.0/html/search/files_0.js

@@ -1,12 +1,0 @@
-var searchData=
-[
-  ['blockmemoryallocator_2eh',['BlockMemoryAllocator.h',['../BlockMemoryAllocator_8h.html',1,'']]],
-  ['branch_5fsdpinfobjective_2ec',['branch_sdpinfobjective.c',['../branch__sdpinfobjective_8c.html',1,'']]],
-  ['branch_5fsdpinfobjective_2eh',['branch_sdpinfobjective.h',['../branch__sdpinfobjective_8h.html',1,'']]],
-  ['branch_5fsdpmostfrac_2ec',['branch_sdpmostfrac.c',['../branch__sdpmostfrac_8c.html',1,'']]],
-  ['branch_5fsdpmostfrac_2eh',['branch_sdpmostfrac.h',['../branch__sdpmostfrac_8h.html',1,'']]],
-  ['branch_5fsdpmostinf_2ec',['branch_sdpmostinf.c',['../branch__sdpmostinf_8c.html',1,'']]],
-  ['branch_5fsdpmostinf_2eh',['branch_sdpmostinf.h',['../branch__sdpmostinf_8h.html',1,'']]],
-  ['branch_5fsdpobjective_2ec',['branch_sdpobjective.c',['../branch__sdpobjective_8c.html',1,'']]],
-  ['branch_5fsdpobjective_2eh',['branch_sdpobjective.h',['../branch__sdpobjective_8h.html',1,'']]]
-];

### 2016-03-10 www/doc-2.0.0/html/search/files_a.js

@@ -1,4 +1,0 @@
-var searchData=
-[
-  ['xternal_2ec',['xternal.c',['../xternal_8c.html',1,'']]]
-];

### 2016-03-10 www/doc-2.0.0/html/search/functions_0.js

@@ -1,6 +1,0 @@
-var searchData=
-[
-  ['address',['address',['../classBlockMemoryAllocator.html#aab61d00366dbb389a729f646df9877c5',1,'BlockMemoryAllocator::address(reference x) const '],['../classBlockMemoryAllocator.html#ad7f6e9aaef497a3afdeb0750a56bbce6',1,'BlockMemoryAllocator::address(const_reference x) const ']]],
-  ['allocate',['allocate',['../classBlockMemoryAllocator.html#aa179675af57d9b44295a579f1ae03d37',1,'BlockMemoryAllocator']]],
-  ['allvarsfixed',['allVarsFixed',['../relax__sdp_8c.html#a53c3e9ea74f8ee83caa5252a1fd923b3',1,'relax_sdp.c']]]
-];

### 2016-03-10 www/doc-2.0.0/html/search/functions_8.js

@@ -1,8 +1,0 @@
-var searchData=
-[
-  ['main',['main',['../main_8cpp.html#a3c04138a5bfe5d72780bb7e82a18e627',1,'main.cpp']]],
-  ['max_5fsize',['max_size',['../classBlockMemoryAllocator.html#aa2909192b7f799f2b115ed99aff4028c',1,'BlockMemoryAllocator']]],
-  ['move_5f1x1_5fblocks_5fto_5flp',['move_1x1_blocks_to_lp',['../cons__sdp_8c.html#a0a6052e778b5ebbe7770a80935098325',1,'cons_sdp.c']]],
-  ['multiaggrvar',['multiaggrVar',['../cons__sdp_8c.html#abc88ebef3100f908bf1822bc71fd28ff',1,'cons_sdp.c']]],
-  ['multiplyconstraintmatrix',['multiplyConstraintMatrix',['../cons__sdp_8c.html#afe2bd14c1755e635d23858cc54ef147b',1,'cons_sdp.c']]]
-];

### 2016-03-10 www/doc-2.0.0/html/search/functions_b.js

@@ -1,4 +1,0 @@
-var searchData=
-[
-  ['runscip',['runSCIP',['../main_8cpp.html#a98aa1ab7fd1b6063603b5adf2db43f62',1,'main.cpp']]]
-];

### 2016-03-10 www/doc-2.0.0/html/search/typedefs_0.js

@@ -1,5 +1,0 @@
-var searchData=
-[
-  ['const_5fpointer',['const_pointer',['../classBlockMemoryAllocator.html#acc3d6ef4de5ab53a9d81a22785aa6552',1,'BlockMemoryAllocator::const_pointer()'],['../classBlockMemoryAllocator_3_01void_01_4.html#ac13a041af178f7e56a22a8ba1e811eef',1,'BlockMemoryAllocator&lt; void &gt;::const_pointer()']]],
-  ['const_5freference',['const_reference',['../classBlockMemoryAllocator.html#a900914c750c80b3df960bf4170126885',1,'BlockMemoryAllocator']]]
-];

### 2016-03-10 www/doc-2.0.0/html/search/variables_7.js

@@ -1,4 +1,0 @@
-var searchData=
-[
-  ['scip_5f',['scip_',['../classBlockMemoryAllocator.html#a6d67035575784f26e270c4cf268a72c5',1,'BlockMemoryAllocator::scip_()'],['../classScipStreamBuffer.html#a4f8e6b65d7c0ade3f9c7ede1d7b0d5e3',1,'ScipStreamBuffer::scip_()']]]
-];

### 2016-03-10 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
removed MATLAB citation

@@
 But the needed proportion between $n$ and $k$ is so small that we can't expect the generated matrices in small dimensions to satisfy the RIP.
-For each type we used \mbox{MATLAB 8.3.0
-\cite{MAT}} to randomly generated nine matrices of three different sizes with rows ranging from 15 to 30, columns from 25 to 40 and order $k$ between three and five.
+For each type we used \mbox{MATLAB 8.3.0}
+to randomly generated nine matrices of three different sizes with rows ranging from 15 to 30, columns from 25 to 40 and order $k$ between three and five.
 

### 2016-03-10 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
corrected order of d'Aspremont citations

@@
 constants for a given matrix~$A$ and~$k$ is NP-hard, see~\cite{PT14}. On
-the other hand, d'Aspremont et al.~\cite{Asp07,Asp08} proposed several
+the other hand, d'Aspremont et al.~\cite{Asp08,Asp07} proposed several
 semidefinite relaxations to compute upper bounds on the restricted isometry
@@
 
-The relaxations in \cite{Asp07,Asp08} were originally derived for the deeply
+The relaxations in \cite{Asp08,Asp07} were originally derived for the deeply
 connected problem of sparse principal component analysis (SPCA). Principal

### 2016-03-10 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
removed sentence about presenting a methodology for MISDP
further changes in the order of d'Aspremont 08 and 07
small corrections

@@
 utilize this MISDP, we will show that the rank constraint can be relaxed
-without changing the optimal value. Moreover, we discuss a solution
-methodology to solve the resulting MISDP. It will turn out that
+without changing the optimal value. It will turn out that
 the resulting problems are quite hard to solve, even for matrices of
@@
 and~\eqref{l1min} in Section~\ref{sec:RIP}. In Section~\ref{sec:SDPrelax},
-we will then review the SDP-relaxations of~\cite{Asp07}
-and~\cite{Asp08}. Afterwards we will construct a mixed integer semidefinite
+we will then review the SDP-relaxations of~\cite{Asp08}
+and~\cite{Asp07}. Afterwards we will construct a mixed integer semidefinite
 programm and show that it produces exact restricted isometry constants in
@@
 we present numerical results and comparisons with the SDP relaxations in
-\cite{Asp07} and \cite{Asp08}.
+\cite{Asp08} and \cite{Asp07}.
 
@@
 
-As one way to solve (\ref{phi}) again the semidefinite lifting technique is used in \cite{Asp08} to compute an upper bound on $\phi$ via
+As one way to solve (\ref{phi}) the semidefinite lifting technique is used again in \cite{Asp08} to compute an upper bound on $\phi$ via
 \begin{align}\label{Asp08}
@@
 
-In the following we will to show that (\ref{MISDP}) not only computes a lower bound but the exact value of $\alpha_k$.
+In the following we will show that (\ref{MISDP}) not only computes a lower bound but the exact value of $\alpha_k$.
 For this we will use results about the geometry of semidefinite programs to show that there always exists an
@@
 \begin{proof}
- Let $\hat{X}, \hat{z}$ be an optimal solution to (\ref{MISDP}) with $\sum_{i=1}^n z_i = \ell \leq k$ 
- (because of $\Tr(X) = 1$ the problem is bounded and $X_{11} = 1, X_{ij} = 0 \ \forall i,j \neq 1,$ \mbox{$z_1 = 1, z_i = 0 \ \forall i \neq 1$} is always feasible for $k > 0$, so an optimal solution exists).
+ Because of the constraint $\Tr(X) = 1$, Problem~\eqref{MISDP} is bounded and $X_{11} = 1, X_{ij} = 0 \ \forall i,j \neq 1,$ \mbox{$z_1 = 1, z_i = 0 \ \forall i \neq 1$} 
+ is always feasible for $k > 0$,  so an optimal solution to~\eqref{MISDP} exists.
+ Let $\hat{X}, \hat{z}$ be such an optimal solution to (\ref{MISDP}) with $\sum_{i=1}^n z_i =: \ell \leq k$.
  Define
@@
  \text{s.t.} \quad & \Tr(\tilde{X}) = 1 \tag{Proj} \\
- & \tilde{X} \succeq 0 , \tilde{X} \in S_\ell, \nonumber 
- \end{align}\comment{$\tilde{X} \in S_\ell$ eigentlich in $\tilde{X} \succeq 0$ enthalten, aber betont halt Dimension, so drinnlassen oder \"andern?}
-where $\tilde{A} = (A_{i_1}, A_{i_2}, ..., A_{i_\ell})$ consists of those
+ & \tilde{X} \succeq 0  \nonumber 
+ \end{align}
+for $\tilde{X} \in S_\ell$, where $\tilde{A} = (A_{i_1}, A_{i_2}, ..., A_{i_\ell})$ consists of those
 columns $\{{i_1}, {i_2}, ..., {i_\ell}\} \enifed T$ of $A$ for which
@@
 
-So the optimal objective values of (\ref{Proj}) and (\ref{MISDP}) agree and with this construction we also get that 
-$\Rk(X^{\prime}) = \Rk(\check{X})$, so it suffices to show that (\ref{Proj}) always allows for a solution of rank one.
+So the optimal objective values of (\ref{Proj}) and (\ref{MISDP}) agree and with this construction also 
+$\Rk(X^{\prime}) = \Rk(\check{X})$, so it suffices to show that (\ref{Proj}) always allows for an optimal solution of rank one.
 Let $\breve{X}$ be an extreme point of (\ref{Proj}) (as the set of optimal points of (\ref{Proj}) is a non-empty, convex, compact subset of $\R^{\ell(\ell+1)/2}$, such a point exists
@@
 
-\begin{remark}\label{HalfConstraint}
+\begin{remark}\label{HalfConstraint} \comment{Als Remark belassen oder ein Lemma draus machen?}
 We can substitute the constraint $-z_j \leq X_{ij} \leq z_j$ for all non-diagonal entries of $X$ by the stronger constraint 

### 2016-03-10 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
added condition alpha_k =!= 0 to RIR
be more explicit that we are ommiting details in explanation of Asp08

@@
  respectively. The \emph{restricted isometry ratio} (RIR) $\gamma_k$ is
- defined as $\gamma_k \define \beta_k^2 / \alpha_k^2$.
+ defined as $\gamma_k \define \beta_k^2 / \alpha_k^2$ for $\alpha_k \neq 0$.
 \end{definition}
@@
 
-As one way to solve (\ref{phi}) the semidefinite lifting technique is used again in \cite{Asp08} to compute an upper bound on $\phi$ via
+As one way to solve (\ref{phi}) the semidefinite lifting technique is used again to compute an upper bound on $\phi$ via
 \begin{align}\label{Asp08}
@@
 \end{align}
-for $B_i = b_ib_i\T - \rho I,$ where $b_i$ is the $i$-th column of a squareroot of $\Sigma$. The corresponding dual problem can then be written as 
+for $B_i = b_ib_i\T - \rho I,$ where $b_i$ is the $i$-th column of a squareroot of $\Sigma$, for details see \cite{Asp08}. The corresponding dual problem can then be written as 
 \begin{align}\label{Asp08Dual}

### 2016-03-11 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
remove boldface around dim
capital letters for section headings
added numerical results

@@
 \usepackage{url}
+\usepackage{sistyle}
 
@@
 \newcommand{\comment}[1]{\marginpar{\raggedright\tiny {#1}}}
+
+% comma after thousands in numbers
+\SIthousandsep{,}
 
@@
 
-\section{The restriced isometry property}
+\section{The Restriced Isometry Property}
 \label{sec:RIP}
@@
 
-\section{SDP relaxations}
+\section{SDP Relaxations}
 \label{sec:SDPrelax}
@@
 
-\section{A MISDP for the restricted isometry property}
+\section{An MISDP for the Restricted Isometry Property}
 \label{sec:MISDP}
@@
  \begin{equation*}
-  \frac{1}{2}\Rk(X)\cdot(\Rk(X)+1) \leq m + \textbf{dim}(F).
+  \frac{1}{2}\Rk(X)\cdot(\Rk(X)+1) \leq m + dim(F).
  \end{equation*}
@@
 \begin{equation*}
- \frac{1}{2}\Rk(\breve{X})\cdot(\Rk(\breve{X})+1) \leq \tilde{m} + \textbf{dim}(F) = 1 + 0 = 1.
+ \frac{1}{2}\Rk(\breve{X})\cdot(\Rk(\breve{X})+1) \leq \tilde{m} + dim(F) = 1 + 0 = 1.
 \end{equation*}
@@
 
-For solving the mixed-integer semidefinite programs we used the MISDP-Solver SCIP-SDP \cite{SCIP-SDP} originally developed in \cite{MS12} and \cite{Mar13}, which combines the branch-and-bound-framework of SCIP 3.1 \cite{SCIP} with
-interior-point SDP-solvers. The continuous SDPs were solved using the command-line version of SDPA~\cite{SDPA6,SDPA7}. The testset consisted of 63 matrices of seven 
-different types, and for each matrix we computed the left- and right-hand restricted isometry constant $\alpha$ and $\beta$ using the problems (\ref{MISDP}), (\ref{Asp07}) and (\ref{Asp08Dual}).
-In the latter case we solved 15
+In this section we want to show the applicability of our proposed MISDP-formulation and use it to
+assess the quality of the relaxations proposed in \cite{Asp08} and \cite{Asp07}.
+For solving the mixed-integer semidefinite programs we use the MISDP-Solver SCIP-SDP \cite{SCIP-SDP} 
+originally developed in \cite{Mar13} and \cite{MS12}, which combines the branch-and-bound-framework of SCIP 3.2.1 \cite{SCIP} with 
+interior-point SDP-solvers, in our case SDPA~7.3.8~\cite{SDPA6,SDPA7}. The continuous SDPs are solved using the command-line version of 
+SDPA~7.3.8. In our tests we compute the left- and right-hand restricted isometry constant 
+$\alpha_k$ and $\beta_k$ either exactly using~\eqref{MISDP} or produce corresponding bounds via (\ref{Asp07}) and (\ref{Asp08Dual}). In the latter case we solve 15 
 SDPs for $\rho$ between $0$ and the maximum diagonal entry of $A\T A$.
-The matrices were band matrices with band size three or five with entries within the band chosen uniformly in $\{0,1\}$, binary matrices with all entries chosen
-uniformly in $\{0,1\}$ as well as matrices with entries chosen with distribution $\mathcal{N}(0,1)$ and rank one matrices $A = aa\T$, with $\mathcal{N}(0,1)$-distributed entries of $a$.
-In addition to these four kinds of matrices,
-we also used three more with theoretical results about their restricted isometry constants, namely matrices with $N(0, 1/m)$-distributed entries, with entries chosen uniformly in $\pm 1/\sqrt{m}$ and finaly with distribution
+
+The testset consists of 63 matrices of seven 
+different types including band matrices 
+with band size three or five with entries within the band chosen uniformly in $\{0,1\}$, binary matrices with all entries chosen 
+uniformly in $\{0,1\}$, Gaussian-distributed matrices and rank one 
+matrices $A = aa\T$ with $\mathcal{N}(0,1)$-distributed entries of $a$. In addition to these four kinds of matrices, 
+we also use three more with theoretical results about their restricted isometry constants, namely matrices with 
+$N(0, 1/m)$-distributed entries, with entries chosen uniformly in $\pm 1/\sqrt{m}$ and finaly with distribution
 \begin{equation}\nonumber
@@
 \end{equation}
-For the latter three kinds of matrices it was shown in \cite{BDDW08}, that for $n \rightarrow \infty$, sufficiently small $k$ and given $\delta$ they satisfy the restricted isometry property (\ref{rip}) of order $k$ for this $\delta$
-with probability converging exponentially to one.
-But the needed proportion between $n$ and $k$ is so small that we can't expect the generated matrices in small dimensions to satisfy the RIP.
-For each type we used \mbox{MATLAB 8.3.0}
-to randomly generated nine matrices of three different sizes with rows ranging from 15 to 30, columns from 25 to 40 and order $k$ between three and five.
+For the latter three kinds of matrices it is shown in \cite{BDDW08} that for $n \rightarrow \infty$, 
+sufficiently small $k$ and given $\delta$, they satisfy the restricted isometry property (\ref{rip}) of order $k$  
+with probability converging exponentially to one. The needed proportion between $n$ and $k$ is so small, however,
+that we cannot expect the generated matrices in small dimensions to satisfy the RIP. For each type we used \mbox{MATLAB 8.3.0} 
+to randomly generated nine matrices of three different sizes with rows ranging from 15 to 30, columns 
+from 25 to 40 and order $k$ between three and five.
 
-The tests were performed on an Intel i7-4770 with 3.4 GHz.
-For each instance we set a time limit of one hour, if a problem did not finish within this time, we counted 3600 seconds and used the dual bound as the result.
-In our test this
-only happened once for the right-hand-side of a $15 \times 30$ binary matrix with order $k=5$ in (\ref{MISDP}).
+The tests were performed on a Linux cluster with Intel i3 CPUs with 3.2GHz, 4MB cache and 8GB memory running
+Linux. Each computation was performed single-threaded with a single process running on each computer and with 
+a time limit of four hours. The code was compiled with gcc 4.4.5 with \texttt{-O3} optimization. For the 
+Problem~\eqref{Asp08Dual} most of the larger instance were aborted because of the memory limit, these are
+counted with the maximum allowed time of four hours.
 
-The average gap of the relaxations (\ref{Asp07}) and (\ref{Asp08Dual}) in contrast to the exact solution computed
-with (\ref{MISDP}) is shown in table \ref{results}. Results for the left hand sides of the band and rank one matrices where omitted, as they equalled zero within numerical exactness for all problems.
+In Table~\ref{lhsGap} we compare the relative gap between the bounds on $\alpha_k$ computed by the two heuristics
+and the exact values given by our
+mixed integer semidefinite program. We give arithmetic means over all of the nine matrices for the given type that
+could be solved within the time- and memory-limit. Note that we ommitted the rank one matrices, since these satisfy
+$\alpha_k = 0$ in all cases, so that the trivial lower bound would already be optimal. For the 
+left-hand-side of the RIP, both relaxations did not perform
+too well. Especially the relaxation given in \cite{Asp08} regularly returned worse solution than the
+trivial lower bound of zero. The relaxation in \cite{Asp07} performed slightly better, but the bounds
+were still not nearly good enough to have any chance of proving $\ell_0$-$\ell_1$-equivalance. The relatively
+bad performance of both relaxations for the left-hand side might not be too surprising, since both relaxations
+were originally designed for sparse principal component analysis, which corresponds to the right-hand side
+of the RIP.
 
-\begin{table}[ht]
-   \centering
-        \caption{Average gap of relaxations for restricted isometry constants $\alpha_k$ and $\beta_k$}
-   \label{results}
-   \begin{tabular*}{\textwidth}{l @{\extracolsep{\fill}} c r r c r r}
-   \toprule
-  & & \multicolumn{2}{c}{$\alpha_k^2$} & \phantom{abc} & \multicolumn{2}{c}{$\beta_k^2$} \\
- \cmidrule{3-4} \cmidrule{6-7} 
-type of matrix & & (\ref{Asp07}) & (\ref{Asp08Dual}) & & (\ref{Asp07}) & (\ref{Asp08Dual}) \\
-\midrule
-$\mathcal{N}(0,1)$ & & $85.45\%$ & todo &  & $23.84\%$ & todo\\
-binary & & $85.28\%$ & todo &  & $47.12\%$ & todo\\
-band matrix & & - & todo &  & $12.97\%$ & todo\\
-rank 1 & & - & todo &  & $50.94\%$ & todo\\
-$\mathcal{N}(0,1/m)$ & & $88.36\%$ & todo &  & $26.05\%$ & todo\\
-$\pm 1/\sqrt{m}$ & & $85.35\%$ & todo &  & $29.53\%$ & todo\\
-$0, \pm \sqrt{3/m}$ & & $87.45\%$ & todo &  & $23.05\%$ & todo\\
-\midrule
-average & & $86.38\%$ & todo &  & $29.92\%$ & todo \\
-\bottomrule
-\end{tabular*}
-\end{table}
+\input{tables/lhsGap}
 
-As we can see from the results, for the right- and especially for the left-hand-side for all of our instances there was quite a big gap for the relaxation in (\ref{Asp07}) which we were able to close with our exact optimization
-problem (\ref{MISDP}).
-For the right-hand-side the relaxation from \cite{Asp07} performed quite well for band matrices with a gap of $13 \%$ and satisfactorly for most other classes of matrices, except for binary matrices, where in
-our testset the average gap reached almost $50\%$.
-On the left-hand-side however, the gap of the relaxation was between $85\%$ and $90\%$ for all classes, which would make it extremely hard to verify the requirements of theorem
-\ref{RIPsqrt2} or theorem \ref{asymRIPtheorem}.
+Results for the right-hand side, i.e. $\beta_k$, are given in Table~\ref{rhsGap}. Here both relaxations
+perform much better. Problem~\eqref{Asp08Dual} constantly produced bounds within 10 to 35\,\% of the optimal
+solution. The formulation in~\eqref{Asp08Dual} showed bigger fluctuation, but was very tight in many cases,
+with a gap of less than two percent for band matrices.
 
-The average solving times of (\ref{MISDP}), (\ref{Asp07}) and (\ref{Asp08Dual}) in seconds are shown in table \ref{times}.
+\input{tables/rhsGap}
 
-\begin{table}[ht]
-   \centering
-        \caption{Average solving times for left- and right-hand-sides (seconds)}
-   \label{times}
-   \begin{tabular*}{\textwidth}{l @{\extracolsep{\fill}} c r r c r r c r r}
-   \toprule
-  & & \multicolumn{2}{c}{(\ref{MISDP})} & \phantom{abc} & \multicolumn{2}{c}{(\ref{Asp07})} & \phantom{abc} & \multicolumn{2}{c}{(\ref{Asp08Dual})} \\
- \cmidrule{3-4} \cmidrule{6-7} \cmidrule{9-10} 
-type of matrix & & l & r &  & l & r &  & l & r \\
-\midrule
-$\mathcal{N}(0,1)$ & & 2145.35 & 1489.05 & & 5.35 & 5.6 & & todo & todo\\
-binary & & 2009.65 & 2586.0 & & 5.05 & 3.85 & & todo & todo\\
-band matrix & & 75.65 & 732.75 & & 3.1 & 5.55 & & todo & todo\\
-rank 1 & & 222.85 & 1658.1 & & 4.4 & 5.05 & & todo & todo\\
-$\mathcal{N}(0,1/m)$ & & 2035.05 & 1488.05& & 7.4 & 5.8 & & todo & todo\\
-$\pm 1/\sqrt{m}$ & & 2106.3 & 1537.45 & & 3.7 & 4.3 & & todo & todo\\
-$0, \pm \sqrt{3/m}$ & & 1731.45 & 1309.8 & & 5.6 & 5.35 & & todo & todo\\
-\midrule
-average & & 1475.19 & 1543.03 & & 4.94 & 5.07 & & todo & todo \\
-\bottomrule
-\end{tabular*}
-\end{table}
+Solving times for the left- and right-hand side are given in Table~\ref{lhsTime} and~\ref{rhsTime}. The relaxation proposed in \cite{Asp07}
+was very fast, taking only a few seconds for all instances. The mixed-integer semidefinite program obviously
+cannot compete with these solving times, although for band-matrices, specifically, the difference was relatively
+small. The Lagrangian Relaxation from \cite{Asp08}, however, took more time to produce bounds than was needed
+by our MISDP to compute exact solutions. The big difference is also caused by the instances running into the 
+memory limit, but even without those, the algorithm still took slightly longer than the MISDP.
 
-Here we can see the advantage of the relaxation proposed in \cite{Asp07}.
-While it won't produce exact results, it can solve instances, where the mixed integer semidefinite problem takes 30 or more minutes, within a few seconds.
-With the mixed integer semidefinite program we could only solve instances up to size $40$ and order $5$ within our time limit of 60 minutes, which is far from the amount needed for the instances used in practical applications.
-But this still gives us a way to compute restricted isometry constants apart from total enumeration, if they need to be known exactly.
-Also mixed integer semidefinite programming is still a relatively new subject of research, and like
-in the mixed integer linear case, there should be hope for improvements in the algorithms in the future, which would make this approach much more feasible.
+\input{tables/lhsTime}
 
-In table \ref{nodes} we compare the number of nodes in the branch-and-bound-tree for (\ref{MISDP}) with the number of eigenvalue problems that would have to be solved to compute the restricted isometry constants via total enumeration.
-For this we computed the ratio of branch-and-bound nodes to submatrices of order $k$ of $A$, of which there exist $\binom{n}{k}$.
-
-\begin{table}[ht]
-   \centering
-        \caption{Branch-and-bound-nodes as percentage of total enumeration}
-   \label{nodes}
-   \begin{tabular*}{\textwidth}{l @{\extracolsep{\fill}} r r}
-   \toprule
-type of matrix & left & right \\
-\midrule
-$\mathcal{N}(0,1)$ & $76,94\%$ & $59.52\%$\\
-binary & $89.40\%$ & $138.87\%$\\
-band matrix & $0.77\%$ & $0.17\%$ \\
-rank 1 & $1.01\%$ & $30.64\%$\\
-$\mathcal{N}(0,1/m)$ & $59.85\%$ & $58.85\%$\\
-$\pm 1/\sqrt{m}$ & $90.74\%$ & $69.49\%$\\
-$0, \pm \sqrt{3/m}$ & $75.96\%$ & $60.68\%$\\
-\midrule
-average & $56.38\%$ & $61.87\%$ \\
-\bottomrule
-\end{tabular*}
-\end{table}
-
-While the mixed integer semidefinite program performed quite well for band and rank one matrices, it didn't perform too well for most other clases, sometimes even needing more branch-and-bound nodes than there are submatrices of $A$
-(which can happen, as the latter only appear as leaves in the branch-and-bound-tree).
-One reason for this is that sparse eigenvalues will usually be lower than non-sparse eigenvalues, leading to fractional solution as long as there
-are still unfixed binary variables, while integer solutions only appear in the leaves.
-But for these computations we only used general heuristics and branching rules as implemented in SCIP, so there might be chances for improvements with
-specially tailored heuristics and branching rules for this kind of problem.
-Also our tests showed, that the ratio gets smaller with larger $k$, so for bigger instances the performance of the algorithm in contrast to total enumeration
-should increase.
+\input{tables/rhsTime}
 
@@
 We have seen that exact restricted isometry constants can be computed not only via total enumeration but also by solving a mixed integer semidefinite program.
-Because of this exactness result, this new problem could drastically
+Because of this exactness result, the new problem could drastically
 improve the bounds computed with the semidefinite relaxations proposed in \cite{Asp07} and \cite{Asp08}, but this came at the cost of a much higher complexity of computation, at least in comparison to \cite{Asp07}.

### 2016-03-11 RIP-MISDP-Paper/ResultScript.py

@@ -1,0 +1,424 @@
+import os
+import sys
+import math
+from decimal import Decimal
+
+#TODO: write tables for online supplement
+
+#which tables/plots for the paper should be prepared?
+texfile = 0 # make the output a compilable texfile or just a figure/table that can be included?
+Lbounds = 0
+Rbounds = 0
+Ltimes = 0
+Rtimes = 1
+completeTable = 0
+
+MISDPfilename = "/local/gally/results/RIP-MISDP-Paper/160311/RIP-results/check.RIPMISDP.scipsdp.linux.x86_64.gnu.opt.sdpa.extra.branchinfobj_nofracdive.out"
+Asp0708path = "/local/gally/results/RIP-MISDP-Paper/160311/SDPA-results/"
+Asp07filenames = ["0+-115305A_07l.out", "0+-115305A_07r.out", "0+-115305B_07l.out", "0+-115305B_07r.out", "0+-115305C_07l.out", "0+-115305C_07r.out", "0+-125354A_07l.out", "0+-125354A_07r.out", "0+-125354B_07l.out", "0+-125354B_07r.out", "0+-125354C_07l.out", "0+-125354C_07r.out", "0+-130403A_07l.out", "0+-130403A_07r.out", "0+-130403B_07l.out", "0+-130403B_07r.out", "0+-130403C_07l.out", "0+-130403C_07r.out", "bina15305A_07l.out", "bina15305A_07r.out", "bina15305B_07l.out", "bina15305B_07r.out", "bina15305C_07l.out", "bina15305C_07r.out", "bina25354A_07l.out", "bina25354A_07r.out", "bina25354B_07l.out", "bina25354B_07r.out", "bina25354C_07l.out", "bina25354C_07r.out", "bina30403A_07l.out", "bina30403A_07r.out", "bina30403B_07l.out", "bina30403B_07r.out", "bina30403C_07l.out", "bina30403C_07r.out", "bern15305A_07l.out", "bern15305A_07r.out", "bern15305B_07l.out", "bern15305B_07r.out", "bern15305C_07l.out", "bern15305C_07r.out", "bern25354A_07l.out", "bern25354A_07r.out", "bern25354B_07l.out", "bern25354B_07r.out", "bern25354C_07l.out", "bern25354C_07r.out", "bern30403A_07l.out", "bern30403A_07r.out", "bern30403B_07l.out", "bern30403B_07r.out", "bern30403C_07l.out", "bern30403C_
[…]

### 2016-03-11 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
reworked introduction

@@
 
+\section{The Restriced Isometry Property}
+\label{sec:RIP}
+
 Several conditions in compressed sensing imply that a solution
@@
 One important condition is the \emph{restricted isometry property} (RIP),
-which was first introduced by Cand\`es and Tao~\cite{CT05}. For each $k \in
-\{1, \dots, n\}$, the RIP involves a \emph{restricted isometry constant}
-$\delta_k$ bounding the $k$-sparse eigenvalues of $A$. If $\delta_k$ is sufficiently small
-and there exists $\tilde{x} \in \R^n$ with $A \tilde{x} = b$ and
-$\Norm{\tilde{x}}{0} \leq k$, then problems~\eqref{l0min} and~\eqref{l1min}
-are equivalent; see Section~\ref{sec:RIP} for more details. There are
-several types of matrices~$A$ for which the RIP holds, for instance random
-matrices, see, e.g., Baraniuk et al.~\cite{BDDW08}. Although the RIP does
-not provide necessary conditions, it is one of the major tools for
-establishing the $\ell_0$-$\ell_1$-equivalence.
-
-We note that recovery results can also be derived for the \emph{denoising}
-case, in which instead of~\eqref{l0min} one considers
-\begin{align}\label{l0minDenoising}
-  \min \quad & \Norm{x}{0} \tag{$P_0^\delta$}\\
-  \text{s.t.} \quad & \Norm{Ax - b}{2} \leq \delta, \nonumber
-\end{align}
-for some $\delta \geq 0$; the $\ell_1$-problem~\eqref{l1min} is changed
-similarly. For more details we again refer the reader to~\cite{FouR13}.
-
-However, one drawback of the RIP is that computing the restricted isometry
-constants for a given matrix~$A$ and~$k$ is NP-hard, see~\cite{PT14}. On
-the other hand, d'Aspremont et al.~\cite{Asp08,Asp07} proposed several
-semidefinite relaxations to compute upper bounds on the restricted isometry
-constants. Note that small upper bounds suffice to guarantee
-$\ell_0$-$\ell_1$-equivalence between~\eqref{l0min} and~\eqref{l1min}, but
-there exist cases in which the bounds are too weak to give a guarantee.
-
-The relaxations in \cite{Asp08,Asp07} were originally derived for the deeply
-connected problem of sparse principal component analysis (SPCA). Principal
-component analysis (PCA) is about finding orthogonal directions maximizing the 
-variance given in matrix $A$. SPCA tries to fix the problem that the components
-of PCA usually consist of small contributions of all axes, which make the results
-hard to interpret, by enforcing sparsity on the components. SPCA was
-originally proposed by Zuo et al. in \cite{SPCA}, where a penalty term was added 
-to the objective. Other authors like Moghaddam et al. in \cite{sbSPCA} also proposed 
-computing a sparse maximal eigenvalue for SPCA, like it is done for the RIP.
-
-The goal of this paper is to compute optimal restricted isometry constants
-via a mixed integer semidefinite program (MISDP). To this end, we will
-introduce an MISDP with a rank-1 constraint. In order to
-utilize this MISDP, we will show that the rank constraint can be relaxed
-without changing the optimal value. It will turn out that
-the resulting problems are quite hard to solve, even for matrices of
-smaller size. However, the results allow an evaluation of the SDP
-relaxations of d'Aspremont et al. Depending on the input matrices, these
-bounds can be quite tight or weak. This complements the theoretical
-approximation guarantees developed by d'Aspremont et
-al. in~\cite{AspBG14}.
-
-In the following, we will start by introducing the restricted isometry
-property and corresponding results about the equivalence of~\eqref{l0min}
-and~\eqref{l1min} in Section~\ref{sec:RIP}. In Section~\ref{sec:SDPrelax},
-we will then review the SDP-relaxations of~\cite{Asp08}
-and~\cite{Asp07}. Afterwards we will construct a mixed integer semidefinite
-programm and show that it produces exact restricted isometry constants in
-Section~\ref{sec:MISDP}. Finally, in Section~\ref{sec:numerical_results},
-we present numerical results and comparisons with the SDP relaxations in
-\cite{Asp08} and \cite{Asp07}.
-
-\section{The Restriced Isometry Property}
-\label{sec:RIP}
-
-In this section we will review some general results about the restricted
-isometry property and its relationship with the optimization problems
-(\ref{l0min}) and (\ref{l1min}). We start with the definition of the RIP as
-in \cite{CT05}:
+which was first introduced by Cand\`es and Tao~\cite{CT05}.
 
@@
 
-As the following results depend on the constant $\delta$ in (\ref{rip}) being
+Since the following results depend on the constant $\delta$ in (\ref{rip}) being
 small enough, we are generally interested in the smallest constant satisfying
@@
 \end{definition}
+
+If the restricted isometry property holds for a sufficiently small
+restricted isometry constant, then the optimal solutions of~\eqref{l0min}
+and~\eqref{l1min} coincide.  The best-known result of this type is probably
+the following:
+
+\begin{theorem}[Cand\'es~\cite{Can08}]\label{RIPsqrt2}
+  Let $A$ satisfy $\delta_{2k} < \sqrt{2}-1 \approx 0.4142$. If there
+  exists $\tilde{x} \in \R^n$ with $\Norm{\tilde{x}}{0} \leq k$ such that
+  $A\tilde{x} = b$, then the optimal solutions of~\eqref{l0min}
+  and~\eqref{l1min} coincide.
+\end{theorem}
+
+There are also many other results similar to Theorem~\ref{RIPsqrt2}: For
+example, Foucart~\cite{fou10} relaxed the condition to $\delta_{2k} <
+0.4652$ and Foucart and Rauhut~\cite{FouR13} to $\delta_{2k} < 0.6246$.
+Moreover, Cai et al.~\cite{CWX10} gave the condition $\delta_k <
+0.307$ for Theorem~\ref{RIPsqrt2} to hold.
 
@@
 
- % If the restricted isometry property holds for a sufficiently small restricted isometry constant, then the optimal solutions of (\ref{l0min}) and (\ref{l1min}) coincide.
-To show this, one needs estimates of the form
- %  \begin{equation}\label{l1ErrorRIP}
- %  \Norm{x^* - x} \leq C \Norm{x - x_k},
- % \end{equation}
- % where $x^*$ is an optimal solution of (\ref{l1min}) with $b = Ax$ and $x_k$ is the best approximation of some vector $x \in \R^n$ with $k$-sparse vectors.
-If we get a result like (\ref{l1ErrorRIP}), and we know that the vector $x$, which in the
- % applications would be the sparse signal that needs to be recovered, is already $k$-sparse itself, then the best approximation of $x$ with $k$-sparse vectors is clearly $x$ itself.
-This means that the right-hand-side in
- % (\ref{l1ErrorRIP}) equals zero, and therefore the unique optimal solution of (\ref{l1min}) has to be $x$ again.
-Now (\ref{l1ErrorRIP}) not only holds for this $x$ exclusively, but for any $k$-sparse vector $y$ with $Ay=Ax=b$.
-And
- % because of $Ay=Ax=b$, we get the same problem in (\ref{l1min}) and therefore also the same solution $x^*$ which needs to equal $y$ because of (\ref{l1ErrorRIP}), so this means that there can only be a single vector $x$ with sparsity
- % at most $k$ for which $Ax=b$ holds, which then also has to solve (\ref{l0min}), which gives us the equivalance of these two problems.
-
-If the restricted isometry property holds for a sufficiently small
-restricted isometry constant, then the optimal solutions of~\eqref{l0min}
-and~\eqref{l1min} coincide.  The best-known result of this type is probably
-the following:
-
-\begin{theorem}[Cand\'es~\cite{Can08}]\label{RIPsqrt2}
-  Let $A$ satisfy $\delta_{2k} < \sqrt{2}-1 \approx 0.4142$. If there
-  exists $\tilde{x} \in \R^n$ with $\Norm{\tilde{x}}{0} \leq k$ such that
-  $A\tilde{x} = b$, then the optimal solutions of~\eqref{l0min}
-  and~\eqref{l1min} coincide.
-\end{theorem}
-
-There are also many other results similar to Theorem~\ref{RIPsqrt2}: For
-example, Foucart~\cite{fou10} relaxed the condition to $\delta_{2k} <
-0.4652$ and Foucart and Rauhut~\cite{FouR13} to $\delta_{2k} < 0.6246$.
-\comment{Ergebnis ist f\"ur $\C$, sollte aber auch f\"ur $\R$
-gelten, da man RIC \"uber $\C$ durch Aufspaltung in Real- und 
-Imagin\"arteil durch reelle RIC absch\"atzen k\"onnen sollte.}
-Moreover, Cai et al.~\cite{CWX10} gave the condition $\delta_k <
-0.307$ for Theorem~\ref{RIPsqrt2} to hold. Note that all of these results
-can be generalized to the denoising case~\eqref{l0minDenoising}.
-
-Similar results hold for the asymmetric case. For instance, we have:
+For the asmmetric version of the RIP in Definition~\ref{asymRIPdef}, we
+can again show $\ell_0$-$\ell_1$-equivalence:
 
@@
 
-Again this result can be generalized to the denoising case. One advantage
+One advantage
 of Definition~\ref{asymRIPdef} and Theorem~\ref{asymRIPtheorem} is that they
 are invariant under scaling of $Ax = b$. The condition of
-Theorem~\ref{RIPsqrt2} holding for~$A x = b$ does not imply it for the scaled
-$\lambda Ax = \lambda b$. However, in Definition \ref{asymRIPdef}, we can
+Theorem~\ref{RIPsqrt2} holding for~$A x = b$ does not imply the same 
+for the scaled equation
+$\lambda Ax = \lambda b$. In Definition \ref{asymRIPdef}, however, we can
 use $\tilde{\alpha}_k = \lambda \alpha_k$, $\tilde{\beta}_k = \lambda
@@
 
-In the following we will design an approach to compute the optimal
-constants $\alpha_k$ and $\beta_k$. These can then be used to compute
-either the restricted isometry ratio $\gamma_k \define \beta_k^2 / \alpha_k^2$
-or the restricted isometry constant $\delta_k = \max \{ 1 - \alpha_k^2,\;
-\beta_k^2 - 1 \}$.
+Furthermore, the recovery results given in this section can also be derived 
+for the \emph{denoising} case, where instead of~\eqref{l0min} one considers
+\begin{align}\label{l0minDenoising}
+  \min \quad & \Norm{x}{0} \tag{$P_0^\delta$}\\
+  \text{s.t.} \quad & \Norm{Ax - b}{2} \leq \delta, \nonumber
+\end{align}
+for some $\delta \geq 0$; the $\ell_1$-problem~\eqref{l1min} is changed
+similarly. For more details we refer the reader to~\cite{FouR13}.
+
+\subsection{Literature Overview and Contribution of this Work}
+
+There are several types of matrices~$A$ for which it is possible
+to show analytically that the RIP holds, for instance random
+matrices, see, e.g., Baraniuk et al.~\cite{BDDW08}.
+Still, a big drawback of the RIP is that computing the restricted isometry
+constants for a given matrix~$A$ and~$k$ is NP-hard, see~\cite{PT14}. 
+On the other hand, d'Aspremont et al.~\cite{Asp08,Asp07} proposed several
+semidefinite relaxations to compute upper bounds on the restricted isometry
+constants. Note that small upper bounds suffice to guarantee
+$\ell_0$-$\ell_1$-equivalence between~\eqref{l0min} and~\eqref{l1min}, but
+there exist cases in which the bounds are too weak to give a guarantee.
+
+The relaxations in \cite{Asp08,Asp07} were originally derived for the deeply
+connected problem of sparse principal component analysis (SPCA). Principal
+component analysis (PCA) is about finding orthogonal directions maximizing the 
+variance given in matrix $A$. SPCA tries to fix the problem that the components
+of PCA usually consist of small contributions of all axes, which make the results
+hard to interpret, by enforcing sparsity on the components. SPCA was
+originally proposed by Zuo et al. in \cite{SPCA}, where a penalty term was added 
+to the objective. Other authors like Moghaddam et al. in \cite{sbSPCA} also proposed 
+computing a sparse maximal eigenvalue for SPCA, like it is done for the RIP.
+
+The goal of this paper is to compute optimal restricted isometry constants
+via a mixed integer semidefinite program (MISDP). To this end, we will
+introduce an MISDP with a rank-1 constraint. In order to
+utilize this MISDP, we will show that the rank constraint can be relaxed
+without changing the optimal value. It will turn out that
+the resulting problems are quite hard to solve, even for matrices of
+smaller size. However, the results allow an evaluation of the SDP
+relaxations of d'Aspremont et al. Depending on the input matrices, these
+bounds can be quite tight or weak. This complements the theoretical
+approximation guarantees developed by d'Aspremont et
+al. in~\cite{AspBG14}.
+
+In the following, we will start by reviewing the SDP-relaxations of~\cite{Asp08}
+and~\cite{Asp07} in Section~\ref{sec:SDPrelax}.
+Afterwards, we will construct a mixed integer semidefinite
+programm and show that it produces exact restricted isometry constants in
+Section~\ref{sec:MISDP}. Finally, in Section~\ref{sec:numerical_results},
+we present numerical results and comparisons with the SDP relaxations in
+\cite{Asp08} and \cite{Asp07}.
 
@@
 
-In this section we will discuss two known SDP relaxations to compute an upper bound on the restricted isometry constant $\delta_k$ in (\ref{rip}), as they have some ideas in common with the mixed integer semidefinite program we want to
+In this section, we will discuss two known SDP relaxations to compute an upper bound on the restricted isometry constant $\delta_k$ in (\ref{rip}), as they have some ideas in common with the mixed integer semidefinite program we want to
 propose and will also be checked against in the section on numerical results.
@@
 we also use three more with theoretical results about their restricted isometry constants, namely matrices with 
-$N(0, 1/m)$-distributed entries, with entries chosen uniformly in $\pm 1/\sqrt{m}$ and finaly with distribution
+$N(0, 1/m)$-distributed entries, with entries chosen uniformly in $\pm 1/\sqrt{m}$ and finally with distribution
 \begin{equation}\nonumber
@@
 that we cannot expect the generated matrices in small dimensions to satisfy the RIP. For each type we used \mbox{MATLAB 8.3.0} 
-to randomly generated nine matrices of three different sizes with rows ranging from 15 to 30, columns 
+to randomly generate nine matrices of three different sizes with rows ranging from 15 to 30, columns 
 from 25 to 40 and order $k$ between three and five.
@@
 left-hand-side of the RIP, both relaxations did not perform
-too well. Especially the relaxation given in \cite{Asp08} regularly returned worse solution than the
+too well. Especially the relaxation given in \cite{Asp08} regularly returned worse solutions than the
 trivial lower bound of zero. The relaxation in \cite{Asp07} performed slightly better, but the bounds
@@
 Results for the right-hand side, i.e. $\beta_k$, are given in Table~\ref{rhsGap}. Here both relaxations
-perform much better. Problem~\eqref{Asp08Dual} constantly produced bounds within 10 to 35\,\% of the optimal
-solution. The formulation in~\eqref{Asp08Dual} showed bigger fluctuation, but was very tight in many cases,
+perform much better. Problem~\eqref{Asp08Dual} constantly produces bounds within 10 to 35\,\% of the optimal
+solution. The formulation in~\eqref{Asp08Dual} shows bigger fluctuation, but is very tight in many cases,
 with a gap of less than two percent for band matrices.
@@
 Solving times for the left- and right-hand side are given in Table~\ref{lhsTime} and~\ref{rhsTime}. The relaxation proposed in \cite{Asp07}
-was very fast, taking only a few seconds for all instances. The mixed-integer semidefinite program obviously
-cannot compete with these solving times, although for band-matrices, specifically, the difference was relatively
-small. The Lagrangian Relaxation from \cite{Asp08}, however, took more time to produce bounds than was needed
+is very fast, taking only a few seconds for all instances. The mixed-integer semidefinite program obviously
+cannot compete with these solving times, although for band-matrices, specifically, the difference is relatively
+small. The Lagrangian Relaxation from \cite{Asp08}, however, takes more time to produce bounds than is needed
 by our MISDP to compute exact solutions. The big difference is also caused by the instances running into the 
-memory limit, but even without those, the algorithm still took slightly longer than the MISDP.
+memory limit, but even without those, the algorithm still takes slightly longer than the MISDP.
 
@@
 We have seen that exact restricted isometry constants can be computed not only via total enumeration but also by solving a mixed integer semidefinite program.
-Because of this exactness result, the new problem could drastically
-improve the bounds computed with the semidefinite relaxations proposed in \cite{Asp07} and \cite{Asp08}, but this came at the cost of a much higher complexity of computation, at least in comparison to \cite{Asp07}.
+Because of this exactness result, the new problem can drastically
+improve the bounds computed with the semidefinite relaxations proposed in \cite{Asp08} and \cite{Asp07}, but this comes at the cost of a much higher complexity of computation, at least in comparison to \cite{Asp07}.
 We could only
-solve these problems for relatively small instances, but this could change with sufficient advances in the area of mixed integer semidefinite programming, especially with the addition of special cutting planes, which proved to be
+solve these problems for relatively small instances, but this might change with sufficient advances in the area of mixed integer semidefinite programming, especially with the addition of special cutting planes, which proved to be
 the main enhancement needed to make mixed integer linear solvers as fast as they are today.

### 2016-03-11 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
made remark lemma and further small corrections

@@
 \newtheorem{remark}{Remark}
+\newtheorem{lemma}[theorem]{Lemma}
 
@@
 
-\title{Computing exact restricted isometry constants via\\ mixed integer semidefinite programming}
+\title{Computing Exact Restricted Isometry Constants via\\ Mixed Integer Semidefinite Programming}
 \author{Tristan~Gally and Marc~E.~Pfetsch
@@
 An important task in compressed sensing is finding the sparsest solution to an underdetermined system of linear equations.
-An important result is that, under certain conditions, this is equivalent to minimizing the 1-Norm and thus
+An outstanding result is that, under certain conditions, this is equivalent to minimizing the 1-Norm and thus
 solving a linear program.
@@
 equations is one key problem in compressed sensing and has been
-thoroughly investigated in recent years. Given a matrix $A \in \R^{m \times n}$ with $m < n$ and 
-right-hand side $b \in \R^m$, the problem is to compute
+thoroughly investigated in recent years. Given a matrix $A \in \R^{m \times n}$ with $m \ n$ and 
+right-hand side $b \in \R^m$, one core problem is to compute
 \begin{align}\label{l0min}
@@
 where $\Norm{x}{0}$ denotes the number of nonzero components of~$x$. This
-problem is \NP-hard, see for example the book by Garey and
+problem is \NP-hard, see Garey and
 Johnson~\cite{GareyJohnson}. One key observation in compressed sensing is
@@
 For a matrix $A \in \R^{m \times n}$ and order~$k$, the smallest constant in 
-(\ref{rip}), meaning
+(\ref{rip}),
 \begin{equation}\label{ric}
@@
 
-There are also many other results similar to Theorem~\ref{RIPsqrt2}: For
+Many results similar to Theorem~\ref{RIPsqrt2} exist: For
 example, Foucart~\cite{fou10} relaxed the condition to $\delta_{2k} <
 0.4652$ and Foucart and Rauhut~\cite{FouR13} to $\delta_{2k} < 0.6246$.
-Moreover, Cai et al.~\cite{CWX10} gave the condition $\delta_k <
+Moreover, Cai et al.~\cite{CWX10} showed the condition $\delta_k <
 0.307$ for Theorem~\ref{RIPsqrt2} to hold.
@@
 For this paper, it is useful to take an asymmetric viewpoint and
-distinguish between the lower and upper bounds in~\eqref{rip}, as was done
-in~\cite{FL09}:
+distinguish between the lower and upper bounds in~\eqref{rip}, as proposed
+by Foucart and Lai~\cite{FL09}:
 
 \begin{definition}\label{asymRIPdef}
- The \emph{lower} and \emph{upper RIC} are defined as
+ The \emph{lower} and \emph{upper restricted isometry constant} are defined as
  \begin{align}
@@
    \leq \Norm{Ax}{2}^2,\; \text{ for all }x \in \Sigma_k\}, \label{lowerRIP}\\
-   \beta_k & \define \argmin_{\beta \geq 1}\; \{ \Norm{Ax}{2}^2 \leq \beta^2
+   \beta_k & \define \argmin_{\beta \geq 0}\; \{ \Norm{Ax}{2}^2 \leq \beta^2
    \Norm{x}{2}^2,\; \text{ for all }x \in \Sigma_k\},\label{upperRIP}
  \end{align}
- respectively. The \emph{restricted isometry ratio} (RIR) $\gamma_k$ is
+ respectively. The \emph{restricted isometry ratio} (RIR) is
  defined as $\gamma_k \define \beta_k^2 / \alpha_k^2$ for $\alpha_k \neq 0$.
@@
 
-In this section, we will discuss two known SDP relaxations to compute an upper bound on the restricted isometry constant $\delta_k$ in (\ref{rip}), as they have some ideas in common with the mixed integer semidefinite program we want to
+In this section, we will introduce two known SDP relaxations to compute an upper bound on the restricted isometry constant $\delta_k$ in (\ref{rip}), as they have some ideas in common with the mixed integer semidefinite program we want to
 propose and will also be checked against in the section on numerical results.
@@
 \end{align}
-The corresponding maximization problem can be solved to compute $\beta_k^2$.
+The corresponding maximization problem allows to compute $\beta_k^2$.
 In \cite{Asp07} the non-convex quadratic equality constraint $\Norm{x}{2}^2 = 1$ is tackled by the technique of semidefinite lifting first introduced in \cite{GW95}
@@
 
-\begin{remark}\label{HalfConstraint} \comment{Als Remark belassen oder ein Lemma draus machen?}
-We can substitute the constraint $-z_j \leq X_{ij} \leq z_j$ for all non-diagonal entries of $X$ by the stronger constraint 
+\begin{lemma}\label{HalfConstraint} 
+The constraint $-z_j \leq X_{ij} \leq z_j$ in~\eqref{MISDP} can be substituted for all non-diagonal entries of $X$ by the stronger constraint 
 \begin{equation}\label{eq:HalfConstraint}
@@
 \end{equation}
+\end{lemma}
+\begin{proof}
 Let $X$ satisfy $\Tr(X)=1$ and $X\succeq 0$, which immediately implies $X_{ii}\geq 0$.
@@
 $-z_j \leq X_{ij} \leq z_j$ and~\eqref{eq:HalfConstraint} for all integer values 
-of $z$, while~\eqref{eq:HalfConstraint} might cut off some solutions for fractional $z$.
-\end{remark}
+of $z$..
+\end{proof}
 
@@
 Because of this exactness result, the new problem can drastically
-improve the bounds computed with the semidefinite relaxations proposed in \cite{Asp08} and \cite{Asp07}, but this comes at the cost of a much higher complexity of computation, at least in comparison to \cite{Asp07}.
+improve the bounds computed by the semidefinite relaxations proposed in \cite{Asp08} and \cite{Asp07}, but this comes at the cost of a much higher complexity of computation, at least in comparison to \cite{Asp07}.
 We could only

### 2016-03-11 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
slightly extended proof of lemma

@@
 \begin{proof}
-Let $X$ satisfy $\Tr(X)=1$ and $X\succeq 0$, which immediately implies $X_{ii}\geq 0$.
-Since all minors of $X$ need to be nonnegative, we get that
+Let $X$ satisfy $\Tr(X)=1$ and $X\succeq 0$. Then $X_{ii}+X_{jj}\leq 1$ and nonnegativity of the diagonal 
+entries imply $X_{ii}X_{jj}\leq 1/4$. 
+Since all minors of $X$ need to be nonnegative, it follows that
 \begin{equation*}
@@
 $-z_j \leq X_{ij} \leq z_j$ and~\eqref{eq:HalfConstraint} for all integer values 
-of $z$..
+of $z$.
 \end{proof}

### 2016-03-12 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
restructured abstract

@@
 
+\noindent
 \begin{abstract}
-An important task in compressed sensing is finding the sparsest solution to an underdetermined system of linear equations.
-An outstanding result is that, under certain conditions, this is equivalent to minimizing the 1-Norm and thus
-solving a linear program.
-One such condition is the restricted isometry property for which no efficient algorithms are known to compute the optimal constants, the so called restricted isometry constants.
-In this paper we propose a
-mixed integer semidefinite program for computing these constants. Numerical results are presented comparing this to ealier SDP relaxations for the same task.
+  One of the fundamental tasks in compressed sensing is finding the
+  sparsest solution to an underdetermined system of linear equations. It is
+  well known that although this problem is \NP-hard, under certain
+  conditions it can be solved by using a linear program which minimizes the
+  1-norm. The restricted isometry property has been one of the key
+  conditions in this context. However, computing the best constants for
+  this condition is itself \NP-hard. In this paper we propose a
+  mixed-integer semidefinite programming approach for computing these
+  optimal constants. This also subsumes sparse principal component
+  analysis. Computational results with this approach allow to evaluate
+  ealier SDP relaxations and show that their quality is limited.
 \end{abstract}

### 2016-03-12 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
paper: some style changes in RIP section

@@
 equations is one key problem in compressed sensing and has been
-thoroughly investigated in recent years. Given a matrix $A \in \R^{m \times n}$ with $m \ n$ and 
-right-hand side $b \in \R^m$, one core problem is to compute
+thoroughly investigated in recent years. Given a matrix $A \in \R^{m \times n}$ with $m \leq n$ and 
+right-hand side $b \in \R^m$, the core problem is
 \begin{align}\label{l0min}
@@
 that a solution of~\eqref{l0min} can be found by solving the following
-convex optimization problem if certain conditions are fulfilled:
+convex optimization problem, if certain conditions are fulfilled:
 \begin{align}\label{l1min}
@@
 available both in theory and practice: as an extremely selected list
-consider the articles~\cite{BecBC11,CheDS99,OsbPT00,vdBF08}. For more
-information we refer the reader to~\cite{FouR13} and the practical
-comparison in~\cite{LorPT15}.
+consider the articles~\cite{BecBC11,CheDS99,OsbPT00,vdBF08} and the
+practical comparison in~\cite{LorPT15}.
 % Note: do not write out author names to keep text short at this place.
@@
 
+\noindent
 Several conditions in compressed sensing imply that a solution
@@
 
-Since the following results depend on the constant $\delta$ in (\ref{rip}) being
+Since the following results depend on the constant $\delta$ in \eqref{rip} being
 small enough, we are generally interested in the smallest constant satisfying
-(\ref{rip}).
+\eqref{rip}.
 
@@
   For a matrix $A \in \R^{m \times n}$ and order~$k$, the smallest constant in 
-(\ref{rip}),
+  \eqref{rip},
   \begin{equation}\label{ric}
     \begin{aligned}
-  \delta_k \define \argmin_{\delta \geq 0} \{&(1 - \delta) \Norm{x}{2}^2 \leq \Norm{Ax}{2}^2 \leq (1+\delta) \Norm{x}{2}^2 ,\\
-  & \text{ for all }x \in \Sigma_k\},
+      \delta_k \define \argmin_{\delta \geq 0} \big\{&(1 - \delta) \Norm{x}{2}^2 \leq \Norm{Ax}{2}^2 \leq (1+\delta) \Norm{x}{2}^2 ,\\
+      & \text{ for all }x \in \Sigma_k \big\},
     \end{aligned}
   \end{equation}
-is called \emph{restricted isometry constant} (RIC).
+  is called \emph{restricted isometry constant} (RIC) of order~$k$.
 \end{definition}
@@
 If the restricted isometry property holds for a sufficiently small
-restricted isometry constant, then the optimal solutions of~\eqref{l0min}
+restricted isometry constant, the optimal solutions of~\eqref{l0min}
 and~\eqref{l1min} coincide.  The best-known result of this type is probably
@@
   Let $A$ satisfy $\delta_{2k} < \sqrt{2}-1 \approx 0.4142$. If there
-  exists $\tilde{x} \in \R^n$ with $\Norm{\tilde{x}}{0} \leq k$ such that
+  exists $\tilde{x} \in \R^n$ with $\Norm{\tilde{x}}{0} \leq k$ and
   $A\tilde{x} = b$, then the optimal solutions of~\eqref{l0min}
@@
  \begin{align}
-   \alpha_k & \define \argmax_{\alpha \geq 0}\; \{ \alpha^2 \Norm{x}{2}^2
-   \leq \Norm{Ax}{2}^2,\; \text{ for all }x \in \Sigma_k\}, \label{lowerRIP}\\
-   \beta_k & \define \argmin_{\beta \geq 0}\; \{ \Norm{Ax}{2}^2 \leq \beta^2
-   \Norm{x}{2}^2,\; \text{ for all }x \in \Sigma_k\},\label{upperRIP}
+   \alpha_k & \define \argmax_{\alpha \geq 0}\; \big\{ \alpha^2\, \Norm{x}{2}^2
+   \leq \Norm{Ax}{2}^2,\; \text{ for all }x \in \Sigma_k\big\}, \label{lowerRIP}\\
+   \beta_k & \define \argmin_{\beta \geq 0}\; \big\{ \Norm{Ax}{2}^2 \leq \beta^2\,
+   \Norm{x}{2}^2,\; \text{ for all }x \in \Sigma_k\big\},\label{upperRIP}
  \end{align}
@@
 
-%
 Definition~\ref{asymRIPdef} is a generalization of Definition~\ref{ripdef}
@@
 
-For the asmmetric version of the RIP in Definition~\ref{asymRIPdef}, we
+For the asmmetric version of the RIP in Definition~\ref{asymRIPdef}, one
 can again show $\ell_0$-$\ell_1$-equivalence:
@@
 for the scaled equation
-$\lambda Ax = \lambda b$. In Definition \ref{asymRIPdef}, however, we can
+$\lambda Ax = \lambda b$, $\lambda \in \R \setminus \{0\}$. In Definition~\ref{asymRIPdef}, however, we can
 use $\tilde{\alpha}_k = \lambda \alpha_k$, $\tilde{\beta}_k = \lambda
 \beta_k$, and therefore $\tilde{\gamma}_k = \lambda^2 \alpha_k^2 /
-\lambda^2 \beta_k^2 = \gamma_k$.
+(\lambda^2 \beta_k^2) = \gamma_k$.
 
-Furthermore, the recovery results given in this section can also be derived 
+Furthermore, we note that the recovery results given in this section can also be derived 
 for the \emph{denoising} case, where instead of~\eqref{l0min} one considers

### 2016-03-12 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
paper: revised our contributions section

@@
 
-\subsection{Literature Overview and Contribution of this Work}
+\subsection{Contribution of this Work}
 
+\noindent
 There are several types of matrices~$A$ for which it is possible
 to show analytically that the RIP holds, for instance random
-matrices, see, e.g., Baraniuk et al.~\cite{BDDW08}.
-Still, a big drawback of the RIP is that computing the restricted isometry
+matrices with high probability, see, e.g., Baraniuk et al.~\cite{BDDW08}.
+However, it turns out that computing the restricted isometry
 constants for a given matrix~$A$ and~$k$ is NP-hard, see~\cite{PT14}. 
-On the other hand, d'Aspremont et al.~\cite{Asp08,Asp07} proposed several
+Note the the unfortunate implication that it is hard to check for a
+concrete random matrix whether it satisfies the RIP, since the theoretical
+results hold asymptotically and with respect to high probabilty.
+
+This motivated d'Aspremont et al.~\cite{Asp08,Asp07} to propose several
 semidefinite relaxations to compute upper bounds on the restricted isometry
-constants. Note that small upper bounds suffice to guarantee
+constants; small upper bounds suffice to guarantee
 $\ell_0$-$\ell_1$-equivalence between~\eqref{l0min} and~\eqref{l1min}, but
@@
 
-The relaxations in \cite{Asp08,Asp07} were originally derived for the deeply
-connected problem of sparse principal component analysis (SPCA). Principal
-component analysis (PCA) is about finding orthogonal directions maximizing the 
-variance given in matrix $A$. SPCA tries to fix the problem that the components
-of PCA usually consist of small contributions of all axes, which make the results
-hard to interpret, by enforcing sparsity on the components. SPCA was
-originally proposed by Zuo et al. in \cite{SPCA}, where a penalty term was added 
-to the objective. Other authors like Moghaddam et al. in \cite{sbSPCA} also proposed 
-computing a sparse maximal eigenvalue for SPCA, like it is done for the RIP.
+The relaxations in \cite{Asp08,Asp07} were originally derived for the
+connected \emph{sparse principal component analysis} (SPCA). Principal
+component analysis (PCA) finds orthogonal directions maximizing the 
+variance given in matrix $A$, solving problems
+\[
+\max \{ \Norm{A x}{2}^2 \suchthat \Norm{x}{2} \leq 1\}
+\]
+along the way. For SPCA, one further adds the condition that $\Norm{x}{0}
+\leq k$ with the goal to reduce the dependency on small components. Note
+that this is then equivalent to computing the upper RIC and it is \NP-hard
+in the strong sense~\cite{PT14}. SPCA was
+originally proposed by Zuo et al.~\cite{SPCA}. Other authors like Moghaddam et al.
+\cite{sbSPCA} also proposed computing a sparse maximal eigenvalue for SPCA.
 
 The goal of this paper is to compute optimal restricted isometry constants
-via a mixed integer semidefinite program (MISDP). To this end, we will
-introduce an MISDP with a rank-1 constraint. In order to
-utilize this MISDP, we will show that the rank constraint can be relaxed
-without changing the optimal value. It will turn out that
-the resulting problems are quite hard to solve, even for matrices of
-smaller size. However, the results allow an evaluation of the SDP
-relaxations of d'Aspremont et al. Depending on the input matrices, these
-bounds can be quite tight or weak. This complements the theoretical
-approximation guarantees developed by d'Aspremont et
+via mixed-integer semidefinite programming (MISDP). To this end, we will
+introduce an MISDP with a rank-1 constraint. In order to utilize this
+MISDP, we will show that the rank constraint can be relaxed without
+changing the optimal value. We can then apply a MISDP solver.
+
+In our computational results we show that the resulting problems are quite
+hard to solve, even for matrices of smaller size. However, this is true for
+the SDP relaxations of d'Aspremont et al.\ as well, and it allows for an
+evaluation of the quality of these bounds.  In fact, these bounds turn out
+to be quite tight or weak, depending on the instance. This complements the
+theoretical approximation guarantees developed by d'Aspremont et
 al. in~\cite{AspBG14}.
 
-In the following, we will start by reviewing the SDP-relaxations of~\cite{Asp08}
-and~\cite{Asp07} in Section~\ref{sec:SDPrelax}.
-Afterwards, we will construct a mixed integer semidefinite
-programm and show that it produces exact restricted isometry constants in
-Section~\ref{sec:MISDP}. Finally, in Section~\ref{sec:numerical_results},
-we present numerical results and comparisons with the SDP relaxations in
-\cite{Asp08} and \cite{Asp07}.
+In the following, we start by reviewing the SDP-relaxations of~\cite{Asp08}
+and~\cite{Asp07} in Section~\ref{sec:SDPrelax}.  Afterwards, in
+Section~\ref{sec:MISDP}, we construct a mixed integer semidefinite
+programm and show that it produces exact restricted isometry
+constants. Finally, in Section~\ref{sec:numerical_results}, we present
+numerical results and comparisons with the SDP relaxations in \cite{Asp08}
+and \cite{Asp07}.
 

### 2016-03-12 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
paper: reworked SDP relaxation section

@@
 
-In this section, we will introduce two known SDP relaxations to compute an upper bound on the restricted isometry constant $\delta_k$ in (\ref{rip}), as they have some ideas in common with the mixed integer semidefinite program we want to
-propose and will also be checked against in the section on numerical results.
-The idea of these two relaxations as well as the MISDP in the next section is that the optimal constant $\alpha_k^2$ in (\ref{lowerRIP}) can be computed
+\noindent
+In this section, we will introduce two known SDP relaxations to compute an upper bound on the restricted isometry constant $\delta_k$ in~\eqref{rip}, since they have some ideas in common with the mixed integer semidefinite program we want to
+propose and will also be checked against in the section on numerical
+results.
+
+The idea is that the optimal constant $\alpha_k^2$ in \eqref{lowerRIP} can be computed
 by the non-convex quadratic optimization problem
@@
  \min \quad & \Norm{Ax}{2}^2 \nonumber \\
- \text{s.t.} \quad & \Norm{x}{2}^2 = 1 \tag{QP} \\
+ \text{s.t.} \quad & \Norm{x}{2}^2 = 1, \tag{QP} \\
  & \Norm{x}{0} \leq k. \nonumber
@@
 The corresponding maximization problem allows to compute $\beta_k^2$.
-In \cite{Asp07} the non-convex quadratic equality constraint $\Norm{x}{2}^2 = 1$ is tackled by the technique of semidefinite lifting first introduced in \cite{GW95}
+In \cite{Asp07}, the non-convex quadratic equality constraint $\Norm{x}{2}^2 = 1$ is tackled by the technique of semidefinite lifting first introduced in \cite{GW95}
 for the max-cut problem.
-The idea is to use a new matrix variable $X=xx\T$ for the quadratic terms, where the condition $X=xx\T$ is then enforced by the equivalent constraints $X \succeq 0$ and $\Rk(X) = 1$, where
-$X \succeq 0$ means that $X$ is symmetric (shortly written \mbox{$X \in S_n$}) and positive semidefinite.
-The objective and the normalization-constraint can then be expressed in $X$ as  $\Norm{Ax}{2}^2 = x\T A\T Ax =
-\Tr(A\T A X)$ and $\Norm{x}{2}^2 = \sum_{i=1}^n x_i^2 = \Tr(X)$. In the
-next step the inequality $\Norm{x}{0} \leq k$ is then substituted by the coarser constraint $\ones\T
-\abs{X} \ones \leq k$, which follows from the fact that $\ones\T\abs{X} \ones = \Norm{\vect(X)}{1}\leq \sqrt{\Norm{\vect(X)}{0}} \Norm{\vect(X)}{2}$ and $\Norm{\vect(X)}{0}=\Norm{x}{0}^2$ in the rank 1 case, where $\vect(X)$ is a vector
+The idea is to use new matrix variables $X=xx\T$ for the quadratic terms, where the condition $X=xx\T$ is enforced by the equivalent constraints $X \succeq 0$ and $\Rk(X) = 1$, where
+$X \succeq 0$ means that $X$ is symmetric (shortly written $X \in S_n$) and positive semidefinite.
+The objective and the normalization-constraint can then be expressed as  $\Norm{Ax}{2}^2 = x\T A\T Ax =
+\Tr(A\T A X)$ and $\Norm{x}{2}^2 = \sum_{i=1}^n x_i^2 = \Tr(X)$.
+
+In the next step, the inequality $\Norm{x}{0} \leq k$ is then substituted by the weaker constraint $\ones\T
+\abs{X} \ones \leq k$, which follows from the fact that $\ones\T\abs{X} \ones = \Norm{\vect(X)}{1}\leq \sqrt{\Norm{\vect(X)}{0}}\, \Norm{\vect(X)}{2}$ and $\Norm{\vect(X)}{0}=\Norm{x}{0}^2$ in the rank-1 case, where $\vect(X)$ is a vector
 in $\R^{n^2}$ consisting of all entries of $X$. 
-Relaxing the non-convex Rank-constraint finally leads to the following relaxation:
+Relaxing the non-convex rank constraint finally leads to the following SDP:
 \begin{align}\label{Asp07}
  \min \quad & \Tr(A\T A X) \nonumber \\
- \text{s.t.} \quad & \Tr(X) = 1 \nonumber \\
- & \ones\T \abs{X} \ones \leq k \tag{A1} \\
+ \text{s.t.} \quad & \Tr(X) = 1, \tag{A1} \\
+ & \ones\T \abs{X} \ones \leq k, \nonumber \\
  & X \succeq 0. \nonumber
@@
 
-Another SDP relaxation was proposed in \cite{Asp08}. There the problem
+Another SDP relaxation was proposed in \cite{Asp08}, where
 \begin{equation}\label{phi}
- \phi(\rho) = \max_{\Norm{x}{2} \leq 1} x\T \Sigma x - \rho \Norm{x}{0},
+  \phi(\rho) = \max \big\{ x\T \Sigma\, x - \rho \Norm{x}{0} \suchthat \Norm{x}{2} \leq 1 \big\},
 \end{equation}
@@
  \begin{aligned}
-  \beta_k^2 & = \max_{\|x\|_2 \leq 1, \|x\|_0 \leq k} x^TA^TAx \\
-	    & \leq \inf_{\rho \geq 0} \max_{\|x\|_2\leq 1} x^TA^TAx - \rho (\|x\|_0-k) \\
-	    & = \inf_{\rho \geq 0} \phi(\rho) + \rho k.
+  \beta_k^2 & = \max \big\{ x\T A\T A x \suchthat \Norm{x}{2} \leq 1,\; \Norm{x}{0} \leq k \big\}\\
+	    & \leq \inf_{\rho \geq 0} \max \big\{ x\T A\T A x - \rho
+            (\Norm{x}{0} - k) \suchthat \Norm{x}{2} \leq 1 \big\}\\
+	    & = \inf_{\rho \geq 0} \phi(\rho) + \rho\, k.
  \end{aligned}
@@
 
-As one way to solve (\ref{phi}) the semidefinite lifting technique is used again to compute an upper bound on $\phi$ via
+As one way to solve \eqref{phi} the semidefinite lifting technique is used again to compute an upper bound on $\phi$ via
 \begin{align}\label{Asp08}
- \max \quad & \sum_{i=1}^n\Tr(P_iB_i) \nonumber \\
- \text{s.t.} \quad & \Tr(X) = 1 \nonumber \\
- & X \succeq 0 \tag{A2-Primal} \\
- & X \succeq P_i \succeq 0 \quad \text{for } i = 1, ..., n, \nonumber
+ \max \quad & \sum_{i=1}^n \Tr(P_i \, B_i) \nonumber \\
+ \text{s.t.} \quad & \Tr(X) = 1, \tag{A2-Primal}\\
+ & X \succeq 0,  \nonumber \\
+ & X \succeq P_i \succeq 0 \quad \text{for } i = 1, \dots, n, \nonumber
 \end{align}
-for $B_i = b_ib_i\T - \rho I,$ where $b_i$ is the $i$-th column of a squareroot of $\Sigma$, for details see \cite{Asp08}. The corresponding dual problem can then be written as 
+for $B_i = b_i\,b_i\T - \rho\, I$, where $b_i$ is the $i$-th column of the
+squareroot of $\Sigma$. For details, we refer the reader to~\cite{Asp08}. The corresponding dual problem can then be written as 
 \begin{align}\label{Asp08Dual}
-  \min \quad & \lambda_{\max}\left(\sum_{i=1}^n Y_i \right) \nonumber \\
-  \text{s.t.} \quad & Y_i \succeq B_i \quad \forall i \leq n \tag{A2-Dual} \\
-  & Y_i \succeq 0 \quad \ \ \text{for } i = 1, ..., n, \nonumber
+  \min \quad & \lambda_{\max}\Big(\sum_{i=1}^n Y_i \Big) \nonumber \\
+  \text{s.t.} \quad & Y_i \succeq B_i \quad \text{for } i = 1, \dots, n \tag{A2-Dual} \\
+  & Y_i \succeq 0 \quad \ \ \text{for } i = 1, \dots, n, \nonumber
 \end{align}
-which only has half as many variables as (\ref{Asp08}) and one less linear and semidefinite constraint.
-Then (\ref{Asp08Dual}) can be solved for non-negative $\rho$, which should be smaller than $\Sigma_{11}$, because otherwise the
-optimal solution for (\ref{phi}) will always be $x=0$.
+which has only half as many variables as \eqref{Asp08} and one fewer linear and semidefinite constraint.
+Then \eqref{Asp08Dual} can be solved for non-negative $\rho$, which should be smaller than $\Sigma_{11}$, because otherwise the
+optimal solution for \eqref{phi} will always be $x=0$.
 
-As mentioned in \cite{Asp08}, a lower bound for $\alpha_k$ can be computed via
+As shown in \cite{Asp08}, a lower bound for $\alpha_k$ can be computed via
 \begin{equation*}
- \alpha_k^2 \geq \sup_{\rho \geq 0} \psi(\rho) - \rho k,
+ \alpha_k^2 \geq \sup_{\rho \geq 0}\; \psi(\rho) - \rho\, k,
 \end{equation*}
@@
 \begin{equation*}
- \psi(\rho) = \min_{\Norm{x}{2} \leq 1} x\T \Sigma x + \rho \Norm{x}{0}.
+  \psi(\rho) = \min \big\{ x\T \Sigma\, x + \rho \Norm{x}{0} \suchthat \Norm{x}{2} \leq 1\big\}.
 \end{equation*}
-In this case problem (\ref{Asp08}) or equivalently (\ref{Asp08Dual}) has to be solved for $\Sigma = M I - A\T A$, where $M$ has to be big enough to make $\Sigma$ positive semidefinite again.
-If $\bar{\Psi}$ is the corresponding optimal solution, this implies
-a lower bound on $\alpha_k$ in
+\comment{$\bar{X}$ L\"osung von \eqref{Asp08}?}
+In this case problem \eqref{Asp08} or equivalently \eqref{Asp08Dual} has to
+be solved for $\Sigma = M I - A\T A$, where $M \in \R$ has to be big enough
+to make $\Sigma$ positive semidefinite again. If $\bar{X}$ is a
+corresponding optimal solution, this implies a lower bound on $\alpha_k$
+in\comment{Ist $M\,I$ statt $M$ korrekt?}
 \begin{equation*}
- \alpha_k^2 \geq \sup_{\rho \geq 0} (M - \bar{\Psi}) - \rho k.
+  \alpha_k^2 \geq \sup_{\rho \geq 0}\; (M\, I - \bar{X}) - \rho\, k.
 \end{equation*}
-
-In \cite{AspBG14} explicit bounds on the quality of the semidefinite relaxation where derived using randomization arguments.
+In \cite{AspBG14} explicit bounds on the quality of this relaxation where derived using randomization arguments.
+\comment{Ist hier die Lagrange-Relaxierung gemeint?}
 

### 2016-03-12 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
rework MISDP section

@@
 
-\section{An MISDP for the Restricted Isometry Property}
+\section{An MISDP for the RIP}
 \label{sec:MISDP}
 
-In this section we want to compute the optimal constants $\alpha_k$ and $\beta_k$ in (\ref{lowerRIP}) and (\ref{upperRIP}) using mixed integer semidefinite programming.
-We will again start with the non-convex quadratic problem
-\begin{align}\label{QP2}
- \min \quad & \Norm{Ax}{2}^2 \nonumber \\
- \text{s.t.} \quad & \Norm{x}{2}^2 = 1 \tag{QP} \\
- & \Norm{x}{0} \leq k. \nonumber
-\end{align}
-For the non-convex quadratic equality constraint $\Norm{x}{2}^2 = 1$ we will
+\noindent
+In this section we propose a MISDP formulation to compute the optimal
+constants $\alpha_k$ and $\beta_k$ in \eqref{lowerRIP} and
+\eqref{upperRIP}, respectively. We will again start with the non-convex quadratic problem~\eqref{QP}.
+For the non-convex quadratic equality constraint $\Norm{x}{2}^2 = 1$, we
 again use semidefinite lifting, introducing the new variable $X=xx\T$ like in the last section. 
-The cardinality-constraint in (\ref{QP}) can equivalently be written using binary variables $z_i$ as $-z_i \leq x_i \leq z_i$ and $\sum_{i=1}^n z_i \leq k$.
-We can thus rewrite (\ref{QP}) equivalently as
-% In this section we want to compute the optimal constants $\alpha_k$ and $\beta_k$ in (\ref{asymRIP}) using mixed integer semidefinite programming.
-A non-convex quadratic optimization problem to compute $\alpha_k$ can be formulated as
-% 
-% \begin{align}\label{QP}
-%  \min \quad & \Norm{Ax}{2}^2 \nonumber \\
-%  \text{s.t.} \quad & \Norm{x}{2}^2 = 1 \tag{QP} \\
-%  & \Norm{x}{0} \leq k. \nonumber
-% \end{align}
-% 
-% By switching $\min$ for $\max$ in (\ref{QP}) we also get a corresponding optimization problem to compute $\beta_k$ and therefore also $\gamma_k$ and $\delta_k$.
-% The cardinality-constraint in (\ref{QP}) can equivalently be written using binary variables $z_i$ as $-z_i \leq x_i \leq z_i$ and $\sum_{i=1}^n z_i \leq k$.
-For the non-convex quadratic equality constraint $\Norm{x}{2}^2 = 1$ we will
-% use the technique of semidefinite lifting first introduced in \cite{GW95} for the max-cut problem.
-The idea is to use a new matrix variable $X=xx\T$ for the quadratic terms, where the condition $X=xx\T$ is then enforced by the
-% equivalent constraints $X \succeq 0$ and $\Rk(X) = 1$, where $X
-% \succeq 0$ means that $X$ is symmetric and positive semidefinite. Using
-% that $\Norm{Ax}{2}^2 = x\T A\T Ax = \Tr(A\T A X)$ and 
-% $\Tr(X) = \sum_{i=1}^n x_i^2$, we can thus rewrite (\ref{QP}) equivalently to
+The cardinality-constraint in \eqref{QP} can equivalently be written using
+binary variables $z_i$ as $-z_i \leq x_i \leq z_i$ and $\sum_{i=1}^n z_i
+\leq k$. We can thus equivalently rewrite \eqref{QP} as
 \begin{align}\label{Rk1MISDP}
  \min \quad & \Tr(A\T A X) \nonumber \\
- \text{s.t.} \quad & \Tr(X) = 1 \nonumber \\
- & -z_j \leq X_{ij} \leq z_j \quad \text{for } j = 1, ..., n \nonumber \\
- & \sum_{i=1}^n z_i \leq k \tag{M1} \\
- & \Rk(X) = 1 \nonumber \\
- & X \succeq 0  \nonumber \\
+ \text{s.t.} \quad & \Tr(X) = 1, \nonumber \\
+ & -z_j \leq X_{ij} \leq z_j \quad \text{for } j = 1, \dots, n, \nonumber \\
+ & \sum_{i=1}^n z_i \leq k,\\
+ & \Rk(X) = 1, \nonumber \\
+ & X \succeq 0,  \nonumber \\
  & z \in \{0,1\}^n. \nonumber
 \end{align}
-The next step would then be to relax the non-convex rank constraint again to arrive at the following relaxation which computes a lower bound on the optimal constant $\alpha_k$ in (\ref{lowerRIP}):
+We then relax the non-convex rank constraint to arrive at the following
+MISDP formulation:
 \begin{align}\label{MISDP}
  \min \quad & \Tr(A\T A X) \nonumber \\
- \text{s.t.} \quad & \Tr(X) = 1 \nonumber \\
- & -z_j \leq X_{ij} \leq z_j \quad \text{for } j = 1, ..., n \nonumber \\
- & \sum_{i=1}^n z_i \leq k \tag{MISDP} \\
- & X \succeq 0  \nonumber \\
- & z \in \{0,1\}^n \nonumber
+ \text{s.t.} \quad & \Tr(X) = 1, \nonumber \\
+ & -z_j \leq X_{ij} \leq z_j \quad \text{for } j = 1, \dots, n, \nonumber \\
+ & \sum_{i=1}^n z_i \leq k, \tag{MISDP} \\
+ & X \succeq 0,  \nonumber \\
+ & z \in \{0,1\}^n. \nonumber
 \end{align}
+In the following, we will show that \eqref{MISDP} actually produces the
+exact value of $\alpha_k$. We will use the following result about the
+relationship between the rank of solutions of semidefinite programs and the
+dimension of corresponding faces of a convex set~$P$, i.e., subsets $F
+\subseteq P$ such that $x \in F$, $y, z \in P$, and $x = \tfrac{1}{2} (y + z)$ imply
+$y,z \in F$:
 
-In the following we will show that (\ref{MISDP}) not only computes a lower bound but the exact value of $\alpha_k$.
-For this we will use results about the geometry of semidefinite programs to show that there always exists an
-optimal solution to (\ref{MISDP}) with rank 1 so that the optimal objective values of (\ref{Rk1MISDP}) and (\ref{MISDP}) have to agree.
-We will use the following result by Pataki \cite{pat98} about the relationship between the rank of
-solutions of semidefinite programs and the dimension of corresponding faces, meaning a subset $F$ of the convex set $P$ such that \mbox{$x \in F, y, z \in P$} and $x = 0.5y + 0.5z$ imply $y,z \in F$:
-
-\begin{theorem}\label{patakiLemma}
+\begin{theorem}[Pataki~\cite{pat98}]\label{patakiLemma}
   Let $X \in F$, where $F$ is a face of
   \begin{equation*}
- P \define \{X \in S_n : X \succeq 0, \Tr(A_i X) = b_i \ \text{for } i = 1, ..., m\}
+    P \define \{X \in S_n \suchthat X \succeq 0,\; \Tr(A_i X) = b_i\; \text{ for } i = 1, \dots, \tilde{m}\}
   \end{equation*}
- for symmetric matrices $A_i \in S_n, i \leq m$. Then
+  for symmetric matrices $A_i \in S_n$, $i = 1, \dots, \tilde{m}$. Then
   \begin{equation*}
-  \frac{1}{2}\Rk(X)\cdot(\Rk(X)+1) \leq m + dim(F).
+    \tfrac{1}{2}\Rk(X)\cdot(\Rk(X)+1) \leq \tilde{m} + dim(F).
   \end{equation*}
@@
 
-Now we will use this theorem on the projection of the feasible set of (\ref{MISDP}) onto the space of all indices $i$, such that $z_i = 1$ for a given optimal solution of (\ref{MISDP}), to show that there always exists an optimal
-solution to (\ref{MISDP}) with rank one.
+Using this result on the projection of~\eqref{MISDP} onto the $X$ variables, yields the following.
 
 \begin{theorem}\label{Rk1thm}
- For every $A \in \R^{m \times n}$ and $k \geq 1$ there exists an optimal solution $X^*, z^*$ to (\ref{MISDP}) such that $\Rk(X^*) = 1$. 
+  For every $A \in \R^{m \times n}$ and $k \in \Z_{>0}$, there exists an optimal solution $(X^\star, z^\star)$ of \eqref{MISDP} with $\Rk(X^\star) = 1$. 
 \end{theorem}
+
 \begin{proof}
- Because of the constraint $\Tr(X) = 1$, Problem~\eqref{MISDP} is bounded and $X_{11} = 1, X_{ij} = 0 \ \forall i,j \neq 1,$ \mbox{$z_1 = 1, z_i = 0 \ \forall i \neq 1$} 
- is always feasible for $k > 0$,  so an optimal solution to~\eqref{MISDP} exists.
- Let $\hat{X}, \hat{z}$ be such an optimal solution to (\ref{MISDP}) with $\sum_{i=1}^n z_i =: \ell \leq k$.
- Define
+  First note that because of the constraint $\Tr(X) = 1$,
+  Problem~\eqref{MISDP} is bounded. Moreover, $X_{11} = 1$, $X_{ij} = 0$
+  for all $i,j \neq 1$, $z_1 = 1$, $z_i = 0$ for all $i = 2, \dots, n$ is
+  feasible for $k > 0$. Thus, an optimal solution to~\eqref{MISDP} exists.
+
+  Let $(X^\star, z^\star)$ be an optimal solution to \eqref{MISDP} with
+  $\ell \define \sum_{i=1}^n z^\star_i \leq k$. Let $T \define \{i_1, i_2,
+  \dots, i_\ell\}$ be the support of $z^\star$, and define $\check{A} =
+  (A_{i_1}, A_{i_2}, \dots, A_{i_\ell})$ to be the submatrix of~$A$ formed by
+  the columns indexed by~$T$. Then consider
   \begin{align}\label{Proj}
- \min \quad & \Tr(\tilde{A}\T\tilde{A}\tilde{X}) \nonumber \\
- \text{s.t.} \quad & \Tr(\tilde{X}) = 1 \tag{Proj} \\
- & \tilde{X} \succeq 0  \nonumber 
+    \min \quad & \Tr(\check{A}\T \check{A} \check{X}), \nonumber \\
+    \text{s.t.} \quad & \Tr(\check{X}) = 1, \\
+    & \check{X} \succeq 0,\; \check{X} \in S_\ell.\nonumber 
   \end{align}
-for $\tilde{X} \in S_\ell$, where $\tilde{A} = (A_{i_1}, A_{i_2}, ..., A_{i_\ell})$ consists of those
-columns $\{{i_1}, {i_2}, ..., {i_\ell}\} \enifed T$ of $A$ for which
-$\hat{z}_{i_k} = 1$.
 
-Now let $\bar{X}$ be the $\ell \times \ell$ submatrix of $\hat{X}$ with all rows and columns outside of $T$ removed.
-Then $\bar{X}$ is feasible for (\ref{Proj}) with the same objective value as the optimal solution $\hat{X}$ of (\ref{MISDP}),
-as we only removed zero rows and columns of $\hat{X}$ and $(A\T A)\hat{X}$, which do not influence the trace or the positive semidefiniteness.
+  Let $\check{X}$ be the $\ell \times \ell$ submatrix of $X^\star$ with all
+  rows and columns outside of $T$ removed. Then $\check{X}$ is feasible for
+  \eqref{Proj} with the same objective value, since we only removed zero
+  rows and columns of $X^\star$ and $(A\T A) X^\star$, which do not
+  influence the trace or the positive semidefiniteness.
 
-On the other hand we can lift any optimal solution $\check{X}$ of (\ref{Proj}) to a feasible point for (\ref{MISDP}) with identical objective value by setting $z^{\prime}_i$ to one if and only if $i \in T$
-and $X^{\prime}_{i_{j_1} i_{j_2}} = \check{X}_{j_1j_2}$ if and only if $i_{j_1}, i_{j_2} \in T$
-and $X^{\prime}_{ij}=0$ otherwise.
-This solution will then be feasible for (\ref{MISDP}), since $\Tr(\check{X}) = 1$ and $\check{X}\in S_\ell$ imply $0 \leq \check{X}_{jj} = X^{\prime}_{i_j i_j} \leq 1$ and therefore
-$X^{\prime}_{ij} \leq z^{\prime}_j$ by diagonal dominance of semidefinite matrices.
+  On the other hand, we can lift any solution $\check{X}$ of~\eqref{Proj}
+  to a point $\hat{X} \in S_n$ by extending with zeros. Defining
+  $\hat{z}_i$ to be 1 if and only if $i \in T$, yields a feasible point
+  $(\hat{X}, \hat{z})$ for \eqref{MISDP} with identical objective value,
+  since $\Tr(\check{X}) = 1$ and $\check{X} \in S_\ell$ imply $0 \leq
+  \check{X}_{jj} = \hat{X}_{i_j i_j} \leq 1$ for $j = 1, \dots,
+  \ell$. Therefore, $\hat{X}_{ij} \leq \hat{z}_j$ by diagonal dominance of
+  semidefinite matrices.
 
-So the optimal objective values of (\ref{Proj}) and (\ref{MISDP}) agree and with this construction also 
-$\Rk(X^{\prime}) = \Rk(\check{X})$, so it suffices to show that (\ref{Proj}) always allows for an optimal solution of rank one.
-Let $\breve{X}$ be an extreme point of (\ref{Proj}) (as the set of optimal points of (\ref{Proj}) is a non-empty, convex, compact subset of $\R^{\ell(\ell+1)/2}$, such a point exists
-\cite[Corollary 18.5.1]{Roc70}) with corresponding face $F=\{\breve{X}\}$. Then using Theorem~\ref{patakiLemma} on (\ref{Proj}) with $\tilde{m} = 1$ LP-constraint gives
+  Thus, the optimal objective values of \eqref{Proj} and \eqref{MISDP}
+  agree and $\Rk(\hat{X}) = \Rk(\check{X})$.  Therefore, it
+  suffices to show that \eqref{Proj} has an optimal solution of rank one.
+
+  Let $\check{X}$ be an extreme point of \eqref{Proj} (since the set of
+  optimal points of~\eqref{Proj} is a nonempty, convex, and compact, such a
+  point exists, see~\cite[Corollary 18.5.1]{Roc70}). Defining the face $F =
+  \{\check{X}\}$ and applying Theorem~\ref{patakiLemma} on \eqref{Proj}
+  gives
   \begin{equation*}
- \frac{1}{2}\Rk(\breve{X})\cdot(\Rk(\breve{X})+1) \leq \tilde{m} + dim(F) = 1 + 0 = 1.
+    \tfrac{1}{2} \Rk(\check{X}) \cdot (\Rk(\check{X})+1) \leq \tilde{m} + dim(F) = 1 + 0 = 1,
   \end{equation*}
-Since $\breve{X}=0$ is infeasible for $\Tr(X)=1$, we must have $\Rk(\breve{X}) = 1$, which can then be lifted to an optimal solution of (\ref{MISDP}) with rank one.
+  since there are $\tilde{m} = 1$ linear constraints. Since $\check{X} = 0$
+  is infeasible for $\Tr(X) = 1$, it follows that $\Rk(\check{X}) = 1$,
+  which can then be lifted to an optimal solution of \eqref{MISDP} of rank
+  one.
 \end{proof}
 
-Since Theorem~\ref{Rk1thm} guarantees that (\ref{MISDP}) always has a solution of rank one and the rank-constraint was the only constraint we relaxed to get (\ref{MISDP}) from (\ref{QP}), this means that their optimal objective values
-have to agree and we can compute the optimal constant $\alpha_k$ for the RIP by solving (\ref{MISDP}).
-Since the same argumentation also holds for the maximization problem for $\beta_k$, this allows us to compute the restricted isometry
-constant and the restricted isometry ratio by mixed integer semidefinite programming.
+Theorem~\ref{Rk1thm} guarantees that \eqref{MISDP} always has a solution of
+rank one. Therefore, the optimal values of \eqref{MISDP} and \eqref{QP}
+agree. We can therefore compute the optimal constant $\alpha_k$ for the RIP
+by solving \eqref{MISDP}. Since the same argumentation also holds for the
+maximization problem for $\beta_k$, this allows us to compute the
+restricted isometry constant and the restricted isometry ratio by
+mixed-integer semidefinite programming.
 
 \begin{remark}\label{Rk1rmk}
-Theorem \ref{Rk1thm} still holds if additional constraints are added fixing some of the binary variables $z_j$ to $0$ or $1$.
-In this case there will always exist an optimal solution to problem (\ref{MISDP}) with these added
-constraints with rank $1$, such that the optimal objective value of this problem will also equal that of (\ref{Rk1MISDP}) with the added constraints.
-This means in particular, that solving problem (\ref{MISDP}) using a
-branch-and-bound-approach is the same as solving the original problem (\ref{Rk1MISDP}) with a branch-and-bound-code.
+  Theorem \ref{Rk1thm} still holds, if additional constraints involving
+  only $z$-variables are added, e.g., fixing some $z_j$ to $0$ or $1$. In
+  particular, \eqref{MISDP} or equivalently \eqref{Rk1MISDP} can be
+  solved by branch and bound.
 \end{remark}
 
-To solve \eqref{MISDP} successfully using a branch-and-bound-approach, it is important to have strong semidefinite relaxations for fractional $z$ variables. 
-Therefore we will further strengthen the constraint coupling the continuous and binary variables to force the nonzero entries of $z$ towards one in the
-semidefinite relaxations.
+Theorem~\ref{Rk1thm} and Remark~\ref{Rk1rmk} do not guarantee that any
+found solution will have rank one. However, branch and bound will still
+produce the correct optimal value. This suffices for our application of
+computing the RIP constants $\alpha_k$ and $\beta_k$. For other
+applications like SPCA, an extreme solution can be generated with an
+algorithm proposed by Pataki~\cite{coneLP}.
+
+In order to successfully apply branch and bound to \eqref{MISDP}, it is
+important to use strong semidefinite relaxations for fractional $z$
+variables. To this end, we can use the following strengthening of the
+constraints $-z_j \leq X_{ij} \leq z_j$.
 
 \begin{lemma}\label{HalfConstraint} 
-The constraint $-z_j \leq X_{ij} \leq z_j$ in~\eqref{MISDP} can be substituted for all non-diagonal entries of $X$ by the stronger constraint 
+  The inequalities
   \begin{equation}\label{eq:HalfConstraint}
-  -\frac{1}{2}z_j \leq X_{ij} \leq \frac{1}{2}z_j.
+    -\tfrac{1}{2} z_j \leq X_{ij} \leq \tfrac{1}{2} z_j,
   \end{equation}
+  for $i, j \in \{1, \dots, n\}$, are valid for~\eqref{MISDP}.
 \end{lemma}
+
 \begin{proof}
-Let $X$ satisfy $\Tr(X)=1$ and $X\succeq 0$. Then $X_{ii}+X_{jj}\leq 1$ and nonnegativity of the diagonal 
-entries imply $X_{ii}X_{jj}\leq 1/4$. 
-Since all minors of $X$ need to be nonnegative, it follows that
+  Let $X \succeq 0$ satisfy $\Tr(X)=1$. Then $X_{ii}+X_{jj}\leq 1$ and
+  nonnegativity of the diagonal entries imply $X_{ii}X_{jj}\leq
+  \tfrac{1}{4}$.  Since all minors of $X$ need to be nonnegative, it
+  follows that
   \begin{equation*}
-  X_{ij}^2 \leq  X_{ii} X_{jj} \leq \frac{1}{4}
+    X_{ij}^2 \leq  X_{ii} X_{jj} \leq \tfrac{1}{4}.
   \end{equation*}
-and therefore $-1/2 \leq X_{ij} \leq 1/2$, which implies the equivalence between
-$-z_j \leq X_{ij} \leq z_j$ and~\eqref{eq:HalfConstraint} for all integer values 
-of $z$.
+  Therefore $- \tfrac{1}{2} \leq X_{ij} \leq \tfrac{1}{2}$, which shows
+  validity of~\eqref{eq:HalfConstraint}.
 \end{proof}
-
-Theorem \ref{Rk1thm} and Remark \ref{Rk1rmk} do not guarantee that any found solution will have rank one, they only guarantee that there exists a rank one solution.
-But this is not a problem for applications in the context of the RIP,
-since we are only interested in the optimal objective value, since it implies the constant $\alpha_k$ or $\beta_k$ for the RIP.
-For other applications like the SPCA such an extreme solution can also be generated with an algorithm
-proposed by Pataki in \cite{coneLP}.
 

### 2016-03-12 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
paper: reworked computational results

@@
 
-In this section we want to show the applicability of our proposed MISDP-formulation and use it to
+\noindent
+In this section we demonstrate the applicability of the proposed MISDP-formulation and use it to
 assess the quality of the relaxations proposed in \cite{Asp08} and \cite{Asp07}.
-For solving the mixed-integer semidefinite programs we use the MISDP-Solver SCIP-SDP \cite{SCIP-SDP} 
-originally developed in \cite{Mar13} and \cite{MS12}, which combines the branch-and-bound-framework of SCIP 3.2.1 \cite{SCIP} with 
-interior-point SDP-solvers, in our case SDPA~7.3.8~\cite{SDPA6,SDPA7}. The continuous SDPs are solved using the command-line version of 
-SDPA~7.3.8. In our tests we compute the left- and right-hand restricted isometry constant 
-$\alpha_k$ and $\beta_k$ either exactly using~\eqref{MISDP} or produce corresponding bounds via (\ref{Asp07}) and (\ref{Asp08Dual}). In the latter case we solve 15 
-SDPs for $\rho$ between $0$ and the maximum diagonal entry of $A\T A$.
+For solving the mixed-integer semidefinite programs we use the MISDP solver SCIP-SDP \cite{SCIP-SDP},
+originally developed in \cite{Mar13} and \cite{MS12}. It combines the
+branch and bound framework SCIP 3.2.1 \cite{SCIP} with 
+interior-point SDP-solvers, in our case SDPA~7.3.8~\cite{SDPA6,SDPA7}.
 
-The testset consists of 63 matrices of seven 
-different types including band matrices 
+The continuous SDPs are solved using the command-line version of
+SDPA~7.3.8. We compute the left- and right-hand restricted isometry
+constant $\alpha_k$ and $\beta_k$ either exactly using~\eqref{MISDP} or
+produce corresponding bounds via \eqref{Asp07} and \eqref{Asp08Dual}. In
+the latter case, we solve 15 SDPs for $\rho$ between~$0$ and the maximum
+diagonal entry of $A\T A$.\comment{As done in \cite{Asp08}?}
+
+We use a testset consisting of 63 matrices with seven 
+different types including band matrices, 
 with band size three or five with entries within the band chosen uniformly in $\{0,1\}$, binary matrices with all entries chosen 
 uniformly in $\{0,1\}$, Gaussian-distributed matrices and rank one 
-matrices $A = aa\T$ with $\mathcal{N}(0,1)$-distributed entries of $a$. In addition to these four kinds of matrices, 
-we also use three more with theoretical results about their restricted isometry constants, namely matrices with 
+matrices $A = aa\T$ with $N(0,1)$-distributed entries of $a$. In addition to these four types,
+we also use three more for which theoretical results on restricted isometry
+constants are available. This includes matrices with 
 $N(0, 1/m)$-distributed entries, with entries chosen uniformly in $\pm 1/\sqrt{m}$ and finally with distribution
 \begin{equation}\nonumber
-A_{ij} = \begin{cases} + \sqrt{3/m} & \text{with probability } \frac{1}{6} \\ 0 & \text{with probability } \frac{2}{3} \\ - \sqrt{3/m} & \text{with probability } \frac{1}{6}. \end{cases}
+  A_{ij} = \begin{cases}
+    + \sqrt{3/m} & \text{with probability } \frac{1}{6}, \\
+    0 & \text{with probability } \frac{2}{3}, \\
+    - \sqrt{3/m} & \text{with probability } \frac{1}{6}.
+  \end{cases}
 \end{equation}
-For the latter three kinds of matrices it is shown in \cite{BDDW08} that for $n \rightarrow \infty$, 
-sufficiently small $k$ and given $\delta$, they satisfy the restricted isometry property (\ref{rip}) of order $k$  
-with probability converging exponentially to one. The needed proportion between $n$ and $k$ is so small, however,
-that we cannot expect the generated matrices in small dimensions to satisfy the RIP. For each type we used \mbox{MATLAB 8.3.0} 
-to randomly generate nine matrices of three different sizes with rows ranging from 15 to 30, columns 
+For each type we used \mbox{MATLAB 8.3.0} to randomly generate nine
+matrices of three different sizes with rows ranging from 15 to 30, columns
 from 25 to 40 and order $k$ between three and five.
 
-The tests were performed on a Linux cluster with Intel i3 CPUs with 3.2GHz, 4MB cache and 8GB memory running
+For the latter three kinds of matrices, it was shown by Baraniuk et al.~\cite{BDDW08} that for $n \rightarrow \infty$, 
+sufficiently small $k$, and given $\delta$, they satisfy the restricted isometry property of order $k$  
+with probability converging exponentially to one. The needed proportion between $n$ and $k$ is so small, however,
+that we cannot expect the generated matrices in small dimensions to satisfy the RIP.
+
+The tests were performed on a Linux cluster with Intel i3 CPUs with 3.2GHz, 4MB cache, and 8GB memory running
 Linux. Each computation was performed single-threaded with a single process running on each computer and with 
 a time limit of four hours. The code was compiled with gcc 4.4.5 with \texttt{-O3} optimization. For the 
-Problem~\eqref{Asp08Dual} most of the larger instance were aborted because of the memory limit, these are
+Problem~\eqref{Asp08Dual} most of the larger instance were aborted because of the memory limit. These are
 counted with the maximum allowed time of four hours.
@@
 In Table~\ref{lhsGap} we compare the relative gap between the bounds on $\alpha_k$ computed by the two heuristics
-and the exact values given by our
-mixed integer semidefinite program. We give arithmetic means over all of the nine matrices for the given type that
-could be solved within the time- and memory-limit. Note that we ommitted the rank one matrices, since these satisfy
-$\alpha_k = 0$ in all cases, so that the trivial lower bound would already be optimal. For the 
-left-hand-side of the RIP, both relaxations did not perform
+and the exact values given by~\eqref{MISDP}. We present arithmetic means over all of the nine matrices for the given type that
+could be solved within the time- and memory-limit. Note that in this case, we ommitted the rank one matrices, since these satisfy
+$\alpha_k = 0$ in all cases, so that the trivial lower bound would already
+be optimal.
+
+For the left-hand-side of the RIP, both relaxations did not perform
 too well. Especially the relaxation given in \cite{Asp08} regularly returned worse solutions than the
@@
 
-Results for the right-hand side, i.e. $\beta_k$, are given in Table~\ref{rhsGap}. Here both relaxations
+Results for the right-hand side, i.e., $\beta_k$, are given in Table~\ref{rhsGap}. Here both relaxations
 perform much better. Problem~\eqref{Asp08Dual} constantly produces bounds within 10 to 35\,\% of the optimal
-solution. The formulation in~\eqref{Asp08Dual} shows bigger fluctuation, but is very tight in many cases,
+solution. The formulation in~\eqref{Asp08Dual} shows bigger fluctuations, but is very tight in many cases,
 with a gap of less than two percent for band matrices.
@@
 
-Solving times for the left- and right-hand side are given in Table~\ref{lhsTime} and~\ref{rhsTime}. The relaxation proposed in \cite{Asp07}
+\comment{Wieder arith. Mean?}
+The solving times in seconds for the left- and right-hand side of the RIP are given in
+Table~\ref{lhsTime} and~\ref{rhsTime}, respectively. The relaxation proposed in \cite{Asp07}
 is very fast, taking only a few seconds for all instances. The mixed-integer semidefinite program obviously
-cannot compete with these solving times, although for band-matrices, specifically, the difference is relatively
+cannot compete with these solving times, although for band-matrices, the difference is relatively
 small. The Lagrangian Relaxation from \cite{Asp08}, however, takes more time to produce bounds than is needed

### 2016-03-12 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
paper: conclusion

@@
 
-We have seen that exact restricted isometry constants can be computed not only via total enumeration but also by solving a mixed integer semidefinite program.
-Because of this exactness result, the new problem can drastically
-improve the bounds computed by the semidefinite relaxations proposed in \cite{Asp08} and \cite{Asp07}, but this comes at the cost of a much higher complexity of computation, at least in comparison to \cite{Asp07}.
-We could only
-solve these problems for relatively small instances, but this might change with sufficient advances in the area of mixed integer semidefinite programming, especially with the addition of special cutting planes, which proved to be
-the main enhancement needed to make mixed integer linear solvers as fast as they are today.
-Also special heuristics and branching rules could improve the performance of the branch-and-bound-algorithm and allow us to solve bigger
-instances of the proposed MISDP.
+\noindent
+This paper shows that exact values of restricted isometry constants can be
+computed via mixed-integer semidefinite programming. Although the used
+matrices are necessarily small, we can significantly improve the bounds
+computed by the semidefinite relaxations proposed in \cite{Asp08} and
+\cite{Asp07}. Moreover, the results of this paper can be used to design
+further approximations of the RIP, which might also be applicable for
+larger matrices. On the negative side, the running time of the mixed-integer
+semidefinite approach is significantly larger than the relaxation
+of~\cite{Asp07}. On the positive side, it is faster and/or uses less memory
+than the relaxation of \cite{Asp08} and it computes optimal constants!
+
+For the future, the hope is that with the advancement of solving techniques
+for mixed-integer semidefinite programs, the sizes of the matrices can be
+significantly increased. The components to be improved include cutting
+planes, primal heuristics, and branching rules.
 

### 2016-03-12 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
paper: spell check

@@
   analysis. Computational results with this approach allow to evaluate
-  ealier SDP relaxations and show that their quality is limited.
+  earlier SDP relaxations and show that their quality is limited.
 \end{abstract}
@@
 
-\section{The Restriced Isometry Property}
+\section{The Restricted Isometry Property}
 \label{sec:RIP}
@@
 
-For the asmmetric version of the RIP in Definition~\ref{asymRIPdef}, one
+For the asymmetric version of the RIP in Definition~\ref{asymRIPdef}, one
 can again show $\ell_0$-$\ell_1$-equivalence:
@@
 concrete random matrix whether it satisfies the RIP, since the theoretical
-results hold asymptotically and with respect to high probabilty.
+results hold asymptotically and with respect to high probability.
 
@@
 Section~\ref{sec:MISDP}, we construct a mixed integer semidefinite
-programm and show that it produces exact restricted isometry
+program and show that it produces exact restricted isometry
 constants. Finally, in Section~\ref{sec:numerical_results}, we present
@@
 for $B_i = b_i\,b_i\T - \rho\, I$, where $b_i$ is the $i$-th column of the
-squareroot of $\Sigma$. For details, we refer the reader to~\cite{Asp08}. The corresponding dual problem can then be written as 
+square-root of $\Sigma$. For details, we refer the reader to~\cite{Asp08}. The corresponding dual problem can then be written as 
 \begin{align}\label{Asp08Dual}
@@
 and the exact values given by~\eqref{MISDP}. We present arithmetic means over all of the nine matrices for the given type that
-could be solved within the time- and memory-limit. Note that in this case, we ommitted the rank one matrices, since these satisfy
+could be solved within the time- and memory-limit. Note that in this case, we omitted the rank one matrices, since these satisfy
 $\alpha_k = 0$ in all cases, so that the trivial lower bound would already
@@
 trivial lower bound of zero. The relaxation in \cite{Asp07} performed slightly better, but the bounds
-were still not nearly good enough to have any chance of proving $\ell_0$-$\ell_1$-equivalance. The relatively
+were still not nearly good enough to have any chance of proving $\ell_0$-$\ell_1$-equivalence. The relatively
 bad performance of both relaxations for the left-hand side might not be too surprising, since both relaxations

### 2016-03-13 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
paper: made conclusions more optimistic

@@
 This paper shows that exact values of restricted isometry constants can be
-computed via mixed-integer semidefinite programming. Although the used
-matrices are necessarily small, we can significantly improve the bounds
-computed by the semidefinite relaxations proposed in \cite{Asp08} and
-\cite{Asp07}. Moreover, the results of this paper can be used to design
-further approximations of the RIP, which might also be applicable for
-larger matrices. On the negative side, the running time of the mixed-integer
-semidefinite approach is significantly larger than the relaxation
-of~\cite{Asp07}. On the positive side, it is faster and/or uses less memory
-than the relaxation of \cite{Asp08} and it computes optimal constants!
+computed via mixed-integer semidefinite programming. The computed bounds
+significantly improve the bounds computed by the semidefinite relaxations
+proposed in~\cite{Asp07} and \cite{Asp08}. In conclusion, the relaxation of
+\cite{Asp07} is fast to compute, but produces a gap of arround 90\,\% for
+$\alpha_k$ and 23\,\% for $\beta_k$. If successfully solved, the relaxation
+of \cite{Asp08} provides better bounds, but its running time is too large
+to be used in practice. Moreover, directly solving~\eqref{MISDP} is often
+faster and produces the exact value. Consequently, the relaxation
+in~\cite{Asp08} should not be used in practice.
 
-For the future, the hope is that with the advancement of solving techniques
-for mixed-integer semidefinite programs, the sizes of the matrices can be
+Still, the matrices currently handable by~\eqref{MISDP} are small. For the
+future, the hope is that with the advancement of solving techniques for
+mixed-integer semidefinite programs, the sizes of the matrices can be
 significantly increased. The components to be improved include cutting

### 2016-03-14 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
fixed typo

@@
 constants for a given matrix~$A$ and~$k$ is NP-hard, see~\cite{PT14}. 
-Note the the unfortunate implication that it is hard to check for a
+Note the unfortunate implication that it is hard to check for a
 concrete random matrix whether it satisfies the RIP, since the theoretical

### 2016-03-14 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
correction in the explanation of Asp08, Psi should be the optimal objective, not solution

@@
 \end{equation*}
-\comment{$\bar{X}$ L\"osung von \eqref{Asp08}?}
 In this case problem \eqref{Asp08} or equivalently \eqref{Asp08Dual} has to
 be solved for $\Sigma = M I - A\T A$, where $M \in \R$ has to be big enough
-to make $\Sigma$ positive semidefinite again. If $\bar{X}$ is a
-corresponding optimal solution, this implies a lower bound on $\alpha_k$
-in\comment{Ist $M\,I$ statt $M$ korrekt?}
+to make $\Sigma$ positive semidefinite again. If $\Psi$ is the
+corresponding optimal objective, this implies a lower bound on $\alpha_k$
+in\comment{Ist $M\,I$ statt $M$ korrekt? \\ $\Rightarrow$ nein, sollten alles Skalare sein, $\Psi$ demenstrechend optimaler Zielfunktionswert statt L\"osung (habe ich jetzt oben angepasst)}
 \begin{equation*}
-  \alpha_k^2 \geq \sup_{\rho \geq 0}\; (M\, I - \bar{X}) - \rho\, k.
+  \alpha_k^2 \geq \sup_{\rho \geq 0}\; (M - \Psi) - \rho\, k.
 \end{equation*}

### 2016-03-14 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
explained in more detail what was show in Asp14

@@
 \end{equation*}
-In \cite{AspBG14} explicit bounds on the quality of this relaxation where derived using randomization arguments.
-\comment{Ist hier die Lagrange-Relaxierung gemeint?}
+In \cite{AspBG14} explicit bounds on the gap between \eqref{Asp08} and the penalized problem \eqref{phi} were derived using randomization arguments.
 

### 2016-03-14 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
made the role of P more clear ahead of the Pataki Lemma

@@
  & -z_j \leq X_{ij} \leq z_j \quad \text{for } j = 1, \dots, n, \nonumber \\
- & \sum_{i=1}^n z_i \leq k,\\
+ & \sum_{i=1}^n z_i \leq k,\tag{Rk1MISDP}\\
  & \Rk(X) = 1, \nonumber \\
@@
 relationship between the rank of solutions of semidefinite programs and the
-dimension of corresponding faces of a convex set~$P$, i.e., subsets $F
+dimension of corresponding faces of the feasible set~$P$, i.e., subsets $F
 \subseteq P$ such that $x \in F$, $y, z \in P$, and $x = \tfrac{1}{2} (y + z)$ imply

### 2016-03-14 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
explain numerical results in contrast to Asp08 in more detail

@@
   Let $\check{X}$ be an extreme point of \eqref{Proj} (since the set of
-  optimal points of~\eqref{Proj} is a nonempty, convex, and compact, such a
+  optimal points of~\eqref{Proj} is nonempty, convex, and compact, such a
   point exists, see~\cite[Corollary 18.5.1]{Roc70}). Defining the face $F =
@@
 the latter case, we solve 15 SDPs for $\rho$ between~$0$ and the maximum
-diagonal entry of $A\T A$.\comment{As done in \cite{Asp08}?}
+diagonal entry of $A\T A$. This is the same approach as in \cite{Asp08},
+except that we reduced the number of problems from 25 to 15, as this did
+not seem to influence the quality of th solutions much in our preliminary
+experiments but brought the solving times more in line with the other
+approaches.
 

### 2016-03-14 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
added explanation of shifted geometric mean

@@
 
-\comment{Wieder arith. Mean?}
+For reporting some the solving times, we will use the \emph{shifted
+  geometric mean} to decrease the influence of easy instances, see
+Achterberg~\cite{SCIP} for more details. The shifted geometric
+mean of values $x_1$, \dots, $x_n$ is computed as
+\begin{equation*}
+  \Big( \prod_{i=1}^n (x_i + s)\Big)^{1/n} - s,
+\end{equation*}
+where we used a shift of $s=10$.
+
 The solving times in seconds for the left- and right-hand side of the RIP are given in

### 2016-03-14 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
introduced short form "SDP"

@@
   analysis. Computational results with this approach allow to evaluate
-  earlier SDP relaxations and show that their quality is limited.
+  earlier semidefinite relaxations and show that their quality is limited.
 \end{abstract}
@@
 This motivated d'Aspremont et al.~\cite{Asp08,Asp07} to propose several
-semidefinite relaxations to compute upper bounds on the restricted isometry
+relaxations based on semidefinite programming (SDP) to compute upper bounds on the restricted isometry
 constants; small upper bounds suffice to guarantee
@@
  \text{s.t.} \quad & \Tr(X) = 1, \tag{A2-Primal}\\
- & X \succeq 0,  \nonumber \\
  & X \succeq P_i \succeq 0 \quad \text{for } i = 1, \dots, n, \nonumber
@@
  & -z_j \leq X_{ij} \leq z_j \quad \text{for } j = 1, \dots, n, \nonumber \\
- & \sum_{i=1}^n z_i \leq k,\tag{Rk1MISDP}\\
+ & \sum_{i=1}^n z_i \leq k,\\
  & \Rk(X) = 1, \nonumber \\

### 2016-03-14 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
small corrections

@@
 \end{align}
-The corresponding maximization problem allows to compute $\beta_k^2$.
-In \cite{Asp07}, the non-convex quadratic equality constraint $\Norm{x}{2}^2 = 1$ is tackled by the technique of semidefinite lifting first introduced in \cite{GW95}
+The corresponding maximization problem allows to compute $\beta_k^2$. D'Aspremont et al. \cite{Asp07} tackled the non-convex quadratic equality constraint 
+$\Norm{x}{2}^2 = 1$ by the technique of semidefinite lifting, first introduced by Goemans and and Williamson~\cite{GW95} 
 for the max-cut problem.
-The idea is to use new matrix variables $X=xx\T$ for the quadratic terms, where the condition $X=xx\T$ is enforced by the equivalent constraints $X \succeq 0$ and $\Rk(X) = 1$, where
+The idea is to use new matrix variables $X=xx\T$ for the quadratic terms, where the condition $X=xx\T$ is enforced by the equivalent constraints $X \succeq 0$ and $\Rk(X) = 1$, where, as usual,
 $X \succeq 0$ means that $X$ is symmetric (shortly written $X \in S_n$) and positive semidefinite.
 The objective and the normalization-constraint can then be expressed as  $\Norm{Ax}{2}^2 = x\T A\T Ax =
-\Tr(A\T A X)$ and $\Norm{x}{2}^2 = \sum_{i=1}^n x_i^2 = \Tr(X)$.
+\Tr(A\T A X)$ and $1 = \Norm{x}{2}^2 = \sum_{i=1}^n x_i^2 = \Tr(X)$.
 
@@
 
-Another SDP relaxation was proposed in \cite{Asp08}, where
+Another SDP relaxation was proposed by d'Aspremont et al. in \cite{Asp08}, where
 \begin{equation}\label{phi}
@@
 \end{equation}
-originally stemming from sparse principal component analysis, is tackled for a symmetric positive semidefinite matrix $\Sigma$. For
+originally stemming from sparse principal component analysis, is approximately solved for a symmetric positive semidefinite matrix $\Sigma$. For
 $\Sigma = A\T A$ we can use this to compute an upper bound on $\beta_k$ via Lagrangian Relaxation as

### 2016-03-14 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
removed the remark, since it is unneccessary

@@
 
-\begin{remark}\label{Rk1rmk}
-  Theorem \ref{Rk1thm} still holds, if additional constraints involving
-  only $z$-variables are added, e.g., fixing some $z_j$ to $0$ or $1$. In
-  particular, \eqref{MISDP} or equivalently \eqref{Rk1MISDP} can be
-  solved by branch and bound.
-\end{remark}
-
-Theorem~\ref{Rk1thm} and Remark~\ref{Rk1rmk} do not guarantee that any
-found solution will have rank one. However, branch and bound will still
-produce the correct optimal value. This suffices for our application of
+Theorem~\ref{Rk1thm} does not guarantee that any
+found solution will have rank one, only that the optimal objective
+values agree. This suffices for our application of
 computing the RIP constants $\alpha_k$ and $\beta_k$. For other

### 2016-03-14 RIP-MISDP-Paper/ResultScript.py
started rewriting the table script to produce a single table for left and right hand side

@@ -332,6 +332,73 @@ def RhsResultsTable(instancesets, instancesetnames, caption, label):
 	else:
 		file.write("- \\\ \n")
 	file.write("\\bottomrule \n \\end{tabular*} \n \end{scriptsize} \n \\end{table} \n")
+
+
+TODO: write this
+def LhsRhsResultsTable(instancesets, instancesetnames, caption, label):
+	file.write("\\begin{table} \n \\begin{scriptsize} \\caption{" + caption + "} \n \\label{" + label + "} \n \\begin{tabular*}{0.48\\textwidth}{@{}l@{\\;\\;\extracolsep{\\fill}}rrrr")
+	file.write("@{}}\\toprule \n")
+	file.write(" type of matrix & d'Asp 07 & d'Asp 08 ")
+	file.write("\\\ \midrule \n")
+	totalDiff07 = 0
+	totalnInstances07 = 0
+	totalDiff08 = 0
+	totalnInstances08 = 0
+	i = 0
+	for instancetype in instancesets:
+		diff07 = 0
+		nInstances07 = 0
+		diff08 = 0
+		nInstances08 = 0
+		for j in instancetype:
+			# we only take instances with alpha_k > 0, since otherwise the trivial bound is optimal and we cannot compute a relative gap
+			if dualresults[1][j] != -1e20 and gaps[0][j] != "infinite" and float(gaps[0][j]) < epsilon:
+				diff07 += (float(dualresults[1][j]) - float(dualresults[0][j])) / float(dualresults[1][j])
+				totalDiff07 += (float(dualresults[1][j]) - float(dualresults[0][j])) / float(dualresults[1][j])
+				nInstances07 += 1
+				totalnInstances07 += 1
+			if dualresults[2][j] != -1e20 and gaps[0][j] != "infinite" and float(gaps[0][j]) < epsilon and dualresults[2][j] != float("inf"):
+				diff08 += (float(dualresults[2][j]) - float(dualresults[0][j])) / float(dualresults[1][j])
+				nInstances08 += 1
+				totalDiff08 += (float(dualresults[2][j]) - float(dualresults[0][j])) / float(dualresults[1][j])
+				totalnInstances08 += 1
+		if nInstances07 > 0:
+			avgdiff07 = 100 * diff07 / nInstances07
+		else:
+			avgdiff07 = "-"
[…]

### 2016-03-14 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
explicitly named the memory limit that was too small for Asp08
combined results for alpha_k, beta_k in one table, also unified computation of rhs gap
rhs gap was relax-MISDP/relax before, now relax-MISDP/MISDP like lhs

@@
 a time limit of four hours. The code was compiled with gcc 4.4.5 with \texttt{-O3} optimization. For the 
-Problem~\eqref{Asp08Dual} most of the larger instance were aborted because of the memory limit. These are
+Problem~\eqref{Asp08Dual} most of the larger instance were aborted because of the memory limit of 8GB. These are
 counted with the maximum allowed time of four hours.
 
-In Table~\ref{lhsGap} we compare the relative gap between the bounds on $\alpha_k$ computed by the two heuristics
+In Table~\ref{lhsRhsGap} we compare the relative gap between the bounds on $\alpha_k$ and $\beta_k$ computed by the two heuristics
 and the exact values given by~\eqref{MISDP}. We present arithmetic means over all of the nine matrices for the given type that
-could be solved within the time- and memory-limit. Note that in this case, we omitted the rank one matrices, since these satisfy
-$\alpha_k = 0$ in all cases, so that the trivial lower bound would already
-be optimal.
+could be solved within the time- and memory-limit. Note that we omitted the rank one matrices for the left-hand side since these satisfy
+$\alpha_k = 0$ in all cases, so that the trivial lower bound would already be optimal.
 
-For the left-hand-side of the RIP, both relaxations did not perform
+For the left-hand-side of the RIP, $\alpha_k$, both relaxations did not perform
 too well. Especially the relaxation given in \cite{Asp08} regularly returned worse solutions than the
@@
 were still not nearly good enough to have any chance of proving $\ell_0$-$\ell_1$-equivalence. The relatively
-bad performance of both relaxations for the left-hand side might not be too surprising, since both relaxations
+bad performance of both relaxations for the left-hand side might not be too surprising though, since both relaxations
 were originally designed for sparse principal component analysis, which corresponds to the right-hand side
@@
 
-\input{tables/lhsGap}
+\input{tables/lhsRhsGap}
 
-Results for the right-hand side, i.e., $\beta_k$, are given in Table~\ref{rhsGap}. Here both relaxations
-perform much better. Problem~\eqref{Asp08Dual} constantly produces bounds within 10 to 35\,\% of the optimal
-solution. The formulation in~\eqref{Asp08Dual} shows bigger fluctuations, but is very tight in many cases,
-with a gap of less than two percent for band matrices.
-
-\input{tables/rhsGap}
+For the right-hand side, i.e., $\beta_k$,  both relaxations
+perform much better. Problem~\eqref{Asp07} constantly produces bounds within 10 to 30\,\% of the optimal
+solution, with only slightly worse results for binary and rank one matrices. The formulation in~\eqref{Asp08Dual}
+shows bigger fluctuations, but produces better results for five of the seven matrix types, while performing even
+worse for binary and rank one matrices. For band matrices, however, the average gap was below 2\,\%, the best 
+result for any matrix type among both relaxations.
 
@@
 \cite{Asp07} is fast to compute, but produces a gap of arround 90\,\% for
-$\alpha_k$ and 23\,\% for $\beta_k$. If successfully solved, the relaxation
-of \cite{Asp08} provides better bounds, but its running time is too large
+$\alpha_k$ and 30\,\% for $\beta_k$. If successfully solved, the relaxation
+of \cite{Asp08} provides better bounds for $\beta_k$ for most types of 
+matrices, but its running time is too large
 to be used in practice. Moreover, directly solving~\eqref{MISDP} is often

### 2016-03-14 RIP-MISDP-Paper/ResultScript.py

@@ -9,8 +9,9 @@ from decimal import Decimal
 texfile = 0 # make the output a compilable texfile or just a figure/table that can be included?
 Lbounds = 0
 Rbounds = 0
+LRbounds = 1
 Ltimes = 0
-Rtimes = 1
+Rtimes = 0
 completeTable = 0
 
 MISDPfilename = "/local/gally/results/RIP-MISDP-Paper/160311/RIP-results/check.RIPMISDP.scipsdp.linux.x86_64.gnu.opt.sdpa.extra.branchinfobj_nofracdive.out"
@@ -286,14 +287,14 @@ def RhsResultsTable(instancesets, instancesetnames, caption, label):
 		for j in instancetype:
 			# we only take instances with alpha_k > 0, since otherwise the trivial bound is optimal and we cannot compute a relative gap
 			if dualresults[1][j] != -1e20 and gaps[0][j] != "infinite" and float(gaps[0][j]) < epsilon:
-				diff07 += (float(dualresults[1][j]) - float(dualresults[0][j])) / float(dualresults[1][j])
-				totalDiff07 += (float(dualresults[1][j]) - float(dualresults[0][j])) / float(dualresults[1][j])
+				diff07 += (float(dualresults[1][j]) - float(dualresults[0][j])) / float(dualresults[0][j])
+				totalDiff07 += (float(dualresults[1][j]) - float(dualresults[0][j])) / float(dualresults[0][j])
 				nInstances07 += 1
 				totalnInstances07 += 1
 			if dualresults[2][j] != -1e20 and gaps[0][j] != "infinite" and float(gaps[0][j]) < epsilon and dualresults[2][j] != float("inf"):
-				diff08 += (float(dualresults[2][j]) - float(dualresults[0][j])) / float(dualresults[1][j])
+				diff08 += (float(dualresults[2][j]) - float(dualresults[0][j])) / float(dualresults[0][j])
 				nInstances08 += 1
-				totalDiff08 += (float(dualresults[2][j]) - float(dualresults[0][j])) / float(dualresults[1][j])
+				totalDiff08 += (float(dualresults[2][j]) - float(dualresults[0][j])) / float(dualresults[0][j])
[…]

### 2016-03-14 RIP-MISDP-Paper/ResultScript.py
corrected size of tables

@@ -9,9 +9,9 @@ from decimal import Decimal
 texfile = 0 # make the output a compilable texfile or just a figure/table that can be included?
 Lbounds = 0
 Rbounds = 0
-LRbounds = 1
+LRbounds = 0
 Ltimes = 0
-Rtimes = 0
+Rtimes = 1
 completeTable = 0
 
 MISDPfilename = "/local/gally/results/RIP-MISDP-Paper/160311/RIP-results/check.RIPMISDP.scipsdp.linux.x86_64.gnu.opt.sdpa.extra.branchinfobj_nofracdive.out"
@@ -204,7 +204,7 @@ def makeCompleteTableCaptionMISDP(shortcaption, caption, label, file, i):
 
 
 def LhsResultsTable(instancesets, instancesetnames, caption, label):
-	file.write("\\begin{table} \n \\begin{scriptsize} \\caption{" + caption + "} \n \\label{" + label + "} \n \\begin{tabular*}{0.48\\textwidth}{@{}l@{\\;\\;\extracolsep{\\fill}}rr")
+	file.write("\\begin{table} \n \\begin{scriptsize} \\caption{" + caption + "} \n \\label{" + label + "} \n \\begin{tabular*}{\\linewidth}{@{}l@{\\;\\;\extracolsep{\\fill}}rr")
 	file.write("@{}}\\toprule \n")
 	file.write(" type of matrix & d'Asp 07 & d'Asp 08 ")
 	file.write("\\\ \midrule \n")
@@ -270,7 +270,7 @@ def LhsResultsTable(instancesets, instancesetnames, caption, label):
 
 
 def RhsResultsTable(instancesets, instancesetnames, caption, label):
-	file.write("\\begin{table} \n \\begin{scriptsize} \\caption{" + caption + "} \n \\label{" + label + "} \n \\begin{tabular*}{0.48\\textwidth}{@{}l@{\\;\\;\extracolsep{\\fill}}rr")
+	file.write("\\begin{table} \n \\begin{scriptsize} \\caption{" + caption + "} \n \\label{" + label + "} \n \\begin{tabular*}{\\linewidth}{@{}l@{\\;\\;\extracolsep{\\fill}}rr")
 	file.write("@{}}\\toprule \n")
 	file.write(" type of matrix & d'Asp 07 & d'Asp 08 ")
 	file.write("\\\ \midrule \n")
[…]

### 2016-03-15 RIP-MISDP-Paper/ResultScript.py
combined soling times into a single table, also removed mathcal for N(0,1)

@@ -11,7 +11,8 @@ Lbounds = 0
 Rbounds = 0
 LRbounds = 0
 Ltimes = 0
-Rtimes = 1
+Rtimes = 0
+LRtimes = 1
 completeTable = 0
 
 MISDPfilename = "/local/gally/results/RIP-MISDP-Paper/160311/RIP-results/check.RIPMISDP.scipsdp.linux.x86_64.gnu.opt.sdpa.extra.branchinfobj_nofracdive.out"
@@ -206,7 +207,7 @@ def makeCompleteTableCaptionMISDP(shortcaption, caption, label, file, i):
 def LhsResultsTable(instancesets, instancesetnames, caption, label):
 	file.write("\\begin{table} \n \\begin{scriptsize} \\caption{" + caption + "} \n \\label{" + label + "} \n \\begin{tabular*}{\\linewidth}{@{}l@{\\;\\;\extracolsep{\\fill}}rr")
 	file.write("@{}}\\toprule \n")
-	file.write(" type of matrix & d'Asp 07 & d'Asp 08 ")
+	file.write(" matrices & d'Asp 07 & d'Asp 08 ")
 	file.write("\\\ \midrule \n")
 	totalDiff07 = 0
 	totalnInstances07 = 0
@@ -272,7 +273,7 @@ def LhsResultsTable(instancesets, instancesetnames, caption, label):
 def RhsResultsTable(instancesets, instancesetnames, caption, label):
 	file.write("\\begin{table} \n \\begin{scriptsize} \\caption{" + caption + "} \n \\label{" + label + "} \n \\begin{tabular*}{\\linewidth}{@{}l@{\\;\\;\extracolsep{\\fill}}rr")
 	file.write("@{}}\\toprule \n")
-	file.write(" type of matrix & d'Asp 07 & d'Asp 08 ")
+	file.write(" matrices & d'Asp 07 & d'Asp 08 ")
 	file.write("\\\ \midrule \n")
 	totalDiff07 = 0
 	totalnInstances07 = 0
@@ -340,7 +341,7 @@ def LhsRhsResultsTable(instancesets, instancesetnames, caption, label):
 	file.write("@{}}\\toprule \n")
 	file.write("  & \\multicolumn{2}{c}{$\\alpha_k$} & \\multicolumn{2}{c}{$\\beta_k$} \\\ \n")
[…]

### 2016-03-15 RIP-MISDP-Paper/ResultScript.py
also ommit rk1 lhs from solving times

@@ -532,18 +532,21 @@ def LhsRhsTimeTable(instancesets, instancesetnames, caption, label):
 		A08timer = 1.0
 		A08numr = 0
 		for j1,j2 in instancetype:
-			MISDPtimel = math.pow(MISDPtimel, float(MISDPnuml) / float(MISDPnuml+1)) * math.pow(float(times[0][j1] + timeshift), 1.0/float(MISDPnuml+1))
-			MISDPnuml += 1
-			totalMISDPtimel = math.pow(totalMISDPtimel, float(totalMISDPnuml) / float(totalMISDPnuml+1)) * math.pow(float(times[0][j1] + timeshift), 1.0/float(totalMISDPnuml+1))
-			totalMISDPnuml += 1
-			A07timel = math.pow(A07timel, float(A07numl) / float(A07numl+1)) * math.pow(float(times[1][j1] + timeshift), 1.0/float(A07numl+1))
-			A07numl += 1
-			totalA07timel = math.pow(totalA07timel, float(totalA07numl) / float(totalA07numl+1)) * math.pow(float(times[1][j1] + timeshift), 1.0/float(totalA07numl+1))
-			totalA07numl += 1
-			A08timel = math.pow(A08timel, float(A08numl) / float(A08numl+1)) * math.pow(float(times[2][j1] + timeshift), 1.0/float(A08numl+1))
-			A08numl += 1
-			totalA08timel = math.pow(totalA08timel, float(totalA08numl) / float(totalA08numl+1)) * math.pow(float(times[2][j1] + timeshift), 1.0/float(totalA08numl+1))
-			totalA08numl += 1
+			if j1 > -1:
+				MISDPtimel = math.pow(MISDPtimel, float(MISDPnuml) / float(MISDPnuml+1)) * math.pow(float(times[0][j1] + timeshift), 1.0/float(MISDPnuml+1))
+				MISDPnuml += 1
+				totalMISDPtimel = math.pow(totalMISDPtimel, float(totalMISDPnuml) / float(totalMISDPnuml+1)) * math.pow(float(times[0][j1] + timeshift), 1.0/float(totalMISDPnuml+1))
+				totalMISDPnuml += 1
[…]

### 2016-03-15 RIP-MISDP-Paper/ResultScript.py
corrected label

@@ -626,7 +626,7 @@ if __name__=="__main__":
 	if Rtimes:
 		TimeTable([[55,57,59,61,63,65,67,69,71],[19,21,23,25,27,29,31,33,35],[91,93,95,97,99,101,103,105,107],[109,111,113,115,117,119,121,123,125],[73,75,77,79,81,83,85,87,89],[37,39,41,43,45,47,49,51,53],[1,3,5,7,9,11,13,15,17]], ["$N(0,1)$", "binary", "band matrix", "rank 1", "$N(0,1/m)$", "$\\pm 1/\\sqrt{m}$", "$0, \\pm \\sqrt{3/m}$"], "Average Solving times for right-hand side of RIP", "rhsTime")
 	if LRtimes:
-		LhsRhsTimeTable([[[54,55],[56,57],[58,59],[60,61],[62,63],[64,65],[66,67],[68,69],[70,71]],[[18,19],[20,21],[22,23],[24,25],[26,27],[28,29],[30,31],[32,33],[34,35]],[[90,91],[92,93],[94,95],[96,97],[98,99],[100,101],[102,103],[104,105],[106,107]],[[-1,109],[-1,111],[-1,113],[-1,115],[-1,117],[-1,119],[-1,121],[-1,123],[-1,125]],[[72,73],[74,75],[76,77],[78,79],[80,81],[82,83],[84,85],[86,87],[88,89]],[[36,37],[38,39],[40,41],[42,43],[44,45],[46,47],[48,49],[50,51],[52,53]],[[0,1],[2,3],[4,5],[6,7],[8,9],[10,11],[12,13],[14,15],[16,17]]], ["$N(0,1)$", "binary", "band matrix", "rank 1", "$N(0,1/m)$", "$\\pm 1/\\sqrt{m}$", "$0, \\pm \\sqrt{3/m}$"], "Average Solving times for RICs", "rhsTime")
[…]

### 2016-03-15 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
combined soling times into a single table, also removed mathcal for N(0,1)
unified references to the different relaxations in the numerical results section
also changed align to equation/aligned and made further small changes

@@
 for the \emph{denoising} case, where instead of~\eqref{l0min} one considers
-\begin{align}\label{l0minDenoising}
-  \min \quad & \Norm{x}{0} \tag{$P_0^\delta$}\\
-  \text{s.t.} \quad & \Norm{Ax - b}{2} \leq \delta, \nonumber
-\end{align}
+\begin{equation}\label{l0minDenoising}\tag{$P_0^\delta$}
+ \begin{aligned}
+  \min \quad & \Norm{x}{0} \\
+  \text{s.t.} \quad & \Norm{Ax - b}{2} \leq \delta, 
+ \end{aligned}
+\end{equation}
 for some $\delta \geq 0$; the $\ell_1$-problem~\eqref{l1min} is changed
@@
 by the non-convex quadratic optimization problem
-\begin{align}\label{QP}
- \min \quad & \Norm{Ax}{2}^2 \nonumber \\
- \text{s.t.} \quad & \Norm{x}{2}^2 = 1, \tag{QP} \\
- & \Norm{x}{0} \leq k. \nonumber
-\end{align}
+\begin{equation}\label{QP}\tag{QP}
+ \begin{aligned}
+  \min \quad & \Norm{Ax}{2}^2  \\
+  \text{s.t.} \quad & \Norm{x}{2}^2 = 1,  \\
+  & \Norm{x}{0} \leq k. 
+ \end{aligned}
+\end{equation}
 The corresponding maximization problem allows to compute $\beta_k^2$. D'Aspremont et al. \cite{Asp07} tackled the non-convex quadratic equality constraint 
-$\Norm{x}{2}^2 = 1$ by the technique of semidefinite lifting, first introduced by Goemans and and Williamson~\cite{GW95} 
+$\Norm{x}{2}^2 = 1$ by the technique of semidefinite lifting, first introduced by Goemans and Williamson~\cite{GW95} 
 for the max-cut problem.
@@
 Relaxing the non-convex rank constraint finally leads to the following SDP:
-\begin{align}\label{Asp07}
- \min \quad & \Tr(A\T A X) \nonumber \\
- \text{s.t.} \quad & \Tr(X) = 1, \tag{A1} \\
- & \ones\T \abs{X} \ones \leq k, \nonumber \\
+\begin{equation}\label{Asp07}\tag{A1}
+ \begin{aligned}
+  \min \quad & \Tr(A\T A X) \\
+  \text{s.t.} \quad & \Tr(X) = 1, \\
+  & \ones\T \abs{X} \ones \leq k, \\
   & X \succeq 0. \nonumber
-\end{align}
+ \end{aligned}
+\end{equation}
 
@@
 $\Sigma = A\T A$ we can use this to compute an upper bound on $\beta_k$ via Lagrangian Relaxation as
-\begin{equation*}
- \begin{aligned}
-  \beta_k^2 & = \max \big\{ x\T A\T A x \suchthat \Norm{x}{2} \leq 1,\; \Norm{x}{0} \leq k \big\}\\
+ \begin{align*}
+  \beta_k^2 & = \max \big\{ x\T A\T A x \suchthat \Norm{x}{2} \leq 1,\; \Norm{x}{0} \leq k \big\} \\
 	    & \leq \inf_{\rho \geq 0} \max \big\{ x\T A\T A x - \rho 
-            (\Norm{x}{0} - k) \suchthat \Norm{x}{2} \leq 1 \big\}\\
+            (\Norm{x}{0} - k) \suchthat \Norm{x}{2} \leq 1 \big\} \\
 	    & = \inf_{\rho \geq 0} \phi(\rho) + \rho\, k. 
- \end{aligned}
-\end{equation*}
+ \end{align*}
 
 As one way to solve \eqref{phi} the semidefinite lifting technique is used again to compute an upper bound on $\phi$ via
-\begin{align}\label{Asp08}
- \max \quad & \sum_{i=1}^n \Tr(P_i \, B_i) \nonumber \\
- \text{s.t.} \quad & \Tr(X) = 1, \tag{A2-Primal}\\
- & X \succeq P_i \succeq 0 \quad \text{for } i = 1, \dots, n, \nonumber
-\end{align}
+\begin{equation}\label{Asp08Primal}\tag{A2-Primal}
+ \begin{aligned}
+  \max \quad & \sum_{i=1}^n \Tr(P_i \, B_i) \\
+  \text{s.t.} \quad & \Tr(X) = 1, \\
+  & X \succeq P_i \succeq 0 \quad \text{for } i = 1, \dots, n,
+ \end{aligned}
+\end{equation}
 for $B_i = b_i\,b_i\T - \rho\, I$, where $b_i$ is the $i$-th column of the
 square-root of $\Sigma$. For details, we refer the reader to~\cite{Asp08}. The corresponding dual problem can then be written as 
-\begin{align}\label{Asp08Dual}
-  \min \quad & \lambda_{\max}\Big(\sum_{i=1}^n Y_i \Big) \nonumber \\
-  \text{s.t.} \quad & Y_i \succeq B_i \quad \text{for } i = 1, \dots, n \tag{A2-Dual} \\
-  & Y_i \succeq 0 \quad \ \ \text{for } i = 1, \dots, n, \nonumber
-\end{align}
-which has only half as many variables as \eqref{Asp08} and one fewer linear and semidefinite constraint.
+\begin{equation}\label{Asp08Dual}\tag{A2-Dual}
+ \begin{aligned}
+  \min \quad & \lambda_{\max}\Big(\sum_{i=1}^n Y_i \Big) \\
+  \text{s.t.} \quad & Y_i \succeq B_i \quad \text{for } i = 1, \dots, n \\
+  & Y_i \succeq 0 \quad \ \ \text{for } i = 1, \dots, n,
+ \end{aligned}
+\end{equation}
+which has only half as many variables as \eqref{Asp08Primal} and one fewer linear and semidefinite constraint.
 Then \eqref{Asp08Dual} can be solved for non-negative $\rho$, which should be smaller than $\Sigma_{11}$, because otherwise the
@@
 As shown in \cite{Asp08}, a lower bound for $\alpha_k$ can be computed via
-\begin{equation*}
+\begin{equation}\tag{A2}\label{Asp08}
  \alpha_k^2 \geq \sup_{\rho \geq 0}\; \psi(\rho) - \rho\, k, 
-\end{equation*}
+\end{equation}
 with 
@@
 \end{equation*}
-In this case problem \eqref{Asp08} or equivalently \eqref{Asp08Dual} has to
+In this case problem \eqref{Asp08Primal} or equivalently \eqref{Asp08Dual} has to
 be solved for $\Sigma = M I - A\T A$, where $M \in \R$ has to be big enough
@@
 \end{equation*}
-In \cite{AspBG14} explicit bounds on the gap between \eqref{Asp08} and the penalized problem \eqref{phi} were derived using randomization arguments.
+In \cite{AspBG14} explicit bounds on the gap between \eqref{Asp08Primal} and the penalized problem \eqref{phi} were derived using randomization arguments.
 
@@
 \leq k$. We can thus equivalently rewrite \eqref{QP} as
-\begin{align}\label{Rk1MISDP}
- \min \quad & \Tr(A\T A X) \nonumber \\
- \text{s.t.} \quad & \Tr(X) = 1, \nonumber \\
- & -z_j \leq X_{ij} \leq z_j \quad \text{for } j = 1, \dots, n, \nonumber \\
+\begin{equation}\label{Rk1MISDP}
+ \begin{aligned}
+  \min \quad & \Tr(A\T A X) \\
+  \text{s.t.} \quad & \Tr(X) = 1, \\
+  & -z_j \leq X_{ij} \leq z_j \quad \text{for } j = 1, \dots, n, \\
   & \sum_{i=1}^n z_i \leq k,\\
- & \Rk(X) = 1, \nonumber \\
- & X \succeq 0,  \nonumber \\
- & z \in \{0,1\}^n. \nonumber
-\end{align}
+  & \Rk(X) = 1, \\
+  & X \succeq 0, \\
+  & z \in \{0,1\}^n.
+ \end{aligned}
+\end{equation}
 We then relax the non-convex rank constraint to arrive at the following
 MISDP formulation:
-\begin{align}\label{MISDP}
- \min \quad & \Tr(A\T A X) \nonumber \\
- \text{s.t.} \quad & \Tr(X) = 1, \nonumber \\
- & -z_j \leq X_{ij} \leq z_j \quad \text{for } j = 1, \dots, n, \nonumber \\
- & \sum_{i=1}^n z_i \leq k, \tag{MISDP} \\
- & X \succeq 0,  \nonumber \\
- & z \in \{0,1\}^n. \nonumber
-\end{align}
+\begin{equation}\label{MISDP}\tag{MISDP}
+ \begin{aligned}
+  \min \quad & \Tr(A\T A X) \\
+  \text{s.t.} \quad & \Tr(X) = 1, \\
+  & -z_j \leq X_{ij} \leq z_j \quad \text{for } j = 1, \dots, n, \\
+  & \sum_{i=1}^n z_i \leq k, \\
+  & X \succeq 0, \\
+  & z \in \{0,1\}^n.
+ \end{aligned}
+\end{equation}
 In the following, we will show that \eqref{MISDP} actually produces the
@@
   the columns indexed by~$T$. Then consider
-  \begin{align}\label{Proj}
-    \min \quad & \Tr(\check{A}\T \check{A} \check{X}), \nonumber \\
+  \begin{equation}\label{Proj}
+   \begin{aligned}
+    \min \quad & \Tr(\check{A}\T \check{A} \check{X}), \\
     \text{s.t.} \quad & \Tr(\check{X}) = 1, \\
-    & \check{X} \succeq 0,\; \check{X} \in S_\ell.\nonumber 
-  \end{align}
+    & \check{X} \succeq 0,\; \check{X} \in S_\ell.
+   \end{aligned}
+  \end{equation}
 
@@
 In this section we demonstrate the applicability of the proposed MISDP-formulation and use it to
-assess the quality of the relaxations proposed in \cite{Asp08} and \cite{Asp07}.
-For solving the mixed-integer semidefinite programs we use the MISDP solver SCIP-SDP \cite{SCIP-SDP},
-originally developed in \cite{Mar13} and \cite{MS12}. It combines the
+assess the quality of the relaxations proposed in~\cite{Asp08} and~\cite{Asp07}.
+For solving the mixed-integer semidefinite programs we use the MISDP solver SCIP-SDP~\cite{SCIP-SDP},
+originally developed in~\cite{Mar13} and~\cite{MS12}. It combines the
 branch and bound framework SCIP 3.2.1 \cite{SCIP} with 
@@
 constant $\alpha_k$ and $\beta_k$ either exactly using~\eqref{MISDP} or
-produce corresponding bounds via \eqref{Asp07} and \eqref{Asp08Dual}. In
-the latter case, we solve 15 SDPs for $\rho$ between~$0$ and the maximum
+produce corresponding bounds via~\eqref{Asp07} and~\eqref{Asp08}. In
+the latter case, we solve the SDP~\eqref{Asp08Dual} 15 times for $\rho$ between~$0$ and the maximum
 diagonal entry of $A\T A$. This is the same approach as in \cite{Asp08},
 except that we reduced the number of problems from 25 to 15, as this did
-not seem to influence the quality of th solutions much in our preliminary
+not seem to influence the quality of the solutions much in our preliminary
 experiments but brought the solving times more in line with the other
@@
 
-We use a testset consisting of 63 matrices with seven 
+We use a testset consisting of 63 matrices of seven 
 different types including band matrices, 
@@
 matrices $A = aa\T$ with $N(0,1)$-distributed entries of $a$. In addition to these four types,
-we also use three more for which theoretical results on restricted isometry
-constants are available. This includes matrices with 
+we also use three more with theoretical results on their restricted isometry
+constants. These include matrices with 
 $N(0, 1/m)$-distributed entries, with entries chosen uniformly in $\pm 1/\sqrt{m}$ and finally with distribution
@@
 a time limit of four hours. The code was compiled with gcc 4.4.5 with \texttt{-O3} optimization. For the 
-Problem~\eqref{Asp08Dual} most of the larger instance were aborted because of the memory limit of 8GB. These are
+Problem~\eqref{Asp08}, most of the larger instance were aborted because of the memory limit of 8GB. These are
 counted with the maximum allowed time of four hours.
@@
 For the left-hand-side of the RIP, $\alpha_k$, both relaxations did not perform
-too well. Especially the relaxation given in \cite{Asp08} regularly returned worse solutions than the
-trivial lower bound of zero. The relaxation in \cite{Asp07} performed slightly better, but the bounds
+too well. Especially~\eqref{Asp08} regularly returned worse solutions than the
+trivial lower bound of zero. Relaxation~\eqref{Asp07} performed slightly better, but the bounds
 were still not nearly good enough to have any chance of proving $\ell_0$-$\ell_1$-equivalence. The relatively
@@
 perform much better. Problem~\eqref{Asp07} constantly produces bounds within 10 to 30\,\% of the optimal
-solution, with only slightly worse results for binary and rank one matrices. The formulation in~\eqref{Asp08Dual}
+solution, with only slightly worse results for binary and rank one matrices. The formulation~\eqref{Asp08}
 shows bigger fluctuations, but produces better results for five of the seven matrix types, while performing even
@@
 The solving times in seconds for the left- and right-hand side of the RIP are given in
-Table~\ref{lhsTime} and~\ref{rhsTime}, respectively. The relaxation proposed in \cite{Asp07}
+Table~\ref{lhsRhsTime}. The relaxation~\eqref{Asp07}
 is very fast, taking only a few seconds for all instances. The mixed-integer semidefinite program obviously
 cannot compete with these solving times, although for band-matrices, the difference is relatively
-small. The Lagrangian Relaxation from \cite{Asp08}, however, takes more time to produce bounds than is needed
-by our MISDP to compute exact solutions. The big difference is also caused by the instances running into the 
-memory limit, but even without those, the algorithm still takes slightly longer than the MISDP.
+small. The Lagrangian Relaxation~\eqref{Asp08}, however, takes more time to produce bounds than is needed
+by~\eqref{MISDP} to compute exact solutions. The big difference is also caused by the instances running into the 
+memory limit, but even without those, the algorithm still takes slightly longer than~\eqref{MISDP}.
 
-\input{tables/lhsTime}
-
-\input{tables/rhsTime}
+\input{tables/lhsRhsTime}
 
@@
 computed via mixed-integer semidefinite programming. The computed bounds
-significantly improve the bounds computed by the semidefinite relaxations
-proposed in~\cite{Asp07} and \cite{Asp08}. In conclusion, the relaxation of
-\cite{Asp07} is fast to compute, but produces a gap of arround 90\,\% for
-$\alpha_k$ and 30\,\% for $\beta_k$. If successfully solved, the relaxation
-of \cite{Asp08} provides better bounds for $\beta_k$ for most types of 
+significantly improve the bounds computed by the semidefinite
+relaxations~\eqref{Asp07} and~\eqref{Asp08}
+proposed in~\cite{Asp07} and \cite{Asp08}, respectively. In conclusion, the relaxation~
+\eqref{Asp07} is fast to compute, but produces a gap of arround 90\,\% for
+$\alpha_k$ and 30\,\% for $\beta_k$. If successfully solved, 
+\eqref{Asp08} provides better bounds for $\beta_k$ for most types of 
 matrices, but its running time is too large
 to be used in practice. Moreover, directly solving~\eqref{MISDP} is often
-faster and produces the exact value. Consequently, the relaxation
-in~\cite{Asp08} should not be used in practice.
+faster and produces the exact value. Consequently, 
+\eqref{Asp08} should not be used in practice.
 

### 2016-03-15 RIP-MISDP-Paper/ResultScript.py

@@ -207,7 +207,7 @@ def makeCompleteTableCaptionMISDP(shortcaption, caption, label, file, i):
 def LhsResultsTable(instancesets, instancesetnames, caption, label):
 	file.write("\\begin{table} \n \\begin{scriptsize} \\caption{" + caption + "} \n \\label{" + label + "} \n \\begin{tabular*}{\\linewidth}{@{}l@{\\;\\;\extracolsep{\\fill}}rr")
 	file.write("@{}}\\toprule \n")
-	file.write(" matrices & d'Asp 07 & d'Asp 08 ")
+	file.write(" matrices & \eqref{Asp07} & \eqref{Asp08} ")
 	file.write("\\\ \midrule \n")
 	totalDiff07 = 0
 	totalnInstances07 = 0
@@ -273,7 +273,7 @@ def LhsResultsTable(instancesets, instancesetnames, caption, label):
 def RhsResultsTable(instancesets, instancesetnames, caption, label):
 	file.write("\\begin{table} \n \\begin{scriptsize} \\caption{" + caption + "} \n \\label{" + label + "} \n \\begin{tabular*}{\\linewidth}{@{}l@{\\;\\;\extracolsep{\\fill}}rr")
 	file.write("@{}}\\toprule \n")
-	file.write(" matrices & d'Asp 07 & d'Asp 08 ")
+	file.write(" matrices & \eqref{Asp07} & \eqref{Asp08} ")
 	file.write("\\\ \midrule \n")
 	totalDiff07 = 0
 	totalnInstances07 = 0
@@ -341,7 +341,7 @@ def LhsRhsResultsTable(instancesets, instancesetnames, caption, label):
 	file.write("@{}}\\toprule \n")
 	file.write("  & \\multicolumn{2}{c}{$\\alpha_k$} & \\multicolumn{2}{c}{$\\beta_k$} \\\ \n")
 	file.write("\\cmidrule(r){2-3} \\cmidrule(l){4-5} \n")
-	file.write("matrices & d'Asp07 & d'Asp08 & d'Asp07 & d'Asp08 \\ \n")
+	file.write("matrices & \eqref{Asp07} & \eqref{Asp08} & \eqref{Asp07} & \eqref{Asp08} \\ \n")
[…]

### 2016-03-16 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
Rank instead of Rk since we never introduce it formally
added table with sizes of used matrices

@@
 \newcommand{\Tr}{\text{Tr}}
-\newcommand{\Rk}{\text{Rk}}
+\newcommand{\Rk}{\text{Rank}}
 \newcommand{\vect}{\text{vec}}
@@
 matrices of three different sizes with rows ranging from 15 to 30, columns
-from 25 to 40 and order $k$ between three and five.
+from 25 to 40 and order $k$ between three and five, see Table~\ref{matrixsizes}.
+
+\begin{table} 
+ \begin{scriptsize} \caption{Sizes of the used matrices, three matrices were
+ randomly generated for each size} 
+ \label{matrixsizes} 
+ \begin{tabular*}{\linewidth}{@{}l@{\;\;\extracolsep{\fill}}rrrrrrrrr@{}}\toprule 
+  & \multicolumn{3}{c}{1} & \multicolumn{3}{c}{2} & \multicolumn{3}{c}{3} \\ 
+\cmidrule(r){2-4} \cmidrule{5-7}  \cmidrule(l){8-10}
+type & $m$ & $n$ & $k$ & $m$ & $n$ & $k$ & $m$ & $n$ & $k$  \\ 
+\midrule 
+$N(0,1)$ & 15 & 30 & 5 & 25 & 35 & 4 & 30 & 40 & 3 \\
+binary & 15 & 30 & 5 & 25 & 35 & 4 & 30 & 40 & 3 \\
+band matrix & 30 & 30 & 5 & 35 & 35 & 4 & 40 & 40 & 3 \\
+rank 1& 30 & 30 & 5 & 35 & 35 & 4 & 40 & 40 & 3 \\
+$N(0,1/m)$ & 15 & 30 & 5 & 25 & 35 & 4 & 30 & 40 & 3 \\
+$\pm 1/\sqrt{m}$ & 15 & 30 & 5 & 25 & 35 & 4 & 30 & 40 & 3 \\
+$0, \pm \sqrt{3/m}$ & 15 & 30 & 5 & 25 & 35 & 4 & 30 & 40 & 3 \\
+\bottomrule 
+ \end{tabular*} 
+ \end{scriptsize} 
+ \end{table}
 

### 2016-03-16 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
explained size as reason of long solving times for Asp08, also explained gap

@@
 In Table~\ref{lhsRhsGap} we compare the relative gap between the bounds on $\alpha_k$ and $\beta_k$ computed by the two heuristics
-and the exact values given by~\eqref{MISDP}. We present arithmetic means over all of the nine matrices for the given type that
+and the exact values given by~\eqref{MISDP}, computed as their difference divided by the exact value. We present arithmetic means over all of the nine matrices for the given type that
 could be solved within the time- and memory-limit. Note that we omitted the rank one matrices for the left-hand side since these satisfy
@@
 small. The Lagrangian Relaxation~\eqref{Asp08}, however, takes more time to produce bounds than is needed
-by~\eqref{MISDP} to compute exact solutions. The big difference is also caused by the instances running into the 
+by~\eqref{MISDP} to compute exact solutions. This is mainly caused by the fact that~\eqref{Asp08Dual} involves
+$n+1$ $n\times n$ matrix variables and SDP-blocks instead of a single one each in~\eqref{MISDP} and~\eqref{Asp07}.
+Furthermore, the big difference is also caused by the instances running into the 
 memory limit, but even without those, the algorithm still takes slightly longer than~\eqref{MISDP}.

### 2016-03-16 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
slight reformulation

@@
 by~\eqref{MISDP} to compute exact solutions. This is mainly caused by the fact that~\eqref{Asp08Dual} involves
-$n+1$ $n\times n$ matrix variables and SDP-blocks instead of a single one each in~\eqref{MISDP} and~\eqref{Asp07}.
+$n+1$ matrix variables of dimension $n\times n$ with corresponding SDP-constraints instead of a 
+single one each in~\eqref{MISDP} and~\eqref{Asp07}.
 Furthermore, the big difference is also caused by the instances running into the 

### 2016-03-16 RIP-MISDP-Paper/ResultScript.py
explicitly state the number of aborted instances for Asp08

@@ -12,7 +12,8 @@ Rbounds = 0
 LRbounds = 0
 Ltimes = 0
 Rtimes = 0
-LRtimes = 1
+LRtimes = 0
+printFails = 1
 completeTable = 0
 
 MISDPfilename = "/local/gally/results/RIP-MISDP-Paper/160311/RIP-results/check.RIPMISDP.scipsdp.linux.x86_64.gnu.opt.sdpa.extra.branchinfobj_nofracdive.out"
@@ -92,11 +93,15 @@ def readFileAsp07(filename, j, rhs):
 	iters = substring[substring.find("   Iteration = ") + 15:].split()[0]
 	sdpiters[1][j] = float(iters)
 	returncode = substring[substring.find("phase.value  = ") + 15:].split()[0]
+	if substring.find("phase.value  = ") == -1:
+		Asp07unsolved[j] = -1
+	else:
+		returncode = substring[substring.find("phase.value  = ") + 15:].split()[0]
+		if returncode != "pdOPT":
+			Asp07unsolved[j] = 1
 	if rhs:
 		primalresults[1][j] *= -1
 		dualresults[1][j] *= -1
-	#if returncode != "pdOPT":
-		#print(filename + " returned " + returncode) 
 	file.close()
 
 def readFileAsp08(basename, datafilename, j, rhs):
@@ -130,9 +135,12 @@ def readFileAsp08(basename, datafilename, j, rhs):
 		else:
 			iters = substring[substring.find("   Iteration = ") + 15:].split()[0]
 			Asp08sdpiters[j][i] = float(iters)
-		returncode = substring[substring.find("phase.value  = ") + 15:].split()[0]
-		#if returncode != "pdOPT":
-			#print(filename + " returned " + returncode) 
+		if substring.find("phase.value  = ") == -1:
+			Asp08unsolved[j][i] = -1
+		else:
+			returncode = substring[substring.find("phase.value  = ") + 15:].split()[0]
+			if returncode != "pdOPT":
+				Asp08unsolved[j][i] = 1	#TODO maybe also reset results, iters etc. or check for gap below
 		file.close()
 	# compute best bound
 	datafilename = basename + "_1" + ".dat-s"
@@ -582,6 +590,34 @@ def LhsRhsTimeTable(instancesets, instancesetnames, caption, label):
[…]

### 2016-03-17 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
explicitly state the number of aborted instances for Asp08
added hyperref

@@
 \usepackage{sistyle}
+
+\usepackage[breaklinks,colorlinks,citecolor=blue,linkcolor=blue,urlcolor=blue,bookmarks=false,hypertexnames=false,
+  pdftitle={Computing Exact Restricted Isometry Constants via Mixed Integer Semidefinite Programming},
+  pdfauthor={T. Gally, and M. E. Pfetsch}]{hyperref}
 
@@
 
-\title{Computing Exact Restricted Isometry Constants via\\ Mixed Integer Semidefinite Programming}
+\title{Computing Exact Restricted Isometry Constants via Mixed Integer Semidefinite Programming}
+%\title{Computing Restricted Isometry Constants Exactly via Mixed Integer Semidefinite Programming}
 \author{Tristan~Gally and Marc~E.~Pfetsch
@@
 Linux. Each computation was performed single-threaded with a single process running on each computer and with 
-a time limit of four hours. The code was compiled with gcc 4.4.5 with \texttt{-O3} optimization. For the 
-Problem~\eqref{Asp08}, most of the larger instance were aborted because of the memory limit of 8GB. These are
-counted with the maximum allowed time of four hours.
+a time limit of four hours. The code was compiled with gcc 4.4.5 with \texttt{-O3} optimization. For 
+Problem~\eqref{Asp08}, the 585 evaluations of~\eqref{Asp08Dual} for 39 matrices and all 15 choices of $\rho$
+were aborted because of the memory limit of 8GB. These are counted with the maximum allowed time of four hours.
 

### 2016-03-17 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
small corrections and readded the missing i =!= j to the 1/2-Lemma

@@
 MISDP, we will show that the rank constraint can be relaxed without
-changing the optimal value. We can then apply a MISDP solver.
+changing the optimal value. We can then apply an MISDP solver.
 
@@
   \min \quad & \lambda_{\max}\Big(\sum_{i=1}^n Y_i \Big) \\
-  \text{s.t.} \quad & Y_i \succeq B_i \quad \text{for } i = 1, \dots, n \\
+  \text{s.t.} \quad & Y_i \succeq B_i \quad \text{for } i = 1, \dots, n, \\
   & Y_i \succeq 0 \quad \ \ \text{for } i = 1, \dots, n,
@@
 \noindent
-In this section we propose a MISDP formulation to compute the optimal
+In this section we propose an MISDP formulation to compute the optimal
 constants $\alpha_k$ and $\beta_k$ in \eqref{lowerRIP} and
@@
 values agree. This suffices for our application of
-computing the RIP constants $\alpha_k$ and $\beta_k$. For other
+computing the restricted isometry constants $\alpha_k$ and $\beta_k$. For other
 applications like SPCA, an extreme solution can be generated with an
@@
   \end{equation}
-  for $i, j \in \{1, \dots, n\}$, are valid for~\eqref{MISDP}.
+  for $i \neq j \in \{1, \dots, n\}$, are valid for~\eqref{MISDP}.
 \end{lemma}
@@
   nonnegativity of the diagonal entries imply $X_{ii}X_{jj}\leq
-  \tfrac{1}{4}$.  Since all minors of $X$ need to be nonnegative, it
+  \tfrac{1}{4}$ for $i \neq j$. Since all minors of $X$ need to be nonnegative, it
   follows that
@@
 
-In Table~\ref{lhsRhsGap} we compare the relative gap between the bounds on $\alpha_k$ and $\beta_k$ computed by the two heuristics
+In Table~\ref{lhsRhsGap} we compare the relative gap between the bounds on $\alpha_k$ and $\beta_k$ produced by the two heuristics
 and the exact values given by~\eqref{MISDP}, computed as their difference divided by the exact value. We present arithmetic means over all of the nine matrices for the given type that
@@
 
-For reporting some the solving times, we will use the \emph{shifted
+For reporting the solving times, we will use the \emph{shifted
   geometric mean} to decrease the influence of easy instances, see

### 2016-03-18 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
spellcheck
remove old comment

@@
 corresponding optimal objective, this implies a lower bound on $\alpha_k$
-in\comment{Ist $M\,I$ statt $M$ korrekt? \\ $\Rightarrow$ nein, sollten alles Skalare sein, $\Psi$ demenstrechend optimaler Zielfunktionswert statt L\"osung (habe ich jetzt oben angepasst)}
+in
 \begin{equation*}
@@
 proposed in~\cite{Asp07} and \cite{Asp08}, respectively. In conclusion, the relaxation~
-\eqref{Asp07} is fast to compute, but produces a gap of arround 90\,\% for
+\eqref{Asp07} is fast to compute, but produces a gap of around 90\,\% for
 $\alpha_k$ and 30\,\% for $\beta_k$. If successfully solved, 

### 2016-03-18 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
paper: reformulation

@@
 diagonal entry of $A\T A$. This is the same approach as in \cite{Asp08},
-except that we reduced the number of problems from 25 to 15, as this did
-not seem to influence the quality of the solutions much in our preliminary
-experiments but brought the solving times more in line with the other
-approaches.
+except that we reduced the number from 25 to 15, as this did
+not significantly seem to influence the quality of the solutions,
+but reduced the solving times.
 

### 2016-03-18 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
paper: reformulation for aborts
mixed-integer -> mixed integer

@@
   pdftitle={Computing Exact Restricted Isometry Constants via Mixed Integer Semidefinite Programming},
+  % pdftitle={Computing Restricted Isometry Constants via Mixed Integer Semidefinite Programming},
   pdfauthor={T. Gally, and M. E. Pfetsch}]{hyperref}
@@
 \title{Computing Exact Restricted Isometry Constants via Mixed Integer Semidefinite Programming}
-%\title{Computing Restricted Isometry Constants Exactly via Mixed Integer Semidefinite Programming}
+%\title{Computing Restricted Isometry Constants via Mixed Integer Semidefinite Programming}
 \author{Tristan~Gally and Marc~E.~Pfetsch
@@
   this condition is itself \NP-hard. In this paper we propose a
-  mixed-integer semidefinite programming approach for computing these
+  mixed integer semidefinite programming approach for computing these
   optimal constants. This also subsumes sparse principal component
@@
 The goal of this paper is to compute optimal restricted isometry constants
-via mixed-integer semidefinite programming (MISDP). To this end, we will
+via mixed integer semidefinite programming (MISDP). To this end, we will
 introduce an MISDP with a rank-1 constraint. In order to utilize this
@@
 restricted isometry constant and the restricted isometry ratio by
-mixed-integer semidefinite programming.
+mixed integer semidefinite programming.
 
@@
 assess the quality of the relaxations proposed in~\cite{Asp08} and~\cite{Asp07}.
-For solving the mixed-integer semidefinite programs we use the MISDP solver SCIP-SDP~\cite{SCIP-SDP},
+For solving the mixed integer semidefinite programs we use the MISDP solver SCIP-SDP~\cite{SCIP-SDP},
 originally developed in~\cite{Mar13} and~\cite{MS12}. It combines the
@@
 Linux. Each computation was performed single-threaded with a single process running on each computer and with 
-a time limit of four hours. The code was compiled with gcc 4.4.5 with \texttt{-O3} optimization. For 
-Problem~\eqref{Asp08}, the 585 evaluations of~\eqref{Asp08Dual} for 39 matrices and all 15 choices of $\rho$
-were aborted because of the memory limit of 8GB. These are counted with the maximum allowed time of four hours.
+a time limit of four hours. The code was compiled with gcc 4.4.5 with \texttt{-O3} optimization. For 39 matrices and all 15 choices of $\rho$,
+the evaluation of~\eqref{Asp08Dual} was aborted because of the memory limit of 8GB. These are counted with the maximum allowed time of four hours.
 
@@
 Table~\ref{lhsRhsTime}. The relaxation~\eqref{Asp07}
-is very fast, taking only a few seconds for all instances. The mixed-integer semidefinite program obviously
+is very fast, taking only a few seconds for all instances. The mixed integer semidefinite program obviously
 cannot compete with these solving times, although for band-matrices, the difference is relatively
@@
 This paper shows that exact values of restricted isometry constants can be
-computed via mixed-integer semidefinite programming. The computed bounds
+computed via mixed integer semidefinite programming. The computed bounds
 significantly improve the bounds computed by the semidefinite
@@
 future, the hope is that with the advancement of solving techniques for
-mixed-integer semidefinite programs, the sizes of the matrices can be
+mixed integer semidefinite programs, the sizes of the matrices can be
 significantly increased. The components to be improved include cutting

### 2016-03-18 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
made sentence about aborted instances more clear; also small changes to tables

@@
 \begin{table} 
- \begin{scriptsize} \caption{Sizes of the used matrices, three matrices were
+ \begin{scriptsize} \caption{Sizes of the used matrices; three matrices were
  randomly generated for each size} 
@@
  \begin{tabular*}{\linewidth}{@{}l@{\;\;\extracolsep{\fill}}rrrrrrrrr@{}}\toprule 
-  & \multicolumn{3}{c}{1} & \multicolumn{3}{c}{2} & \multicolumn{3}{c}{3} \\ 
+  & \multicolumn{3}{c}{A} & \multicolumn{3}{c}{B} & \multicolumn{3}{c}{C} \\ 
 \cmidrule(r){2-4} \cmidrule{5-7}  \cmidrule(l){8-10}
@@
 Linux. Each computation was performed single-threaded with a single process running on each computer and with 
-a time limit of four hours. The code was compiled with gcc 4.4.5 with \texttt{-O3} optimization. For 39 matrices and all 15 choices of $\rho$,
-the evaluation of~\eqref{Asp08Dual} was aborted because of the memory limit of 8GB. These are counted with the maximum allowed time of four hours.
+a time limit of four hours. The code was compiled with gcc 4.4.5 with \texttt{-O3} optimization. For 39 matrices 
+the evaluation of~\eqref{Asp08Dual} was aborted for all 15 choices of $\rho$ because of the memory limit of 8GB. 
+These are counted with the maximum allowed time of four hours.
 

### 2016-03-18 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
included number of instances used to compute average in results table

@@
 and the exact values given by~\eqref{MISDP}, computed as their difference divided by the exact value. We present arithmetic means over all of the nine matrices for the given type that
-could be solved within the time- and memory-limit. Note that we omitted the rank one matrices for the left-hand side since these satisfy
-$\alpha_k = 0$ in all cases, so that the trivial lower bound would already be optimal.
+could be solved within the time- and memory-limit. Note that for the left-hand side we also omitted all instances with $\alpha_k=0$, 
+since we cannot compute a relative gap in that case. This included all rank one matrices. The number of instances used to compute the
+average is given in brackets.
 

### 2016-03-18 RIP-MISDP-Paper/ResultScript.py

@@ -9,11 +9,11 @@ from decimal import Decimal
 texfile = 0 # make the output a compilable texfile or just a figure/table that can be included?
 Lbounds = 0
 Rbounds = 0
-LRbounds = 0
+LRbounds = 1
 Ltimes = 0
 Rtimes = 0
 LRtimes = 0
-printFails = 1
+printFails = 0
 completeTable = 0
 
 MISDPfilename = "/local/gally/results/RIP-MISDP-Paper/160311/RIP-results/check.RIPMISDP.scipsdp.linux.x86_64.gnu.opt.sdpa.extra.branchinfobj_nofracdive.out"
@@ -409,21 +409,21 @@ def LhsRhsResultsTable(instancesets, instancesetnames, caption, label):
 			avgdiff08r = "-"
 		file.write(instancesetnames[i])
 		if avgdiff07l != "-":
-			file.write(" & \\num{" + "%.2f" % avgdiff07l + "}\,\% &")
+			file.write(" & \\num{" + "%.2f" % avgdiff07l + "}\,\% (" + str(nInstances07l) + ") &")
 		else:
-			file.write("& - &")
+			file.write("& -- (0) &")
 		if avgdiff08l != "-":
-			file.write(" \\num{" + "%.2f" % avgdiff08l + "}\,\% &")
+			file.write(" \\num{" + "%.2f" % avgdiff08l + "}\,\% (" + str(nInstances08l) + ") &")
 		else:
-			file.write("- &")
+			file.write("-- (0) &")
 		if avgdiff07r != "-":
-			file.write(" \\num{" + "%.2f" % avgdiff07r + "}\,\% &")
+			file.write(" \\num{" + "%.2f" % avgdiff07r + "}\,\% (" + str(nInstances07r) + ") &")
 		else:
-			file.write(" - &")
+			file.write(" -- (0) &")
 		if avgdiff08r != "-":
-			file.write(" \\num{" + "%.2f" % avgdiff08r + "}\,\% \\\ \n")
+			file.write(" \\num{" + "%.2f" % avgdiff08r + "}\,\% (" + str(nInstances08r) + ") \\\ \n")
 		else:
-			file.write("- \\\ \n")
+			file.write("-- (0) \\\ \n")
 		i += 1
 	file.write("\\midrule \n")
[…]

### 2016-03-18 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
noted that rnk1-lhs are not included in the averages, but all other instances are

@@
 The solving times in seconds for the left- and right-hand side of the RIP are given in
-Table~\ref{lhsRhsTime}. The relaxation~\eqref{Asp07}
+Table~\ref{lhsRhsTime}, again exlucding the left-hand sides for the rank one matrices
+but including all others. The relaxation~\eqref{Asp07}
 is very fast, taking only a few seconds for all instances. The mixed integer semidefinite program obviously

### 2016-03-20 src/scipsdp/branch_sdpinfobjective.c
minor style improvements in code

@@ -98,18 +98,18 @@ SCIP_DECL_BRANCHCOPY(branchCopySdpinfobjective)
 static
 SCIP_DECL_BRANCHEXECEXT(branchExecextSdpinfobjective)
 {/*lint --e{715}*/
-   int i;
-   int ncands;
-   SCIP_VAR** cands = NULL;
-   SCIP_Real* candssol; /* solution values of all candidates */
-   SCIP_Real* candsscore; /* scores of all candidates */
-   SCIP_VAR* maxtargetvar = NULL; /* variable with currently highest target value, meaning product of integer infeasibility and absolute objective */
-   SCIP_Real maxtargettarget; /* target value of the current candidate with highest target value, meaning product of integer infeasibility and absolute objective */
-   SCIP_Real maxtargetval; /* value of the current candidate with highest target value, meaning product of integer infeasibility and absolute objective */
-   SCIP_Real maxtargetscore; /* score of the current candidate with highest target value, meaning product of integer infeasibility and absolute objective */
-   SCIP_Real currentfrac; /* fractionality of the current candidate */
-   SCIP_Real currenttarget; /* target value, meaning product of integer infeasibility and absolute objective, of the current candidate */
    SCIP_BRANCHRULEDATA* branchruledata;
+   SCIP_VAR** cands = NULL;
+   SCIP_Real* candssol;              /* solution values of all candidates */
+   SCIP_Real* candsscore;            /* scores of all candidates */
+   SCIP_VAR* maxtargetvar = NULL;    /* variable with currently highest target value, i.e., product of integer infeasibility and absolute objective */
+   SCIP_Real maxtargettarget = -1.0; /* target value of current candidate with highest target value, i.e., product of integer infeasibility and absolute objective */
+   SCIP_Real maxtargetval = 0.0;     /* value of current candidate with highest target value, i.e., product of integer infeasibility and absolute objective */
+   SCIP_Real maxtargetscore = 0.0;   /* score of current candidate with highest target value, i.e., product of integer infeasibility and absolute objective */
+   SCIP_Real currentfrac;            /* fractionality of the current candidate */
+   SCIP_Real currenttarget;          /* target value, i.e., product of integer infeasibility and absolute objective, of the current candidate */
+   int ncands;
+   int i;
 
[…]

### 2016-03-20 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
minor improvement in abstract

@@
   analysis. Computational results with this approach allow to evaluate
-  earlier semidefinite relaxations and show that their quality is limited.
+  earlier semidefinite relaxations and show that the quality that can be
+  obtained in reasonable time is limited.
 \end{abstract}

### 2016-03-20 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
minor improvement in address

@@
   \thanks{This work was supported by the German Research Foundation (DFG) through Collaborative Research Center 805 and Priority Program 1798.}
-  \thanks{The  authors  are  with  the  Research  Group  Optimization,  TU  Darmstadt,
-  Darmstadt  64293,  Germany  (e-mail:  gally@mathematik.tu-darmstadt.de;
-  pfetsch@mathematik.tu-darmstadt.de).}}
+  \thanks{The  authors  are  with  the  Research  Group  Optimization,  TU
+    Darmstadt, Dolivostr.\ 5,
+  Darmstadt  64293,  Germany  (e-mail:  gally@opt.tu-darmstadt.de;
+  pfetsch@opt.tu-darmstadt.de).}}
 

### 2016-03-20 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
paper: minor modifications w.r.t. RIC

@@
 Since the following results depend on the constant $\delta$ in \eqref{rip} being
-small enough, we are generally interested in the smallest constant satisfying
+small enough, one is generally interested in the smallest constant satisfying
 \eqref{rip}.
@@
 \begin{theorem}[Cand\'es~\cite{Can08}]\label{RIPsqrt2}
-  Let $A$ satisfy $\delta_{2k} < \sqrt{2}-1 \approx 0.4142$. If there
+  Let the RIC of order $2k$ of $A$ satisfy $\delta_{2k} < \sqrt{2}-1 \approx 0.4142$. If there
   exists $\tilde{x} \in \R^n$ with $\Norm{\tilde{x}}{0} \leq k$ and

### 2016-03-20 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
paper: several minor changes

@@
 Definition~\ref{asymRIPdef} is a generalization of Definition~\ref{ripdef}
-in the following sense: if the matrix~$A$ satisfies the RIP of order $k$
+in the following sense: if the matrix~$A$ satisfies the RIP of order~$k$
 with RIC $\delta_k$, then it has a RIR of at most $(1 + \delta_k) / (1 -
@@
 \begin{theorem}[Foucart and Lai~\cite{FL09}]\label{asymRIPtheorem}
-  Let $A$ satisfy $\gamma_{2k} \leq 4 \sqrt{2} - 3 \approx 2.6569$. If
+  Let the RIR of order $2k$ of $A$ satisfy $\gamma_{2k} \leq 4 \sqrt{2} - 3 \approx 2.6569$. If
   there exists $\tilde{x} \in \R^n$ with $\Norm{\tilde{x}}{0} \leq k$ such
@@
 of Definition~\ref{asymRIPdef} and Theorem~\ref{asymRIPtheorem} is that they
-are invariant under scaling of $Ax = b$. The condition of
+are invariant under scaling of $Ax = b$. In fact, the condition of
 Theorem~\ref{RIPsqrt2} holding for~$A x = b$ does not imply the same 
@@
 \[
-\max \{ \Norm{A x}{2}^2 \suchthat \Norm{x}{2} \leq 1\}
+\max\; \{ \Norm{A x}{2}^2 \suchthat \Norm{x}{2} \leq 1\}
 \]
@@
 
-In our computational results we show that the resulting problems are quite
+Our computational results show that the resulting problems are quite
 hard to solve, even for matrices of smaller size. However, this is true for

### 2016-03-20 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
paper: make "contribution of this work" a section
paper: minor changes

@@
 
-\subsection{Contribution of this Work}
+\section{Contribution of this Work}
 
@@
 The objective and the normalization-constraint can then be expressed as  $\Norm{Ax}{2}^2 = x\T A\T Ax =
-\Tr(A\T A X)$ and $1 = \Norm{x}{2}^2 = \sum_{i=1}^n x_i^2 = \Tr(X)$.
+\Tr(A\T A X)$ and $\Norm{x}{2}^2 = \sum_{i=1}^n x_i^2 = \Tr(X) = 1$.
 
-In the next step, the inequality $\Norm{x}{0} \leq k$ is then substituted by the weaker constraint $\ones\T
+In the next step, the inequality $\Norm{x}{0} \leq k$ is substituted by the weaker constraint $\ones\T
 \abs{X} \ones \leq k$, which follows from the fact that $\ones\T\abs{X} \ones = \Norm{\vect(X)}{1}\leq \sqrt{\Norm{\vect(X)}{0}}\, \Norm{\vect(X)}{2}$ and $\Norm{\vect(X)}{0}=\Norm{x}{0}^2$ in the rank-1 case, where $\vect(X)$ is a vector
@@
 
-Another SDP relaxation was proposed by d'Aspremont et al. in \cite{Asp08}, where
+Another SDP relaxation was proposed by d'Aspremont et al. in \cite{Asp08}. Here,
 \begin{equation}\label{phi}
@@
 
-As one way to solve \eqref{phi} the semidefinite lifting technique is used again to compute an upper bound on $\phi$ via
+As one way to solve \eqref{phi}, the semidefinite lifting technique is used again to compute an upper bound on $\phi$ via
 \begin{equation}\label{Asp08Primal}\tag{A2-Primal}
@@
 be solved for $\Sigma = M I - A\T A$, where $M \in \R$ has to be big enough
-to make $\Sigma$ positive semidefinite again. If $\Psi$ is the
-corresponding optimal objective, this implies a lower bound on $\alpha_k$
-in
+to make $\Sigma$ positive semidefinite. If $\Psi$ is the
+corresponding optimal objective, this implies the following lower bound:
 \begin{equation*}

### 2016-03-20 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
paper: \psi(\rho) in lower bound on \alpha_k
paper: MISDP -> MISDP formulation in section title
paper: reformulated reference to QP

@@
 The idea is that the optimal constant $\alpha_k^2$ in \eqref{lowerRIP} can be computed
-by the non-convex quadratic optimization problem
+by the non-convex quadratic optimization problem with a cardinality constraint
 \begin{equation}\label{QP}\tag{QP}
@@
 be solved for $\Sigma = M I - A\T A$, where $M \in \R$ has to be big enough
-to make $\Sigma$ positive semidefinite. If $\Psi$ is the
-corresponding optimal objective, this implies the following lower bound:
+to make $\Sigma$ positive semidefinite. This implies the following lower bound:
 \begin{equation*}
-  \alpha_k^2 \geq \sup_{\rho \geq 0}\; (M - \Psi) - \rho\, k.
+  \alpha_k^2 \geq \sup_{\rho \geq 0}\; (M - \psi(\rho)) - \rho\, k.
 \end{equation*}
@@
 
-\section{An MISDP for the RIP}
+\section{An MISDP Formulation for the RIP}
 \label{sec:MISDP}
@@
 constants $\alpha_k$ and $\beta_k$ in \eqref{lowerRIP} and
-\eqref{upperRIP}, respectively. We will again start with the non-convex quadratic problem~\eqref{QP}.
+\eqref{upperRIP}, respectively. We will again start with the problem~\eqref{QP}.
 For the non-convex quadratic equality constraint $\Norm{x}{2}^2 = 1$, we

### 2016-03-20 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
paper: changed notation in Pataki's result

@@
 relationship between the rank of solutions of semidefinite programs and the
-dimension of corresponding faces of the feasible set~$P$, i.e., subsets $F
-\subseteq P$ such that $x \in F$, $y, z \in P$, and $x = \tfrac{1}{2} (y + z)$ imply
-$y,z \in F$:
+dimension of corresponding faces of the feasible set~$S$, i.e., subsets $F
+\subseteq S$ such that $\tfrac{1}{2} (p + q) \in F$ for $p, q \in S$ implies
+$p,q \in F$:
 
@@
   \begin{equation*}
-    P \define \{X \in S_n \suchthat X \succeq 0,\; \Tr(A_i X) = b_i\; \text{ for } i = 1, \dots, \tilde{m}\}
+    S \define \{X \in S_n \suchthat X \succeq 0,\; \Tr(A_i X) = b_i\; \text{ for } i = 1, \dots, \tilde{m}\}
   \end{equation*}

### 2016-03-20 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
paper: several minor style changes

@@
 
-We use a testset consisting of 63 matrices of seven 
-different types including band matrices, 
+We use a testset consisting of 63 matrices of the following seven
+different types: band matrices,
 with band size three or five with entries within the band chosen uniformly in $\{0,1\}$, binary matrices with all entries chosen 
-uniformly in $\{0,1\}$, Gaussian-distributed matrices and rank one 
+uniformly in $\{0,1\}$, Gaussian-distributed matrices, and rank one 
 matrices $A = aa\T$ with $N(0,1)$-distributed entries of $a$. In addition to these four types,
@@
 constants. These include matrices with 
-$N(0, 1/m)$-distributed entries, with entries chosen uniformly in $\pm 1/\sqrt{m}$ and finally with distribution
+$N(0, 1/m)$-distributed entries, with entries chosen uniformly in $\pm 1/\sqrt{m}$, and finally with distribution
 \begin{equation}\nonumber
@@
 For each type we used \mbox{MATLAB 8.3.0} to randomly generate nine
-matrices of three different sizes with rows ranging from 15 to 30, columns
-from 25 to 40 and order $k$ between three and five, see Table~\ref{matrixsizes}.
+matrices of three different sizes with the number of rows ranging from 15
+to 30, number of columns
+from 25 to 40, and order $k$ between three and five, see Table~\ref{matrixsizes}.
 
@@
 For the latter three kinds of matrices, it was shown by Baraniuk et al.~\cite{BDDW08} that for $n \rightarrow \infty$, 
-sufficiently small $k$, and given $\delta$, they satisfy the restricted isometry property of order $k$  
+sufficiently small $k$, and given $\delta$, they satisfy the restricted
+isometry property of order~$k$ and constant~$\delta$
 with probability converging exponentially to one. The needed proportion between $n$ and $k$ is so small, however,
@@
 a time limit of four hours. The code was compiled with gcc 4.4.5 with \texttt{-O3} optimization. For 39 matrices 
-the evaluation of~\eqref{Asp08Dual} was aborted for all 15 choices of $\rho$ because of the memory limit of 8GB. 
+the evaluation of~\eqref{Asp08Dual} was aborted for all 15 choices of $\rho$, because of the memory limit of 8GB. 
 These are counted with the maximum allowed time of four hours.
 
-In Table~\ref{lhsRhsGap} we compare the relative gap between the bounds on $\alpha_k$ and $\beta_k$ produced by the two heuristics
+In Table~\ref{lhsRhsGap}, we compare the relative gap between the bounds on $\alpha_k$ and $\beta_k$ produced by the two heuristics
 and the exact values given by~\eqref{MISDP}, computed as their difference divided by the exact value. We present arithmetic means over all of the nine matrices for the given type that
@@
 since we cannot compute a relative gap in that case. This included all rank one matrices. The number of instances used to compute the
-average is given in brackets.
+average is given in parentheses.
 
-For the left-hand-side of the RIP, $\alpha_k$, both relaxations did not perform
-too well. Especially~\eqref{Asp08} regularly returned worse solutions than the
+For the left-hand-side~$\alpha_k$ of the RIP, the quality of both
+relaxations is bad: Especially~\eqref{Asp08} regularly returned worse solutions than the
 trivial lower bound of zero. Relaxation~\eqref{Asp07} performed slightly better, but the bounds
@@
 
-For the right-hand side, i.e., $\beta_k$,  both relaxations
+In fact, for the right-hand side~$\beta_k$,  both relaxations
 perform much better. Problem~\eqref{Asp07} constantly produces bounds within 10 to 30\,\% of the optimal
@@
 The solving times in seconds for the left- and right-hand side of the RIP are given in
-Table~\ref{lhsRhsTime}, again exlucding the left-hand sides for the rank one matrices
+Table~\ref{lhsRhsTime}, again exlucding the left-hand sides for the rank one matrices,
 but including all others. The relaxation~\eqref{Asp07}
@@
 relaxations~\eqref{Asp07} and~\eqref{Asp08}
-proposed in~\cite{Asp07} and \cite{Asp08}, respectively. In conclusion, the relaxation~
-\eqref{Asp07} is fast to compute, but produces a gap of around 90\,\% for
+proposed in~\cite{Asp07} and \cite{Asp08}, respectively. In conclusion, the
+relaxation~\eqref{Asp07}
+is fast to compute, but produces a gap of around 90\,\% for
 $\alpha_k$ and 30\,\% for $\beta_k$. If successfully solved, 

### 2016-03-21 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
explicitly write psi(sigma,rho)

@@
 The corresponding maximization problem allows to compute $\beta_k^2$. D'Aspremont et al. \cite{Asp07} tackled the non-convex quadratic equality constraint 
-$\Norm{x}{2}^2 = 1$ by the technique of semidefinite lifting, first introduced by Goemans and Williamson~\cite{GW95} 
-for the max-cut problem.
+$\Norm{x}{2}^2 = 1$ by the technique of semidefinite lifting, see, e.g., Goemans and Williamson~\cite{GW95}.
 The idea is to use new matrix variables $X=xx\T$ for the quadratic terms, where the condition $X=xx\T$ is enforced by the equivalent constraints $X \succeq 0$ and $\Rk(X) = 1$, where, as usual, 
@@
 \begin{equation}\label{phi}
-  \phi(\rho) = \max \big\{ x\T \Sigma\, x - \rho \Norm{x}{0} \suchthat \Norm{x}{2} \leq 1 \big\},
+  \phi(\Sigma, \rho) = \max \big\{ x\T \Sigma\, x - \rho \Norm{x}{0} \suchthat \Norm{x}{2} \leq 1 \big\},
 \end{equation}
@@
             (\Norm{x}{0} - k) \suchthat \Norm{x}{2} \leq 1 \big\} \\
-	    & = \inf_{\rho \geq 0} \phi(\rho) + \rho\, k. 
+	    & = \inf_{\rho \geq 0} \phi(A\T A, \rho) + \rho\, k. 
  \end{align*}
@@
 \begin{equation}\tag{A2}\label{Asp08}
- \alpha_k^2 \geq \sup_{\rho \geq 0}\; \psi(\rho) - \rho\, k, 
+ \alpha_k^2 \geq \sup_{\rho \geq 0}\; \psi(A\T A, \rho) - \rho\, k, 
 \end{equation}
@@
 \begin{equation*}
-  \psi(\rho) = \min \big\{ x\T \Sigma\, x + \rho \Norm{x}{0} \suchthat \Norm{x}{2} \leq 1\big\}.
+  \psi(\Sigma, \rho) = \min \big\{ x\T \Sigma\, x + \rho \Norm{x}{0} \suchthat \Norm{x}{2} \leq 1\big\}.
 \end{equation*}
@@
 \begin{equation*}
-  \alpha_k^2 \geq \sup_{\rho \geq 0}\; (M - \psi(\rho)) - \rho\, k.
+  \alpha_k^2 \geq \sup_{\rho \geq 0}\; (M - \psi(M I - A\T A, \rho)) - \rho\, k.
 \end{equation*}

### 2016-03-21 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
added explanation about solved instances for MISDP and quality of exact RICS in appendix

@@
 the evaluation of~\eqref{Asp08Dual} was aborted for all 15 choices of $\rho$, because of the memory limit of 8GB.
+\color{red}We also failed to solve \eqref{MISDP} for the left-hand side of the three medium sized rank one matrices with default
+parameters because SDPA failed to solve the root node relaxation, however with different settings or a different
+SDP-solver we could verify that $\alpha_k$ equaled zero up to the numerical tolerance for these instances.\color{black}
 These are counted with the maximum allowed time of four hours.
@@
 result for any matrix type among both relaxations.
+
+\color{red}The restricted isometry constants of all matrices can be found in the appendix. Note that for the unscaled matrices
+the restricted isometry constants can be expected to be much larger than for the scaled ones. Furthermore, since we solved
+the instances with a feasibility tolerance of $10^{-6}$, which influences the optimal objective by a factor of the
+largest objective coefficients and the number of variables, the computed constants are only exact up to a factor
+of about $10^{-2}$ for these matrices, while the precision is much better for the normalized ones because of their
+smaller objective coefficients.\color{black}
 

### 2016-03-21 RIP-MISDP-Paper/ResultScript.py
added table of all matrices in the testset to the appendix
spacing and correct style of caption are currently hacked

@@ -9,12 +9,13 @@ from decimal import Decimal
 texfile = 0 # make the output a compilable texfile or just a figure/table that can be included?
 Lbounds = 0
 Rbounds = 0
-LRbounds = 1
+LRbounds = 0
 Ltimes = 0
 Rtimes = 0
 LRtimes = 0
 printFails = 0
 completeTable = 0
+CompleteRICtable = 1
 
 MISDPfilename = "/local/gally/results/RIP-MISDP-Paper/160311/RIP-results/check.RIPMISDP.scipsdp.linux.x86_64.gnu.opt.sdpa.extra.branchinfobj_nofracdive.out"
 Asp0708path = "/local/gally/results/RIP-MISDP-Paper/160311/SDPA-results/"
@@ -590,6 +591,55 @@ def LhsRhsTimeTable(instancesets, instancesetnames, caption, label):
 	file.write("& \\num{%.1f" % totalMISDPtimer + "} & " + "\\num{%.1f" % totalA07timer + "} & " + "\\num{%.1f" % totalA08timer + "} \\\ \n ")
 	file.write("\\bottomrule \n \\end{tabular*} \n \end{scriptsize} \n \\end{table} \n")
 
+def RICtable(instances, caption, label, tabularx):
+	if tabularx:
+		file.write("\\newpage \n \\begin{scriptsize} \n  \\setlength{\\tabcolsep}{2pt} \n \\tablehead{\\toprule \n")
+		file.write("matrix & $m$ & $n$ & $k$ & $\\alpha_k^2$ & $\\beta_k^2$ & $\\gamma_k^2$ & $\\delta_k^2$")
+		file.write("\\\ \\midrule} \n \\tabletail{ \\midrule \\multicolumn{3}{@{}l}{continued on next page \\dots}\\\ \\bottomrule} \n \\tablelasttail{\\bottomrule} \n \\tablecaption[")
+		file.write(caption)
+		file.write("]{")
+		file.write(caption)
+		file.write("}\label{")
+		file.write(label)
+		file.write("}\n")
+		file.write("\\begin{xtabular*}{\\linewidth}{@{\extracolsep{\\fill}}lrrrrrrr@{}} \n ")
+	else:
[…]

### 2016-03-22 RIP-MISDP-Paper/ResultScript.py
-- instad of - in matrix list

@@ -630,7 +630,7 @@ def RICtable(instances, caption, label, tabularx):
 			delta = beta-1
 		file.write(names[0][j1].split("_")[0] + " & \\num{%.0f" % float(m[j1/2]) + "} & \\num{%.0f" % float(n[j1/2]) + "} & \\num{%.0f" % float(k[j1/2]) + "} & \\num{%.2f" % float(alpha) + "} & \\num{%.2f" % float(beta) + "} & ")
 		if gamma == "-":
-			file.write("- & ")
+			file.write("-- & ")
 		else:
 			file.write("\\num{%.2f" % float(gamma) + "} & ")
 		file.write("\\num{%.2f" % float(delta) + "} \\\ \n")

### 2016-03-22 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
added table of all matrices in the testset to the appendix
spacing and correct style of caption are currently hacked
added some commas after fors

@@
 \usepackage{sistyle}
+\usepackage{tabularx}
+\usepackage{booktabs}
+\usepackage{xtab}
 
@@
 \SIthousandsep{,}
+
+% prevent tables from being broken too early
+\xentrystretch{-0.2}
 
@@
 assess the quality of the relaxations proposed in~\cite{Asp08} and~\cite{Asp07}.
-For solving the mixed integer semidefinite programs we use the MISDP solver SCIP-SDP~\cite{SCIP-SDP},
+For solving the mixed integer semidefinite programs, we use the MISDP solver SCIP-SDP~\cite{SCIP-SDP},
 originally developed in~\cite{Mar13} and~\cite{MS12}. It combines the
@@
 Linux. Each computation was performed single-threaded with a single process running on each computer and with 
-a time limit of four hours. The code was compiled with gcc 4.4.5 with \texttt{-O3} optimization. For 39 matrices 
+a time limit of four hours. The code was compiled with gcc 4.4.5 with \texttt{-O3} optimization. For 39 matrices, 
 the evaluation of~\eqref{Asp08Dual} was aborted for all 15 choices of $\rho$, because of the memory limit of 8GB.
@@
 parameters because SDPA failed to solve the root node relaxation, however with different settings or a different
-SDP-solver we could verify that $\alpha_k$ equaled zero up to the numerical tolerance for these instances.\color{black}
+SDP-solver we could verify that $\alpha_k$ equaled zero up to the numerical tolerance for these instances. \color{black}
 These are counted with the maximum allowed time of four hours.
@@
 
-\color{red}The restricted isometry constants of all matrices can be found in the appendix. Note that for the unscaled matrices
+\color{red}The restricted isometry constants of all matrices can be found in the appendix. Note that for the unscaled matrices,
 the restricted isometry constants can be expected to be much larger than for the scaled ones. Furthermore, since we solved
@@
 \bibliography{Paperbib}
+
+\appendix{ \input{tables/matrixList}}
+
 \end{document}

### 2016-03-22 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
explain which matrices are scaled
added sentence about RIP not even holding for scaled matrices

@@
 matrices $A = aa\T$ with $N(0,1)$-distributed entries of $a$. In addition to these four types,
-we also use three more with theoretical results on their restricted isometry
+we also use three types of scaled matrices with theoretical results on their restricted isometry
 constants. These include matrices with 
@@
 
-\color{red}The restricted isometry constants of all matrices can be found in the appendix. Note that for the unscaled matrices,
-the restricted isometry constants can be expected to be much larger than for the scaled ones. Furthermore, since we solved
+The restricted isometry constants of all matrices can be found in the appendix. Note that for the unscaled matrices,
+the restricted isometry constants can be expected to be much larger than for the scaled ones. Even for the scaled
+matrices, the restricted isometry property did not hold for any of the matrices in the testset, since the instances
+were too small to satisfy the needed relation between $n$ and $k$.
+\color{red} Furthermore, since we solved
 the instances with a feasibility tolerance of $10^{-6}$, which influences the optimal objective by a factor of the
 largest objective coefficients and the number of variables, the computed constants are only exact up to a factor
-of about $10^{-2}$ for these matrices, while the precision is much better for the normalized ones because of their
+of about $10^{-2}$ for the unscaled matrices, while the precision is much better for the normalized ones because of their
 smaller objective coefficients.\color{black}

### 2016-03-22 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
adjusted length of xtab-table
spellcheck

@@
 % prevent tables from being broken too early
-\xentrystretch{-0.2}
+\xentrystretch{-0.22}
 
@@
 The solving times in seconds for the left- and right-hand side of the RIP are given in
-Table~\ref{lhsRhsTime}, again exlucding the left-hand sides for the rank one matrices,
+Table~\ref{lhsRhsTime}, again excluding the left-hand sides for the rank one matrices,
 but including all others. The relaxation~\eqref{Asp07}

### 2016-03-22 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
several minor things, handled marked paragraphs, added framework paper

@@
 assess the quality of the relaxations proposed in~\cite{Asp08} and~\cite{Asp07}.
-For solving the mixed integer semidefinite programs, we use the MISDP solver SCIP-SDP~\cite{SCIP-SDP},
+For solving the mixed integer semidefinite programs, we use the MISDP solver SCIP-SDP~\cite{SCIP-SDP,GalPU16},
 originally developed in~\cite{Mar13} and~\cite{MS12}. It combines the
@@
 matrices $A = aa\T$ with $N(0,1)$-distributed entries of $a$. In addition to these four types,
-we also use three types of scaled matrices with theoretical results on their restricted isometry
-constants. These include matrices with 
+we also use three types of matrices with scaled parameters. These include matrices with 
 $N(0, 1/m)$-distributed entries, with entries chosen uniformly in $\pm 1/\sqrt{m}$, and finally with distribution
@@
 the evaluation of~\eqref{Asp08Dual} was aborted for all 15 choices of $\rho$, because of the memory limit of 8GB.
-\color{red}We also failed to solve \eqref{MISDP} for the left-hand side of the three medium sized rank one matrices with default
-parameters because SDPA failed to solve the root node relaxation, however with different settings or a different
-SDP-solver we could verify that $\alpha_k$ equaled zero up to the numerical tolerance for these instances. \color{black}
+We also failed to solve \eqref{MISDP} for the left-hand side of the three medium sized rank one matrices with default
+parameters, because SDPA failed to solve the root node relaxation. With different settings or a different
+SDP-solver, however, we could verify that $\alpha_k$ is zero for these instances.
 These are counted with the maximum allowed time of four hours.
@@
 
-The restricted isometry constants of all matrices can be found in the appendix. Note that for the unscaled matrices,
-the restricted isometry constants can be expected to be much larger than for the scaled ones. Even for the scaled
-matrices, the restricted isometry property did not hold for any of the matrices in the testset, since the instances
-were too small to satisfy the needed relation between $n$ and $k$.
-\color{red} Furthermore, since we solved
-the instances with a feasibility tolerance of $10^{-6}$, which influences the optimal objective by a factor of the
-largest objective coefficients and the number of variables, the computed constants are only exact up to a factor
-of about $10^{-2}$ for the unscaled matrices, while the precision is much better for the normalized ones because of their
-smaller objective coefficients.\color{black}
+The restricted isometry constants of all matrices can be found in the
+appendix. Note that for the unscaled matrices (the first 36 instances until
+instance \texttt{rnk140403C}),
+we expect the restricted isometry constants to be much larger than for the scaled ones. Even for the scaled
+matrices, the restricted isometry property never holds for the matrices in the testset, since the instances
+are too small to satisfy the needed relation between $n$ and $k$.
 
@@
 to be used in practice. Moreover, directly solving~\eqref{MISDP} is often
-faster and produces the exact value. Consequently, 
-\eqref{Asp08} should not be used in practice.
+faster and produces the exact value!
 
@@
 
+\newpage
 \appendix{ \input{tables/matrixList}}

### 2016-03-22 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
branch and bound -> branch-and-bound and updated table

@@
 
-In order to successfully apply branch and bound to \eqref{MISDP}, it is
+In order to successfully apply branch-and-bound to \eqref{MISDP}, it is
 important to use strong semidefinite relaxations for fractional $z$
@@
 originally developed in~\cite{Mar13} and~\cite{MS12}. It combines the
-branch and bound framework SCIP 3.2.1 \cite{SCIP} with 
+branch-and-bound framework SCIP 3.2.1 \cite{SCIP} with 
 interior-point SDP-solvers, in our case SDPA~7.3.8~\cite{SDPA6,SDPA7}.
@@
 \newpage
-\appendix{ \input{tables/matrixList}}
+\appendix{\input{tables/matrixList}}
 

### 2016-03-23 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
paper: changed title and added - to mixed-integer

@@
 \usepackage[breaklinks,colorlinks,citecolor=blue,linkcolor=blue,urlcolor=blue,bookmarks=false,hypertexnames=false,
-  pdftitle={Computing Exact Restricted Isometry Constants via Mixed Integer Semidefinite Programming},
+  pdftitle={Computing Restricted Isometry Constants via Mixed-Integer Semidefinite Programming},
   % pdftitle={Computing Restricted Isometry Constants via Mixed Integer Semidefinite Programming},
@@
 
-\title{Computing Exact Restricted Isometry Constants via Mixed Integer Semidefinite Programming}
-%\title{Computing Restricted Isometry Constants via Mixed Integer Semidefinite Programming}
+\title{Computing Restricted Isometry Constants via Mixed-Integer Semidefinite Programming}
 \author{Tristan~Gally and Marc~E.~Pfetsch
@@
   this condition is itself \NP-hard. In this paper we propose a
-  mixed integer semidefinite programming approach for computing these
+  mixed-integer semidefinite programming approach for computing these
   optimal constants. This also subsumes sparse principal component
@@
 The goal of this paper is to compute optimal restricted isometry constants
-via mixed integer semidefinite programming (MISDP). To this end, we will
+via mixed-integer semidefinite programming (MISDP). To this end, we will
 introduce an MISDP with a rank-1 constraint. In order to utilize this
@@
 and~\cite{Asp07} in Section~\ref{sec:SDPrelax}.  Afterwards, in
-Section~\ref{sec:MISDP}, we construct a mixed integer semidefinite
+Section~\ref{sec:MISDP}, we construct a mixed-integer semidefinite
 program and show that it produces exact restricted isometry
@@
 \noindent
-In this section, we will introduce two known SDP relaxations to compute an upper bound on the restricted isometry constant $\delta_k$ in~\eqref{rip}, since they have some ideas in common with the mixed integer semidefinite program we want to
+In this section, we will introduce two known SDP relaxations to compute an upper bound on the restricted isometry constant $\delta_k$ in~\eqref{rip}, since they have some ideas in common with the mixed-integer semidefinite program we want to
 propose and will also be checked against in the section on numerical
@@
 restricted isometry constant and the restricted isometry ratio by
-mixed integer semidefinite programming.
+mixed-integer semidefinite programming.
 
@@
 assess the quality of the relaxations proposed in~\cite{Asp08} and~\cite{Asp07}.
-For solving the mixed integer semidefinite programs, we use the MISDP solver SCIP-SDP~\cite{SCIP-SDP,GalPU16},
+For solving the mixed-integer semidefinite programs, we use the MISDP solver SCIP-SDP~\cite{SCIP-SDP,GalPU16},
 originally developed in~\cite{Mar13} and~\cite{MS12}. It combines the
@@
 but including all others. The relaxation~\eqref{Asp07}
-is very fast, taking only a few seconds for all instances. The mixed integer semidefinite program obviously
+is very fast, taking only a few seconds for all instances. The mixed-integer semidefinite program obviously
 cannot compete with these solving times, although for band-matrices, the difference is relatively
@@
 This paper shows that exact values of restricted isometry constants can be
-computed via mixed integer semidefinite programming. The computed bounds
+computed via mixed-integer semidefinite programming. The computed bounds
 significantly improve the bounds computed by the semidefinite
@@
 future, the hope is that with the advancement of solving techniques for
-mixed integer semidefinite programs, the sizes of the matrices can be
+mixed-integer semidefinite programs, the sizes of the matrices can be
 significantly increased. The components to be improved include cutting

### 2016-03-23 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
paper: added biographies

@@
 
+\begin{IEEEbiographynophoto}{Tristan Gally}
+obtained a B.Sc. and an M.Sc. degree in mathematics with economics
+from TU Darmstadt, Germany, in 2011 and 2013, respectively. Since
+02/2014 he is a doctoral candidate at TU Darmstadt working as a
+Researcher in the Collaborative Research Center 805, funded by the
+German Research Foundation (DFG). His research interests are mainly
+in mixed-integer semidefinite programming with applications in
+compressed sensing and truss topology design.
+\end{IEEEbiographynophoto}
+
+\begin{IEEEbiographynophoto}{Marc E. Pfetsch}
+  obtained a Diploma degree in mathematics from the University of
+  Heidelberg, Germany, in 1997. He received a PhD degree in
+  mathematics in 2002 and the Habilitation degree in 2008 from TU
+  Berlin, Germany. From 2008 to 2012 he was a Full Professor for
+  Mathematical Optimization at TU Braunschweig, Germany. Since 04/2012
+  he has been a Full Professor for Discrete Optimization at TU
+  Darmstadt, Germany. His research interests are mostly in discrete
+  optimization, in particular symmetry in integer programs, compressed
+  sensing, and algorithms for mixed integer programs.
+\end{IEEEbiographynophoto}
+
 \newpage

### 2016-03-23 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
corrected adress

@@
   \thanks{The  authors  are  with  the  Research  Group  Optimization,  TU
-    Darmstadt, Dolivostr.\ 5,
+    Darmstadt, Dolivostr.\ 15,
   Darmstadt  64293,  Germany  (e-mail:  gally@opt.tu-darmstadt.de;

### 2016-03-23 RIP-MISDP-Paper/IEEE_RIP-MISDP_peerreview.tex
added single column peer-review version of paper including keywords
corrected adress

--- 2016-03-23 RIP-MISDP-Paper/IEEE_RIP-MISDP.tex
@@
-\documentclass[journal]{IEEEtran}
+\documentclass[peerreviewca,onecolumn]{IEEEtran}
 
@@
 \maketitle
+
+\IEEEpeerreviewmaketitle
 
@@
 \end{abstract}
+
+\begin{IEEEkeywords}
+ Compressed Sensing, Restricted Isometry Property, Mixed-Integer Semidefinite Programming
+\end{IEEEkeywords}
 

### 2016-03-23 RIP-MISDP-Paper/IEEE_RIP-MISDP_peerreview.tex
adapted to submission

@@
-\documentclass[peerreviewca,onecolumn]{IEEEtran}
+\documentclass[journal,onecolumn]{IEEEtran}
 
@@
 
+\begin{abstract}
   \noindent
-\begin{abstract}
   One of the fundamental tasks in compressed sensing is finding the
@@
     (1 - \delta) \Norm{x}{2}^2 \leq \Norm{Ax}{2}^2 \leq (1+\delta) \Norm{x}{2}^2
+    \quad\text{for all }x \in \Sigma_k \define \{x \in \R^n \suchthat \Norm{x}{0} \leq k\}.
   \end{equation}
-  for all $x \in \Sigma_k \define \{x \in \R^n \suchthat \Norm{x}{0} \leq k\}$.
 \end{definition}
@@
     \begin{aligned}
-      \delta_k \define \argmin_{\delta \geq 0} \big\{&(1 - \delta) \Norm{x}{2}^2 \leq \Norm{Ax}{2}^2 \leq (1+\delta) \Norm{x}{2}^2 ,\\
-      & \text{ for all }x \in \Sigma_k \big\},
+      \delta_k \define \argmin_{\delta \geq 0} \big\{(1 - \delta) \Norm{x}{2}^2 \leq \Norm{Ax}{2}^2 \leq (1+\delta) \Norm{x}{2}^2 
+      \quad\text{ for all }x \in \Sigma_k \big\},
     \end{aligned}
@@
 in the following sense: if the matrix~$A$ satisfies the RIP of order~$k$
-with RIC $\delta_k$, then it has a RIR of at most $(1 + \delta_k) / (1 -
+with RIC~$\delta_k$, then it has a RIR of at most $(1 + \delta_k) / (1 -
 \delta_k)$. Conversely, if $A$ has lower/upper RICs $\alpha_k$ and
@@
 \noindent
-In this section, we will introduce two known SDP relaxations to compute an upper bound on the restricted isometry constant $\delta_k$ in~\eqref{rip}, since they have some ideas in common with the mixed-integer semidefinite program we want to
+In this section, we will introduce two known SDP relaxations to compute an upper bound on the restricted isometry constant~$\delta_k$ in~\eqref{rip}, since they have some ideas in common with the mixed-integer semidefinite program we want to
 propose and will also be checked against in the section on numerical
@@
   agree and $\Rk(\hat{X}) = \Rk(\check{X})$.  Therefore, it
-  suffices to show that \eqref{Proj} has an optimal solution of rank one.
+  suffices to show that~\eqref{Proj} has an optimal solution of rank one.
 
@@
 \end{IEEEbiographynophoto}
+\vfill
+\mbox{}
 

### 2016-03-23 RIP-MISDP-Paper/RIP-MISDP_optimization-online.tex
added amsart-version for optimization online

--- 2016-03-23 RIP-MISDP-Paper/IEEE_RIP-MISDP_peerreview.tex
@@
-\documentclass[journal,onecolumn]{IEEEtran}
+\documentclass[11pt,a4paper]{amsart}
 
@@
   pdftitle={Computing Restricted Isometry Constants via Mixed-Integer Semidefinite Programming},
-  % pdftitle={Computing Restricted Isometry Constants via Mixed Integer Semidefinite Programming},
   pdfauthor={T. Gally, and M. E. Pfetsch}]{hyperref}
@@
 \newtheorem{definition}{Definition}
-\newtheorem{remark}{Remark}
-\newtheorem{lemma}[theorem]{Lemma}
+\newtheorem{lemma}{Lemma}
+\newtheorem{proposition}{Proposition}
+\newtheorem{corollary}{Corollary}
 
@@
 % prevent tables from being broken too early
-\xentrystretch{-0.22}
+\xentrystretch{-0.4}
 
@@
 
-\title{Computing Restricted Isometry Constants via Mixed-Integer Semidefinite Programming}
-\author{Tristan~Gally and Marc~E.~Pfetsch
-  \thanks{This work was supported by the German Research Foundation (DFG) through Collaborative Research Center 805 and Priority Program 1798.}
-  \thanks{The  authors  are  with  the  Research  Group  Optimization,  TU
-    Darmstadt, Dolivostr.\ 15,
-  Darmstadt  64293,  Germany  (e-mail:  gally@opt.tu-darmstadt.de;
-  pfetsch@opt.tu-darmstadt.de).}}
+\title[Computing Restricted Isometry Constants via MISDP]{Computing Restricted Isometry Constants via \\ Mixed-Integer Semidefinite Programming}
+\author[T. Gally]{Tristan Gally}
+\author[M. E. Pfetsch]{Marc E. Pfetsch}
 
-\maketitle
-
-\IEEEpeerreviewmaketitle
+\address{Tristan Gally \\ Technische Universit\"at Darmstadt, Department of Mathematics\\ Dolivostra\ss e 15, 64293 Darmstadt, Germany}
+\email{gally@mathematik.tu-darmstadt.de}
+\address{Marc E. Pfetsch \\ Technische Universit\"at Darmstadt, Department of Mathematics\\ Dolivostra\ss e 15, 64293 Darmstadt, Germany}
+\email{pfetsch@mathematik.tu-darmstadt.de}
 
@@
 
-\begin{IEEEkeywords}
+\maketitle
+
+\smallskip
+\noindent \textbf{Keywords.}
 Compressed Sensing, Restricted Isometry Property, Mixed-Integer Semidefinite Programming
-\end{IEEEkeywords}
 
@@
 
-\IEEEPARstart{F}{inding} a sparsest solution of an underdetermined system of linear
+Finding a sparsest solution of an underdetermined system of linear
 equations is one key problem in compressed sensing and has been
@@
     (1 - \delta) \Norm{x}{2}^2 \leq \Norm{Ax}{2}^2 \leq (1+\delta) \Norm{x}{2}^2
-    \quad\text{for all }x \in \Sigma_k \define \{x \in \R^n \suchthat \Norm{x}{0} \leq k\}.
   \end{equation}
+  for all $x \in \Sigma_k \define \{x \in \R^n \suchthat \Norm{x}{0} \leq k\}$.
 \end{definition}
@@
 
-\begin{IEEEbiographynophoto}{Tristan Gally}
-obtained a B.Sc. and an M.Sc. degree in mathematics with economics
-from TU Darmstadt, Germany, in 2011 and 2013, respectively. Since
-02/2014 he is a doctoral candidate at TU Darmstadt working as a
-Researcher in the Collaborative Research Center 805, funded by the
-German Research Foundation (DFG). His research interests are mainly
-in mixed-integer semidefinite programming with applications in
-compressed sensing and truss topology design.
-\end{IEEEbiographynophoto}
-
-\begin{IEEEbiographynophoto}{Marc E. Pfetsch}
-  obtained a Diploma degree in mathematics from the University of
-  Heidelberg, Germany, in 1997. He received a PhD degree in
-  mathematics in 2002 and the Habilitation degree in 2008 from TU
-  Berlin, Germany. From 2008 to 2012 he was a Full Professor for
-  Mathematical Optimization at TU Braunschweig, Germany. Since 04/2012
-  he has been a Full Professor for Discrete Optimization at TU
-  Darmstadt, Germany. His research interests are mostly in discrete
-  optimization, in particular symmetry in integer programs, compressed
-  sensing, and algorithms for mixed integer programs.
-\end{IEEEbiographynophoto}
-\vfill
-\mbox{}
-
-\newpage
-\appendix{\input{tables/matrixList}}
+\appendix
+\section*{Appendix}
+\input{tables/matrixList_optimization-online}
 

### 2016-03-23 RIP-MISDP-Paper/RIP-MISDP_optimization-online.tex
slight reformulations for acceptable line breaks

@@
 \end{equation}
-The corresponding maximization problem allows to compute $\beta_k^2$. D'Aspremont et al. \cite{Asp07} tackled the non-convex quadratic equality constraint 
+The corresponding maximization problem allows to compute the constant $\beta_k^2$ in~\eqref{upperRIP}. D'Aspremont et al. \cite{Asp07} tackled the non-convex quadratic equality constraint 
 $\Norm{x}{2}^2 = 1$ by the technique of semidefinite lifting, see, e.g., Goemans and Williamson~\cite{GW95}.
@@
   First note that because of the constraint $\Tr(X) = 1$,
-  Problem~\eqref{MISDP} is bounded. Moreover, $X_{11} = 1$, $X_{ij} = 0$
+  \eqref{MISDP} is bounded. Moreover, $X_{11} = 1$, $X_{ij} = 0$
   for all $i,j \neq 1$, $z_1 = 1$, $z_i = 0$ for all $i = 2, \dots, n$ is
@@
   $\ell \define \sum_{i=1}^n z^\star_i \leq k$. Let $T \define \{i_1, i_2,
-  \dots, i_\ell\}$ be the support of $z^\star$, and define $\check{A} =
-  (A_{i_1}, A_{i_2}, \dots, A_{i_\ell})$ to be the submatrix of~$A$ formed by
+  \dots, i_\ell\}$ be the support of $z^\star$, and $\check{A} =
+  (A_{i_1}, A_{i_2}, \dots, A_{i_\ell})$ the submatrix of~$A$ formed by
   the columns indexed by~$T$. Then consider
@@
   Thus, the optimal objective values of \eqref{Proj} and \eqref{MISDP}
-  agree and $\Rk(\hat{X}) = \Rk(\check{X})$.  Therefore, it
+  agree, and, furthermore, $\Rk(\hat{X}) = \Rk(\check{X})$.  Therefore, it
   suffices to show that~\eqref{Proj} has an optimal solution of rank one.
@@
 
-The continuous SDPs are solved using the command-line version of
-SDPA~7.3.8. We compute the left- and right-hand restricted isometry
+For solving the continuous SDPs, the command-line version of
+SDPA~7.3.8 is used. We compute the left- and right-hand restricted isometry
 constant $\alpha_k$ and $\beta_k$ either exactly using~\eqref{MISDP} or

### 2016-03-23 RIP-MISDP-Paper/RIP-MISDP_optimization-online.tex
paper: several minor optical corrections

@@
 
+\usepackage[T1]{fontenc}
 \usepackage{amsmath}
@@
 \usepackage{xtab}
+\usepackage[labelfont={footnotesize,bf},textfont=footnotesize]{caption}
 
@@
 
+\setlength{\abovecaptionskip}{1ex plus0.5ex minus0.5ex}
+\setlength{\belowcaptionskip}{1ex plus0.5ex minus0.5ex}
+
 \begin{document}
@@
 
+\noindent
 Finding a sparsest solution of an underdetermined system of linear
@@
 available both in theory and practice: as an extremely selected list
-consider the articles~\cite{BecBC11,CheDS99,OsbPT00,vdBF08} and the
+consider the articles \cite{BecBC11,CheDS99,OsbPT00,vdBF08} and the
 practical comparison in~\cite{LorPT15}.
@@
   \eqref{rip},
-  \begin{equation}\label{ric}
-    \begin{aligned}
+  \[
   \delta_k \define \argmin_{\delta \geq 0} \big\{(1 - \delta) \Norm{x}{2}^2 \leq \Norm{Ax}{2}^2 \leq (1+\delta) \Norm{x}{2}^2 
   \quad\text{ for all }x \in \Sigma_k \big\},
-    \end{aligned}
-  \end{equation}
+  \]
   is called \emph{restricted isometry constant} (RIC) of order~$k$.
@@
 \begin{theorem}[Cand\'es~\cite{Can08}]\label{RIPsqrt2}
-  Let the RIC of order $2k$ of $A$ satisfy $\delta_{2k} < \sqrt{2}-1 \approx 0.4142$. If there
+  Let the RIC of order $2k$ of $A$ satisfy the condition $\delta_{2k} < \sqrt{2}-1 \approx 0.4142$. If there
   exists $\tilde{x} \in \R^n$ with $\Norm{\tilde{x}}{0} \leq k$ and
@@
 are invariant under scaling of $Ax = b$. In fact, the condition of
-Theorem~\ref{RIPsqrt2} holding for~$A x = b$ does not imply the same 
+Theorem~\ref{RIPsqrt2} holding for $A x = b$ does not imply the same 
 for the scaled equation
-$\lambda Ax = \lambda b$, $\lambda \in \R \setminus \{0\}$. In Definition~\ref{asymRIPdef}, however, we can
+$\lambda Ax = \lambda b$ with $\lambda \in \R \setminus \{0\}$. In Definition~\ref{asymRIPdef}, however, we can
 use $\tilde{\alpha}_k = \lambda \alpha_k$, $\tilde{\beta}_k = \lambda
@@
 which has only half as many variables as \eqref{Asp08Primal} and one fewer linear and semidefinite constraint.
-Then \eqref{Asp08Dual} can be solved for non-negative $\rho$, which should be smaller than $\Sigma_{11}$, because otherwise the
+Then \eqref{Asp08Dual} can be solved for non-negative~$\rho$, which should be smaller than $\Sigma_{11}$, because otherwise the
 optimal solution for \eqref{phi} will always be $x=0$.
@@
   Let $\check{X}$ be the $\ell \times \ell$ submatrix of $X^\star$ with all
-  rows and columns outside of $T$ removed. Then $\check{X}$ is feasible for
+  rows and columns outside of~$T$ removed. Then $\check{X}$ is feasible for
   \eqref{Proj} with the same objective value, since we only removed zero
@@
 
-In order to successfully apply branch-and-bound to \eqref{MISDP}, it is
+For efficient application of branch-and-bound to \eqref{MISDP}, it is
 important to use strong semidefinite relaxations for fractional $z$
@@
 \section*{Appendix}
+
+\vspace{-3ex}
 \input{tables/matrixList_optimization-online}

### 2016-08-29 RIP-MISDP-Paper/IEEE_RIP-MISDP_peerreview.tex
PAPER: added reference to Bandeira et al
PAPER: explain that MISDP is still non-convex

@@
 However, it turns out that computing the restricted isometry
-constants for a given matrix~$A$ and~$k$ is NP-hard, see~\cite{PT14}. 
+constants for a given matrix~$A$ and~$k$ is NP-hard, see~\cite{BDDS13,PT14}. 
 Note the unfortunate implication that it is hard to check for a
@@
 \end{equation}
-We then relax the non-convex rank constraint to arrive at the following
+We then relax the non-convex rank constraint to arrive at the following, still non-convex,
 MISDP formulation:

### 2016-08-29 RIP-MISDP-Paper/IEEE_RIP-MISDP_peerreview.tex
PAPER: Changed formulation to no longer impy that d'Aspremont was motivated by newer paper

@@
 
-This motivated d'Aspremont et al.~\cite{Asp08,Asp07} to propose several
+D'Aspremont et al.~\cite{Asp08,Asp07} proposed several
 relaxations based on semidefinite programming (SDP) to compute upper bounds on the restricted isometry

### 2016-08-29 RIP-MISDP-Paper/IEEE_RIP-MISDP_peerreview.tex
use bar instead of check in model to distinguish variable and given solution
PAPER: emphasize that the extreme point should attain the optimal objective in the rank-1-proof

@@
   $\ell \define \sum_{i=1}^n z^\star_i \leq k$. Let $T \define \{i_1, i_2,
-  \dots, i_\ell\}$ be the support of $z^\star$, and define $\check{A} =
+  \dots, i_\ell\}$ be the support of $z^\star$, and define $\bar{A} =
   (A_{i_1}, A_{i_2}, \dots, A_{i_\ell})$ to be the submatrix of~$A$ formed by
@@
    \begin{aligned}
-    \min \quad & \Tr(\check{A}\T \check{A} \check{X}), \\
-    \text{s.t.} \quad & \Tr(\check{X}) = 1, \\
-    & \check{X} \succeq 0,\; \check{X} \in S_\ell.
+    \min \quad & \Tr(\bar{A}\T \bar{A} \bar{X}), \\
+    \text{s.t.} \quad & \Tr(\bar{X}) = 1, \\
+    & \bar{X} \succeq 0,\; \bar{X} \in S_\ell.
    \end{aligned}
@@
 
-  Let $\check{X}$ be an extreme point of \eqref{Proj} (since the set of
+  Let $\check{X}$ be an extreme point of \eqref{Proj} that attains the
+  optimal objective value (since the set of
   optimal points of~\eqref{Proj} is nonempty, convex, and compact, such a

### 2017-03-06 RIP-MISDP-Paper/IEEE_RIP-MISDP_peerreview.tex
PAPER: no longer use the term "diagonal dominance"
PAPER: number I, II, III instead of A,B,C in table 1 since A,B,C are used differently in table 4
merged master

@@
 Our computational results show that the resulting problems are quite
-hard to solve, even for matrices of smaller size. However, this is true for
-the SDP relaxations of d'Aspremont et al.\ as well, and it allows for an
-evaluation of the quality of these bounds.  In fact, these bounds turn out
-to be quite tight or weak, depending on the instance. This complements the
+hard to solve, even for matrices of smaller size. However, they are
+still solvable for larger instances than a brute-force approach computing
+the eigenvalues of all submatrices of corresponding size.
+
+Even if the techniques for solving MISDPs are not yet capable of solving
+large enough instances to be used for computing RICs in practical applications,
+by computing exact restricted isometry constants, our approach allows to evaluate
+relaxations like those proposed by d'Aspremont et al. in \cite{Asp08,Asp07}.
+In fact, these bounds turn out
+to be quite tight or weak, depending on the instance, while in parts requiring
+solution times comparable to those of the MISDP-approach. This complements the
 theoretical approximation guarantees developed by d'Aspremont et
@@
   \check{X}_{jj} = \hat{X}_{i_j i_j} \leq 1$ for $j = 1, \dots,
-  \ell$. Therefore, $\hat{X}_{ij} \leq \hat{z}_j$ by diagonal dominance of
-  semidefinite matrices.
+  \ell$. Therefore, $\hat{X}_{ij} \leq \sqrt{\hat{X}_{ii}\hat{X}_{jj}} \leq \hat{z}_j$ by 
+  nonnegativity of the minors of positive semidefinite matrices.
 
@@
  \begin{tabular*}{\linewidth}{@{}l@{\;\;\extracolsep{\fill}}rrrrrrrrr@{}}\toprule 
-  & \multicolumn{3}{c}{A} & \multicolumn{3}{c}{B} & \multicolumn{3}{c}{C} \\ 
+  & \multicolumn{3}{c}{I} & \multicolumn{3}{c}{II} & \multicolumn{3}{c}{III} \\ 
 \cmidrule(r){2-4} \cmidrule{5-7}  \cmidrule(l){8-10}

### 2017-03-06 RIP-MISDP-Paper/Paper_v1.tex
merged master

@@
 Because of this exactness result, this new problem could drastically
-improve the bounds computed with the semidefinite relaxations proposed in \cite{Asp07} and \cite{Asp08}, but this came at the cost of a much higher complexity of computation.
-We could only solve these problems for relatively small
-instances, but this could change with sufficient advances in the area of mixed integer semidefinite programming, especially with the addition of special cutting planes, which proved to be the main enhancement needed to make mixed
-integer linear solvers as fast as they are today.
-Also special heuristics and branching rules could improve the performance of the branch-and-bound-algorithm and allow us to solve bigger instances of the proposed MISDP.
+improve the bounds computed with the semidefinite relaxations proposed in \cite{Asp07} and \cite{Asp08}, but this came at the cost of a much higher complexity of computation, at least in comparison to \cite{Asp07}.
+We could only
+solve these problems for relatively small instances, but this could change with sufficient advances in the area of mixed integer semidefinite programming, especially with the addition of special cutting planes, which proved to be
+the main enhancement needed to make mixed integer linear solvers as fast as they are today.
+Also special heuristics and branching rules could improve the performance of the branch-and-bound-algorithm and allow us to solve bigger
+instances of the proposed MISDP.
 

### 2017-03-06 RIP-MISDP-Paper/ResultScript.py

@@ -1,0 +1,727 @@
+import os
+import sys
+import math
+from decimal import Decimal
+
+#TODO: write tables for online supplement
+
+#which tables/plots for the paper should be prepared?
+texfile = 0 # make the output a compilable texfile or just a figure/table that can be included?
+Lbounds = 0
+Rbounds = 0
+LRbounds = 0
+Ltimes = 0
+Rtimes = 0
+LRtimes = 0
+printFails = 0
+completeTable = 0
+CompleteRICtable = 1
+
+MISDPfilename = "/local/gally/results/RIP-MISDP-Paper/160311/RIP-results/check.RIPMISDP.scipsdp.linux.x86_64.gnu.opt.sdpa.extra.branchinfobj_nofracdive.out"
+Asp0708path = "/local/gally/results/RIP-MISDP-Paper/160311/SDPA-results/"
+Asp07filenames = ["0+-115305A_07l.out", "0+-115305A_07r.out", "0+-115305B_07l.out", "0+-115305B_07r.out", "0+-115305C_07l.out", "0+-115305C_07r.out", "0+-125354A_07l.out", "0+-125354A_07r.out", "0+-125354B_07l.out", "0+-125354B_07r.out", "0+-125354C_07l.out", "0+-125354C_07r.out", "0+-130403A_07l.out", "0+-130403A_07r.out", "0+-130403B_07l.out", "0+-130403B_07r.out", "0+-130403C_07l.out", "0+-130403C_07r.out", "bina15305A_07l.out", "bina15305A_07r.out", "bina15305B_07l.out", "bina15305B_07r.out", "bina15305C_07l.out", "bina15305C_07r.out", "bina25354A_07l.out", "bina25354A_07r.out", "bina25354B_07l.out", "bina25354B_07r.out", "bina25354C_07l.out", "bina25354C_07r.out", "bina30403A_07l.out", "bina30403A_07r.out", "bina30403B_07l.out", "bina30403B_07r.out", "bina30403C_07l.out", "bina30403C_07r.out", "bern15305A_07l.out", "bern15305A_07r.out", "bern15305B_07l.out", "bern15305B_07r.out", "bern15305C_07l.out", "bern15305C_07r.out", "bern25354A_07l.out", "bern25354A_07r.out", "bern25354B_07l.out", "bern25354B_07r.out", "bern25354C_07l.out", "bern25354C_07r.out", "bern30403A_07l.out", "bern30403A_07r.out", "ber
[…]

### 2017-03-06 doc/xternal.c

@@ -1,11 +1,12 @@
 /* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
 /*                                                                           */
-/*                  This file is part of the                                 */
-/*      SDP-Package for SCIP: a solving framework for                        */
-/*                            mixed-integer semidefinite programms           */
+/* This file is part of SCIPSDP - a solving framework for mixed-integer      */
+/* semidefinite programms based on SCIP.                                     */
 /*                                                                           */
-/* Copyright (C) 2011-2014 Discrete Optimization, TU Darmstadt               */
+/* Copyright (C) 2011-2013 Discrete Optimization, TU Darmstadt               */
 /*                         EDOM, FAU Erlangen-Nürnberg                       */
+/*               2014-2016 Discrete Optimization, TU Darmstadt               */
+/*                                                                           */
 /*                                                                           */
 /* This program is free software; you can redistribute it and/or             */
 /* modify it under the terms of the GNU Lesser General Public License        */
@@ -21,6 +22,12 @@
 /* along with this program; if not, write to the Free Software               */
 /* Foundation, Inc., 51 Franklin St, Fifth Floor, Boston, MA 02110-1301, USA.*/
 /*                                                                           */
+/*                                                                           */
+/* Based on SCIP - Solving Constraint Integer Programs                       */
+/* Copyright (C) 2002-2016 Zuse Institute Berlin                             */
+/* SCIP is distributed under the terms of the SCIP Academic Licence,         */
+/* see file COPYING in the SCIP distribution.                                */
+/*                                                                           */
 /* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
 
 /**@file   xternal.c
@@ -32,13 +39,82 @@
 
 /**@mainpage Overview
  *
- * @version 1.0
- * @author Sonja Mars, Lars Schewe, Tristan Gally, Marc Pfetsch
- * @date 2011-2014
+ * @version 3.0.0
[…]

### 2017-03-06 src/scipsdp/BlockMemoryAllocator.h

@@ -1,11 +1,11 @@
 /* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
 /*                                                                           */
 /* This file is part of SCIPSDP - a solving framework for mixed-integer      */
-/* semidefinite programms based on SCIP.                                     */
+/* semidefinite programs based on SCIP.                                      */
 /*                                                                           */
 /* Copyright (C) 2011-2013 Discrete Optimization, TU Darmstadt               */
 /*                         EDOM, FAU Erlangen-Nürnberg                       */
-/*               2014      Discrete Optimization, TU Darmstadt               */
+/*               2014-2017 Discrete Optimization, TU Darmstadt               */
 /*                                                                           */
 /*                                                                           */
 /* This program is free software; you can redistribute it and/or             */
@@ -24,13 +24,13 @@
 /*                                                                           */
 /*                                                                           */
 /* Based on SCIP - Solving Constraint Integer Programs                       */
-/* Copyright (C) 2002-2014 Zuse Institute Berlin                             */
+/* Copyright (C) 2002-2017 Zuse Institute Berlin                             */
 /* SCIP is distributed under the terms of the SCIP Academic Licence,         */
 /* see file COPYING in the SCIP distribution.                                */
 /*                                                                           */
 /* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
 
-/**@file   BlockMemoryAllocator.cpp
+/**@file   BlockMemoryAllocator.h
  * @brief  An STL allocator class using SCIP block memory
  * @author Lars Schewe
  */

### 2017-03-06 src/scipsdp/ScipStreamBuffer.cpp

@@ -1,11 +1,11 @@
 /* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
 /*                                                                           */
 /* This file is part of SCIPSDP - a solving framework for mixed-integer      */
-/* semidefinite programms based on SCIP.                                     */
+/* semidefinite programs based on SCIP.                                      */
 /*                                                                           */
 /* Copyright (C) 2011-2013 Discrete Optimization, TU Darmstadt               */
 /*                         EDOM, FAU Erlangen-Nürnberg                       */
-/*               2014      Discrete Optimization, TU Darmstadt               */
+/*               2014-2017 Discrete Optimization, TU Darmstadt               */
 /*                                                                           */
 /*                                                                           */
 /* This program is free software; you can redistribute it and/or             */
@@ -24,7 +24,7 @@
 /*                                                                           */
 /*                                                                           */
 /* Based on SCIP - Solving Constraint Integer Programs                       */
-/* Copyright (C) 2002-2014 Zuse Institute Berlin                             */
+/* Copyright (C) 2002-2017 Zuse Institute Berlin                             */
 /* SCIP is distributed under the terms of the SCIP Academic Licence,         */
 /* see file COPYING in the SCIP distribution.                                */
 /*                                                                           */
@@ -37,10 +37,8 @@
 
 #include "ScipStreamBuffer.h"
 
-#include <cstdio>                       // for EOF
 #include <algorithm>                    // for min
 #include <cstring>                      // for memcpy
-//#include "scip/def.h"                   // for SCIP_CALL_ABORT
 #include "scip/scip.h"                  // for SCIPallocBufferArray, etc
 
 #define SCIPSTREAMBUFFERSIZE 256

### 2017-03-06 src/scipsdp/ScipStreamBuffer.h

@@ -1,11 +1,11 @@
 /* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
 /*                                                                           */
 /* This file is part of SCIPSDP - a solving framework for mixed-integer      */
-/* semidefinite programms based on SCIP.                                     */
+/* semidefinite programs based on SCIP.                                      */
 /*                                                                           */
 /* Copyright (C) 2011-2013 Discrete Optimization, TU Darmstadt               */
 /*                         EDOM, FAU Erlangen-Nürnberg                       */
-/*               2014      Discrete Optimization, TU Darmstadt               */
+/*               2014-2017 Discrete Optimization, TU Darmstadt               */
 /*                                                                           */
 /*                                                                           */
 /* This program is free software; you can redistribute it and/or             */
@@ -24,13 +24,13 @@
 /*                                                                           */
 /*                                                                           */
 /* Based on SCIP - Solving Constraint Integer Programs                       */
-/* Copyright (C) 2002-2014 Zuse Institute Berlin                             */
+/* Copyright (C) 2002-2017 Zuse Institute Berlin                             */
 /* SCIP is distributed under the terms of the SCIP Academic Licence,         */
 /* see file COPYING in the SCIP distribution.                                */
 /*                                                                           */
 /* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
 
-/**@file   ScipStreamBuffer.cpp
+/**@file   ScipStreamBuffer.h
  * @brief  An std::streambuf that uses SCIP I/O routines (suitable for reading)
  * @author Lars Schewe
  */
@@ -41,8 +41,6 @@
 #include <streambuf>                    // for streamsize, streambuf
 
 #include <cstddef>                      // for size_t
-//#include "scip/pub_fileio.h"            // for SCIP_FILE
-//#include "scip/type_scip.h"             // for SCIP
 #include "scip/scip.h"
 
 class ScipStreamBuffer : public std::streambuf
@@ -52,7 +50,7 @@ class ScipStreamBuffer : public std::streambuf
 
    ~ScipStreamBuffer();
 
- private:
+ protected:
    /// the underflow function is responsible for the refilling of the buffer
    virtual int underflow();
 

### 2017-03-06 src/scipsdp/SdpVarfixer.c

@@ -1,11 +1,11 @@
 /* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
 /*                                                                           */
 /* This file is part of SCIPSDP - a solving framework for mixed-integer      */
-/* semidefinite programms based on SCIP.                                     */
+/* semidefinite programs based on SCIP.                                      */
 /*                                                                           */
 /* Copyright (C) 2011-2013 Discrete Optimization, TU Darmstadt               */
 /*                         EDOM, FAU Erlangen-Nürnberg                       */
-/*               2014      Discrete Optimization, TU Darmstadt               */
+/*               2014-2017 Discrete Optimization, TU Darmstadt               */
 /*                                                                           */
 /*                                                                           */
 /* This program is free software; you can redistribute it and/or             */
@@ -24,48 +24,34 @@
 /*                                                                           */
 /*                                                                           */
 /* Based on SCIP - Solving Constraint Integer Programs                       */
-/* Copyright (C) 2002-2014 Zuse Institute Berlin                             */
+/* Copyright (C) 2002-2017 Zuse Institute Berlin                             */
 /* SCIP is distributed under the terms of the SCIP Academic Licence,         */
 /* see file COPYING in the SCIP distribution.                                */
 /*                                                                           */
 /* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
 
-/**@file   SdpVarfixer.h
+/**@file   SdpVarfixer.c
  * @brief  adds the main functionality to fix/unfix/(multi-)aggregate variables by merging two three-tuple-arrays of row/col/val together
  * @author Tristan Gally
  */
-
-/* somehow doesn't seem to work
-#ifndef __SDPVARMAPPER_H__
-#define __SDPVARMAPPER_H__
-*/
 
 #include "scip/type_misc.h"
 #include "scip/def.h"
 #include "scip/pub_misc.h" /* for sorting */
 #include "SdpVarfixer.h"
 
-static double epsilon    = 1e-6; /**< only values bigger than this are counted as nonzeros */
-
-/** Checks if a BMSallocMemory-call was successfull, otherwise returns SCIP_NOMEMRY */
- #define BMS_CALL(x) do \
-  { \
-  if( NULL == (x) ) \
-  { \
-  SCIPerrorMessage("No memory in function call\n"); \
-  return SCIP_NOMEMORY; \
-  } \
-  } \
-  while( FALSE )
[…]

### 2017-03-06 src/scipsdp/SdpVarfixer.h

@@ -1,11 +1,11 @@
 /* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
 /*                                                                           */
 /* This file is part of SCIPSDP - a solving framework for mixed-integer      */
-/* semidefinite programms based on SCIP.                                     */
+/* semidefinite programs based on SCIP.                                      */
 /*                                                                           */
 /* Copyright (C) 2011-2013 Discrete Optimization, TU Darmstadt               */
 /*                         EDOM, FAU Erlangen-Nürnberg                       */
-/*               2014      Discrete Optimization, TU Darmstadt               */
+/*               2014-2017 Discrete Optimization, TU Darmstadt               */
 /*                                                                           */
 /*                                                                           */
 /* This program is free software; you can redistribute it and/or             */
@@ -24,7 +24,7 @@
 /*                                                                           */
 /*                                                                           */
 /* Based on SCIP - Solving Constraint Integer Programs                       */
-/* Copyright (C) 2002-2014 Zuse Institute Berlin                             */
+/* Copyright (C) 2002-2017 Zuse Institute Berlin                             */
 /* SCIP is distributed under the terms of the SCIP Academic Licence,         */
 /* see file COPYING in the SCIP distribution.                                */
 /*                                                                           */
@@ -35,10 +35,8 @@
  * @author Tristan Gally
  */
 
-/* somehow doesn't seem to work
 #ifndef __SDPVARFIXER_H__
 #define __SDPVARFIXER_H__
-*/
 
 #include "scip/type_misc.h"
 
@@ -47,10 +45,11 @@ extern "C" {
 #endif
 
 /**
- * sort the given row, col and val arrays first by non-decreasing row-indices, than for those by identical row-indices by non-increasing col-indices
+ * sort the given row, col and val arrays first by non-decreasing row-indices, then for those with identical
+ * row-indices by non-decreasing col-indices
  */
 EXTERN
-void SdpVarfixerSortRowCol(
+void SCIPsdpVarfixerSortRowCol(
    int*                  row,                /* row indices */
    int*                  col,                /* column indices */
    SCIP_Real*            val,                /* values */
@@ -58,17 +57,20 @@ void SdpVarfixerSortRowCol(
    );
 
 /**
- * Merges two three-tuple-arrays together. The original arrays (which may have multiple entries for the same row and col) will be mulitplied with
+ * merges two three-tuple-arrays together
+ *
[…]

### 2017-03-06 src/scipsdp/SdpVarmapper.c

@@ -1,11 +1,11 @@
 /* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
 /*                                                                           */
 /* This file is part of SCIPSDP - a solving framework for mixed-integer      */
-/* semidefinite programms based on SCIP.                                     */
+/* semidefinite programs based on SCIP.                                      */
 /*                                                                           */
 /* Copyright (C) 2011-2013 Discrete Optimization, TU Darmstadt               */
 /*                         EDOM, FAU Erlangen-Nürnberg                       */
-/*               2014      Discrete Optimization, TU Darmstadt               */
+/*               2014-2017 Discrete Optimization, TU Darmstadt               */
 /*                                                                           */
 /*                                                                           */
 /* This program is free software; you can redistribute it and/or             */
@@ -24,7 +24,7 @@
 /*                                                                           */
 /*                                                                           */
 /* Based on SCIP - Solving Constraint Integer Programs                       */
-/* Copyright (C) 2002-2014 Zuse Institute Berlin                             */
+/* Copyright (C) 2002-2017 Zuse Institute Berlin                             */
 /* SCIP is distributed under the terms of the SCIP Academic Licence,         */
 /* see file COPYING in the SCIP distribution.                                */
 /*                                                                           */
@@ -44,6 +44,9 @@
 #include "scip/type_misc.h" /* for SCIP Hashmap */
 #include "SdpVarmapper.h"
 
+/* turn off lint warnings for whole file: */
+/*lint --e{788,818}*/
+
 struct Sdpvarmapper
 {
    SCIP_VAR**            sdptoscip;          /**< array of SCIP variables indexed by their SDP indices */
@@ -51,10 +54,10 @@ struct Sdpvarmapper
[…]

### 2017-03-06 src/scipsdp/SdpVarmapper.h

@@ -1,11 +1,11 @@
 /* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
 /*                                                                           */
 /* This file is part of SCIPSDP - a solving framework for mixed-integer      */
-/* semidefinite programms based on SCIP.                                     */
+/* semidefinite programs based on SCIP.                                      */
 /*                                                                           */
 /* Copyright (C) 2011-2013 Discrete Optimization, TU Darmstadt               */
 /*                         EDOM, FAU Erlangen-Nürnberg                       */
-/*               2014      Discrete Optimization, TU Darmstadt               */
+/*               2014-2017 Discrete Optimization, TU Darmstadt               */
 /*                                                                           */
 /*                                                                           */
 /* This program is free software; you can redistribute it and/or             */
@@ -24,7 +24,7 @@
 /*                                                                           */
 /*                                                                           */
 /* Based on SCIP - Solving Constraint Integer Programs                       */
-/* Copyright (C) 2002-2014 Zuse Institute Berlin                             */
+/* Copyright (C) 2002-2017 Zuse Institute Berlin                             */
 /* SCIP is distributed under the terms of the SCIP Academic Licence,         */
 /* see file COPYING in the SCIP distribution.                                */
 /*                                                                           */
@@ -47,87 +47,87 @@ extern "C" {
 
 typedef struct Sdpvarmapper SdpVarmapper;
 
-/** creates the SDP Varmapper */
+/** creates the SDP varmapper */
 EXTERN
-SCIP_RETCODE SdpVarmapperCreate(
+SCIP_RETCODE SCIPsdpVarmapperCreate(
    SCIP*                 scip,               /**< SCIP data structure */
-   SdpVarmapper**        varmapper,          /**< Pointer to the Varmapper that should be created */
+   SdpVarmapper**        varmapper,          /**< Pointer to the varmapper that should be created */
    int                   size                /**< initial size of the sciptosdp-hashmap */
    );
 
-/** frees the SDP Varmapper */
+/** frees the SDP varmapper */
 EXTERN
-SCIP_RETCODE SdpVarmapperFree(
-   SCIP*                 scip,              /**< SCIP data structure */
[…]

### 2017-03-06 src/scipsdp/config.h

@@ -1,38 +1,0 @@
-/* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
-/*                                                                           */
-/* This file is part of SCIPSDP - a solving framework for mixed-integer      */
-/* semidefinite programms based on SCIP.                                     */
-/*                                                                           */
-/* Copyright (C) 2011-2013 Discrete Optimization, TU Darmstadt               */
-/*                         EDOM, FAU Erlangen-Nürnberg                       */
-/*               2014      Discrete Optimization, TU Darmstadt               */
-/*                                                                           */
-/*                                                                           */
-/* This program is free software; you can redistribute it and/or             */
-/* modify it under the terms of the GNU Lesser General Public License        */
-/* as published by the Free Software Foundation; either version 3            */
-/* of the License, or (at your option) any later version.                    */
-/*                                                                           */
-/* This program is distributed in the hope that it will be useful,           */
-/* but WITHOUT ANY WARRANTY; without even the implied warranty of            */
-/* MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the             */
-/* GNU Lesser General Public License for more details.                       */
-/*                                                                           */
-/* You should have received a copy of the GNU Lesser General Public License  */
-/* along with this program; if not, write to the Free Software               */
-/* Foundation, Inc., 51 Franklin St, Fifth Floor, Boston, MA 02110-1301, USA.*/
-/*                                                                           */
-/*                                                                           */
-/* Based on SCIP - Solving Constraint Integer Programs                       */
-/* Copyright (C) 2002-2014 Zuse Institute Berlin                             */
-/* SCIP is distributed under the terms of the SCIP Academic Licence,         */
-/* see file COPYING in the SCIP distribution.                                */
-/*                                                                           */
-/* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
-
-/** Define to a macro mangling the given C identifier (in lower and upper
-   case), which must not contain underscores, for linking with Fortran. */
-#define F77_FUNC(name,NAME) name ## _
-
-/** As F77_FUNC, but for C identifiers containing underscores. */
-#define F77_FUNC_(name,NAME) name ## _

### 2017-03-06 src/scipsdp/cons_sdp.c

@@ -1,0 +1,2961 @@
+/* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
+/*                                                                           */
+/* This file is part of SCIPSDP - a solving framework for mixed-integer      */
+/* semidefinite programs based on SCIP.                                      */
+/*                                                                           */
+/* Copyright (C) 2011-2013 Discrete Optimization, TU Darmstadt               */
+/*                         EDOM, FAU Erlangen-Nürnberg                       */
+/*               2014-2017 Discrete Optimization, TU Darmstadt               */
+/*                                                                           */
+/*                                                                           */
+/* This program is free software; you can redistribute it and/or             */
+/* modify it under the terms of the GNU Lesser General Public License        */
+/* as published by the Free Software Foundation; either version 3            */
+/* of the License, or (at your option) any later version.                    */
+/*                                                                           */
+/* This program is distributed in the hope that it will be useful,           */
+/* but WITHOUT ANY WARRANTY; without even the implied warranty of            */
+/* MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the             */
+/* GNU Lesser General Public License for more details.                       */
+/*                                                                           */
+/* You should have received a copy of the GNU Lesser General Public License  */
+/* along with this program; if not, write to the Free Software               */
+/* Foundation, Inc., 51 Franklin St, Fifth Floor, Boston, MA 02110-1301, USA.*/
+/*                                                                           */
+/*                                                                           */
+/* Based on SCIP - Solving Constraint Integer Programs                       */
+/* Copyright (C) 2002-2017 Zuse Institute Berlin                             */
+/* SCIP is distributed under the terms of the SCIP Academic Licence,         */
+/* see file COPYING in the SCIP distribution.                                */
[…]

### 2017-03-06 src/scipsdp/cons_sdp.cpp

@@ -1,2159 +1,0 @@
-/* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
-/*                                                                           */
-/* This file is part of SCIPSDP - a solving framework for mixed-integer      */
-/* semidefinite programms based on SCIP.                                     */
-/*                                                                           */
-/* Copyright (C) 2011-2013 Discrete Optimization, TU Darmstadt               */
-/*                         EDOM, FAU Erlangen-Nürnberg                       */
-/*               2014      Discrete Optimization, TU Darmstadt               */
-/*                                                                           */
-/*                                                                           */
-/* This program is free software; you can redistribute it and/or             */
-/* modify it under the terms of the GNU Lesser General Public License        */
-/* as published by the Free Software Foundation; either version 3            */
-/* of the License, or (at your option) any later version.                    */
-/*                                                                           */
-/* This program is distributed in the hope that it will be useful,           */
-/* but WITHOUT ANY WARRANTY; without even the implied warranty of            */
-/* MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the             */
-/* GNU Lesser General Public License for more details.                       */
-/*                                                                           */
-/* You should have received a copy of the GNU Lesser General Public License  */
-/* along with this program; if not, write to the Free Software               */
-/* Foundation, Inc., 51 Franklin St, Fifth Floor, Boston, MA 02110-1301, USA.*/
-/*                                                                           */
-/*                                                                           */
-/* Based on SCIP - Solving Constraint Integer Programs                       */
-/* Copyright (C) 2002-2014 Zuse Institute Berlin                             */
-/* SCIP is distributed under the terms of the SCIP Academic Licence,         */
[…]

### 2017-03-06 src/scipsdp/cons_sdp.h

@@ -1,11 +1,11 @@
 /* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
 /*                                                                           */
 /* This file is part of SCIPSDP - a solving framework for mixed-integer      */
-/* semidefinite programms based on SCIP.                                     */
+/* semidefinite programs based on SCIP.                                      */
 /*                                                                           */
 /* Copyright (C) 2011-2013 Discrete Optimization, TU Darmstadt               */
 /*                         EDOM, FAU Erlangen-Nürnberg                       */
-/*               2014      Discrete Optimization, TU Darmstadt               */
+/*               2014-2017 Discrete Optimization, TU Darmstadt               */
 /*                                                                           */
 /*                                                                           */
 /* This program is free software; you can redistribute it and/or             */
@@ -24,17 +24,22 @@
 /*                                                                           */
 /*                                                                           */
 /* Based on SCIP - Solving Constraint Integer Programs                       */
-/* Copyright (C) 2002-2014 Zuse Institute Berlin                             */
+/* Copyright (C) 2002-2017 Zuse Institute Berlin                             */
 /* SCIP is distributed under the terms of the SCIP Academic Licence,         */
 /* see file COPYING in the SCIP distribution.                                */
 /*                                                                           */
 /* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
 
 /**@file   cons_sdp.h
- * @brief  constraint handler for SDPs
+ * @brief  Constraint handler for SDP-constraints
  * @author Sonja Mars
  * @author Lars Schewe
  * @author Tristan Gally
+ *
+ * Constraint handler for semidefinite constraints of the form \f$ \sum_{j=1}^n A_j y_j - A_0 \succeq 0 \f$,
+ * where the matrices \f$A_j\f$ and \f$A_0\f$ need to be symmetric. Only the nonzero entries of the matrices
+ * are stored.
+ *
  */
 
 #ifndef __SCIP_CONSHDLR_SDP_H__
@@ -47,74 +52,80 @@
 extern "C" {
 #endif
 
-/** creates the handler for sdp constraints and includes it in SCIP */
+/** creates the handler for SDP constraints and includes it in SCIP */
 EXTERN
 SCIP_RETCODE SCIPincludeConshdlrSdp(
    SCIP*                 scip                /**< SCIP data structure */
    );
[…]

### 2017-03-06 src/scipsdp/main.cpp

@@ -1,11 +1,11 @@
 /* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
 /*                                                                           */
 /* This file is part of SCIPSDP - a solving framework for mixed-integer      */
-/* semidefinite programms based on SCIP.                                     */
+/* semidefinite programs based on SCIP.                                      */
 /*                                                                           */
 /* Copyright (C) 2011-2013 Discrete Optimization, TU Darmstadt               */
 /*                         EDOM, FAU Erlangen-Nürnberg                       */
-/*               2014      Discrete Optimization, TU Darmstadt               */
+/*               2014-2017 Discrete Optimization, TU Darmstadt               */
 /*                                                                           */
 /*                                                                           */
 /* This program is free software; you can redistribute it and/or             */
@@ -24,167 +24,143 @@
 /*                                                                           */
 /*                                                                           */
 /* Based on SCIP - Solving Constraint Integer Programs                       */
-/* Copyright (C) 2002-2014 Zuse Institute Berlin                             */
+/* Copyright (C) 2002-2017 Zuse Institute Berlin                             */
 /* SCIP is distributed under the terms of the SCIP Academic Licence,         */
 /* see file COPYING in the SCIP distribution.                                */
 /*                                                                           */
 /* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
 
 /**@file   main.cpp
- * @brief  driver-file for solving MISDPs
+ * @brief  main file for solving MISDPs
  * @author Sonja Mars
+ * @author Tristan Gally
  */
-// standard library includes
-#include <string>
[…]

### 2017-03-06 src/scipsdp/objreader_sdpa.cpp

@@ -1,11 +1,11 @@
 /* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
 /*                                                                           */
 /* This file is part of SCIPSDP - a solving framework for mixed-integer      */
-/* semidefinite programms based on SCIP.                                     */
+/* semidefinite programs based on SCIP.                                      */
 /*                                                                           */
 /* Copyright (C) 2011-2013 Discrete Optimization, TU Darmstadt               */
 /*                         EDOM, FAU Erlangen-Nürnberg                       */
-/*               2014      Discrete Optimization, TU Darmstadt               */
+/*               2014-2017 Discrete Optimization, TU Darmstadt               */
 /*                                                                           */
 /*                                                                           */
 /* This program is free software; you can redistribute it and/or             */
@@ -24,7 +24,7 @@
 /*                                                                           */
 /*                                                                           */
 /* Based on SCIP - Solving Constraint Integer Programs                       */
-/* Copyright (C) 2002-2014 Zuse Institute Berlin                             */
+/* Copyright (C) 2002-2017 Zuse Institute Berlin                             */
 /* SCIP is distributed under the terms of the SCIP Academic Licence,         */
 /* see file COPYING in the SCIP distribution.                                */
 /*                                                                           */
@@ -32,10 +32,13 @@
 
 /**@file   objreader_sdpa.cpp
  * @brief  Reader for SDPA-files
- * @author Jakob Schelbert, Sonja Mars, Tristan Gally
+ * @author Jakob Schelbert
+ * @author Sonja Mars
+ * @author Tristan Gally
  */
 
-//#define SCIP_DEBUG
+/*#define SCIP_DEBUG*/
+/*#define SCIP_MORE_DEBUG*/
 
 /*---+----1----+----2----+----3----+----4----+----5----+----6----+----7----+----8----+----9----+----0----+----1----+----2*/
 #include "objreader_sdpa.h"
@@ -43,28 +46,29 @@
 #include <cassert>                     // for assert
 #include <cctype>                      // for isspace
 #include <cstdio>                      // for printf
[…]

### 2017-03-06 src/scipsdp/objreader_sdpa.h

@@ -1,11 +1,11 @@
 /* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
 /*                                                                           */
 /* This file is part of SCIPSDP - a solving framework for mixed-integer      */
-/* semidefinite programms based on SCIP.                                     */
+/* semidefinite programs based on SCIP.                                      */
 /*                                                                           */
 /* Copyright (C) 2011-2013 Discrete Optimization, TU Darmstadt               */
 /*                         EDOM, FAU Erlangen-Nürnberg                       */
-/*               2014      Discrete Optimization, TU Darmstadt               */
+/*               2014-2017 Discrete Optimization, TU Darmstadt               */
 /*                                                                           */
 /*                                                                           */
 /* This program is free software; you can redistribute it and/or             */
@@ -24,7 +24,7 @@
 /*                                                                           */
 /*                                                                           */
 /* Based on SCIP - Solving Constraint Integer Programs                       */
-/* Copyright (C) 2002-2014 Zuse Institute Berlin                             */
+/* Copyright (C) 2002-2017 Zuse Institute Berlin                             */
 /* SCIP is distributed under the terms of the SCIP Academic Licence,         */
 /* see file COPYING in the SCIP distribution.                                */
 /*                                                                           */
@@ -32,7 +32,8 @@
 
 /**@file   objreader_sdpa.h
  * @brief  Reader for SDPA-Files
- * @author Jakob Schelbert, Sonja Mars
+ * @author Jakob Schelbert
+ * @author Sonja Mars
  */
 
 /*---+----1----+----2----+----3----+----4----+----5----+----6----+----7----+----8----+----9----+----0----+----1----+----2*/
@@ -51,12 +52,13 @@ class SdpVarMapper;
 
 namespace scip
 {
-   /**struct with the lp-rows-data*/
+   /** struct with the lp-rows-data */
    struct LProw
    {
-      std::vector< std::pair<int, double> > data;
+      std::vector< std::pair<int, SCIP_Real> > data;
    };
-   /**class of one sdpblock, very similar to sdpcone and the sdpa-format*/
+
+   /** class of one sdpblock, very similar to sdpcone and the sdpa-format */
    class SDPBlock
    {
    public:
@@ -68,15 +70,15 @@ namespace scip
[…]

### 2017-03-06 src/scipsdp/relax_sdp.c

@@ -1,0 +1,1989 @@
+/* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
+/*                                                                           */
+/* This file is part of SCIPSDP - a solving framework for mixed-integer      */
+/* semidefinite programs based on SCIP.                                      */
+/*                                                                           */
+/* Copyright (C) 2011-2013 Discrete Optimization, TU Darmstadt               */
+/*                         EDOM, FAU Erlangen-Nürnberg                       */
+/*               2014-2017 Discrete Optimization, TU Darmstadt               */
+/*                                                                           */
+/*                                                                           */
+/* This program is free software; you can redistribute it and/or             */
+/* modify it under the terms of the GNU Lesser General Public License        */
+/* as published by the Free Software Foundation; either version 3            */
+/* of the License, or (at your option) any later version.                    */
+/*                                                                           */
+/* This program is distributed in the hope that it will be useful,           */
+/* but WITHOUT ANY WARRANTY; without even the implied warranty of            */
+/* MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the             */
+/* GNU Lesser General Public License for more details.                       */
+/*                                                                           */
+/* You should have received a copy of the GNU Lesser General Public License  */
+/* along with this program; if not, write to the Free Software               */
+/* Foundation, Inc., 51 Franklin St, Fifth Floor, Boston, MA 02110-1301, USA.*/
+/*                                                                           */
+/*                                                                           */
+/* Based on SCIP - Solving Constraint Integer Programs                       */
+/* Copyright (C) 2002-2017 Zuse Institute Berlin                             */
+/* SCIP is distributed under the terms of the SCIP Academic Licence,         */
+/* see file COPYING in the SCIP distribution.                                */
+/*                                                                           */
+/* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
+
+/**@file   relax_sdp.c
+ * @ingroup RELAXATORS
+ * @brief  SDP-relaxator
+ * @author Sonja Mars
+ * @author Tristan Gally
+ */
+
+/* #define SCIP_DEBUG*/
+/* #define SCIP_MORE_DEBUG   *//* displays complete solution for each relaxation */
[…]

### 2017-03-06 src/scipsdp/relax_sdp.cpp

@@ -1,1058 +1,0 @@
-/* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
-/*                                                                           */
-/* This file is part of SCIPSDP - a solving framework for mixed-integer      */
-/* semidefinite programms based on SCIP.                                     */
-/*                                                                           */
-/* Copyright (C) 2011-2013 Discrete Optimization, TU Darmstadt               */
-/*                         EDOM, FAU Erlangen-Nürnberg                       */
-/*               2014      Discrete Optimization, TU Darmstadt               */
-/*                                                                           */
-/*                                                                           */
-/* This program is free software; you can redistribute it and/or             */
-/* modify it under the terms of the GNU Lesser General Public License        */
-/* as published by the Free Software Foundation; either version 3            */
-/* of the License, or (at your option) any later version.                    */
-/*                                                                           */
-/* This program is distributed in the hope that it will be useful,           */
-/* but WITHOUT ANY WARRANTY; without even the implied warranty of            */
-/* MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the             */
-/* GNU Lesser General Public License for more details.                       */
-/*                                                                           */
-/* You should have received a copy of the GNU Lesser General Public License  */
-/* along with this program; if not, write to the Free Software               */
-/* Foundation, Inc., 51 Franklin St, Fifth Floor, Boston, MA 02110-1301, USA.*/
-/*                                                                           */
[…]

### 2017-03-06 src/scipsdp/relax_sdp.h

@@ -1,11 +1,11 @@
 /* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
 /*                                                                           */
 /* This file is part of SCIPSDP - a solving framework for mixed-integer      */
-/* semidefinite programms based on SCIP.                                     */
+/* semidefinite programs based on SCIP.                                      */
 /*                                                                           */
 /* Copyright (C) 2011-2013 Discrete Optimization, TU Darmstadt               */
 /*                         EDOM, FAU Erlangen-Nürnberg                       */
-/*               2014      Discrete Optimization, TU Darmstadt               */
+/*               2014-2017 Discrete Optimization, TU Darmstadt               */
 /*                                                                           */
 /*                                                                           */
 /* This program is free software; you can redistribute it and/or             */
@@ -24,7 +24,7 @@
 /*                                                                           */
 /*                                                                           */
 /* Based on SCIP - Solving Constraint Integer Programs                       */
-/* Copyright (C) 2002-2014 Zuse Institute Berlin                             */
+/* Copyright (C) 2002-2017 Zuse Institute Berlin                             */
 /* SCIP is distributed under the terms of the SCIP Academic Licence,         */
 /* see file COPYING in the SCIP distribution.                                */
 /*                                                                           */
@@ -32,9 +32,18 @@
 
 /**@file   relax_sdp.h
  * @ingroup RELAXATORS
- * @brief  SDP relaxator
+ * @brief  SDP-relaxator
  * @author Sonja Mars
  * @author Tristan Gally
+ *
+ * Relaxator to solve semidefinite programs of the form
+ * \f{eqnarray*}{
+ *    \min & & b^T y \\
+ *    \mbox{s.t.} & & \sum_{j=1}^n A_j^i y_j - A_0^i \succeq 0 \quad \forall i \leq m \\
+ *    & & Dy \geq d \\
+ *    & & \ell \leq y \leq u
+ * \f}
+ * for symmetric matrices \f$ A_j^i \in S_{k_i} \f$ and a matrix \f$ D \in \mathbb{R}^{k_0 \times n}. \f$
  */
 
[…]

### 2017-03-06 src/sdpi/sdpi.h

@@ -1,11 +1,11 @@
 /* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
 /*                                                                           */
 /* This file is part of SCIPSDP - a solving framework for mixed-integer      */
-/* semidefinite programms based on SCIP.                                     */
+/* semidefinite programs based on SCIP.                                      */
 /*                                                                           */
 /* Copyright (C) 2011-2013 Discrete Optimization, TU Darmstadt               */
 /*                         EDOM, FAU Erlangen-Nürnberg                       */
-/*               2014      Discrete Optimization, TU Darmstadt               */
+/*               2014-2017 Discrete Optimization, TU Darmstadt               */
 /*                                                                           */
 /*                                                                           */
 /* This program is free software; you can redistribute it and/or             */
@@ -24,29 +24,26 @@
 /*                                                                           */
 /*                                                                           */
 /* Based on SCIP - Solving Constraint Integer Programs                       */
-/* Copyright (C) 2002-2014 Zuse Institute Berlin                             */
+/* Copyright (C) 2002-2017 Zuse Institute Berlin                             */
 /* SCIP is distributed under the terms of the SCIP Academic Licence,         */
 /* see file COPYING in the SCIP distribution.                                */
 /*                                                                           */
 /* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
 
 /**@file   sdpi.h
- * @brief  interface methods for specific SDP solvers
- * @author Marc Pfetsch
- * @author Leif Naundorf
+ * @brief  General interface methods for SDP-preprocessing (mainly fixing variables and removing empty rows/cols)
  * @author Tristan Gally
  *
- * This file specifies a generic SDP solver interface used by SCIP to create, modify, and solve semidefinite programs of
+ * This file specifies a generic SDP-solver interface used by SCIP to create, modify, and solve semidefinite programs of
  * the (dual) form
  *
  *   \f{eqnarray*}{
[…]

### 2017-03-06 src/sdpi/sdpi_dsdp.c

@@ -1,1972 +1,0 @@
-/* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
-/*                                                                           */
-/* This file is part of SCIPSDP - a solving framework for mixed-integer      */
-/* semidefinite programms based on SCIP.                                     */
-/*                                                                           */
-/* Copyright (C) 2011-2013 Discrete Optimization, TU Darmstadt               */
-/*                         EDOM, FAU Erlangen-Nürnberg                       */
-/*               2014      Discrete Optimization, TU Darmstadt               */
-/*                                                                           */
-/*                                                                           */
-/* This program is free software; you can redistribute it and/or             */
-/* modify it under the terms of the GNU Lesser General Public License        */
-/* as published by the Free Software Foundation; either version 3            */
-/* of the License, or (at your option) any later version.                    */
-/*                                                                           */
-/* This program is distributed in the hope that it will be useful,           */
-/* but WITHOUT ANY WARRANTY; without even the implied warranty of            */
-/* MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the             */
-/* GNU Lesser General Public License for more details.                       */
-/*                                                                           */
-/* You should have received a copy of the GNU Lesser General Public License  */
-/* along with this program; if not, write to the Free Software               */
-/* Foundation, Inc., 51 Franklin St, Fifth Floor, Boston, MA 02110-1301, USA.*/
-/*                                                                           */
-/*                                                                           */
-/* Based on SCIP - Solving Constraint Integer Programs                       */
-/* Copyright (C) 2002-2014 Zuse Institute Berlin                             */
-/* SCIP is distributed under the terms of the SCIP Academic Licence,         */
-/* see file COPYING in the SCIP distribution.                                */
-/*                                                                           */
-/* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
-
-//#define SCIP_DEBUG
-//#define SCIP_MORE_DEBUG
-
-/**@file   sdpi_dsdp.c
- * @brief  interface for dsdp
- * @author Tristan Gally
- */
-
-#include <assert.h>
-
-#include "sdpi/sdpi.h"
-
-#include "dsdp5.h"                           /* for DSDPUsePenalty, etc */
-
-#include "blockmemshell/memory.h"            /* for memory allocation */
[…]

### 2017-03-06 src/sdpi/sdpi_general.c

@@ -1,3098 +1,0 @@
-/* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
-/*                                                                           */
-/* This file is part of SCIPSDP - a solving framework for mixed-integer      */
-/* semidefinite programms based on SCIP.                                     */
-/*                                                                           */
-/* Copyright (C) 2011-2013 Discrete Optimization, TU Darmstadt               */
-/*                         EDOM, FAU Erlangen-Nürnberg                       */
-/*               2014      Discrete Optimization, TU Darmstadt               */
-/*                                                                           */
-/*                                                                           */
-/* This program is free software; you can redistribute it and/or             */
-/* modify it under the terms of the GNU Lesser General Public License        */
-/* as published by the Free Software Foundation; either version 3            */
-/* of the License, or (at your option) any later version.                    */
-/*                                                                           */
-/* This program is distributed in the hope that it will be useful,           */
-/* but WITHOUT ANY WARRANTY; without even the implied warranty of            */
-/* MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the             */
-/* GNU Lesser General Public License for more details.                       */
-/*                                                                           */
-/* You should have received a copy of the GNU Lesser General Public License  */
-/* along with this program; if not, write to the Free Software               */
-/* Foundation, Inc., 51 Franklin St, Fifth Floor, Boston, MA 02110-1301, USA.*/
-/*                                                                           */
-/*                                                                           */
-/* Based on SCIP - Solving Constraint Integer Programs                       */
-/* Copyright (C) 2002-2014 Zuse Institute Berlin                             */
-/* SCIP is distributed under the terms of the SCIP Academic Licence,         */
-/* see file COPYING in the SCIP distribution.                                */
-/*                                                                           */
-/* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
-
-#define SCIP_DEBUG
-#define SCIP_MORE_DEBUG
-
-/**@file   sdpi_dsdp.c
- * @brief  interface for dsdp
- * @author Tristan Gally
- */
-
-#include <assert.h>
-
-#include "sdpi/sdpi.h"
-#include "sdpi/sdpi_general.h"
-#include "scipsdp/SdpVarfixer.h"
-
-#include "blockmemshell/memory.h"            /* for memory allocation */
[…]

### 2017-03-06 src/sdpi/sdpi_general.h

@@ -1,829 +1,0 @@
-/* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
-/*                                                                           */
-/* This file is part of SCIPSDP - a solving framework for mixed-integer      */
-/* semidefinite programms based on SCIP.                                     */
-/*                                                                           */
-/* Copyright (C) 2011-2013 Discrete Optimization, TU Darmstadt               */
-/*                         EDOM, FAU Erlangen-Nürnberg                       */
-/*               2014      Discrete Optimization, TU Darmstadt               */
-/*                                                                           */
-/*                                                                           */
-/* This program is free software; you can redistribute it and/or             */
-/* modify it under the terms of the GNU Lesser General Public License        */
-/* as published by the Free Software Foundation; either version 3            */
-/* of the License, or (at your option) any later version.                    */
-/*                                                                           */
-/* This program is distributed in the hope that it will be useful,           */
-/* but WITHOUT ANY WARRANTY; without even the implied warranty of            */
-/* MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the             */
-/* GNU Lesser General Public License for more details.                       */
-/*                                                                           */
-/* You should have received a copy of the GNU Lesser General Public License  */
-/* along with this program; if not, write to the Free Software               */
-/* Foundation, Inc., 51 Franklin St, Fifth Floor, Boston, MA 02110-1301, USA.*/
-/*                                                                           */
-/*                                                                           */
-/* Based on SCIP - Solving Constraint Integer Programs                       */
-/* Copyright (C) 2002-2014 Zuse Institute Berlin                             */
-/* SCIP is distributed under the terms of the SCIP Academic Licence,         */
-/* see file COPYING in the SCIP distribution.                                */
-/*                                                                           */
[…]

### 2017-03-06 src/sdpi/sdpi_msk.c

@@ -1,4618 +1,0 @@
-/* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
-/*                                                                           */
-/*                  This file is part of the program and library             */
-/*         SCIP --- Solving Constraint Integer Programs                      */
-/*                                                                           */
-/*    Copyright (C) 2002-2013 Konrad-Zuse-Zentrum                            */
-/*                            fuer Informationstechnik Berlin                */
-/*                                                                           */
-/*  SCIP is distributed under the terms of the ZIB Academic License.         */
-/*                                                                           */
-/*  You should have received a copy of the ZIB Academic License              */
-/*  along with SCIP; see the file COPYING. If not email to scip@zib.de.      */
-/*                                                                           */
-/* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
-
-/**@file   sdpi_msk.c
- * @ingroup SDPIS
- * @brief  SDP interface for MOSEK
- * @author Bo Jensen
- * @author Leif Naundorf
- */
-
-#include <assert.h>
-
-#define MSKCONST const
-#include "mosek.h"
-
-#include "sdpi/sdpi.h"
-#include "scip/bitencode.h"
-#include <string.h>
-
-/* check version */
-#if (MSK_VERSION_MAJOR < 7)
-#error "MOSEK version 7 or higher is required for the SDP functionality"
-#endif
-
-/* do defines for windows directly her to make the sdpi more independent*/
-#if defined(_WIN32) || defined(_WIN64)
-#define snprintf _snprintf
-#endif
-
-#define scipmskobjsen MSKobjsensee
-#define SENSE2MOSEK(objsen) (((objsen)==SCIP_OBJSEN_MINIMIZE)?(MSK_OBJECTIVE_SENSE_MINIMIZE):(MSK_OBJECTIVE_SENSE_MAXIMIZE))
-
-#define MOSEK_CALL(x)  do                                                                                     \
[…]

### 2017-03-06 src/sdpi/sdpi_none.c

@@ -1,1534 +1,0 @@
-/* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
-/*                                                                           */
-/*                  This file is part of the program and library             */
-/*         SCIP --- Solving Constraint Integer Programs                      */
-/*                                                                           */
-/*    Copyright (C) 2002-2013 Konrad-Zuse-Zentrum                            */
-/*                            fuer Informationstechnik Berlin                */
-/*                                                                           */
-/*  SCIP is distributed under the terms of the ZIB Academic License.         */
-/*                                                                           */
-/*  You should have received a copy of the ZIB Academic License              */
-/*  along with SCIP; see the file COPYING. If not email to scip@zib.de.      */
-/*                                                                           */
-/* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
-
-/**@file   sdpi_none.c
- * @ingroup SDPIS
- * @brief  dummy interface for the case no SDP solver is needed
- * @author Stefan Heinz
- * @author Leif Naundorf
- */
-
-/*--+----1----+----2----+----3----+----4----+----5----+----6----+----7----+----8----+----9----+----0----+----1----+----2*/
-
-#include <assert.h>
-
-#include "sdpi/sdpi.h"
-#include "scip/pub_message.h"
-
-#define SDPINAME          "NONE"              /**< name of the SDPI interface */
-#define SDPIINFINITY       1e20               /**< infinity value */
-
-/** SDP interface
- *
- *  Store several statistic values about the SDP. These values are only needed in order to provide a rudimentary
- *  communication, e.g., there are asserts that check the number of rows and columns.
- */
-struct SCIP_SDPi
-{
-   int                   nrows;              /**< number of rows */
-   int                   ncols;              /**< number of columns */
-   int                   nsdpvars;           /**< number of semidefinite variables */
-};
-
-
-/*
- * Local Methods
- */
-
-/** error handling method */
-static
-void errorMessageAbort(
-   void
-   )
-{
-   SCIPerrorMessage("No SDP solver available (SDPS=none).\n");
[…]

### 2017-03-06 src/sdpi/sdpi_sdpa.cpp

@@ -1,2938 +1,0 @@
-/* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
-/*                                                                           */
-/* This file is part of SCIPSDP - a solving framework for mixed-integer      */
-/* semidefinite programms based on SCIP.                                     */
-/*                                                                           */
-/* Copyright (C) 2011-2013 Discrete Optimization, TU Darmstadt               */
-/*                         EDOM, FAU Erlangen-Nürnberg                       */
-/*               2014      Discrete Optimization, TU Darmstadt               */
-/*                                                                           */
-/*                                                                           */
-/* This program is free software; you can redistribute it and/or             */
-/* modify it under the terms of the GNU Lesser General Public License        */
-/* as published by the Free Software Foundation; either version 3            */
-/* of the License, or (at your option) any later version.                    */
-/*                                                                           */
-/* This program is distributed in the hope that it will be useful,           */
-/* but WITHOUT ANY WARRANTY; without even the implied warranty of            */
-/* MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the             */
-/* GNU Lesser General Public License for more details.                       */
-/*                                                                           */
-/* You should have received a copy of the GNU Lesser General Public License  */
-/* along with this program; if not, write to the Free Software               */
-/* Foundation, Inc., 51 Franklin St, Fifth Floor, Boston, MA 02110-1301, USA.*/
-/*                                                                           */
-/*                                                                           */
-/* Based on SCIP - Solving Constraint Integer Programs                       */
-/* Copyright (C) 2002-2014 Zuse Institute Berlin                             */
-/* SCIP is distributed under the terms of the SCIP Academic Licence,         */
[…]

### 2017-03-06 src/sdpi/sdpsolchecker.c

@@ -1,0 +1,270 @@
+/* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
+/*                                                                           */
+/* This file is part of SCIPSDP - a solving framework for mixed-integer      */
+/* semidefinite programs based on SCIP.                                      */
+/*                                                                           */
+/* Copyright (C) 2011-2013 Discrete Optimization, TU Darmstadt               */
+/*                         EDOM, FAU Erlangen-Nürnberg                       */
+/*               2014-2017 Discrete Optimization, TU Darmstadt               */
+/*                                                                           */
+/*                                                                           */
+/* This program is free software; you can redistribute it and/or             */
+/* modify it under the terms of the GNU Lesser General Public License        */
+/* as published by the Free Software Foundation; either version 3            */
+/* of the License, or (at your option) any later version.                    */
+/*                                                                           */
+/* This program is distributed in the hope that it will be useful,           */
+/* but WITHOUT ANY WARRANTY; without even the implied warranty of            */
+/* MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the             */
+/* GNU Lesser General Public License for more details.                       */
+/*                                                                           */
+/* You should have received a copy of the GNU Lesser General Public License  */
+/* along with this program; if not, write to the Free Software               */
+/* Foundation, Inc., 51 Franklin St, Fifth Floor, Boston, MA 02110-1301, USA.*/
+/*                                                                           */
+/*                                                                           */
+/* Based on SCIP - Solving Constraint Integer Programs                       */
+/* Copyright (C) 2002-2017 Zuse Institute Berlin                             */
+/* SCIP is distributed under the terms of the SCIP Academic Licence,         */
+/* see file COPYING in the SCIP distribution.                                */
+/*                                                                           */
[…]

### 2017-03-06 src/sdpi/type_sdpi.h

@@ -1,21 +1,38 @@
 /* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
 /*                                                                           */
-/*                  This file is part of the program and library             */
-/*         SCIP --- Solving Constraint Integer Programs                      */
+/* This file is part of SCIPSDP - a solving framework for mixed-integer      */
+/* semidefinite programs based on SCIP.                                      */
 /*                                                                           */
-/*    Copyright (C) 2002-2013 Konrad-Zuse-Zentrum                            */
-/*                            fuer Informationstechnik Berlin                */
+/* Copyright (C) 2011-2013 Discrete Optimization, TU Darmstadt               */
+/*                         EDOM, FAU Erlangen-Nürnberg                       */
+/*               2014-2017 Discrete Optimization, TU Darmstadt               */
 /*                                                                           */
-/*  SCIP is distributed under the terms of the ZIB Academic License.         */
 /*                                                                           */
-/*  You should have received a copy of the ZIB Academic License              */
-/*  along with SCIP; see the file COPYING. If not email to scip@zib.de.      */
+/* This program is free software; you can redistribute it and/or             */
+/* modify it under the terms of the GNU Lesser General Public License        */
+/* as published by the Free Software Foundation; either version 3            */
+/* of the License, or (at your option) any later version.                    */
+/*                                                                           */
+/* This program is distributed in the hope that it will be useful,           */
+/* but WITHOUT ANY WARRANTY; without even the implied warranty of            */
+/* MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the             */
+/* GNU Lesser General Public License for more details.                       */
[…]

### 2017-03-06 tools/makerelease.sh

@@ -1,6 +1,6 @@
 #!/bin/sh
 
-VERSION="1.0"
+VERSION="3.0.0"
 NAME="scipsdp-$VERSION"
 rm -f $NAME
 
@@ -17,7 +17,18 @@ tar --no-recursion --ignore-failed-read -cvzhf $NAME.tgz \
     $NAME/Makefile \
     $NAME/doc/scipsdp.dxy \
     $NAME/doc/xternal.c \
-    $NAME/src/*.c $NAME/src/*.cpp $NAME/src/*.h
+    $NAME/doc/layout.xml \
+    $NAME/instances/* \
+    $NAME/INSTALL \
+    $NAME/license.txt \
+    $NAME/changelog.txt \
+    $NAME/sdpa_format.txt \
+    $NAME/src/scipsdp/*.c $NAME/src/scipsdp/*.cpp $NAME/src/scipsdp/*.h \
+    $NAME/src/sdpi/*.c $NAME/src/sdpi/*.cpp $NAME/src/sdpi/*.h \
+    $NAME/src/scipsdpgithash.c \
+    $NAME/check/testset/short.test \
+    $NAME/check/testset/short.solu \
+    $NAME/settings/lp_approx.set
 
 echo ""
 echo "check version numbers in doc/scipsdp.dxy, doc/xternal.c, and Makefile:"

### 2017-03-06 src/scipsdp/relax_sdp.c
readd changes in relax_sdp which were lost due to delete/modify conflict

@@ -51,6 +51,7 @@
 #include "SdpVarmapper.h"
 #include "sdpi/sdpi.h"
 #include "scipsdp/cons_sdp.h"
+#include "scipsdp/cons_savesdpsol.h"
 #include "scipsdp/cons_savedsdpsettings.h"
 
 /* turn off lint warnings for whole file: */
@@ -667,7 +668,57 @@ SCIP_RETCODE calcRelax(
    enforceslater = SCIPisInfinity(scip, -1 * SCIPnodeGetLowerbound(SCIPgetCurrentNode(scip)));
 
    /* solve the problem */
-   SCIP_CALL( SCIPsdpiSolve(sdpi, NULL, startsetting, enforceslater, timelimit) );
+   if (!SCIPnodeGetParent(SCIPgetCurrentNode(scip)))
+      SCIP_CALL(SCIPsdpiSolve(sdpi, NULL, startsettings, enforceslater, timelimit)); /* we are in the root node, so there is no chance for a warmstart */
+   else
+   {
+      SCIP_CONSHDLR* conshdlr;
+      SCIP_Real* start;
+      int length;
+
+      /* find starting solution as optimal solution of parent node */
+
+      /* allocate memory */
+      SCIP_CALL( SCIPallocBufferArray(scip, &start, nvars) );
+
+      /* get constraint handler */
+      conshdlr = SCIPfindConshdlr(scip, "Savesdpsol");
+      if ( conshdlr == NULL )
+      {
+         SCIPerrorMessage("Savesdpsol constraint handler not found\n");
+         return SCIP_PLUGINNOTFOUND;
+      }
+
+      /* get constraints */
+      conss = SCIPconshdlrGetConss(conshdlr);
+
+      assert ( conss != NULL );
+      assert ( conss[0] != NULL );
+
+      /* because of stickingtonode we should only have one constraint of this type */
+      //assert( SCIPconshdlrGetNConss(conshdlr) == 1 );
+
+      /* get the solution */
+      length = nvars;
+      SCIP_CALL( getStartingPoint(scip, conss[0], start, &length) );
+
+      /* make sure that the memory was sufficient (this has to be the case as length = nvars */
+      assert ( length <= nvars );
+
+      /* check if start is still feasible for the variable bounds, otherwise round it */
+      for (i = 0; i < nvars; i++)
+      {
[…]

### 2017-03-06 src/scipsdp/relax_sdp.c
took care of some more version mismatches

@@ -669,12 +669,13 @@ SCIP_RETCODE calcRelax(
 
    /* solve the problem */
    if (!SCIPnodeGetParent(SCIPgetCurrentNode(scip)))
-      SCIP_CALL(SCIPsdpiSolve(sdpi, NULL, startsettings, enforceslater, timelimit)); /* we are in the root node, so there is no chance for a warmstart */
+      SCIP_CALL(SCIPsdpiSolve(sdpi, NULL, startsetting, enforceslater, timelimit)); /* we are in the root node, so there is no chance for a warmstart */
    else
    {
       SCIP_CONSHDLR* conshdlr;
       SCIP_Real* start;
       int length;
+      int v;
 
       /* find starting solution as optimal solution of parent node */
 
@@ -706,16 +707,16 @@ SCIP_RETCODE calcRelax(
       assert ( length <= nvars );
 
       /* check if start is still feasible for the variable bounds, otherwise round it */
-      for (i = 0; i < nvars; i++)
+      for (v = 0; v < nvars; v++)
       {
-         if (start[i] < SCIPvarGetLbLocal(SdpVarmapperGetSCIPvar(varmapper, i)))
-            start[i] = SCIPvarGetLbLocal(SdpVarmapperGetSCIPvar(varmapper, i));
-         else if (start[i] > SCIPvarGetUbLocal(SdpVarmapperGetSCIPvar(varmapper, i)))
-            start[i] = SCIPvarGetUbLocal(SdpVarmapperGetSCIPvar(varmapper, i));
+         if (start[v] < SCIPvarGetLbLocal(SCIPsdpVarmapperGetSCIPvar(relaxdata->varmapper, v)))
+            start[v] = SCIPvarGetLbLocal(SCIPsdpVarmapperGetSCIPvar(relaxdata->varmapper, v));
+         else if (start[v] > SCIPvarGetUbLocal(SCIPsdpVarmapperGetSCIPvar(relaxdata->varmapper, v)))
+            start[v] = SCIPvarGetUbLocal(SCIPsdpVarmapperGetSCIPvar(relaxdata->varmapper, v));
       }
 
       /* solve with given starting point */
-      SCIP_CALL(SCIPsdpiSolve(sdpi, start));
+      SCIP_CALL(SCIPsdpiSolve(sdpi, start, startsetting, enforceslater, timelimit));
 
       SCIPfreeBufferArray(scip, &start);
    }

### 2020-06-27 src/scipsdp/cons_sdp.c
fix bug in computeScalingFactor()

@@ -524,7 +524,7 @@ SCIP_RETCODE computeScalingFactor(
          for (j = 0; j < blocksize; ++j)
          {
             pos = i * blocksize + j;
-            matrix[pos] = scalar * A[pos] + Aconst[pos];
+            matrix[pos] = scalar * A[pos] - Aconst[pos];
          }
       }
 

### 2020-06-27 src/scipsdp/cons_sdp.c
do not need that all variables are binary

@@ -821,16 +821,6 @@ SCIP_RETCODE tightenMatrices(
       if ( ! consdata->allmatricespsd )
          continue;
 
-      /* make sure that all variables are binary */
-      nvars = consdata->nvars;
-      for (i = 0; i < nvars; ++i)
-      {
-         if ( ! SCIPvarIsBinary(consdata->vars[i]) )
-            break;
-      }
-      if ( i < nvars )
-         continue;
-
       SCIPdebugMsg(scip, "Trying to tighten matrices for constraint <%s>.\n", SCIPconsGetName(conss[c]));
 
       /* get matrices */
@@ -840,8 +830,12 @@ SCIP_RETCODE tightenMatrices(
 
       SCIP_CALL( SCIPconsSdpGetFullConstMatrix(scip, conss[c], constmatrix) );
 
+      nvars = consdata->nvars;
       for (i = 0; i < nvars; ++i)
       {
+         if ( ! SCIPvarIsBinary(consdata->vars[i]) )
+            continue;
+
          SCIP_CALL( SCIPconsSdpGetFullAj(scip, conss[c], i, matrix) );
 
          SCIP_CALL( computeScalingFactor(scip, blocksize, matrix, constmatrix, 0.0, 1.0, &factor) );

### 2020-06-27 tex/MISDP-Presolving/MISDPpresolving.tex
add document on presolving

\documentclass[10pt, a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[american]{babel}
\usepackage[dvipsnames]{xcolor}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{dsfont}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage[lined,ruled,noend,linesnumbered]{algorithm2e}
\usepackage[color=red!50!blue!50,textsize=footnotesize,textwidth=4cm]{todonotes}
\usepackage[breaklinks,colorlinks,citecolor=blue,linkcolor=blue]{hyperref}
% \usepackage[tmargin=2.5cm,bmargin=2.5cm,lmargin=2.8cm,rmargin=2.8cm]{geometry}
\usepackage{cleveref}
\usepackage{xspace}

% operators
\DeclareMathOperator{\conv}{conv}
\DeclareMathOperator{\cone}{cone}
\DeclareMathOperator{\aff}{aff}
\DeclareMathOperator{\lin}{lin}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}
\DeclareMathOperator{\supp}{supp}
\DeclareMathOperator{\relint}{relint}
\DeclareMathOperator{\diag}{diag}
\DeclareMathOperator{\Diag}{Diag}
\DeclareMathOperator{\sgn}{sgn}
\DeclareMathOperator{\tr}{tr}
\renewcommand{\Re}{\text{Re}}

% macros
\newcommand{\suchthat}{\,:\,}
\newcommand{\abs}[1]{\lvert{#1}\rvert}
\newcommand{\card}[1]{\lvert{#1}\rvert}
\newcommand{\ceil}[1]{\lceil{#1}\rceil}
\newcommand{\floor}[1]{\lfloor{#1}\rfloor}
\newcommand{\define}{\coloneqq}
\newcommand{\enifed}{\eqqcolon}
\newcommand{\norm}[1]{\lVert{#1}\rVert}
\newcommand{\Norm}[2]{\lVert{#1}\rVert_{#2}}
\newcommand{\skal}[2]{\langle{#1},{#2}\rangle}
\newcommand{\order}[1]{O\left({#1}\right)}
\newcommand{\T}{^{\top}}

% sets
\newcommand{\R}{\mathds{R}}
\newcommand{\Z}{\mathds{Z}}
\newcommand{\Q}{\mathds{Q}}
\newcommand{\N}{\mathds{N}}
\newcommand{\C}{\mathds{C}}
\newcommand{\E}{\mathds{E}}
\newcommand{\ones}{\mathds{1}}
\renewcommand{\P}{\mathds{P}}
\renewcommand{\E}{\mathds{E}}

% environments
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{observation}[theorem]{Observation}
\newtheorem{example}[theorem]{Example}
\newtheorem{question}[theorem]{Question}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{assumption}[theorem]{Assumption}

% adjust itemize
\setlist[itemize]{topsep=0.5ex,partopsep=0ex,parsep=0ex,itemsep=0.5ex}

% other defines
\newcommand{\MP}[1]{\todo{#1}}
\newcommand{\MPin}[1]{\todo[inline]{#1}}

% title
\title{Presolving for\\ Mixed-Integer Semidefinite Optimization}
\author{Marc E. Pfetsch\thanks{Department of Mathematics, TU Darmstadt, Germany}}
\date{\today}

% -------------------------------------------------------------------------
% -------------------------------------------------------------------------
\begin{document}

\maketitle

\section{Introduction}

We consider mixed-integer semidefinite programs (MISDP) of the following
form:
\begin{equation}\label{MISDP}
  \begin{aligned}
    \inf \quad & b\T y \\
    \text{s.t.} \quad & \sum_{i=1}^m A^i\, y_i - A^0 \succeq 0, \\
    & \ell_i \leq y_i \leq u_i && \forall\, i \in [m], \\
    & y_i \in \Z && \forall\, i \in I,
  \end{aligned}
\end{equation}
with symmetric matrices $A^i \in \R^{n \times n}$ for
$i \in [m]_0 \define \{0, \dots, m\}$, $b \in \R^m$,
$\ell_i \in \R \cup \{- \infty\}$, $u_i \in \R \cup \{\infty\}$ for all
$i \in [m] \define \{1, \dots, m\}$. The set of indices of integer
variables is given by $I \subseteq [m]$.

We also consider the following form -- often called ``primal'' form:
\begin{equation}\label{MISDP-P}
  \begin{aligned}
    \inf \quad & \skal{A^0}{X} \\
    \text{s.t.} \quad & \skal{A^i}{X} = b_i && \forall\, i \in [m] \\
    & L_{ij} \leq X_{ij} \leq U_{ij} && \forall\, i,\,j \in [n], \\
    & X_{ij} \in \Z && \forall\, i,\,j \in I \times I,\\
    & X \succeq 0.
  \end{aligned}
\end{equation}
Here the bounds are given by $L_{ij} \in \R \cup \{- \infty\}$, $U_{ij} \in \R \cup \{\infty\}$ for all
$i$, $j \in [m]$.

In this manuscript we try to derive presolving methods for the above two
problems. For instance we try to generalize the presolving techniques from
the MIP to the MISDP case. For instance, we can consider the techniques
demonstrated by Savelsbergh~\cite{Sav94}.

\paragraph{Notation}

We use the notation $A(y) \define \sum_{i=1}^m A^i\, y_i - A^0$ for $y \in \R^m$.

% -------------------------------------------------------------------------
% -------------------------------------------------------------------------
\section{Possibly Useful Procedures}

One basic argument that we will use repeatedly is the following. Consider a
positive semidefinite (psd) matrix $A \in \R^{n \times n}$. Then the
diagonal entries are nonnegative (since $0 \leq e_i\T A e_i =
A_{ii}$). Then taking the $2 \times 2$-minor for $i$ and $j \in [n]$, we
get $A_{ii}\, A_{jj} - A_{ij}^2 \geq 0$. This implies
\begin{equation}\label{eq:TwoByTwoMinorInequality}
  \abs{A_{ij}} \leq \sqrt{A_{ii} A_{jj}} \leq \frac{A_{ii} + A_{jj}}{2} \leq \max\,\{A_{ii},
  A_{jj}\},
\end{equation}
where we used inequality of geometric and arithmetic means for the second inequality.

This has the following two implications (see, e.g., Observation 1.1.5 and
1.1.6 in Helmberg~\cite{Hel00}):
\begin{enumerate}
\item There the always exists a diagonal entry of maximal absolute value,
  i.e., there exists $r \in [n]$ with
  \[
    A_{rr} = \max\, \{\abs{A_{ij}} \suchthat i,\,j \in [n]\}.
  \]
\item If $A_{ii} = 0$ then $A_{ij} = 0$ for all $j \in [n]$.
\end{enumerate}

\begin{question}
  Is the second condition enforced somehow -- in the sense that the bounds
  are fixed? For example if we branch variables in~\eqref{MISDP} such that
  a diagonal entry becomes 0, are the other entries in the column and row
  fixed to 0 as well?
\end{question}

% -------------------------------------------------------------------------
\subsection{Implied Linear Inequalities}

The following inequalities are known from the literature, see
Gally~\cite{Gal19}, and can be added as linear inequalities w.r.t.\
\eqref{MISDP}:
\begin{itemize}
\item Any feasible $y \in \R^m$ satisfies $A(y) \succeq 0$, which implies
  that the diagonal entries of $A(y)$ are nonnegative, i.e., for all
  $i \in [n]$:
  \[
    \sum_{k = 1}^m (A^k)_{ii}\, y_k - (A^0)_{ii} \geq 0.
  \]
\item If for some $i$, $j \in [n]$, $A^0_{ij} \neq 0$, and $A^0_{ii} = 0$,
  $A^k_{ij} = 0$ for all $k \in [m]$, $A^k_{ii} = 0$ for all continuous
  variables and $\ell_i \geq 0$ for all integer variables,
  the following inequality is valid:
  \[
    \sum_{\substack{k \in \mathcal{I}:\\ A^k_{ii} > 0}} y_k \geq 1.
  \]
\item For all $X \succeq 0$, we have
  \[
    X_{ii} + X_{jj} - 2\, X_{ij} \geq 0,\qquad
    X_{ii} + X_{jj} + 2\, X_{ij} \geq 0.
  \]
  This follows by using a 2 by 2 minor and multiplying from left
  and right by the all-ones vector and $[1,-1]$, respectively. It also
  follows from~\eqref{eq:TwoByTwoMinorInequality}.
\item The previous inequalities translated to the matrix pencil are:
  \begin{align*}
    & \sum_{k=1}^m A^k_{ii}\, y_k - A^0_{ii} + \sum_{k=1}^m A^k_{jj}\, y_k - A^0_{jj} - 2 \Big(\sum_{k=1}^m A^k_{ij}\, y_k - A^0_{ij}\Big) \geq 0\\
    \quad\Leftrightarrow\quad&
    \sum_{k=1}^m \Big(A ^k_{ii} + A^k_{jj} - 2\, A^k_{ij}\Big)\, y_k \geq A^0_{ii} + A^0_{jj} - 2 A^0_{ij}.
  \end{align*}
\end{itemize}

All the above methods are implemented and help for some instances.

% -------------------------------------------------------------------------
\subsection{Presolving for $\boldsymbol{2 \times 2}$-minors}
\label{sec:2by2Minors}

\begin{lemma}
  Consider a solution $X$ to~\eqref{MISDP-P}. Then
  \[
    -\sqrt{U_{ii}\, U_{jj}} \leq X_{ij} \leq \sqrt{U_{ij}\, U_{jj}}
  \]
  holds for all $i$, $j \in [n]$.
\end{lemma}

\begin{proof}
  Since $X$ is positive semidefinite, we have
  \[
    X_{ii} X_{jj} - X_{ij}^2 \geq 0
    \quad\Leftrightarrow\quad
    X_{ij}^2 \leq X_{ii} X_{jj}.
  \]
  Moreover, since the diagonal terms of $X$ are nonnegative, we have
  $U_{ii} \geq L_{ii} \geq 0$ for all $i \in [n]$. Upper bounding via
  McCormick inequalities yields:
  \begin{align*}
    X_{ii} X_{jj} & \leq \min \{L_{jj}\, X_{ii} + U_{ii}\, X_{jj} - U_{ii}\, L_{jj},\;
                    U_{jj}\, X_{ii} + L_{ii}\, X_{jj} - L_{ii}\, U_{jj}\}  \\
    & \leq U_{ii}\, U_{jj}.
  \end{align*}
  Taking the square root gives the claim.
\end{proof}

This result can partly be translated to the matrix pencil format by defining
\[
  \tilde{U}_{ij} \define \sum_{k \in [m]: A^k_{ij} > 0} A^k_{ij}\, u_i -
  \sum_{k \in [m]: A^k_{ij} < 0} A^k_{ij}\, \ell_i - A^0_{ij}.
\]
Thus for any $y \in [\ell,u]$ we have $A(y)_{ij} \leq \tilde{U}_{ij}$. This
directly yields:

\begin{lemma}
  For any solution $y \in \R^m$ of~\eqref{MISDP}, we have
  \[
    -\sqrt{\tilde{U}_{ii}\,\tilde{U}_{jj}} \leq A(y)_{ij} \leq \sqrt{\tilde{U}_{ii}\,\tilde{U}_{jj}}
  \]
  for all $i$, $j \in [n]$.
\end{lemma}

Currently this method is implemented for the case in which each entry
$(i,j)$ is contained in exactly on matrix, i.e., there exists
$k = k(i,j) \in [m]$ such that
\[
  A(y)_{ij} = A^k_{ij}\, y_k - A^0_{ij}.
\]
In this case, we have
\[
  \frac{-\sqrt{\tilde{U}_{ii}\,\tilde{U}_{jj}} + A^0_{ij}}{A^k_{ij}} \leq y_k \leq
  \frac{\sqrt{\tilde{U}_{ii}\,\tilde{U}_{jj}} + A^0_{ij}}{A^k_{ij}}.
\]
Unfortunately, this has no effect on our test instances, i.e., the bounds
could never be strengthened across the large testset.

\begin{question}
  Can this be effective on some instance? Should this be applied
  nevertheless? Does it make sense to extend the method for the general
  case in~\eqref{MISDP}?
\end{question}

% -------------------------------------------------------------------------
\subsection{Implications of Trace Constraints}

The following is an extension of an observation in the context of computing
the RIP, see~\cite{GalP16}.

\begin{lemma}
  Consider~\eqref{MISDP-P} in which $\tr(X) \leq \alpha$ has to hold. Then
  \[
    -\tfrac{\alpha}{2} \leq X_{ij} \leq \tfrac{\alpha}{2}
  \]
  holds for all $i$, $j \in [n]$.
\end{lemma}

\begin{proof}
  Since $X$ is positive semidefinite, we have
  \[
    X_{ii} X_{jj} - X_{ij}^2 \geq 0
    \quad\Leftrightarrow\quad
    X_{ij}^2 \leq X_{ii} X_{jj}.
  \]
  We observe that $X_{ii} + X_{jj} \leq \alpha$ and $X_{ii}$,
  $X_{jj} \geq 0$. Thus
  \[
    X_{ii} X_{jj} \leq X_{ii} (\alpha - X_{ii}) = \alpha X_{ii} - X_{ii}^2.
  \]
  Taking the derivative and equating 0 yields a maximal point
  $X_{ii}^\star = \tfrac{\alpha}{2}$. Consequently,
  \[
    X_{ij}^2 \leq X_{ii} X_{jj} \leq \alpha X_{ii}^\star - (X_{ii}^\star)^2 =
    \tfrac{\alpha^2}{2} - \tfrac{\alpha^2}{4} = \tfrac{\alpha^2}{4}.
  \]
  Taking the square root shows the claim.
\end{proof}

\begin{question}
  Should this be transferred to~\eqref{MISDP} and how?
\end{question}

This is also implemented if each matrix entry arises from a single matrix
$A^k$ and $A^0$ is 0 for all involved entries. It does not find a
strengthening for the testset, but it strengthens the bounds for RIP
instances if the bounds have not been tightened before.

% -------------------------------------------------------------------------
\subsection{``Coefficient Tightening''}
\label{sec:CoefficientTightening}

For a linear constraint, coefficient tightening can be explained by the
following example: Consider two binary variables~$x$ and $y$ and the
constraint $x + 2\, y \geq 1$. Then the coefficient of $y$ can be tightened
to $1$ yielding $x + y \geq 1$.

Assume that all matrices $A^k$, $k \in [m]$, are psd and $y_k$ is binary
for $k \in [m]$. Then we can compute
\[
  \hat{\mu}_k = \min\, \{ \mu \geq 0 \suchthat \mu A^k - A^0 \succeq 0\}
\]
adn define $\mu_k \define \min\, \{\hat{\mu}_k, 1\}$. Note that $\mu_k = 1$ if
$\hat{\mu}_k = \infty$, i.e., if the above optimization problem is infeasible.

\begin{lemma}\label{lem:Tightening1}
  Let $k \in [m]$ with $A^k \succeq 0$. Then
  \[
    A^k - A^0 \succeq 0
    \quad\Leftrightarrow\quad
    \mu_k\, A^k - A^0 \succeq 0.
  \]
\end{lemma}

\begin{proof}
  Assume that $A^k - A^0 \succeq 0$. This implies that $\hat{\mu}_k \leq 1$
  and thus $\mu_k = \hat{\mu}_k$. This implies that
  $\mu_k\, A^k - A^0 \succeq 0$ by definition of $\hat{\mu}_k$.

  Conversely, assume that $\mu_k\, A^k - A^0 \succeq 0$. If $\mu_k = 1$, the
  statement is clear, so assume that $\mu_k = \hat{\mu}_k < 1$. Then for
  all $x \in \R^n$:
  \[
    0 \leq x\T \big(\mu_k\, A^k - A^0\big) x = \mu_k\,
    \underbrace{x\T A^k x}_{\geq 0} -
    x\T A^0 x \leq x\T (A^k - A^0) x,
  \]
  which implies that $A^k - A^0 \succeq 0$.
\end{proof}

\begin{lemma}\label{lem:TightenedProblem}
  Let $A^k \succeq 0$ for all $k \in [m]$ and $y \in \R^m$ with
  $y_i \in \{0,1\}$ for all integral variables $i \in I$ and $y_i \geq 0$
  for $i \notin I$. Then
  \[
    A(y) \succeq 0 \quad\Leftrightarrow\quad
    \sum_{k=1}^m \mu_k\, A^k\, y_k - A^0 \succeq 0,
  \]
  where we define $\mu_k = 1$ for $i \notin I$.
\end{lemma}

\begin{proof}
  First assume that
  $\sum_{k=1}^m \mu_k\, A^k\, y_k - A^0 \succeq 0$. Then for every
  $x \in \R^n$
  \[
    0 \leq x\T \bigg(\sum_{k=1}^m \mu_k\, A^k - A^0\bigg) x =
    \sum_{k=1}^m \mu_k\, \underbrace{x\T A^k x}_{\geq 0} -
    x\T A^0 x \leq x\T (\sum_{k=1}^m A^k - A^0) x,
  \]
  which implies that $A(y) \succeq 0$.

  We now assume that $A(y) \succeq 0$. By removing terms with $y_k = 0$ for
  $k \in I$, we can assume that $y_k = 1 $ for all $k \in I$. Thus,
  $\sum_{k=1}^m A^k - A^0 \succeq 0$. If $\mu_k = 1$ for all $k \in [m]$,
  then the statement is directly clear. Therefore assume that there exists
  $k \in I$ with $\mu_k = \hat{\mu}_k < 1$. But then already
  $\mu_k\, A^k - A^0 \succeq 0$. Adding the psd matrices $A^{\ell}$ for
  $\ell \in [m] \setminus \{k\}$ does not change this, which shows the
  claim.
\end{proof}

We now turn to the computation of $\mu_k$.

\begin{lemma}\label{lem:TightenPosDef}
  If $A^k \succ 0$ is positive definite, then
  \[
    \mu_k = \min \big\{ \lambda_{\max}(V^{-\top} A^0 V^{-1}),\, 1\big\}.
  \]
  where $A^k = V\T V$ for some invertible matrix $V \in \R^{n \times n}$.
\end{lemma}

\begin{proof}
  For some matrix $A \in \R^{n \times n}$ and invertible matrix
  $B \in \R^{n \times n}$ we have $A \succeq 0$ if and only if
  $B\T A B \succeq 0$ (see, e.g., Helmberg~\cite[Prop.~1.1.7]{Hel00}).
  Thus, $\mu\, A^k - A^0 \succeq 0$ if and only if
  \[
    0 \preceq V^{-\top} (\mu\, A^k - A^0) V^{-1} = \mu\,
    V^{-\top} A^k V^{-1} - V^{-\top} A^0 V^{-1} = \mu\, I - V^{-\top} A^0 V^{-1}.
  \]
  The eigenvalues of $\mu\, I - B$ for some matrix $B$ are $\mu - \lambda$
  for eigenvalues $\mu$ of $B$.\footnote{This follows, for example, by the
    general theorem that if $g(t) \in \R[t]$ is a scalar polynomial and
    $\lambda_1, \dots, \lambda_n$ are the eigenvalues of $B$, then the
    eigenvalues of $g(B)$ are $g(\lambda_1), \dots, g(\lambda_n)$, see
    Gantmacher~\cite[Ch.~IV, Thm.~3]{Gan59I}. We here use
    $g(t) = \mu - t = \mu\, t^0 - t^1$. (Alternatively, $(\mu I - B)v =
    \lambda v \Leftrightarrow Bv = (\mu - \lambda)v$, so $v$ is an
    eigenvector for eigenvalue $\mu - \lambda$.)}  thus, to make
  $\mu\, I - V^{-\top} A^0 V^{-1}$ psd, we need
  $\mu \geq \lambda_{\max}(V^{-\top} A^0 V^{-1})$.
\end{proof}

If $A^k$ is ``only'' positive semi-definite, one needs other methods that
we discuss now. Note that we want to find the largest $\mu$ such that
$\lambda_{\min}(\mu\, A^k - A^0) = 0$. It is therefore helpful to consider
$f(\mu) = \lambda_{\min}(\mu\, A^k - A^0)$.

The following is inspired by~\cite{Str16,HigSS16} and only stated here for completeness.

\begin{lemma}
  $f(\mu)$ is concave.
\end{lemma}

\begin{proof}
  We have for $\alpha \in [0,1]$ and $\mu_1$, $\mu_2 \geq 0$:
  \begin{align*}
    f\big(\alpha \mu_1 + (1 - \alpha) \mu_2\big)
    & = \lambda_{\min}\big((\alpha \mu_1 + (1 - \alpha) \mu_2) A^k - A^0\big)\\
    & = \lambda_{\min}\big(\alpha (\mu_1\, A^k - A^0) + (1 - \alpha) (\mu_2\, A^k - A^0)\big)\\
    & \geq \lambda_{\min}\big(\alpha (\mu_1\, A^k - A^0)\big) + \lambda_{\min}\big((1 - \alpha) (\mu_2\, A^k - A^0)\big)\\
    & = \alpha\, \lambda_{\min}\big(\mu_1\, A^k - A^0\big) + (1 - \alpha)\, \lambda_{\min}\big(\mu_2\, A^k - A^0\big)\\
    & = \alpha\, f(\mu_1) + (1 - \alpha)\, f(\mu_2),
  \end{align*}
  which shows the claim.
\end{proof}

\noindent
There are at least three choices to compute $\mu_k$:
\begin{itemize}[leftmargin=3ex]
\item One can use bisection within the interval $[0,1]$, checking whether
  $f(\mu) \geq 0$ in each step.
\item One can use Newton's method to compute $\mu_k$ similar
  to~\cite{Str16,HigSS16}.
\item One can solve the generalized eigenvalue problem $\mu A^k x = \lambda
  A^0 x$. It seems to be the case that most algorithms for this problem
  work best when one of the matrices is positive definite. But in this case
  we can apply Lemma~\ref{lem:TightenPosDef}. Otherwise, one can use the
  technique described in~\cite{Str16,HigSS16}. 
\end{itemize}

According to~\cite{Str16,HigSS16}, solving the generalized eigenvalue
problem is fastest, followed by bisection, and then Newton's method. But
note that the problems in~\cite{Str16,HigSS16} are slightly different
($f(\alpha) = \lambda_{\min}(\alpha A^k + (1 - \alpha) A^0)$).

The methods are implemented (using bisection) but none of the instances in
the testset allows a tightening.

\begin{question}
  Can these methods be successful and are they relevant?
\end{question}

% -------------------------------------------------------------------------
\subsection{Bound Tightening}

One can also use the methods in Section~\ref{sec:CoefficientTightening} to
compute bounds on each $y_k$ variable. For this we only need that all $A^k$
are psd, but $y_k$ does not need to be binary. We need the following:

\begin{lemma}
  Let $A^k \succeq 0$ for all $k \in [m]$ and $\ell \in [m]$. Then for
  $y \in \R^m$, $y \geq 0$:
  \[
    A(y) \succeq 0 \quad\Leftrightarrow\quad
    \alpha\, A^{\ell} + \sum_{\substack{k=1\\k \neq \ell}}^m A^k\, y_k - A^0 \succeq 0,
  \]
  where $\alpha = \min\,\{y_{\ell},\mu_{\ell}\}$.
\end{lemma}

\begin{proof}
  Observe that if $\alpha = y_{\ell}$, then the statement is trivially
  true. Thus, let $\alpha = \hat{\mu}_k < y_{\ell}$.

  First assume that
  $\alpha\, A^{\ell} + \sum_{k=1,k \neq \ell}^m \mu_k\, A^k\, y_k - A^0
  \succeq 0$.  Then for every $x \in \R_+^n$
  \begin{align*}
    0 & \leq x\T \bigg(\alpha\, A^{\ell} + \sum_{\substack{k=1\\k \neq \ell}}^m A^k\, y_k - A^0\bigg) x =
        \alpha\, \underbrace{x\T A^{\ell} x}_{\geq 0} + \sum_{\substack{k=1\\k \neq \ell}}^m x\T A^k
        x\, y_k - x\T A^0 x\\
      & \leq y_{\ell}\, x\T A^{\ell} x + \sum_{\substack{k=1\\k \neq \ell}}^m x\T A^k x\, y_k - x\T A^0 x
        = x\T A(y)x,
  \end{align*}
  which implies that $A(y) \succeq 0$.

  We now assume that $A(y) \succeq 0$. By definition of $\hat{\mu}_{\ell}$
  we have $\hat{\mu}_{\ell}\, A^{\ell} - A^0 \succeq 0$. Note that
  $A^k \, y_k$ is psd, since $A^k$ is psd and $y_k \geq 0$. Then adding
  these matrices for all for $k \in [m] \setminus \{\ell\}$ does not change
  the property of being psd, which shows the claim.
\end{proof}

From the lemma, we conclude that one can tighten the upper bounds of all
variables $y_{\ell}$ to $\min\, \{u_{\ell}, \hat{\mu}_{\ell}\}$. This is
implemented, but has not tightened any bound in presolving.

% -------------------------------------------------------------------------
\subsection{Other Presolving Methods}

We collect some more techniques here, all of which are currently not implemented.

\paragraph{Removing 0 rows/columns}

If for some $i \in [m]$ we have $A_{ij}^k = 0$ for all $j \in [n]$ and
$k \in [m]_0$, then row/column $i$ is not used and can be removed from the
matrices $A^k$.

\paragraph{Dual Fixing}

If for some $k \in [m]$ the matrix $A^k$ is psd and disjoint from the rest
(i.e., $A^k$ has no common nonzero with the other matrices), depending on
the objective coefficient one can fix $y_k$ to its upper bound.

% -------------------------------------------------------------------------
% -------------------------------------------------------------------------
\section{Useless or Already Implied Methods}

% -------------------------------------------------------------------------
\subsection{Implications of Primal Bounds}

The implications in the following results are automatically exploited for
linear constraints in problems of the form~\eqref{MISDP-P}. The argument
is standard argument, but we repeat it to see whether it can be exploited
somewhere.

\begin{lemma}
  Let $A \in \R^{n \times n}$, $\beta \in \R$ and define
  \[
    \mathcal{X} \define \{ X \in \R^{n\times n} \suchthat \skal{A}{X} \leq
    \beta,\; L_{ij} \leq X_{ij} \leq U_{ij}\; \forall (i,j) \in [n]^2\}.
  \]
  Then for any $X \in \mathcal{X}$ the inequality
  \[
    X_{st} \leq \frac{\beta - \alpha_{st}}{A_{st}}
  \]
  holds for all $(s,t) \in [n]^2$, where
  \begin{align*}
    & \alpha_{st} \define \sum_{(i,j) \in S^-_{st}} A_{ij} u_{ij} - \sum_{(i,j) \in S^+_{st}} A_{ij} \ell_{ij},\\
    & S^+_{st} = \{(i,j) \in [n]^2 \setminus \{(s,t)\} \suchthat A^k_{ij} > 0\},\\
    & S^-_{st} = \{(i,j) \in [n]^2 \setminus \{(s,t)\} \suchthat A^k_{ij} < 0\}.
  \end{align*}
\end{lemma}

\begin{proof}
  For any~$X \in \mathcal{X}$ we have
  \[
    \sum_{(i,j) \in [n]^2} A_{ij} X_{ij} \leq \beta
    \quad\Leftrightarrow\quad A_{st} X_{st} \leq \beta - \sum_{(i,j) \in [n]^2
      \setminus \{(s,t)\}} A_{ij} X_{ij}.
  \]
  Then
  \[
    A_{st} X_{st} \leq \beta - \sum_{(i,j) \in S^-_{st}} A_{ij} u_{ij} - \sum_{(i,j) \in S^+_{st}} A_{ij} \ell_{ij}.
  \]
  This yields the claim.
\end{proof}

%-------------------------------------------------------------------------
\subsection{Gershgorin-based Presolving}

Gershgorin's circle theorem says that each eigenvalue of a matrix
$A \in \R^{n \times n}$ lies in (at least) one of the Gershgorin discs
\[
  D_i \define \Big\{z \in \C \suchthat \abs{z - a_{ii}} \leq \sum_{j=1, j \neq
    i}^n \abs{a_{ij}} \Big\}.
\]
Thus, given a vector $y \in \R^m$, for $A(y) \succeq 0$ to hold, we need
that
\[
  \bigcup_{i=1}^n \bigg\{z \in \C \suchthat \abs{z - A(y)_{ii}} \leq \sum_{j=1, j \neq
    i}^n \abs{A(y)_{ij}}\bigg\} \cap \{z \in \C \suchthat \Re(z) \geq 0\}
  \neq \varnothing.
\]
But since we need to have $A(y)_{ii} \geq 0$ for all $y$ with
$A(y) \succeq 0$, the above condition is always fulfilled: The center
points of the circles are always in the nonnegative complex halfplane.

% -------------------------------------------------------------------------

% -------------------------------------------------------------------------
\subsection{SOCP-inspired Presolving}

Consider a second order cone constraint $(x,t) \in L_{n+1}$, where
\[
  L_{n+1} \define \Big\{x \in \R^{n+1} \suchthat \sqrt{x_1^2 + \dots + x_n^2}
  \leq x_{n+1}\Big\}
\]
is the Lorentz cone.  Presolving/propagation of
$\sqrt{x_1^2 + \dots + x_n^2} \leq t$ with bounds $\ell \leq x \leq u$
yields for some $k \in [n]$:
\begin{equation}\label{eq:SOCPpre}
  x_k^2 \leq t^2 - \sum_{\substack{i=1\\ i \neq k}}^n x_i^2 \leq t^2 -
  \sum_{\substack{i=1\\ i \neq k}} \gamma_i \leq t^2,\qquad
  \gamma_i \define \min_{\gamma \in [\ell_i, u_i]} \gamma^2.
\end{equation}
Note that $\ell_i < 0 < u_i$ is possible. One main application is the case
in which there is an upper bound on~$t$. We would like to have a similar
presolving (propagation) for SDPs -- at least if we would write an SOCP as
an SDP, but possibly for more general cases.

Indeed, SOCP is special case of SDP, because:
\[
  (x,t) \in L_{n+1} \quad\Leftrightarrow\quad
  \tilde{X} \define
  \begin{pmatrix}
    t & x\T\\
    x & I_n
  \end{pmatrix}
  \succeq 0.
\]
To see this, we need the following result adapted
from~\cite[Section~A.5.5]{BoyV09}. Let
\[
  D = \begin{pmatrix}
    A & B\T\\
    B & C
  \end{pmatrix}
\]
with a positive definite $C \succ 0$. Then $D \succeq 0$ if and only if
$S = A - B\T C^{-1} B \succeq 0$ and $A \succeq 0$\footnote{Note that the
  condition $A \succeq 0$ is missing in~\cite{BoyV09}.}. For $D = \tilde{X}$, we
get: $\tilde{X} \succeq 0$ if and only if $t \geq 0$, $t - x\T x \geq
0$. The latter is equivalent to $(x,t) \in L_{n+!}$.

Now consider a more general matrix
\begin{equation}\label{eq:GeneralPSD}
  \begin{pmatrix}
    t & x\T\\
    x & X
  \end{pmatrix}
  \succeq 0
\end{equation}
with an upper bound $\bar{t}$ on $t$. This seems to be the most natural
generalization of the SOCP case.

Taking $2 \times 2$ minors for some $i \in [n]$, we get
$t X_{ii} - x_i^2 \geq 0$. Thus, $x_i^2 \leq \bar{t} X_{ii}$, which leads
to
\[
  - \sqrt{\bar{t} X_{ii}} \leq x_i \leq \sqrt{\bar{t} X_{ii}}.
\]
Note that this is covered by the bounds considered in
Section~\ref{sec:2by2Minors}. In the SOCP case, we obtain
$x_i \leq \sqrt{\bar{t}}$, which only provides the weakest upper bound
in~\eqref{eq:SOCPpre}.

Thus, one question is whether one can derive strengthened bounds. This is
currently unclear. As one idea, one can consider generalized Schur
polynomials (since $X$ is not necessarily positive definite, one needs the
generalization), which we adapt from~\cite[Section~A.5.5]{BoyV09}: Consider
again the matrix $D$ from above. Then
\[
  D \succeq 0 \quad\Leftrightarrow\quad
  A \succeq 0,\; (I - AA^{+})B\T = 0,\; C - B A^{+} B\T \succeq 0,
\]
where $A^+$ is the pseudo-inverse of $C$. Reversing the roles of $A$ and
$C$ we get:
\[
  D \succeq 0 \quad\Leftrightarrow\quad
  C \succeq 0,\; (I - CC^{+})B = 0,\; A - B\T C^{+} B \succeq 0.
\]
Applied to~\eqref{eq:GeneralPSD}, we obtain from the first:
\begin{equation}\label{eq:GenSchur1}
  t \geq 0,\; (1 - t t^{-1}) x\T = 0,\; X - x t^{-1} x\T \succeq 0,\; 
\end{equation}
if $t > 0$ and from the second:
\begin{equation}\label{eq:GenSchur2}
  X \succeq 0,\; (I - XX^{+})x = 0,\; t - x\T X^{+} x \succeq 0.
\end{equation}

Using
$X - \tfrac{1}{t} x x\T \succeq 0 \Leftrightarrow t X - x x\T \succeq 0$
from~\eqref{eq:GenSchur1}, we get for the $2 \times 2$-minors with $i$,
$j \in [n]$, $i \neq j$:
\begin{align*}
  & (t\, X_{ii} - x_i^2) (t\, X_{jj} - x_j^2) -  (t\, X_{ij} - x_i\, x_j)^2 \geq 0
  \\
  \Leftrightarrow\quad & t \, X_{ij} - x_i\, x_j \leq \sqrt{t\, X_{ii} - x_i^2} \sqrt{t\, X_{jj} - x_j^2}
  \\
  \Leftrightarrow\quad & X_{ij} \leq \frac{1}{t} \bigg(\sqrt{t\, X_{ii} - x_i^2}
                         \sqrt{t\, X_{jj} - x_j^2} + x_i\, x_j\bigg).
\end{align*}
This, for instance, yields
\[
  X_{ij} \leq \sqrt{U_{ii}} \sqrt{U_{jj}} + \tfrac{1}{t} x_i\, x_j.
\]
which does not seem to help. If we would know that $X_{ij} = x_i\, x_j$,
i.e., $X$ is rank-1, then we obtain
\[
  X_{ij} \leq \frac{\sqrt{t\, X_{ii} - x_i^2} \sqrt{t\, X_{jj} - x_j^2}}{t-1}.
\]

\begin{question}
  Can any of this be made useful?
\end{question}

% -------------------------------------------------------------------------
\section{Specialized Arguments for Computing the RIP}

\subsection{Sparse Principal Component Analysis}

Consider the sparse principal component analysis (PCA) problem
\[
  \max\; \{\Norm{Ax}{2}^2 \suchthat \Norm{x}{2}^2 = 1,\; \Norm{x}{0} \leq k\}. 
\]
with $A \in \R^{m \times n}$.  Let $x^*$ be an optimal solution and
$S = \supp(x^*)$ be its support with $k \define \Norm{x}{0}$. Consider the
submatrix~$A_S \in \R^{m \times k}$ indexed by columns in $S$. Then
$\tilde{A} = A_S\T A_S^{\phantom{T}} \in \R^{k \times k}$ is symmetric positive
semidefinite. By the Rayleigh-Ritz theorem (see
Helmberg~\cite[Thm.~A.0.4]{Hel00}) we have
\[
  \max_{y \in \R^k} \; \{\Norm{\tilde{A}y}{2}^2 \suchthat \Norm{y}{2}^2 =
  1\} = \lambda_{\max}(\tilde{A}),
\]
i.e., the sparse PCA problem looks for a $k$-sparse vector yielding the
largest eigenvalue of the supporting submatrix. A similar statement holds
for minimization variant and the minimal eigenvalue.

A formulation for the sparse PCA problem is
\begin{subequations}\label{eq:SPCA}
  \begin{align}
    \max \quad & \skal{A\T A}{X} \\
    \text{s.t.} \quad & \tr(X) = 1, \\
    & -z_j \leq X_{ij} \leq z_j \quad \text{for } i,\; j \in [n],\label{eq:SPCAcouple} \\
    & \sum_{i=1}^n z_i \leq k,\\
    & X \succeq 0, \\
    & z \in \{0,1\}^n.
  \end{align}
\end{subequations}
In~\cite{GalP16} (and~\cite{LiX20}) it is proved that there exists an
optimal solution rank-1 solution~$X^*$. Thus, $X^* = x^* (x^*)\T$ for some
$x^* \in \R^n$ with $\Norm{x^*}{0} \leq k$. Let $S = \supp(x^*)$. Then
$x^*_S$ is an eigenvector for a maximal eigenvalue of
$A_S\T A_S^{\phantom{T}}$.

\begin{question}
  Since $X_{ii} = 0$ implies $X_{ij} = 0$ for all $j \in [n]$, it would
  suffice to only consider the inequalities $-z_i \leq X_{ii} \leq x_i$ for
  all $i \in [n]$ in~\eqref{eq:SPCAcouple}. Can we detect that the other
  inequalities are implied and remove them? Does this speed up the solution
  process?
\end{question}

We apply the Perron-Frobenius theorem. To this end, we define a matrix $A$ to
be reducible if there exists a permutation matrix $P$ such that
\[
  P\T A P = \begin{pmatrix} B & C \\ 0 & D \end{pmatrix}.
\]
Otherwise it is \emph{irreducble}. We have the following:

\begin{theorem}[{Frobenius Theorem, see Gantmacher~\cite[Chapter 2, Thm.~2]{Gan59II}}]\label{thm:PerronFrobenius}
  An irreducible nonnegative matrix $A$ has a positive maximal eigenvalue
  $\lambda > 0$. The corresponding eigenvectors have positive entries.
\end{theorem}

For general (i.e., also reducible) matrices the following holds:
\begin{theorem}[{Gantmacher~\cite[Chapter 2, Thm.~3]{Gan59II}}]\label{thm:PerronFrobeniusNonnegative}
  A nonnegative matrix $A$ has a nonnegative maximal eigenvalue
  $\lambda \geq 0$. The corresponding eigenvectors have nonnegative
  entries.
\end{theorem}

Moreover, Gantmacher proves the following.
\begin{lemma}[{\cite[Chapter 2, Remark 3]{Gan59II}}]\label{lemma:PFNegative}
  An irreducible nonnegative matrix $A$ cannot have two linearly
  independent nonnegative eigenvectors.
\end{lemma}
This implies that all eigenvectors to all other eigenvalues cannot be nonnegative!
For oscillaritory matrices some more can be said about the signs, see
\cite[Chapter 2, Theorem~13]{Gan59II}.

\begin{lemma}
  If $A_S \geq 0$, then $x^* \geq 0$.
\end{lemma}

\begin{proof}
  Since $A_S \geq 0$, Theorem~\ref{thm:PerronFrobeniusNonnegative} implies
  that $A_S$ has a maximal eigenvalue with corresponding \emph{nonnegative}
  eigenvector. This eigenvector is unique up to scaling.
\end{proof}

Thus, one can restrict $X \geq 0$ and write
\[
  0 \leq X_{ij} \leq z_j \quad \text{for } i,\; j \in [n]
\]
instead of~\eqref{eq:SPCAcouple}.

\paragraph{A Valid SOCP Constraint}

Li and Xie~\cite{LiX20} show that the following inequalitiy is valid for
sparse PCA:
\[
  \sum_{j \in [n]} X_{ij}^2 \leq X_{ii}\, z_i \quad\text{for all } i \in [n].
\]
This holds, since in this case $X = x x\T$ and $\sum_{j=1}^m x_j^2 =
1$. Thus,
\[
  \sum_{j \in [n]} X_{ij}^2 \leq \sum_{j \in [n]} x_i^2\, x_j^2 = x_i^2
  \leq z_i\, X_{ii}.
\]
Note that these inequalities can be reformulated by subtracting $X_{ii}^2$
from both sides to yield
\[
  \sum_{\substack{j \in [n]\\ j \neq i}} X_{ij}^2 \leq z_i\, (X_{ii} - X_{ii}^2).
\]

These inequalities are SOCP representable as
\begin{align*}
  & (X_{i1}, \dots, X_{in}, \tfrac{X_{ii} - z_i}{2}, \tfrac{X_{ii} + z_i}{2}) \in L_{n+2},\\
  \Leftrightarrow\quad & \sum_{j=1}^n X_{ij}^2 + \tfrac{1}{4}(X_{ii}^2 - 2\,
                    z_i\, X_{ii} + z_i^2) \leq \tfrac{1}{4}(X_{ii}^2 + 2\, z_i\, X_{ii} + z_i^2)\\
  \Leftrightarrow\quad & \sum_{j=1}^n X_{ij}^2 \leq z_i\, X_{ii}.
\end{align*}
Formulated as an SDP, we obtain
\[
  \begin{pmatrix}
    \tfrac{X_{ii} + z_i}{2} & X_{i1} & \dots & X_{in} & \tfrac{X_{ii} - z_i}{2}\\
    X_{i1} & 1 &  & 0 & 0\\
    \vdots & & \ddots & & \vdots\\
    X_{in} & &    & 1 & 0\\
    \tfrac{X_{ii} - z_i}{2} & & \dots & 0 & 1
  \end{pmatrix}
  \succeq 0,
\]
for all $i \in [n]$.  These SDP blocks can in principle be added
to~\eqref{eq:SPCA}. Alternatively, one can encode the weaker inequality
$X_{ij}^2 \leq z_i\, X_{ii}$ as
\[
  \begin{pmatrix}
    z_i & X_{ij}\\
    X_{ij} & X_{ii}
  \end{pmatrix}
  \succeq 0,
\]
but this would require quadratically many SDP-constraints.

Moreover, since $X_{ii} \leq x_i$, we have the weaker
$\sum_{j=1}^n X_{ij}^2 \leq z_i^2$, which is equivalent to
\[
  \begin{pmatrix}
    z_i & X_{i1} & \dots & X_{in} \\
    X_{i1} & 1 &  & 0 \\
    \vdots & & \ddots & \vdots \\
    X_{in} & &    & 1 \\
  \end{pmatrix}
  \succeq 0.
\]

\begin{question}
  Can these inequalities be useful within SCIP-SDP?
\end{question}

\begin{question}
  Is there any inequality that exploits the fact that $z_i^2 = z_i$ for
  feasible solutions?
\end{question}

\subsection{Minimization Variant}

Now consider the lower problem
\[
  \min_{y \in \R^k} \; \{\Norm{\tilde{A}y}{2}^2 \suchthat \Norm{y}{2}^2 =
  1\} = \lambda_{\min}(\tilde{A}).
\]
Unfortunately, by Lemma~\ref{lemma:PFNegative}, a vector for an eigenvalue
that is not equal to the (unique) maximal eigenvalue always has positive
and negative entries.

Nevertheless if the matrix $A\T A$ is \emph{positive definite}, we get
solve the maximization version in~\eqref{eq:SPCA} where $A\T A$ is replaced
by $(A\T A)^{-1}$. This holds since $\lambda$ is an eigenvalue of $A$ if
and only if $1/\lambda$ is an eigenvalue of $A^{-1}$. This transformation
might have the advantage that the maximization version is sometimes easier
to solve, but this might only hold for nonnegative matrices $A$. Note,
however, that $(A\T A)^{-1}$ is very unlikely to be nonnegative, so we
cannot benefit from the application of the Perron-Frobenius theorem here.

\begin{question}
  Can the transformation help to speed up the solution process? How often
  are the matrices positive definite?
\end{question}

\begin{small}
  \bibliographystyle{abbrv}
  \bibliography{MISDPpresolving.bib}
\end{small}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2020-09-01 tex/MISDP-Presolving/MISDPpresolving.tex
fix some errors in presolving document

@@
   are fixed? For example if we branch variables in~\eqref{MISDP} such that
-  a diagonal entry becomes 0, are the other entries in the column and row
+  a diagonal entry of~$A(y)$ becomes 0, are the other entries in the column and row
   fixed to 0 as well?
@@
   \[
-    -\sqrt{U_{ii}\, U_{jj}} \leq X_{ij} \leq \sqrt{U_{ij}\, U_{jj}}
+    -\sqrt{U_{ii}\, U_{jj}} \leq X_{ij} \leq \sqrt{U_{ii}\, U_{jj}}
   \]
@@
 \[
-  \tilde{U}_{ij} \define \sum_{k \in [m]: A^k_{ij} > 0} A^k_{ij}\, u_i -
-  \sum_{k \in [m]: A^k_{ij} < 0} A^k_{ij}\, \ell_i - A^0_{ij}.
+  \tilde{U}_{ij} \define \sum_{k \in [m]: A^k_{ij} > 0} A^k_{ij}\, u_k -
+  \sum_{k \in [m]: A^k_{ij} < 0} A^k_{ij}\, \ell_k - A^0_{ij}.
 \]
@@
 Currently this method is implemented for the case in which each entry
-$(i,j)$ is contained in exactly on matrix, i.e., there exists
+$(i,j)$ is contained in exactly one matrix, i.e., there exists
 $k = k(i,j) \in [m]$ such that
@@
   Thus, $\mu\, A^k - A^0 \succeq 0$ if and only if
-  \[
-    0 \preceq V^{-\top} (\mu\, A^k - A^0) V^{-1} = \mu\,
-    V^{-\top} A^k V^{-1} - V^{-\top} A^0 V^{-1} = \mu\, I - V^{-\top} A^0 V^{-1}.
-  \]
+  \begin{align*}
+    0 \preceq (V^{-1})^\top (\mu\, A^k - A^0) V^{-1} &= \mu\,
+    (V^{-1})^\top A^k V^{-1} - (V^{-1})^\top A^0 V^{-1} \\
+    &= \mu\, I - (V^{-1})^\top A^0 V^{-1}.
+  \end{align*}
   The eigenvalues of $\mu\, I - B$ for some matrix $B$ are $\mu - \lambda$
-  for eigenvalues $\mu$ of $B$.\footnote{This follows, for example, by the
+  for eigenvalues $\lambda$ of $B$.\footnote{This follows, for example, by the
     general theorem that if $g(t) \in \R[t]$ is a scalar polynomial and
@@
   \]
-  where $\alpha = \min\,\{y_{\ell},\mu_{\ell}\}$.
+  where $\alpha = \min\,\{y_{\ell},\hat{\mu}_{\ell}\}$.
 \end{lemma}
@@
   First assume that
-  $\alpha\, A^{\ell} + \sum_{k=1,k \neq \ell}^m \mu_k\, A^k\, y_k - A^0
-  \succeq 0$.  Then for every $x \in \R_+^n$
+  $\alpha\, A^{\ell} + \sum_{k=1,k \neq \ell}^m \hat{\mu}_k\, A^k\, y_k - A^0
+  \succeq 0$.  Then for every $x \in \R^n$
   \begin{align*}
@@
 \]
-where $A^+$ is the pseudo-inverse of $C$. Reversing the roles of $A$ and
+where $A^+$ is the pseudo-inverse of $A$. Reversing the roles of $A$ and
 $C$ we get:
@@
 In~\cite{GalP16} (and~\cite{LiX20}) it is proved that there exists an
-optimal solution rank-1 solution~$X^*$. Thus, $X^* = x^* (x^*)\T$ for some
+optimal rank-1 solution~$X^*$. Thus, $X^* = x^* (x^*)\T$ for some
 $x^* \in \R^n$ with $\Norm{x^*}{0} \leq k$. Let $S = \supp(x^*)$. Then
@@
 
-Nevertheless if the matrix $A\T A$ is \emph{positive definite}, we get
+Nevertheless if the matrix $A\T A$ is \emph{positive definite}, we can
 solve the maximization version in~\eqref{eq:SPCA} where $A\T A$ is replaced

### 2020-09-01 RIP-MISDP-Paper/scripts/RandomInstancesMISDP.tex
Add documentation of generation of random instances (general MISDP and RIP)

\documentclass[11pt,a4paper]{article}

\usepackage[T1]{fontenc}
\usepackage[american]{babel}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{dsfont}
\usepackage{mathtools}
\usepackage{booktabs}
\usepackage{url}
\usepackage[inline]{enumitem}
\usepackage{xcolor}
\usepackage{todonotes}
\usepackage{graphicx}
\usepackage{float}
\usepackage{tikz}
\usepackage{etoolbox}
% \usepackage{enumerate}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[top=35mm,bottom=30mm,left=30mm,right=30mm]{geometry}
\usepackage[breaklinks,colorlinks,citecolor=blue,linkcolor=blue,
  pdftitle={Random MISDP Instances},pdfauthor={Frederic Matter}]{hyperref}

\usepackage[breaklinks,colorlinks,citecolor=blue,linkcolor=blue]{hyperref}
\usepackage{multirow}
\usepackage{xspace}         % must be loaded last

% % number types
\newcommand{\R}{\mathds{R}}
\newcommand{\C}{\mathds{C}}
\newcommand{\Q}{\mathds{Q}}
\newcommand{\N}{\mathds{N}}
\newcommand{\Z}{\mathds{Z}}
\newcommand{\E}{\mathds{E}}
\newcommand{\ones}{\mathds{1}}

\newcommand{\classP}{\mathcal{P}}
\newcommand{\NP}{\mathcal{N}\mathcal{P}}

% % general commands
\newcommand{\define}{\coloneqq}
\newcommand{\enifed}{\eqqcolon}
\newcommand{\bs}{\backslash}
\newcommand{\card}[1]{\lvert{#1}\rvert}
\newcommand{\abs}[1]{\lvert{#1}\rvert}
\newcommand{\norm}[1]{\lVert{#1}\rVert}
\newcommand{\sym}{\mathcal{S}}
\newcommand{\psd}{\mathcal{S}_{+}}
\newcommand{\pd}{\mathcal{S}_{++}}
\newcommand{\suchthat}{\, : \,}
\newcommand{\T}{^\top}
\newcommand{\ceil}[1]{\lceil{#1}\rceil}
\newcommand{\floor}[1]{\lfloor{#1}\rfloor}
\newcommand{\size}[1]{\langle{#1}\rangle}
\newcommand{\sprod}[2]{{#1} \bullet {#2}}
\newcommand{\charac}[1]{\ensuremath{\chi^{#1}}}
\newcommand{\encode}[1]{\langle #1 \rangle}
\renewcommand{\L}{\mathcal{L}}
\newcommand{\A}{\mathcal{A}}
\renewcommand{\Re}{\textnormal{Re}}
\renewcommand{\Im}{\textnormal{Im}}
\newcommand{\id}{\mathds{I}} % identity matrix (since $I$ collides with
% index set $I$)
\newcommand{\nullspace}{\textnormal{null}} % null space
\newcommand{\sphere}{\mathds{S}} % unit sphere
\newcommand{\ball}{\mathds{B}}   % unit ball
\newcommand{\prob}{\mathds{P}}   % probability
\newcommand{\dc}{\mathcal{T}}    % descent cone
\newcommand{\normal}{\mathcal{N}}

% math operators
\DeclareMathOperator{\conv}{conv}
\DeclareMathOperator{\cone}{cone}
\DeclareMathOperator{\supp}{supp}
\DeclareMathOperator{\dist}{dist}
\DeclareMathOperator{\lin}{lin}
\DeclareMathOperator{\aff}{aff}
\DeclareMathOperator{\poly}{poly}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator{\rank}{rank}
\DeclareMathOperator{\sign}{sign}
\DeclareMathOperator{\spark}{spark}
\DeclareMathOperator{\tr}{tr}
\DeclareMathOperator{\linspan}{span}
\DeclareMathOperator{\diag}{diag}
\DeclareMathOperator{\blkdiag}{blkdiag}
\DeclareMathOperator{\unif}{Unif}

% useful environments
\newenvironment{dense_itemize}{%
  \begin{list}{$\circ$}%
    {\setlength{\topsep}{1mm}%
      \setlength{\partopsep}{0mm}%
      \setlength{\parskip}{0mm}%
      \setlength{\parsep}{0mm}%
      \setlength{\itemsep}{0mm}%
      \setlength{\labelwidth}{4mm}%
      \setlength{\leftmargin}{0mm}%
      \addtolength{\leftmargin}{\labelwidth}%
      \addtolength{\leftmargin}{\labelsep}%
      \setlength{\itemindent}{0mm}}}%
  {\end{list}}

% % alternative definitions for enumerate, itemize, etc.
% \setlist[itemize]{label=$\triangleright$,leftmargin=2ex,itemsep=0ex,parsep=0ex,topsep=0.5ex}
% \setlist[enumerate]{leftmargin=3ex,itemsep=0ex,parsep=0.25ex,topsep=0.5ex}
% \setlist[description]{labelwidth=0pt,leftmargin=30pt,itemindent=\dimexpr-20pt-\labelsep\relax}

\newtheorem{lemma}{Lemma}
\newtheorem{proposition}[lemma]{Proposition}
\newtheorem{corollary}[lemma]{Corollary}
\newtheorem{theorem}[lemma]{Theorem}
\newtheorem{corrollary}[lemma]{Corrollary}
\newtheorem{remark}[lemma]{Remark}
\newtheorem{conjecture}[lemma]{Conjecture}
\newtheorem*{questions}{Questions}
\theoremstyle{definition}
\newtheorem{definition}[lemma]{Definition}
\newtheorem{observation}[lemma]{Observation}
\newtheorem{fact}[lemma]{Fact}
\newtheorem*{note}{Note}
\newtheorem{example}[lemma]{Example}

% TO DO's
% \renewcommand{\todo}{\footnote}
\newcommand{\ter}[1]{\textcolor{red}{#1}}

% remove indentation at beginning of section
\makeatletter
\patchcmd{\@startsection}
  {\@afterindenttrue}
  {\@afterindentfalse}
  {}{}
\makeatother

\title{Documentation of Scripts for Generating Random MISDP Instances}

\author{Frederic Matter}

\date{\today}

\begin{document}

\maketitle

This document describes the various routines used for generating random
instances of general (dual) MISDP, see Section~\ref{sec:RandomMISDP} and
random instances for computing the RIP, see Section~\ref{sec:RandomRIP}.

\section{General Random (Dual) MISDP}
\label{sec:RandomMISDP}
We consider the following general (dual) form of an MISDP:
\begin{align}\label{eq:DualMISDP}
  \max\, \{b\T y \suchthat \sum_{k=1}^m A^{(k)}y_k - A^{(0)} \succeq 0,\; y
  \in \{0,1\}^{m_b} \times \R^{m_c}\},
\end{align}
where~$m = m_b + m_c$ and~$A^{(k)} \in \sym^n$ for~$0\leq k\leq m$.

Kobayashi and Takano~\cite{Kob20} create random instances
of~\eqref{eq:DualMISDP} as follows:
\begin{align*}
  y_k &\sim \unif(\{0,1\}), \, k \leq m_b,\\
  y_k &\sim \unif([0,1]),\, k > m_b, \\
  A^{(k)}_{i,j} &\sim \unif([-1,1]),\, 1\leq j \leq i \leq n, k \in [m],\\
  A^{(0)} &= \sum_{i=1}^m A^{(k)}y_k - \alpha \id,\\
  b_k &= A^{(k)}\bullet \id, \; k \in [m],
\end{align*}
so that~$\sum_{k=1}^m A^{(k)}y_k - A^{(0)} = \alpha\id \succeq 0$
for~$\alpha \geq 0$. They use the following parameters:
\begin{table}[h]
  \centering
  \begin{tabular}{cccc}
    $n$ & $m_c$ & $m_b$ & $\alpha$ \\
    \toprule
    30 & 30 & 30 & $\{1.0,0.1,10\}$ \\
    45 & 45 & 45 & 1.0 \\
    60 & 60 & 60 & 1.0 \\
    $\{15,60\}$ & 30 & 30 & 1.0 \\
    30 & $\{15,60\}$ & 30 & 1.0 \\
    30 & 30 & $\{15,60\}$ & 1.0
  \end{tabular}
\end{table}

The
script~\textsf{SCIPSDP/tools/InstanceGeneration/generateRandomMISDP.prl}
takes as input~$(n,m_b,m_c,\alpha)$ and generates the
instance~\textsf{randomMISDP\_n\_mb\_mc\_alpha.cbf} of the
form~\eqref{eq:DualMISDP} in the CBF-format,
see~\url{http://cblib.zib.de/}.

The
script~\textsf{SCIPSDP/tools/InstanceGeneration/generateRandomMISDP-PSD.prl}
generates random instances of~\eqref{eq:DualMISDP} where all
matrices~$A^{(k)}$ are positive semidefinite. This is ensured by
using~$A^{(k)} = aa\T$, where~$a_i \sim \unif([-1,1])$.

\section{MISDP for RIP}
\label{sec:RandomRIP}

The RIP can be computed by computing~$\alpha_k$ and~$\beta_k$ defined as
\begin{align}
  \alpha_k &= \min\, \{\norm{Ax}_2^2 \suchthat \norm{x}_2^2 = 1, \;
             \norm{x}_0 \leq k\}, \label{eq:RIPalpha} \\
  \beta_k &=  \max\, \{\norm{Ax}_2^2 \suchthat \norm{x}_2^2 = 1, \;
             \norm{x}_0 \leq k\}. \label{eq:RIPbeta}
\end{align}
Note that the problem for computing~$\beta_k$ is also called sparse PCA (SPCA).
The exact MISDP-formulation reads
\begin{align}\label{eq:MISDPRIP}
  \max/\min\, \Big\{\tr(A\T AX) \suchthat
  &\tr(X) = 1, \, -z_i \leq X_{ii} \leq
  z_i,\, -\tfrac{1}{2}z_j \leq X_{ij} \leq \tfrac{1}{2}z_j, \text{ for } i
    \neq j, \\
  &\sum_{i=1}^k z_i\leq k,\, X \succeq 0,\, z_i \in \{0,1\}\Big\}.
\end{align}
Note that there always is an optimal solution~$X \succeq 0$ that has
rank~$1$, so that no rank-$1$-constraint is needed in the
MISDP-formulation, see~\cite[Theorem 4]{GalP16}. Moreover, for the
validity of the stronger bounds~$-\tfrac{1}{2}z_j \leq X_{ij} \leq
\tfrac{1}{2}z_j$ for $i \neq j$, see~\cite[Lemma 1]{GalP16}.

Kobayashi and Takano~\cite{Kob20} create random instances
with~$A_{i,j} \sim \normal(0,1)$. They use the following parameters with
three instances per each triple~$(m,n,k)$:
\begin{table}[h]
  \centering
  \begin{tabular}{ccc}
    $m$ & $n$ & $k$ \\
    \toprule
    10 & $\{10,15,20,25,30,35\}$ & $\{3,5\}$
  \end{tabular}
\end{table}

Li and Xie~\cite{LiX20} propose two valid inequalities for strengthening the MISDP:
\begin{align*}
  \sum_{j=1}^n X_{ij}^2 &\leq X_{ii}z_i,\quad \forall \; i \in [n],\\
  \big(\sum_{j=1}^n \abs{X_{ij}}\big)^2 &\leq kX_{ii}z_i, \quad \forall \;
                                          i \in [n].
\end{align*}
Both inequalities are SOC-representable and can thus be formulated as an SDP-constraint:
\begin{align}\label{eq:ValidSOCPineqSDP}
  \sum_{j=1}^n X_{ij}^2 \leq X_{ii}z_i \; \Leftrightarrow \;
  \begin{pmatrix}
    \tfrac{X_{ii}+z_I}{2} & X_{i1} & \cdots & X_{in} & \tfrac{X_{ii}-z_i}{2}
    \\
    X_{i1} & 1 & & 0 & 0 \\
    \vdots &\vdots & \ddots & \vdots & \vdots \\
    X_{in} & 0 & & 1 & 0 \\
    \tfrac{X_{ii}-z_i}{2} & 0 & \cdots & 0 & 1
  \end{pmatrix} \succeq 0.
\end{align}
This gives rise to the following weaker inequalities:
\begin{align*}
  \sum_{j=1}^n X_{ij}^2 \leq z_i^2 \; \Leftrightarrow \;
  \begin{pmatrix}
    z_i & X_{i1} & \cdots & X_{in} \\
    X_{i1} & 1 & & 0 \\
    \vdots &\vdots & \ddots & \vdots \\
    X_{in} & 0 & & 1
  \end{pmatrix} \succeq 0, \qquad
                   X_{ij}^2 \leq z_iX_{ii}  \; \Leftrightarrow \;
                   \begin{pmatrix}
                     z_i & X_{ij} \\
                     X_{ij} & X_{ii}
                   \end{pmatrix} \succeq 0,
\end{align*}
where the second inequality can be added for all~$0\leq j < i\leq n$.

The
script~\textsf{SCIPSDP/tools/InstanceGeneration/generateRandomRIP-MISDP.prl}
takes as input~$(m,n,k)$ and generates the
instance~\textsf{randomRIP\_m\_n\_k.cbf}, which is the cbf-file for the
MISDP-formulation of~$\alpha_k$, i.e., the lower bound for the RIP.

\begin{table}
  \begin{center}
    \begin{tabular}[h]{l c}
      0+-1 & $A_{ij} = 0$ (probability~$\tfrac{2}{3}$) or
             $A_{ij} = \pm \sqrt{3/m}$ (probability~$\tfrac{1}{6}$) \\
      band & $A_{ij}\sim$ Unif($\{0,1\}$), band matrix with specified bandwidth \\
      bern & $A_{ij}\sim$ Unif($\{\pm \sqrt{}\}$) \\
      bina & $A_{ij}\sim$ Unif($\{0,1\}$) \\
      norm & $A_{ij}\sim$ $\mathcal{N}(0,1)$ \\
      rnk1 & $A = aa^\top$ $a_i \sim
             \mathcal{N}(0,1)$ \\
      wish & $A_{ij}\sim$ $\mathcal{N}(0,\tfrac{1}{m})$
    \end{tabular}
  \end{center}
  \caption{The seven types of random matrices used in~\cite{GalP16}.}
  \label{tab:MatrixTypesRIP}
\end{table}

Gally and Pfetsch~\cite{GalP16} use seven types of random matrices for the
computation of the RIP, which are summarized in
Table~\ref{tab:MatrixTypesRIP}. Additionally, they test the exact
MISDP-formulation~\eqref{eq:MISDPRIP} against two relaxations used
by~\cite{dAspEtAl08,dAspEtAl07}. The scripts to generate such matrices in
SDPA- or CBF-format, and for computing the two relaxations
from~\cite{dAspEtAl08,dAspEtAl07} are contained in the
folder~\textsf{SCIPSDP/RIP-MISDP-Paper/scripts}. In the following, there is
a short documentation of these scripts:
\begin{itemize}
\item \textsf{Aspr07SDPA.m}: This scripts writes the SDP-relaxation of~\cite{dAspEtAl07} in the SDPA-format to the file given by ``file''.\\
  Input: Matrix~$A$; order~$k$ of the RIC~$\alpha_k$, $\beta_k$; side
  $\in \{"\ell",\, "r"\}$ for the left side/lower bound/$\alpha_k$ or the
  right side/upper bound/$\beta_k$, respectively; Rank $\in \{0,1\}$, where
  Rank = 1 adds a rank-1 constraint for the positive semidefinite
  matrix~$X$.
\item \textsf{Aspr08LagrangeRelaxSDP.m}: This script computes a single
  SDP-relaxation for the Lagrange-relaxation of the RIP, as described
  in~\cite{dAspEtAl08}.\\
  Input: $A,\, k,\, \rho,\, \alpha$ as in~\cite{dAspEtAl08}; ``side''
  (``l'' for left side/lower bound/$\alpha_k$; ``r'' for right
  side/upper bound/$\beta_k$)
\item \textsf{Aspr08Subgradient.m}: This script computes a bound for
  $alpha_k, \, beta_k$ using a subgradient approach for the
  Lagrange-relaxation of the RIP described in~\cite{dAspEtAl08}.\\
  Input: $A,\, k,\, \rho,\, \alpha$ as in~\cite{dAspEtAl08}; ``side''
  (``l'' for left side/lower bound/$\alpha_k$; ``r'' for right
  side/upper bound/$\beta_k$)
\item \textsf{generateRandomMatricesRIP.m}: This script
  calls~\textsf{generateRIPA.m}. Therefore, it defines values
  for~$m,\, n,\, k$, the number of instances of each type to be generated,
  and a random seed to be used.
\item \textsf{generateRIPA.m}: This script generates random matrices~$A$
  for computing the RIP. Input:
  \begin{dense_itemize}
  \item $m$ = number of rows of the random matrix~$A$. 
  \item $n$ = number of columns of the random matrix~$A$. 
  \item $k$ = order~$k$ for~$\alpha_k, \, \beta_k$.
  \item ``seed'' = random seed used in \textsf{generateRandomMatricesRIP.m}.
  \item ``instances'' = number of matrices~$A$ to be generated.
  \item ``type'' = type of randomness used for generating~$A$, see Table~\ref{tab:MatrixTypesRIP}.
  \item ``bandwidth'' = bandwidth of~$A$, if ``type'' is ``band''.
  \end{dense_itemize}
  Output: Instances~\textsf{Matrices/type.m.n.k\_MISDP.side.pd.rank.socp.strgbnds[.nb]}, where
  \begin{dense_itemize}
  \item ``type'', $m$, $n$, $k$ : as above
  \item ``side''          : ``l'' (left side/$\alpha_k$), ``r'' (right side/$\beta_k$)
  \item ``pd''            : p (primal form using matrix variables), d (dual
    form using scalar variables)
  \item ``rank''          : 1 (with rank-1 constraint), 0 (without)
  \item ``socp''          : 1 (with valid inequalities from~\cite{LiX20}, see~\eqref{eq:ValidSOCPineqSDP}), 0 (without)
  \item ``strgbnds''      : 1 (with stronger lower bounds~$X_{ij} \geq 0$
    if~$A \geq 0$ entrywise), 0 (with)
  \item ``nobnds'' : ``nb'' (without stronger bounds~$-\tfrac{1}{2}z_j \leq X_{ij} \leq
\tfrac{1}{2}z_j$ for $i \neq j$), else nothing is added to the filename.
  \end{dense_itemize}
\item \textsf{generateTestsets.sh}: This script generates testset-files
  \textsf{\*.test} for the instances generated
  by~\textsf{generateRandomMatricesRIP.m}.
\item \textsf{NSPAZimpl.m}: This script generates a zimpl-file for
  computing~$\alpha_k$ for the NSP of a given matrix~$A$ and order~$k$ as
  MINLP.
\item \textsf{RIPAZimpl.m}: This script generates a zimpl-file for
  computing~$\alpha_k,\, \beta_k$ for the RIP of a given matrix~$A$ and
  order~$k$ as MINLP.
\item \textsf{RIPCBFdual.m}: This script generates the MISDP formulation
  for computing~$\alpha_k,\, \beta_k$ of the RIP for a given matrix~$A$ and
  a given order~$k$ in the CBF-format using scalar variables (dual
  form). The instance is written to ``file''. See Section~\ref{sec:CBFdual}
  for the structure of
  this cbf-file. \\
  Input:
  \begin{dense_itemize}
  \item $A$: matrix 
  \item $k$: order
  \item ``side'', ``Rank'', ``socp'', ``strgbnds'', ``nobnds'': see Output
    of \textsf{generateRIPA.m} or Section~\ref{sec:CBFdual}.
  \end{dense_itemize}
  Attention: The instances generated by this script are written as lower
  triangular matrices.
\item \textsf{RIPCBFprimal.m}: This script generates the MISDP formulation
  for computing~$\alpha_k,\, \beta_k$ of the RIP for a given matrix~$A$ and
  a given order~$k$ in the CBF-format using matrix variables (primal
  form). The instance is written to ``file''. See
  Section~\ref{sec:CBFprimal} for the structure of
  this cbf-file. \\
  Input:
  \begin{dense_itemize}
  \item $A$: matrix 
  \item $k$: order
  \item ``side'', ``Rank'', ``socp'', ``strgbnds'', ``nobnds'': see Output
    of \textsf{generateRIPA.m}, or Section~\ref{sec:CBFprimal}.
  \end{dense_itemize}
  Attention: The instances generated by this script are written as lower
  triangular matrices. 
\item \textsf{RIPSDPAcont.m}: This script
  generates the continous relaxation of the MISDP formulation for
  computing~$\alpha_k,\, \beta_k$ of the RIP for a given matrix~$A$ and a
  given order~$k$ in the SDPA-format. That is, the integral
  constraints~$z_j \in \{0,1\}$ are replaced by~$z_j \in [0,1]$.  The
  instance is written to ``file''. \\
  Input:
  \begin{dense_itemize}
  \item $A$: matrix 
  \item $k$: order
  \item ``side'', ``Rank'': see Output of \textsf{generateRIPA.m}
  \end{dense_itemize}

\item \textsf{RIPSDPA.m}: This script generates the MISDP formulation
  for computing~$\alpha_k,\, \beta_k$ of the RIP for a given matrix~$A$ and
  a given order~$k$ in the SDPA-format. The instance is written to
  ``file''. \\
  Input:
  \begin{dense_itemize}
  \item $A$: matrix 
  \item $k$: order
  \item ``side'', ``Rank'': see Output of \textsf{generateRIPA.m}
  \end{dense_itemize}
\end{itemize}

\subsection{CBF-file for dual MISDP-formulation of RIP}
\label{sec:CBFdual}
This section describes the structure of the CBF-file for the
MISDP-formulation~\eqref{eq:MISDPRIP} in dual form (using scalar variables)
generated by
\textsf{SCIPSDP/RIP-MISDP-Paper/scripts/RIPCBFdual.m}.\\

{\footnotesize
\noindent
\textbf{VER} \\
1 \\
\textbf{OBJSENSE} \\
MIN (if ``side'' = l) \; or \; MAX (if ``side'' = r) \\
\textbf{VAR} \\
\# vars:  $n+\tfrac{1}{2}n(n+1)$, \qquad \# cones: $2$ \\
\[
  \begin{aligned}
    &\text{L+} &n \qquad & z_0,\dots,z_{n-1} \\
    &\text{F}\quad{\color{blue} (L+)}  &\tfrac{1}{2}n(n+1) \qquad & X_{ij},\;
    {\color{blue} (X_{ij} \geq 0)} \qquad 0 \leq j \leq i \leq n-1
  \end{aligned}
\]
\textbf{CON} \\
\# cons: $n(n+1)+n+2 \quad {\color{blue}(\tfrac{1}{2}n(n+1)+n+2)}$, \qquad
\# cones = $5 \quad {\color{blue} (4)}$\\
\[
  \begin{aligned}
    &\text{L+} &n \qquad & -z_i + 1 \geq 0, \quad & i = 0,\dots,n-1 \\
    &\text{L-} &\tfrac{1}{2}n(n+1) \qquad & {\color{blue}-z_i - X_{ij} \leq 0},\quad & 0
    \leq j \leq i \leq n-1 &\qquad {\color{blue} \text{(this line is cancelled)}}\\
    &\text{L+} &\tfrac{1}{2}n(n+1) \qquad & z_i - X_{ij} \geq 0,\quad & 0
    \leq j \leq i \leq n-1 \\
    &\text{L=}  &1 \qquad & \tr(X) - 1 = 0 \\
    &\text{L-}  &1 \qquad & \sum z_i - k \leq 0
  \end{aligned}
\]
\textbf{INT} \\
$n$ \\
$i \qquad z_i \in \{0,1\}, \quad i = 0,\dots,n-1$ \\
\textbf{PSDCON} \\
$1$ \quad {\color{red} ($n+1$)} \\
$n \qquad \qquad X \succeq 0$ \\
{\color{red}($n+2 \qquad \; P_i \succeq 0 \qquad i = 0,\dots,n-1$)} \\
{\color{green}\textbf{PSDCONRANK1} (optional, Rank=1) \\
1 \\
0\\}
\textbf{OBJACOORD} (index scalar var | coeff) \\
\# entries: $\tfrac{1}{2}n(n+1)$\\
\[
  \begin{aligned}
    &n+\tfrac{1}{2}i(i+1)+j \qquad & B_{ij} \; (i=j) \text{ or } B_{ij} +
    B_{ji} \; (i\neq j) \qquad & 0 \leq j \leq i \leq n-1
  \end{aligned}
\]
\textbf{ACOORD} (scalar cons $i$ | scalar var $j$ | val) \\
\# entries: $2n(n+1)+3n \quad {\color{blue} (- n(n+1))}$ \\
{\color{blue} If strgbnds = 1, then $\tfrac{1}{2}n(n+1)$ constraints are cancelled, so
  that $\tfrac{1}{2}n(n+1)$ needs to be subtracted from the first column
  (scalar cons $i$) in the fourth to seventh line.}
\[
  \begin{aligned}
    & i \quad & i \quad & -1.0 \quad & -z_i + 1 \geq 0 \quad & i =
    0,\dots,n-1 \\
    & n+\tfrac{1}{2}i(i+1)+j \quad & i \quad & -1.0\; (i = j),\,
    {\color{orange} -\tfrac{1}{2}\; (i\neq j)} \quad & {\color{blue}-z_j -
      X_{ij} \leq 0}
    \quad & 0 \leq j \leq i \leq n-1\\
    & n+\tfrac{1}{2}i(i+1)+j \quad & n+\tfrac{1}{2}i(i+1)+j \quad & -1.0
    \quad & {\color{blue}-z_j - X_{ij} \leq 0} \quad & 0 \leq j \leq i \leq n-1 \\
    & n+\tfrac{1}{2}n(n+1)+\tfrac{1}{2}i(i+1)+j \quad & i \quad & 1.0\; (i
    = j),\, {\color{orange} \tfrac{1}{2}\; (i\neq j)} \quad & z_j - X_{ij} \geq 0
    \quad & 0
    \leq j \leq i \leq n-1 \\
    & n+\tfrac{1}{2}n(n+1)+\tfrac{1}{2}i(i+1)+j \quad &
    n+\tfrac{1}{2}i(i+1)+j \quad & -1.0 \quad & z_j - X_{ij} \geq 0 \quad &
    0 \leq j \leq i \leq n-1 \\
    & n+n(n+1) \quad & n+\tfrac{1}{2}i(i+1)+i \quad & 1.0 \quad & \tr(X) -
    1 = 0 \quad & i = 0,\dots,n-1 \\
    & n+n(n+1)+1 \quad & i \quad & 1.0 \quad & \sum z_i - k \geq 0 \quad &
    i = 0,\dots,n-1
  \end{aligned}
\]
\textbf{BCOORD} (scalar cons $i$ | val) \\
\# entries: $n+2$ \\
{\color{blue} If strgbnds = 1, then $\tfrac{1}{2}n(n+1)$ constraints are cancelled, so
  that $\tfrac{1}{2}n(n+1)$ needs to be subtracted from the first column
  (scalar cons $i$) in the second and third line.}
\[
  \begin{aligned}
    & i \quad & -1.0 \quad & -z_i + 1 \geq 0 \quad & i = 0,\dots,n-1 \\
    & n+n(n+1) \quad & 1.0 \quad & \tr(X) - 1 = 0 \quad & i = 0,\dots,n-1
    \\
    & n+n(n+1)+1 \quad & 1.0 \quad & \sum z_i - k \geq 0 \quad & i =
    0,\dots,n-1
  \end{aligned}
\]
\textbf{HCOORD} (psd cons $i$ | scalar var $j$ | row | col | val) \\
\# entries: $\tfrac{1}{2}n(n+1)$ {\color{red} $(+n(n+4))$} \\
\[
  \begin{aligned}
    &0 \qquad &n+\tfrac{1}{2}i(i+1)+j \qquad &i \qquad &j \qquad &1.0
    \qquad &X \succeq 0 \qquad &0 \leq j \leq i \leq n-1 \\
    &{\color{red}i+1} \qquad &{\color{red}i} \qquad &{\color{red}0} \qquad
    &{\color{red}0} \qquad &{\color{red}\tfrac{1}{2}} \qquad
    &{\color{red}P_i \succeq 0} \qquad & \\
    &{\color{red}i+1} \qquad &{\color{red}i} \qquad &{\color{red}n+1} \qquad
    &{\color{red}0} \qquad &{\color{red}-\tfrac{1}{2}} \qquad
    &{\color{red}P_i \succeq 0} \qquad &\\
    &{\color{red}i+1} \qquad &{\color{red}n+\tfrac{1}{2}i(i+1)+i} \qquad &{\color{red}0} \qquad
    &{\color{red}0} \qquad &{\color{red}\tfrac{1}{2}} \qquad
    &{\color{red}P_i \succeq 0} \qquad &\\
    &{\color{red}i+1} \qquad &{\color{red}n+\tfrac{1}{2}i(i+1)+i} \qquad &{\color{red}n+1} \qquad
    &{\color{red}0} \qquad &{\color{red}\tfrac{1}{2}} \qquad
    &{\color{red}P_i \succeq 0} \qquad &\\
    &{\color{red}i+1} \qquad &{\color{red}i} \qquad &{\color{red}0} \qquad
    &{\color{red}0} \qquad &{\color{red}\tfrac{1}{2}} \qquad
    &{\color{red}P_i \succeq 0} \qquad &\\
    &{\color{red}i+1} \qquad &{\color{red}n+\tfrac{1}{2}i(i+1)+j} \qquad &{\color{red}j+1} \qquad
    &{\color{red}0} \qquad &{\color{red}1.0} \qquad
    &{\color{red}P_i \succeq 0} \qquad &{\color{red}0 \leq j \leq i}\\ 
    &{\color{red}i+1} \qquad &{\color{red}n+\tfrac{1}{2}j(j+1)+i} \qquad &{\color{red}j+1} \qquad
    &{\color{red}0} \qquad &{\color{red}1.0} \qquad
    &{\color{red}P_i \succeq 0} \qquad &{\color{red}i < j \leq
      n-1 }
  \end{aligned}
\]
{\color{red}\textbf{DCOORD} (psd cons $i$ | row | col | val) \\
  \# entries: $n(n+1)$\\
  \[
    \begin{aligned}
      {\color{red} i+1} \qquad & {\color{red} j+1} \qquad & {\color{red}
        1.0} \qquad & {\color{red} P_i \succeq 0} \qquad & {\color{red} i =
        0, \dots, n-1}
    \end{aligned}
    \]

  }
}

Options in \textsf{SCIPSDP/RIP-MISDP-Paper/scripts/RIPCBFdual.m}:
\begin{itemize}
\item If Rank = 1: \textcolor{green}{PSDCONRANK1} is added.
\item If nobnds = 1: {\color{orange}
    $-\tfrac{1}{2}z_j \leq X_{ij} \leq \tfrac{1}{2}z_j$ for $i \neq j$} is
  weakened to {\color{orange}$-z_j \leq X_{ij} \leq z_j$}, so that the
  {\color{orange} $-\tfrac{1}{2}$} is changed to $-1.0$ in ACOORD.
\item If strgbnds = 1: If A is nonnegative, then the stronger lower
  bounds~$X_{ij} \geq 0$ for all $i,j =0,\dots,n-1$ are used. See the
  changes in \textcolor{blue}{blue} for VAR, CON. Additionally, the two
  lines in ACOORD with a \textcolor{blue}{blue constraint} are cancelled
  as well.
\item If socp = 1: The valid SOCP inequalities
  $\sum_{j=0}^{n-1}X_{ij}^2 \leq X_{ii}z_i$ for $i = 0,\dots,n-1$
  from~\cite{LiX20} are added. See the changes in PSDCON, HCOORD, DCOORD in
  \textcolor{red}{red}. These inequalities are SDP-representable via
  \begin{align*}
    \sum_{j=1}^n X_{ij}^2 \leq X_{ii}z_i \; \Leftrightarrow \;
    P_i \define \begin{pmatrix}
      \tfrac{X_{ii}+z_I}{2} & X_{i1} & \cdots & X_{in} & \tfrac{X_{ii}-z_i}{2}
      \\
      X_{i1} & 1 & & 0 & 0 \\
      \vdots &\vdots & \ddots & \vdots & \vdots \\
      X_{in} & 0 & & 1 & 0 \\
      \tfrac{X_{ii}-z_i}{2} & 0 & \cdots & 0 & 1
    \end{pmatrix} \succeq 0.
  \end{align*}
  See the last two columns in FCOORD, ACOORD, BCOORD for a description on
  how to write these inequalities in the CBF-format. Here, $E_{ij}$ is a
  symmetric matrix of suitable dimension with ones at positions $(i,j)$ and
  $(j,i)$, and zeros else.
\end{itemize}

\subsection{CBF-file for primal MISDP-formulation of RIP}
\label{sec:CBFprimal}
This section describes the structure of the CBF-file for the
MISDP-formulation~\eqref{eq:MISDPRIP} in primal form (using matrix variables)
generated by
\textsf{SCIPSDP/RIP-MISDP-Paper/scripts/RIPCBFdual.m}.\\

{\footnotesize
\noindent
\textbf{VER} \\
1 \\
\textbf{OBJSENSE} \\
MIN (if ``side'' = l) \; or \; MAX (if ``side'' = r) \\
\textbf{VAR} \\
\# vars:  $n$, \qquad \# cones: $1$ \\
\[
  \begin{aligned}
    &\text{L+} &n \qquad & z_0,\dots,z_{n-1}
  \end{aligned}
\]
\textbf{PSDVAR} \\
1 \qquad {\color{red}(n+1)} \\
$n \qquad \qquad X$ \\
{\color{red} ($n+2 \qquad \; P_i, \qquad i = 0,\dots,n-1$)}\\
{\color{green}\textbf{PSDVARRANK1}  \\
1 \\
0\\}
\textbf{INT} \\
$n$ \\
$i \qquad z_i \in \{0,1\}, \quad i = 0,\dots,n-1$ \\
\textbf{OBJFCOORD} (psd var $i$ | row | col | val) \\
\# entries: $\tfrac{1}{2}n(n+1)$\\
\[
  \begin{aligned}
    &0 \qquad &i \qquad &j \qquad B_{i,j} \qquad & 0\leq j \leq i \leq n-1
  \end{aligned}
\]
\textbf{CON} \\
\# cons: $n(n+1)+n+2 \quad {\color{red}(+\tfrac{1}{2}n(n+3)(n+2))}$, \qquad
\# cones = $5 \quad {\color{red} (6)}$\\
\[
  \begin{aligned}
    &{\color{red}\text{L=}} & {\color{red}\tfrac{1}{2}n(n+3)(n+2)} \qquad 
    &{\color{red} \sum_{j=0}^{n-1} X_{ij}^2 \leq X_{ii}z_i}, \quad &
    {\color{red} i = 0,\dots, n-1} \\
    &\text{L+} &n \qquad & -z_i + 1 \geq 0, \quad & i = 0,\dots,n-1 \\
    &\text{L-} &\tfrac{1}{2}n(n+1) \qquad & -z_i - X_{ij} \leq 0,\;
    {\color{blue} (-X_{ij} \leq 0)}\quad & 0
    \leq j \leq i \leq n-1 \\
    &\text{L+} &\tfrac{1}{2}n(n+1) \qquad & z_i - X_{ij} \geq 0,\quad & 0
    \leq j \leq i \leq n-1 \\
    &\text{L=}  &1 \qquad & \tr(X) - 1 = 0 \\
    &\text{L-}  &1 \qquad & \sum z_i - k \leq 0
  \end{aligned}
\]
\textbf{FCOORD} (scalar cons $i$ | psd var $j$ | row | col | val) \\
\# entries: $n(n+1)+n$ \quad {\color{red} $+
  n\cdot(3n+5+\tfrac{1}{2}n(n+1))$ }\\
{\color{red} The following entries need to be added for all~$i = 0,\dots,n-1$}:
\[
  \begin{aligned}
    &{\color{red}j} \quad &{\color{red}i+1} \quad & {\color{red}j+1} \quad
    & {\color{red}j+1} \quad & {\color{red}1.0} \quad & {\color{red}
      \langle P_i,
      E_{j+1,j+1} \rangle - 1 = 0} \quad & {\color{red} j = 0,\dots,n} \\
    &{\color{red}n+1} \quad &{\color{red}i+1} \quad & {\color{red}0} \quad
    & {\color{red}0} \quad & {\color{red}1.0} \quad & {\color{red} \langle
      P_i, E_{00} \rangle - \tfrac{1}{2}\langle X,E_{ii}\rangle -
      \tfrac{1}{2}z_i = 0} \quad & \\
    &{\color{red}n+1} \quad &{\color{red}0} \quad & {\color{red}i} \quad
    & {\color{red}i} \quad & {\color{red}-\tfrac{1}{2}} \quad & {\color{red} \langle
      P_i, E_{00} \rangle - \tfrac{1}{2}\langle X,E_{ii}\rangle -
      \tfrac{1}{2}z_i = 0} \quad & \\
    &{\color{red}n+2+k} \quad &{\color{red}i+1} \quad & {\color{red}k+1} \quad
    & {\color{red}0} \quad & {\color{red}1.0\, (i\neq k),\; 0.5\, (i=k)} \quad & {\color{red} \langle
      P_i, E_{k+1,0} \rangle - \langle X,E_{ki}\rangle = 0} \quad &
    {\color{red} k = 0,\dots,n-1} \\
    &{\color{red}n+2+k} \quad &{\color{red}0} \quad & {\color{red}i} \quad
    & {\color{red}k} \quad & {\color{red}-1.0} \quad & {\color{red} \langle
      P_i, E_{k+1,0} \rangle - \langle X,E_{ki}\rangle = 0} \quad &
    {\color{red} k = 0,\dots,i} \\
    &{\color{red}n+2+k} \quad &{\color{red}0} \quad & {\color{red}k} \quad
    & {\color{red}i} \quad & {\color{red}-1.0} \quad & {\color{red} \langle
      P_i, E_{k+1,0} \rangle - \langle X,E_{ki}\rangle = 0} \quad &
    {\color{red} k = i+1,\dots,n-1} \\    
    &{\color{red}2n+2} \quad &{\color{red}i+1} \quad & {\color{red}n+1} \quad
    & {\color{red}0} \quad & {\color{red}\tfrac{1}{2}} \quad & {\color{red}
      \tfrac{1}{2}\langle P_i, E_{n+1,0} \rangle - \tfrac{1}{2} \langle
      X,E_{ii}\rangle + \tfrac{1}{2}z_i = 0} \quad & \\
    &{\color{red}2n+2} \quad &{\color{red}0} \quad & {\color{red}i} \quad
    & {\color{red}i} \quad & {\color{red}-\tfrac{1}{2}} \quad & {\color{red}
      \tfrac{1}{2}\langle P_i, E_{n+1,0} \rangle - \tfrac{1}{2} \langle
      X,E_{ii}\rangle + \tfrac{1}{2}z_i = 0} \quad & \\
    &{\color{red}2n+3+...} \quad &{\color{red}i+1} \quad & {\color{red}k+1} \quad
    & {\color{red}j+1} \quad & {\color{red}1.0} \quad & {\color{red}
     \langle P_i, E_{k+1,j+1} \rangle = 0} \quad & {\color{red} 0 \leq j < k \leq n}
  \end{aligned}
\]
\noindent
{\color{red} If the valid SOCP-inequality is added, then $\tfrac{1}{2}n(n+3)(n+2)$ needs
to be added to the entries in the first column (scalar cons $i$).}
\[
  \begin{aligned}
    &n+\tfrac{1}{2}i(i+1)+j \quad & 0 \quad & i \quad & j \quad & -1.0\,
    (i\neq j),\; -\tfrac{1}{2}\, (i=j) \quad & -z_j - X_{ij} \leq 0 \quad &
    0 \leq j \leq i \leq n-1 \\
    &n+\tfrac{1}{2}n(n+1)+\tfrac{1}{2}i(i+1)+j \quad & 0 \quad & i \quad &
    j \quad & -1.0\, (i\neq j),\; -\tfrac{1}{2}\, (i=j) \quad & +z_j -
    X_{ij} \geq 0 \quad & 0 \leq j \leq i \leq n-1 \\
    &n+n(n+1)\quad & 0 \quad & i \quad & i \quad & 1.0 \quad & \tr(X) - 1
    \leq 0 \quad & i = 0,\dots,n-1
  \end{aligned}
\]
\textbf{ACOORD} (scalar cons $i$ | scalar var $j$ val) \\
\# entries: $n(n+1)+2n$ \qquad {\color{red} $(+2n)$} \qquad {\color{blue}
  $-(\tfrac{1}{2}n(n+1))$} \\
\[
  \begin{aligned}
    &{\color{red}i\cdot\tfrac{1}{2}(n+3)(n+2) + n+1} \quad & {\color{red}i} \quad &
    {\color{red}-\tfrac{1}{2}} \quad & {\color{red} \langle P_i, E_{00}
      \rangle - \tfrac{1}{2}\langle X,E_{ii}\rangle - \tfrac{1}{2}z_i = 0}
    \quad &{\color{red} i = 0,\dots,n-1} \\
    &{\color{red}i\cdot\tfrac{1}{2}(n+3)(n+2) + 2n+2} \quad & {\color{red}i} \quad &
    {\color{red}\tfrac{1}{2}} \quad & {\color{red} \langle P_i, E_{00}
      \rangle - \tfrac{1}{2}\langle X,E_{ii}\rangle + \tfrac{1}{2}z_i = 0}
    \quad &{\color{red} i = 0,\dots,n-1} \\
  \end{aligned}
\]
{\color{red} If the valid SOCP-inequality is added, then $\tfrac{1}{2}n(n+3)(n+2)$ needs
  to be added to the entries in the first column (scalar cons $i$).}
\[
  \begin{aligned}
    &i \quad & i \quad & -1.0 \quad & -z_i + 1 = 0 \quad & i = 0,\dots,n-1
    \\
    &n+\tfrac{1}{2}i(i+1)+j \quad & i \quad & -1.0\, (i =
    j),\; {\color{orange} -\tfrac{1}{2},\,(i \neq j)} \quad & {\color{blue}-z_i - X_{ij}
    \leq 0} \quad & 0 \leq j \leq i \leq n-1 \quad {\color{blue}
    (\text{this line is cancelled})} \\
    &n+\tfrac{1}{2}n(n+1)+\tfrac{1}{2}i(i+1)+j \quad & i \quad & 1.0\, (i =
    j),\; {\color{orange} \tfrac{1}{2},\,(i \neq j)}\quad & z_i - X_{ij}
    \geq 0 \quad & 0 \leq j \leq i \leq n-1 \\
    &in+n(n+1)+1 \quad & i \quad & 1.0 \quad & \sum z_i - k \leq 0 \quad &
    i = 0,\dots,n-1 \\
  \end{aligned}
\]
\textbf{BCOORD} (scalar cons $i$ | val) \\
\# entries: $n+2 \qquad {\color{red} (+n(n+1))}$ \\
\[
\begin{aligned}
  &{\color{red}i\cdot\tfrac{1}{2}(n+3)(n+2) + j} \quad & {\color{red}-1.0}
  \quad & {\color{red} \langle P_i, E_{j+1,j+1} \rangle - 1 = 0} \quad & {\color{red} j
  = 0,\dots,n,\; i = 0,\dots,n-1} \\
  \end{aligned}
\]
{\color{red} If the valid SOCP-inequality is added, then $\tfrac{1}{2}n(n+3)(n+2)$ needs
  to be added to the entries in the first column (scalar cons $i$).}
\[
  \begin{aligned}
    &i \quad & 1.0 \quad & -z_i + 1 \geq 0 \quad & i = 0,\dots,n-1 \\
    &n(n+1)+n \quad & -1.0 \quad & \tr(X) -1 = 0 \quad & \\
    &n(n+1)+n+1 \quad & k \quad & \sum z_i - k \leq 0 \quad & \\
  \end{aligned}
\]
}

Options in \textsf{SCIPSDP/RIP-MISDP-Paper/scripts/RIPCBFdual.m}:
\begin{itemize}
\item If Rank = 1: \textcolor{green}{PSDVARRANK1} is added.
\item If nobnds = 1: {\color{orange}
    $-\tfrac{1}{2}z_j \leq X_{ij} \leq \tfrac{1}{2}z_j$ for $i \neq j$} is
  weakened to {\color{orange}$-z_j \leq X_{ij} \leq z_j$}, so that the
  {\color{orange} $-\tfrac{1}{2}$} is changed to $-1.0$ in ACOORD.
\item If strgbnds = 1: If A is nonnegative, then the stronger lower
  bounds~$X_{ij} \geq 0$ for all $i,j =0,\dots,n-1$ are used. See the
  changes in \textcolor{blue}{blue} in ACOORD.
\item If socp = 1: The valid SOCP inequalities
  $\sum_{j=0}^{n-1}X_{ij}^2 \leq X_{ii}z_i$ for $i = 0,\dots,n-1$
  from~\cite{LiX20} are added. See the changes in PSDVAR, CON, FCOORD,
  ACOORD and BCOORD in \textcolor{red}{red}. These inequalities are
  SDP-representable via
  \begin{align*}
    \sum_{j=1}^n X_{ij}^2 \leq X_{ii}z_i \; \Leftrightarrow \;
    P_i \define \begin{pmatrix}
      \tfrac{X_{ii}+z_I}{2} & X_{i1} & \cdots & X_{in} & \tfrac{X_{ii}-z_i}{2}
      \\
      X_{i1} & 1 & & 0 & 0 \\
      \vdots &\vdots & \ddots & \vdots & \vdots \\
      X_{in} & 0 & & 1 & 0 \\
      \tfrac{X_{ii}-z_i}{2} & 0 & \cdots & 0 & 1
    \end{pmatrix} \succeq 0.
  \end{align*}
  See the last two columns in FCOORD, ACOORD, BCOORD for a description on
  how to write these inequalities in the CBF-format. Here, $E_{ij}$ is a
  symmetric matrix of suitable dimension with ones at positions $(i,j)$ and
  $(j,i)$, and zeros else.
\end{itemize}

\begin{small}
  \bibliographystyle{abbrv}
  \bibliography{RandomInstancesMISDP.bib}
\end{small}

\end{document}

### 2020-09-09 RIP-MISDP-Paper/scripts/RandomInstancesMISDP.tex
Correct cbf-generation of RIP instances

@@
     &\text{L+} &n \qquad & z_0,\dots,z_{n-1} \\
-    &\text{F}\quad{\color{blue} (L+)}  &\tfrac{1}{2}n(n+1) \qquad & X_{ij},\;
+    &\text{F}\quad{\color{blue} (\text{L}+)}  &\tfrac{1}{2}n(n+1) \qquad & X_{ij},\;
     {\color{blue} (X_{ij} \geq 0)} \qquad 0 \leq j \leq i \leq n-1
@@
 \textbf{CON} \\
-\# cons: $n(n+1)+n+2 \quad {\color{blue}(\tfrac{1}{2}n(n+1)+n+2)}$, \qquad
+\# cons: $2n^2+n+2 \quad {\color{blue}(n^2+n+2)}$, \qquad
 \# cones = $5 \quad {\color{blue} (4)}$\\
@@
     &\text{L+} &n \qquad & -z_i + 1 \geq 0, \quad & i = 0,\dots,n-1 \\
-    &\text{L-} &\tfrac{1}{2}n(n+1) \qquad & {\color{blue}-z_i - X_{ij} \leq 0},\quad & 0
-    \leq j \leq i \leq n-1 &\qquad {\color{blue} \text{(this line is cancelled)}}\\
-    &\text{L+} &\tfrac{1}{2}n(n+1) \qquad & z_i - X_{ij} \geq 0,\quad & 0
-    \leq j \leq i \leq n-1 \\
+    &\text{L-} &n^2 \qquad & {\color{blue}-z_i - X_{ij} \leq 0},\quad & 0
+    \leq i,\, j \leq n-1 &\qquad {\color{blue} \text{(this line is cancelled)}}\\
+    &\text{L+} &n^2 \qquad & z_i - X_{ij} \geq 0,\quad & 0
+    \leq i, \, j \leq n-1 \\
     &\text{L=}  &1 \qquad & \tr(X) - 1 = 0 \\
@@
 \textbf{ACOORD} (scalar cons $i$ | scalar var $j$ | val) \\
-\# entries: $2n(n+1)+3n \quad {\color{blue} (- n(n+1))}$ \\
-{\color{blue} If strgbnds = 1, then $\tfrac{1}{2}n(n+1)$ constraints are cancelled, so
-  that $\tfrac{1}{2}n(n+1)$ needs to be subtracted from the first column
+\# entries: $4n^2+3n \quad {\color{blue} (- 2n^2)}$ \\
+{\color{blue} If strgbnds = 1, then $n^2$ constraints are cancelled, so
+  that $n^2$ needs to be subtracted from the first column
   (scalar cons $i$) in the fourth to seventh line.}
@@
     0,\dots,n-1 \\
-    & n+\tfrac{1}{2}i(i+1)+j \quad & i \quad & -1.0\; (i = j),\,
+    & n+in+j \quad & i \quad & -1.0\; (i = j),\,
     {\color{orange} -\tfrac{1}{2}\; (i\neq j)} \quad & {\color{blue}-z_j -
       X_{ij} \leq 0}
-    \quad & 0 \leq j \leq i \leq n-1\\
-    & n+\tfrac{1}{2}i(i+1)+j \quad & n+\tfrac{1}{2}i(i+1)+j \quad & -1.0
-    \quad & {\color{blue}-z_j - X_{ij} \leq 0} \quad & 0 \leq j \leq i \leq n-1 \\
-    & n+\tfrac{1}{2}n(n+1)+\tfrac{1}{2}i(i+1)+j \quad & i \quad & 1.0\; (i
+    \quad & 0 \leq i,\, j \leq n-1\\
+    & n+in+j \quad & n+\tfrac{1}{2}i(i+1)+j \quad & -1.0
+    \quad & {\color{blue}-z_j - X_{ij} \leq 0} \quad & 0 \leq i,\, j \leq
+    n-1, \; i \geq j \\
+    & n+in+j \quad & n+\tfrac{1}{2}j(j+1)+i \quad & -1.0
+    \quad & {\color{blue}-z_j - X_{ij} \leq 0} \quad & 0 \leq i,\, j \leq
+    n-1, \; i < j \\
+    & n+n^2+in+j \quad & i \quad & 1.0\; (i
     = j),\, {\color{orange} \tfrac{1}{2}\; (i\neq j)} \quad & z_j - X_{ij} \geq 0
     \quad & 0
-    \leq j \leq i \leq n-1 \\
-    & n+\tfrac{1}{2}n(n+1)+\tfrac{1}{2}i(i+1)+j \quad &
+    \leq i,\, j \leq n-1 \\
+    & n+n^2+in+j \quad &
     n+\tfrac{1}{2}i(i+1)+j \quad & -1.0 \quad & z_j - X_{ij} \geq 0 \quad &
-    0 \leq j \leq i \leq n-1 \\
-    & n+n(n+1) \quad & n+\tfrac{1}{2}i(i+1)+i \quad & 1.0 \quad & \tr(X) -
+    0 \leq i,\, j \leq n-1,\; i \geq j \\
+    & n+n^2+in+j \quad &
+    n+\tfrac{1}{2}j(j+1)+i \quad & -1.0 \quad & z_j - X_{ij} \geq 0 \quad &
+    0 \leq i,\, j \leq n-1,\; i < j \\
+    & n+2n^2 \quad & n+\tfrac{1}{2}i(i+1)+i \quad & 1.0 \quad & \tr(X) -
     1 = 0 \quad & i = 0,\dots,n-1 \\
-    & n+n(n+1)+1 \quad & i \quad & 1.0 \quad & \sum z_i - k \geq 0 \quad &
+    & n+2n^2+1 \quad & i \quad & 1.0 \quad & \sum z_i - k \geq 0 \quad &
     i = 0,\dots,n-1
@@
 \# entries: $n+2$ \\
-{\color{blue} If strgbnds = 1, then $\tfrac{1}{2}n(n+1)$ constraints are cancelled, so
-  that $\tfrac{1}{2}n(n+1)$ needs to be subtracted from the first column
+{\color{blue} If strgbnds = 1, then $n^2$ constraints are cancelled, so
+  that $n^2$ needs to be subtracted from the first column
   (scalar cons $i$) in the second and third line.}
@@
   \begin{aligned}
-    & i \quad & -1.0 \quad & -z_i + 1 \geq 0 \quad & i = 0,\dots,n-1 \\
-    & n+n(n+1) \quad & 1.0 \quad & \tr(X) - 1 = 0 \quad & i = 0,\dots,n-1
+    & i \quad & 1.0 \quad & -z_i + 1 \geq 0 \quad & i = 0,\dots,n-1 \\
+    & n+2n^2 \quad & -1.0 \quad & \tr(X) - 1 = 0 \quad & i = 0,\dots,n-1
     \\
-    & n+n(n+1)+1 \quad & 1.0 \quad & \sum z_i - k \geq 0 \quad & i =
+    & n+2n^2+1 \quad & -k  \quad & \sum z_i - k \geq 0 \quad & i =
     0,\dots,n-1
@@
 \textbf{CON} \\
-\# cons: $n(n+1)+n+2 \quad {\color{red}(+\tfrac{1}{2}n(n+3)(n+2))}$, \qquad
+\# cons: $2n^2+n+2 \quad {\color{red}(+\tfrac{1}{2}n(n+3)(n+2))} \quad
+{\color{blue} (-\tfrac{1}{2}n(n-1))}$, \qquad
 \# cones = $5 \quad {\color{red} (6)}$\\
@@
     &\text{L+} &n \qquad & -z_i + 1 \geq 0, \quad & i = 0,\dots,n-1 \\
-    &\text{L-} &\tfrac{1}{2}n(n+1) \qquad & -z_i - X_{ij} \leq 0,\;
-    {\color{blue} (-X_{ij} \leq 0)}\quad & 0
-    \leq j \leq i \leq n-1 \\
-    &\text{L+} &\tfrac{1}{2}n(n+1) \qquad & z_i - X_{ij} \geq 0,\quad & 0
+    &\text{L-} &n^2\quad {\color{blue}(\tfrac{1}{2}n(n+1))} \qquad & -z_i - X_{ij} \leq 0,\;
+    {\color{blue} (-X_{ij} \leq 0)}\quad & 0 \leq i,\, j \leq n-1\quad {\color{blue}(0
+    \leq j \leq i \leq n-1)} \\
+    &\text{L+} &n^2 \qquad & z_i - X_{ij} \geq 0,\quad & 0
     \leq j \leq i \leq n-1 \\
@@
 \textbf{FCOORD} (scalar cons $i$ | psd var $j$ | row | col | val) \\
-\# entries: $n(n+1)+n$ \quad {\color{red} $+
-  n\cdot(3n+5+\tfrac{1}{2}n(n+1))$ }\\
+\# entries: $2n^2+n$ \quad {\color{red} $+
+  n\cdot(3n+5+\tfrac{1}{2}n(n+1))$ } \quad {\color{blue} $- \tfrac{1}{2}n(n-1)$}\\
 {\color{red} The following entries need to be added for all~$i = 0,\dots,n-1$}:
@@
   \begin{aligned}
-    &n+\tfrac{1}{2}i(i+1)+j \quad & 0 \quad & i \quad & j \quad & -1.0\,
-    (i\neq j),\; -\tfrac{1}{2}\, (i=j) \quad & -z_j - X_{ij} \leq 0 \quad &
-    0 \leq j \leq i \leq n-1 \\
-    &n+\tfrac{1}{2}n(n+1)+\tfrac{1}{2}i(i+1)+j \quad & 0 \quad & i \quad &
-    j \quad & -1.0\, (i\neq j),\; -\tfrac{1}{2}\, (i=j) \quad & +z_j -
-    X_{ij} \geq 0 \quad & 0 \leq j \leq i \leq n-1 \\
-    &n+n(n+1)\quad & 0 \quad & i \quad & i \quad & 1.0 \quad & \tr(X) - 1
+    &n+in+j \quad & 0 \quad & i \quad & j \quad & -1.0\,
+    (i = j),\; -\tfrac{1}{2}\, (i\neq j) \quad & -z_j - X_{ij} \leq 0 \quad &
+    0 \leq i,\, j \leq n-1,\; i \geq j \\
+    &n+in+j \quad & 0 \quad & j \quad & i \quad & -1.0\,
+    (i = j),\; -\tfrac{1}{2}\, (i\neq j) \quad & -z_j - X_{ij} \leq 0 \quad &
+    0 \leq i,\, j \leq n-1,\; i < j \\
+    &n+n^2+in+j \quad & 0 \quad & i \quad &
+    j \quad & -1.0\, (i = j),\; -\tfrac{1}{2}\, (i\neq j) \quad & +z_j -
+    X_{ij} \geq 0 \quad & 0 \leq i,\, j \leq n-1,\; i \geq j \\
+    &n+n^2+in+j \quad & 0 \quad & j \quad &
+    i \quad & -1.0\, (i = j),\; -\tfrac{1}{2}\, (i\neq j) \quad & +z_j -
+    X_{ij} \geq 0 \quad & 0 \leq i,\, j \leq n-1,\; i < j \\
+    &n+2n^2 \quad & 0 \quad & i \quad & i \quad & 1.0 \quad & \tr(X) - 1
     \leq 0 \quad & i = 0,\dots,n-1
@@
 \]
+\noindent
+{\color{blue} If strgbnds = 1, then the first two lines are replaced by the
+  following line:
+  \[
+    n+\tfrac{1}{2}i(i+1)+j \quad  0 \quad i \quad  j \quad  -1.0
+    \quad - X_{ij} \leq 0 \quad 0 \leq j \leq i \leq n-1,
+  \]
+  so that $\tfrac{1}{2}n(n-1)$ needs to be subtracted from the first column
+  (scalar cons $i$) in the third to fifth line.\\
+}
+
+\noindent
 \textbf{ACOORD} (scalar cons $i$ | scalar var $j$ val) \\
-\# entries: $n(n+1)+2n$ \qquad {\color{red} $(+2n)$} \qquad {\color{blue}
-  $-(\tfrac{1}{2}n(n+1))$} \\
+\# entries: $2n^2+2n$ \qquad {\color{red} $(+2n)$} \qquad {\color{blue}
+  $(-n^2)$} \\
 \[
@@
   \begin{aligned}
-    &i \quad & i \quad & -1.0 \quad & -z_i + 1 = 0 \quad & i = 0,\dots,n-1
-    \\
-    &n+\tfrac{1}{2}i(i+1)+j \quad & i \quad & -1.0\, (i =
+    &i \quad & i \quad & -1.0 \quad & -z_i + 1 \geq 0 \quad & i = 0, \dots,
+    n-1 \\
+    &n+in+j \quad & i \quad & -1.0\, (i =
     j),\; {\color{orange} -\tfrac{1}{2},\,(i \neq j)} \quad & {\color{blue}-z_i - X_{ij}
-    \leq 0} \quad & 0 \leq j \leq i \leq n-1 \quad {\color{blue}
+    \leq 0} \quad & 0 \leq i,\, j \leq n-1 \quad {\color{blue}
     (\text{this line is cancelled})} \\
-    &n+\tfrac{1}{2}n(n+1)+\tfrac{1}{2}i(i+1)+j \quad & i \quad & 1.0\, (i =
+    &n+n^2+in+j \quad & i \quad & 1.0\, (i =
     j),\; {\color{orange} \tfrac{1}{2},\,(i \neq j)}\quad & z_i - X_{ij}
-    \geq 0 \quad & 0 \leq j \leq i \leq n-1 \\
-    &in+n(n+1)+1 \quad & i \quad & 1.0 \quad & \sum z_i - k \leq 0 \quad &
+    \geq 0 \quad & 0 \leq i,\, j \leq n-1 \\
+    &n+2n^2+1 \quad & i \quad & 1.0 \quad & \sum z_i - k \leq 0 \quad &
     i = 0,\dots,n-1 \\
@@
 \]
+\noindent
+{\color{blue} If strgbnds = 1, then $\tfrac{1}{2}n(n-1)$ constraints are
+  cancelled, so that $\tfrac{1}{2}n(n-1)$ needs to be subtracted from the
+  first column (scalar cons $i$) in the third and fourth line.\\
+}
+
 \textbf{BCOORD} (scalar cons $i$ | val) \\
@@
     &i \quad & 1.0 \quad & -z_i + 1 \geq 0 \quad & i = 0,\dots,n-1 \\
-    &n(n+1)+n \quad & -1.0 \quad & \tr(X) -1 = 0 \quad & \\
-    &n(n+1)+n+1 \quad & k \quad & \sum z_i - k \leq 0 \quad & \\
+    &2n^2+n \quad & -1.0 \quad & \tr(X) -1 = 0 \quad & \\
+    &2n^2+n+1 \quad & k \quad & \sum z_i - k \leq 0 \quad & \\
   \end{aligned}
 \]
+\noindent
+{\color{blue} If strgbnds = 1, then $\tfrac{1}{2}n(n-1)$ constraints are
+  cancelled, so that $\tfrac{1}{2}n(n-1)$ needs to be subtracted from the
+  first column (scalar cons $i$) in the second and third line.\\
+}
 }

### 2020-09-09 RIP-MISDP-Paper/scripts/RIPCBFdual.m

@@ -120,14 +120,14 @@ function [] = RIPCBFdual(A, order, side, file, Rank, socp, strgbnds,nobnds)
     % ------------------------------------------
     % constraints
     fprintf(fid, "CON\n");
-    ncons = n*(n+1)+n+2 - strgbnds*0.5*n*(n+1);
+    ncons = 2*n^2+n+2 - strgbnds*n^2;
     ncones = 5 - strgbnds;
     fprintf(fid, "%d %d\n", ncons,ncones);
-    fprintf(fid, "L+ %d\n", n);         % -z_j + 1 >= 0
+    fprintf(fid, "L+ %d\n", n);                % -z_j + 1 >= 0
     if strgbnds == 0
-        fprintf(fid, "L- %d\n", 0.5*n*(n+1));  % -z_j - X_{ij} <= 0
+        fprintf(fid, "L- %d\n", n^2);           % -z_j - X_{ij} <= 0
     end
-    fprintf(fid, "L+ %d\n", 0.5*n*(n+1));       % z_j - X_{ij} >= 0
+    fprintf(fid, "L+ %d\n", n^2);               % z_j - X_{ij} >= 0
     fprintf(fid, "L= 1\n");                    % \sum_j X_{jj} == 1
     fprintf(fid, "L- 1\n");                    % \sum_j z_j <= k
     fprintf(fid, "\n");
@@ -135,7 +135,7 @@ function [] = RIPCBFdual(A, order, side, file, Rank, socp, strgbnds,nobnds)
     
     % write ACOORD
     fprintf(fid, "ACOORD\n");
-    nACOORD = 2*n*(n+1)+3*n - strgbnds*n*(n+1);
+    nACOORD = 4*n^2+3*n - strgbnds*2*n^2;
     fprintf(fid, "%d\n", nACOORD);
     cnt = 0;       % counts number of specified entries
     conscnt = 0;   % counts number of specified constraints 
@@ -150,8 +150,9 @@ function [] = RIPCBFdual(A, order, side, file, Rank, socp, strgbnds,nobnds)
     % add coupling constraints -z_j - X_{ij} <= 0 (only if strgbnds = 0)
     if strgbnds == 0
         for i = 0:n-1
-            for j = 0:i
-                if ( j ~= i && nobnds == 0)
+            % case i > j
+            for j = 0:i-1
+                if nobnds == 0
                     fprintf(fid, "%d %d -0.5\n", conscnt, j);
                 else
                     fprintf(fid, "%d %d -1.0\n", conscnt, j);
[…]

### 2020-09-09 RIP-MISDP-Paper/scripts/RIPCBFprimal.m

@@ -101,7 +101,7 @@ function [] = RIPCBFprimal(A, order, side, file, Rank, socp, strgbnds,nobnds)
     %% constraints
     fprintf(fid, "CON\n");
     nsocpcons = 0.5*n*(n+3)*(n+2);
-    ncons = n*(n+1)+n+2 + socp*nsocpcons;
+    ncons = 2*n^2+n+2 + socp*nsocpcons - strgbnds*0.5*n*(n-1);
     ncones = 5 + socp;
     fprintf(fid, "%d %d\n", ncons, ncones);
     if socp == 1
@@ -109,15 +109,18 @@ function [] = RIPCBFprimal(A, order, side, file, Rank, socp, strgbnds,nobnds)
         fprintf(fid, "L= %d\n", nsocpcons);  % \sum_j X_{ij}^2 <= X_{ii}z_i
     end
     fprintf(fid, "L+ %d\n", n);              % -z_j + 1 >= 0
-    fprintf(fid, "L- %d\n", 0.5*n*(n+1));    % -z_j - X_{ij} <= 0
-    fprintf(fid, "L+ %d\n", 0.5*n*(n+1));    % z_j - X_{ij} >= 0
+    fprintf(fid, "L- %d\n", n^2 - strgbnds*0.5*n*(n-1));    % -z_j - X_{ij} <= 0
+    fprintf(fid, "L+ %d\n", n^2);             % z_j - X_{ij} >= 0
     fprintf(fid, "L= 1\n");                  % \sum_j X_{jj} == 1
     fprintf(fid, "L- 1\n");                  % \sum_j z_j <= k
     fprintf(fid, "\n");
 
     %% write FCOORD
     fprintf(fid, "FCOORD\n");
-    nFCOORD = n*(n+1)+n + socp*(n*(3*n+5+0.5*n*(n+1)));
+    if strgbnds ~= 0 && strgbnds ~= 1
+        error("Error: Option <%s> for parameter strgbnds not valid!\n", strgbnds);
+    end
+    nFCOORD = 2*n^2+n + socp*(n*(3*n+5+0.5*n*(n+1))) - strgbnds*0.5*n*(n-1);
     fprintf(fid, "%d\n", nFCOORD);
     cnt = 0;
     conscnt = 0;
@@ -170,27 +173,44 @@ function [] = RIPCBFprimal(A, order, side, file, Rank, socp, strgbnds,nobnds)
     % no FCOORD for -z_j + 1 >= 0
[…]

### 2020-09-09 RIP-MISDP-Paper/scripts/generateRandomMatricesRIP.m

@@ -1,9 +1,14 @@
-valsm = [15,25,30];
-valsn = [30,35,40];
-valsk = [5,4,3];
+% valsm = [15,25,30];
+% valsn = [30,35,40];
+% valsk = [5,4,3];
+% 
+valsm = [15];
+valsn = [30];
+valsk = [5];
+
 
 seed = 1234;
-instances = 3; % number of instances per (type,m,n,k)
+instances = 1; % number of instances per (type,m,n,k)
 types = ['0+-1';'band';'bern';'bina';'norm';'rnk1';'wish'];
 
 mkdir Matrices

### 2020-09-09 src/scipsdp/sorttpllex.c
add lexicographic sorting macros

@@ -1,0 +1,589 @@
+/* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
+/*                                                                           */
+/* This file is part of SCIPSDP - a solving framework for mixed-integer      */
+/* semidefinite programs based on SCIP.                                      */
+/*                                                                           */
+/* Copyright (C) 2011-2013 Discrete Optimization, TU Darmstadt               */
+/*                         EDOM, FAU Erlangen-Nürnberg                       */
+/*               2014-2020 Discrete Optimization, TU Darmstadt               */
+/*                                                                           */
+/*                                                                           */
+/* This program is free software; you can redistribute it and/or             */
+/* modify it under the terms of the GNU Lesser General Public License        */
+/* as published by the Free Software Foundation; either version 3            */
+/* of the License, or (at your option) any later version.                    */
+/*                                                                           */
+/* This program is distributed in the hope that it will be useful,           */
+/* but WITHOUT ANY WARRANTY; without even the implied warranty of            */
+/* MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the             */
+/* GNU Lesser General Public License for more details.                       */
+/*                                                                           */
+/* You should have received a copy of the GNU Lesser General Public License  */
+/* along with this program; if not, write to the Free Software               */
+/* Foundation, Inc., 51 Franklin St, Fifth Floor, Boston, MA 02110-1301, USA.*/
+/*                                                                           */
+/*                                                                           */
[…]

### 2020-09-09 unittests/src/lexsort.c
add unit test for sorting

@@ -1,0 +1,97 @@
+/* * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * * */
+/*                                                                           */
+/* This file is part of SCIPSDP - a solving framework for mixed-integer      */
+/* semidefinite programs based on SCIP.                                      */
+/*                                                                           */
+/* Copyright (C) 2011-2013 Discrete Optimization, TU Darmstadt               */
+/*                         EDOM, FAU Erlangen-Nürnberg                       */
+/*               2014-2020 Discrete Optimization, TU Darmstadt               */
+/*                                                                           */
+/*                                                                           */
+/* This program is free software; you can redistribute it and/or             */
+/* modify it under the terms of the GNU Lesser General Public License        */
+/* as published by the Free Software Foundation; either version 3            */
+/* of the License, or (at your option) any later version.                    */
+/*                                                                           */
+/* This program is distributed in the hope that it will be useful,           */
+/* but WITHOUT ANY WARRANTY; without even the implied warranty of            */
+/* MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the             */
+/* GNU Lesser General Public License for more details.                       */
+/*                                                                           */
+/* You should have received a copy of the GNU Lesser General Public License  */
+/* along with this program; if not, write to the Free Software               */
+/* Foundation, Inc., 51 Franklin St, Fifth Floor, Boston, MA 02110-1301, USA.*/
+/*                                                                           */
+/*                                                                           */
+/* Based on SCIP - Solving Constraint Integer Programs                       */
+/* Copyright (C) 2002-2020 Zuse Institute Berlin                             */
+/* SCIP is distributed under the terms of the SCIP Academic Licence,         */
+/* see file COPYING in the SCIP distribution.                                */
+/*                                                                           */
[…]

### 2020-09-09 src/scipsdp/SdpVarfixer.c
use lexicographic sorting for matrix entries

@@ -44,6 +44,13 @@
 /* turn off lint warnings for whole file: */
 /*lint --e{788,818}*/
 
+/* defines for lexicographic sorting macros */
+#define SORTTPL_NAMEEXT     IntIntReal
+#define SORTTPL_KEYTYPE     int
+#define SORTTPL_FIELD1TYPE  SCIP_Real
+#include "sorttpllex.c" /*lint !e451*/
+
+
 /** sort the given row, col and val arrays first by non-decreasing row-indices, then for those with identical
  *  row-indices by non-decreasing col-indices
  */
@@ -54,26 +61,18 @@ void SCIPsdpVarfixerSortRowCol(
    int                   length              /**< length of the given arrays */
    )
 {
-   int firstentry;
-   int nextentry;
+   SCIPlexSortIntIntReal(row, col, val, length);
 
-   /* first sort by row indices */
-   SCIPsortIntIntReal(row, col, val, length);
-
-   /* for those with identical row-indices now sort by non-decreasing col-index, first find all entries with the same row-index */
-   nextentry = 0;
-   while (nextentry < length)
+#ifndef NDEBUG
    {
-      firstentry = nextentry; /* the next row starts where the last one ended*/
-
-      while (nextentry < length && row[nextentry] == row[firstentry]) /* as long as the row still matches, increase nextentry */
+      int j;
+      for (j = 0; j < length-1; ++j)
       {
-         nextentry++;
+         assert( row[j] <= row[j] );
+         assert( row[j] != row[j] || col[j] <= col[j] );
       }
-
-      /* now sort all entries between firstentry and nextentry-1 by their col-indices */
-      SCIPsortIntReal(col + firstentry, val + firstentry, nextentry - firstentry);
    }
+#endif
 }
 
 /** merges two three-tuple-arrays together

### 2020-09-09 src/scipsdp/sorttpllex.c
lint warnings

@@ -159,12 +159,12 @@ void SORTTPL_NAME(sorttpl_lexShellSort, SORTTPL_NAMEEXT)
 (
    SORTTPL_KEYTYPE*      key1,               /**< pointer to data array that defines the primary order */
    SORTTPL_KEYTYPE*      key2,               /**< pointer to data array that defines the secondary order */
-   SORTTPL_HASFIELD1PAR(  SORTTPL_FIELD1TYPE*    field1 )      /**< additional field that should be sorted in the same way */
-   SORTTPL_HASFIELD2PAR(  SORTTPL_FIELD2TYPE*    field2 )      /**< additional field that should be sorted in the same way */
-   SORTTPL_HASFIELD3PAR(  SORTTPL_FIELD3TYPE*    field3 )      /**< additional field that should be sorted in the same way */
-   SORTTPL_HASFIELD4PAR(  SORTTPL_FIELD4TYPE*    field4 )      /**< additional field that should be sorted in the same way */
-   SORTTPL_HASFIELD5PAR(  SORTTPL_FIELD5TYPE*    field5 )      /**< additional field that should be sorted in the same way */
-   SORTTPL_HASFIELD6PAR(  SORTTPL_FIELD6TYPE*    field6 )      /**< additional field that should be sorted in the same way */
+   SORTTPL_HASFIELD1PAR(  SORTTPL_FIELD1TYPE*    field1 )      /**< additional field that should be sorted in the same way */ /*lint !e665*/
+   SORTTPL_HASFIELD2PAR(  SORTTPL_FIELD2TYPE*    field2 )      /**< additional field that should be sorted in the same way */ /*lint !e665*/
+   SORTTPL_HASFIELD3PAR(  SORTTPL_FIELD3TYPE*    field3 )      /**< additional field that should be sorted in the same way */ /*lint !e665*/
+   SORTTPL_HASFIELD4PAR(  SORTTPL_FIELD4TYPE*    field4 )      /**< additional field that should be sorted in the same way */ /*lint !e665*/
+   SORTTPL_HASFIELD5PAR(  SORTTPL_FIELD5TYPE*    field5 )      /**< additional field that should be sorted in the same way */ /*lint !e665*/
+   SORTTPL_HASFIELD6PAR(  SORTTPL_FIELD6TYPE*    field6 )      /**< additional field that should be sorted in the same way */ /*lint !e665*/
    int                   start,              /**< starting index */
[…]

### 2020-09-13 tex/MISDP-Presolving/MISDPpresolving.tex
some rework, added other Li and Xie inequalities

@@
 \author{Marc E. Pfetsch\thanks{Department of Mathematics, TU Darmstadt, Germany}}
-\date{\today}
+\date{September 2020}
 
@@
 The following inequalities are known from the literature, see
-Gally~\cite{Gal19}, and can be added as linear inequalities w.r.t.\
+Mars~\cite{Mar13} and  Gally~\cite{Gal19}, and can be added as linear inequalities w.r.t.\
 \eqref{MISDP}:
@@
   \]
+  Note that the conditions imply that $A^0$ is not psd.
 \item For all $X \succeq 0$, we have
@@
 
-All the above methods are implemented and help for some instances.
+All the above methods are implemented and seem to help for some instances.
 
@@
   Moreover, since the diagonal terms of $X$ are nonnegative, we have
-  $U_{ii} \geq L_{ii} \geq 0$ for all $i \in [n]$. Upper bounding via
+  $U_{ii} \geq L_{ii} \geq 0$ for all $i \in [n]$. Thus,
+
+  Upper bounding via
   McCormick inequalities yields:
@@
                     U_{jj}\, X_{ii} + L_{ii}\, X_{jj} - L_{ii}\, U_{jj}\}  \\
-    & \leq U_{ii}\, U_{jj}.
+    & \leq \min \{L_{jj}\, U_{ii} + U_{ii}\, U_{jj} - U_{ii}\, L_{jj},\;
+                    U_{jj}\, U_{ii} + L_{ii}\, U_{jj} - L_{ii}\, U_{jj}\}  \\
+    & = \min \{U_{ii}\, U_{jj},\;  U_{jj}\, U_{ii}\} = U_{ii}\, U_{jj}.
   \end{align*}
@@
 constraint $x + 2\, y \geq 1$. Then the coefficient of $y$ can be tightened
-to $1$ yielding $x + y \geq 1$.
+to $1$, yielding $x + y \geq 1$.
 
@@
 for $k \in [m]$. Then we can compute
-\[
+\begin{equation}\label{eq:ComputeHatMu}
   \hat{\mu}_k = \min\, \{ \mu \geq 0 \suchthat \mu A^k - A^0 \succeq 0\}
-\]
-adn define $\mu_k \define \min\, \{\hat{\mu}_k, 1\}$. Note that $\mu_k = 1$ if
+\end{equation}
+and define $\mu_k \define \min\, \{\hat{\mu}_k, 1\}$. Note that $\mu_k = 1$ if
 $\hat{\mu}_k = \infty$, i.e., if the above optimization problem is infeasible.
@@
     \lambda v \Leftrightarrow Bv = (\mu - \lambda)v$, so $v$ is an
-    eigenvector for eigenvalue $\mu - \lambda$.)}  thus, to make
+    eigenvector for the eigenvalue $\mu - \lambda$.)}  Thus, to make
   $\mu\, I - V^{-\top} A^0 V^{-1}$ psd, we need
@@
 \end{itemize}
+
+Note that~\eqref{eq:ComputeHatMu} might be infeasible, e.g., if $A^0$ is
+not negative definite and $A^k$ is psd with minimal eigenvalue 0. An
+extreme example is if $A^k = 0$, but $A^0 \neq 0$. Thus, the implementation
+of these methods have to be safe-guarded against infeasibility.
+
+\begin{question}
+  Is there an easy way to determine whether~\eqref{eq:ComputeHatMu} is
+  infeasible, e.g., by computing eigenvalues once?  \footnote{For example
+    is the following true: If $\mu A^k - A^0$ is not psd for
+    $\mu \geq \lambda_{\max}(A^0)/\lambda_{\max}(A^k)$ then~\eqref{eq:ComputeHatMu} is
+    infeasible? This would require that $\mu A^k - A^0 \succeq 0$ for all
+    $\mu \geq \lambda_{\max}(A^0) / \lambda_{\max}(A^k)$, e.g., if $A^0$ is
+    psd and if there exists any $\mu'$ such that $\mu' A^k - A^0 \succeq 0$.}
+\end{question}
 
@@
 get: $\tilde{X} \succeq 0$ if and only if $t \geq 0$, $t - x\T x \geq
-0$. The latter is equivalent to $(x,t) \in L_{n+!}$.
+0$. The latter is equivalent to $(x,t) \in L_{n+1}$.
 
@@
 
-\paragraph{A Valid SOCP Constraint}
+\paragraph{Two Valid SOCP Constraint}
 
@@
 \]
-This holds, since in this case $X = x x\T$ and $\sum_{j=1}^m x_j^2 =
-1$. Thus,
+This holds, since in this case on can assume that $X = x x\T$ and $\sum_{j=1}^m x_j^2 =
+1$.\footnote{Note that this might cut off optimal solutions that do not
+  have rank 1.} Thus,
 \[
@@
 
+A second type of inequality of Li and Xie~\cite{LiX20} is
+\[
+  \bigg(\sum_{j \in [n]} \abs{X_{ij}}\bigg)^2 \leq k\, X_{ii}\, z_i\quad
+  \forall\, i \in [n].
+\]
+These inequalities seem to be nontrivial
+\footnote{Note that the inequality $\sum_{j \in [n]} \abs{X_{ij}} \leq k$ is trivial,
+since $\abs{X_{ij}} \leq z_j$ and thus
+\[
+  \sum_{j \in [n]} \abs{X_{ij}} \leq \sum_{j \in [n]} z_j \leq k.
+\]
+Similarly, $\sum_{j \in [n]} \abs{X_{ij}} \leq z_i$ holds, but it seems
+that $\sum_{j \in [n]} \abs{X_{ij}} \leq X_{ii}\, z_i$ does not hold in general.}
+and follow, because
+\[
+  \sum_{j \in [n]} \abs{X_{ij}} = \sum_{j \in [n]} \abs{x_i} \abs{x_j} =
+  \abs{x_i} \sum_{j \in [n]} \abs{x_j} \leq \abs{x_i} \sqrt{k},
+\]
+where we used that $\norm{x}_1 \leq \sqrt{k}$ if $\norm{x}_2 = 1$ and
+$\norm{x}_0 \leq k$. Moreover, $\abs{x_i} = \sqrt{X_{ii}} = \sqrt{X_{ii}
+  z_i}$, because if $z_i = 0$ then $X_{ii} = 0$.
+
+These inequalities are SOC representable, but one needs to express the
+absolute values in $\sum_{j \in [n]} \abs{X_{ij}}$, e.g., by variable
+splitting.
+
 \subsection{Minimization Variant}

### 2020-09-15 RIP-MISDP-Paper/scripts/RandomInstancesMISDP.tex
update documentation of RIP-CBF-formulation

@@
   \end{dense_itemize}
-  Output: Instances~\textsf{Matrices/type.m.n.k\_MISDP.side.pd.rank.socp.strgbnds[.nb]}, where
+  Output: Instances~\textsf{Matrices/type.m.n.k\_MISDP.side.pd.rank.socp.strgbnds.trineq.boundver}, where
   \begin{dense_itemize}
   \item ``type'', $m$, $n$, $k$ : as above
-  \item ``side''          : ``l'' (left side/$\alpha_k$), ``r'' (right side/$\beta_k$)
+  \item ``side''          : ``$\ell$'' (left side/$\alpha_k$), ``r'' (right side/$\beta_k$)
   \item ``pd''            : p (primal form using matrix variables), d (dual
@@
     if~$A \geq 0$ entrywise), 0 (with)
-  \item ``nobnds'' : ``nb'' (without stronger bounds~$-\tfrac{1}{2}z_j \leq X_{ij} \leq
-\tfrac{1}{2}z_j$ for $i \neq j$), else nothing is added to the filename.
+  \item ``trineq''        : 1 (add the trace constraint as inequality
+    constraint, that is $\tr(X) \leq 1$ for right side, resp. $\tr(X) \geq
+    1$ for th left side), 0 (use $\tr(X) == 1$ for both sides)
+  \item  ``boundver''      :
+    \begin{dense_itemize}
+    \item 0: $-z_i \leq X_{ii} \leq z_i,\; i \in [n]$, 
+    \item 1: $-z_j \leq X_{ij} \leq z_j,\; i,j \in [n]$,
+    \item 2: as 1 plus strengthened off-diagonal bounds
+      $-0.5z_j \leq X_{ij} \leq 0.5z_j$ for $i \neq j$.
+    \end{dense_itemize}
   \end{dense_itemize}
@@
   \item $k$: order
-  \item ``side'', ``Rank'', ``socp'', ``strgbnds'', ``nobnds'': see Output
+  \item ``side'', ``Rank'', ``socp'', ``strgbnds'', ``trineq'', ``boundver'': see Output
     of \textsf{generateRIPA.m} or Section~\ref{sec:CBFdual}.
@@
   \item $k$: order
-  \item ``side'', ``Rank'', ``socp'', ``strgbnds'', ``nobnds'': see Output
+  \item ``side'', ``Rank'', ``socp'', ``strgbnds'', ``trineq'', ``boundver'': see Output
     of \textsf{generateRIPA.m}, or Section~\ref{sec:CBFprimal}.
@@
 \textbf{OBJSENSE} \\
-MIN (if ``side'' = l) \; or \; MAX (if ``side'' = r) \\
+MIN (if ``side'' = $\ell$) \; or \; MAX (if ``side'' = r) \\
 \textbf{VAR} \\
@@
 \textbf{CON} \\
-\# cons: $2n^2+n+2 \quad {\color{blue}(n^2+n+2)}$, \qquad
-\# cones = $5 \quad {\color{blue} (4)}$\\
+\# cons: $2n^2+n+2 - {\color{blue}\text{strgbnds}\cdot n^2}$, \qquad
+\# cones = $5 - {\color{blue} \text{strgbnds}}$\\
+{\color{violet} If ``boundver'' = 0 : \# cons: $3n+2 - \text{strgbnds}\cdot n$} \\
 \[
   \begin{aligned}
-    &\text{L+} &n \qquad & -z_i + 1 \geq 0, \quad & i = 0,\dots,n-1 \\
-    &\text{L-} &n^2 \qquad & {\color{blue}-z_i - X_{ij} \leq 0},\quad & 0
-    \leq i,\, j \leq n-1 &\qquad {\color{blue} \text{(this line is cancelled)}}\\
-    &\text{L+} &n^2 \qquad & z_i - X_{ij} \geq 0,\quad & 0
-    \leq i, \, j \leq n-1 \\
-    &\text{L=}  &1 \qquad & \tr(X) - 1 = 0 \\
-    &\text{L-}  &1 \qquad & \sum z_i - k \leq 0
+    &\text{L+} &n\; {\color{violet}(n)} \qquad & -z_i + 1 \geq 0, \quad & i = 0,\dots,n-1 \\
+    &\text{L-} &n^2\; {\color{violet}(n)} \qquad & {\color{blue}-z_i - X_{ij} \leq 0},\quad & 0
+    \leq i,\, j \leq n-1 &\qquad {\color{blue} \text{(this line is
+        cancelled)}} \qquad {\color{violet} (\text{only for } i = j)}\\
+    &\text{L+} &n^2\; {\color{violet}(n)} \qquad & z_i - X_{ij} \geq 0,\quad & 0
+    \leq i, \, j \leq n-1 &\qquad {\color{violet} (\text{only for } i = j)}\\
+    &\text{L=}  &1\; {\color{violet}(1)} \qquad & \tr(X) - 1 = 0 \quad & \qquad & \qquad
+    {\color{cyan} \text{L}+ (\text{if ``side''} = \ell),\; \text{L}-
+      (\text{if ``side'' = r})} \\
+    &\text{L-}  &1\; {\color{violet}(1)} \qquad & \sum z_i - k \leq 0
   \end{aligned}
@@
 \textbf{ACOORD} (scalar cons $i$ | scalar var $j$ | val) \\
-\# entries: $4n^2+3n \quad {\color{blue} (- 2n^2)}$ \\
+\# entries: $4n^2+3n - {\color{blue} \text{strgbnds}\cdot 2n^2}$ \\
+{\color{violet} If ``boundver'' = 0 : \# entries: $7n -
+  \text{strgbnds}\cdot 2n$. In this case, all constraints are only added
+  for~$i=j$, so that the first column (scalar cons $i$) changes accordingly.} \\
 {\color{blue} If strgbnds = 1, then $n^2$ constraints are cancelled, so
@@
 \# entries: $n+2$ \\
+{\color{violet} If ``boundver'' = 0 : then all constraints are only added
+  for~$i=j$, so that the first column (scalar cons $i$) changes accordingly.} \\
 {\color{blue} If strgbnds = 1, then $n^2$ constraints are cancelled, so
@@
 \item If Rank = 1: \textcolor{green}{PSDCONRANK1} is added.
-\item If nobnds = 1: {\color{orange}
+\item If trineq = 1: In CON the definition of the cone for the trace
+  constraint is changed from~$\text{L}=$ to~\textcolor{cyan}{$\text{L-}$}
+  or~\textcolor{cyan}{$\text{L}+$}, respectively.
+\item If boundver = 0: The coupling constraints are only added for~$i = j$,
+  i.e., {\color{violet} $-z_i \leq X_{ii} \leq z_i$, $i \in [n]$}. This
+  changes the number of constraints in CON, and the number of entries in
+  ACOORD, since all entries with $i \neq j$ in ACOORD need to be
+  cancelled. Additionally, the first column (scalar cons $i$) in ACOORD and
+  BCOORD needs to be adapted. See the changes in
+  \textcolor{violet}{violet}.
+\item If boundver = 2: {\color{orange}
     $-\tfrac{1}{2}z_j \leq X_{ij} \leq \tfrac{1}{2}z_j$ for $i \neq j$} is
@@
 \textbf{CON} \\
-\# cons: $2n^2+n+2 \quad {\color{red}(+\tfrac{1}{2}n(n+3)(n+2))} \quad
-{\color{blue} (-\tfrac{1}{2}n(n-1))}$, \qquad
+\# cons: $2n^2+n+2 + {\color{red}\text{socp}\cdot(\tfrac{1}{2}n(n+3)(n+2))} - {\color{blue}
+    \text{strgbnds} \cdot (\tfrac{1}{2}n(n-1))}$, \qquad
 \# cones = $5 \quad {\color{red} (6)}$\\
+{\color{violet} If ``boundver'' = 0: \# cons: $3n+2 +
+  \text{socp}\cdot(\tfrac{1}{2}n(n+3)(n+2))$} \\
 \[
@@
     &\text{L+} &n \qquad & -z_i + 1 \geq 0, \quad & i = 0,\dots,n-1 \\
-    &\text{L-} &n^2\quad {\color{blue}(\tfrac{1}{2}n(n+1))} \qquad & -z_i - X_{ij} \leq 0,\;
+    &\text{L-} &n^2\; {\color{blue}(\tfrac{1}{2}n(n+1))}\; {\color{violet}(n)} \qquad & -z_i - X_{ij} \leq 0,\;
     {\color{blue} (-X_{ij} \leq 0)}\quad & 0 \leq i,\, j \leq n-1\quad {\color{blue}(0
-    \leq j \leq i \leq n-1)} \\
-    &\text{L+} &n^2 \qquad & z_i - X_{ij} \geq 0,\quad & 0
-    \leq j \leq i \leq n-1 \\
-    &\text{L=}  &1 \qquad & \tr(X) - 1 = 0 \\
+    \leq j \leq i \leq n-1)}\quad {\color{violet} (\text{only for } i=j)} \\
+    &\text{L+} &n^2\; {\color{violet}(n)} \qquad & z_i - X_{ij} \geq 0,\quad & 0
+    \leq j \leq i \leq n-1 \quad {\color{violet} (\text{only for } i=j)} \\
+    &\text{L=}  &1 \qquad & \tr(X) - 1 = 0 \quad & \qquad
+    {\color{cyan} \text{L+ (if ``side''} = \ell),\; \text{L- (if ``side'' = r})} \\
     &\text{L-}  &1 \qquad & \sum z_i - k \leq 0
@@
 \textbf{FCOORD} (scalar cons $i$ | psd var $j$ | row | col | val) \\
-\# entries: $2n^2+n$ \quad {\color{red} $+
-  n\cdot(3n+5+\tfrac{1}{2}n(n+1))$ } \quad {\color{blue} $- \tfrac{1}{2}n(n-1)$}\\
+\# entries: $2n^2+n$ + {\color{red} $\text{socp} \cdot
+  n\cdot(3n+5+\tfrac{1}{2}n(n+1))$ } - {\color{blue} $\text{strgbnds} \cdot
+  \tfrac{1}{2}n(n-1)$}\\
+{\color{violet} If ``boundver'' = 0: \# entries: $3n + \text{socp} \cdot
+  n\cdot(3n+5+\tfrac{1}{2}n(n+1)) - \text{strgbnds} \cdot
+  \tfrac{1}{2}n(n-1)$.}\\
 {\color{red} The following entries need to be added for all~$i = 0,\dots,n-1$}:
@@
 {\color{red} If the valid SOCP-inequality is added, then $\tfrac{1}{2}n(n+3)(n+2)$ needs
-to be added to the entries in the first column (scalar cons $i$).}
+  to be added to the entries in the first column (scalar cons $i$).}\\
+{\color{violet} If ``boundver = 0'', then all constraints are only added
+  for~$i=j$, so that the first column (scalar cons $i$) changes accordingly.}\\
 \[
@@
 \textbf{ACOORD} (scalar cons $i$ | scalar var $j$ val) \\
-\# entries: $2n^2+2n$ \qquad {\color{red} $(+2n)$} \qquad {\color{blue}
-  $(-n^2)$} \\
+\# entries: $2n^2+2n$ + {\color{red} $\text{socp}\cdot 2n$} - {\color{blue}
+  $\text{strgbnds} \cdot n^2$} \\
+{\color{violet} If ``boundver'' = 0: \# entries: $4n +\text{socp}\cdot 2n
+  - \text{strgbnds} \cdot n$} \\
 \[
@@
 {\color{red} If the valid SOCP-inequality is added, then $\tfrac{1}{2}n(n+3)(n+2)$ needs
-  to be added to the entries in the first column (scalar cons $i$).}
+  to be added to the entries in the first column (scalar cons $i$).}\\
+{\color{violet} If ``boundver = 0'', then all constraints are only added
+  for~$i=j$, so that the first column (scalar cons $i$) changes accordingly.}\\
 \[
@@
 \item If Rank = 1: \textcolor{green}{PSDVARRANK1} is added.
-\item If nobnds = 1: {\color{orange}
+\item If trineq = 1: In CON the definition of the cone for the trace
+  constraint is changed from~$\text{L}=$ to~\textcolor{cyan}{$\text{L-}$}
+  or~\textcolor{cyan}{$\text{L}+$}, respectively.
+\item If boundver = 0: The coupling constraints are only added for~$i = j$,
+  i.e., {\color{violet} $-z_i \leq X_{ii} \leq z_i$, $i \in [n]$}. This
+  changes the number of constraints in CON, and the number of entries in
+  FCOORD and ACOORD, since all entries with $i \neq j$ in FCOORD and ACOORD
+  need to be cancelled. Additionally, the first column (scalar cons $i$) in
+  FCOORD, ACOORD and BCOORD needs to be adapted. See the changes in
+  \textcolor{violet}{violet}.
+\item If boundver = 2: {\color{orange}
     $-\tfrac{1}{2}z_j \leq X_{ij} \leq \tfrac{1}{2}z_j$ for $i \neq j$} is

### 2020-11-08 tex/MISDP-Presolving/MISDPpresolving.tex
typo
add more potenitallly usefull inequalities

@@
 
+\paragraph{Bounds for Rank 1 Motivated Structures}
+
+If the constraints contain
+\[
+  \begin{pmatrix}
+    1 & x\T \\
+    x & X
+  \end{pmatrix}
+  \succeq 0,
+  \quad
+  \ell \leq x \leq u,
+\]
+then Nohra et al.~\cite{NohRS20} propose the following inequalities for all
+$i \in [n]$:
+\begin{align*}
+  X_{ii} & \geq 2\, \ell_i x_i - \ell_i^2,\\ 
+  X_{ii} & \geq 2\, u_i x_i - u_i^2.
+\end{align*}
+For example, the validity of the first type can be seen as follows:
+\[
+  (x_i - \ell_i)^2 \geq 0 \quad\Leftrightarrow\quad x_i^2 \geq 2\,
+  \ell_i x_i - \ell_i^2.
+\]
+Combining with the $2 \times 2$-minor inequality $X_{ii} - x_i^2 \geq 0$
+yields the inequality. The inequality $X_{ii} \geq 2\, u_i x_i - u_i^2$
+arises similarly.
+
+The given inequalities are not yet implemented, but the results
+in~\cite{NohRS20} suggest that they might be usefull if the LP-relaxation
+is solved.
+
+The above structure arises from relaxing quadratic problems. In this case,
+additionally $X = x x\T$ and the rank of $X$ has to be 1. Then additional
+inequalities have to hold, for example McCormick inequalities.
+
 % -------------------------------------------------------------------------
@@
 We apply the Perron-Frobenius theorem. To this end, we define a matrix $A$ to
-be reducible if there exists a permutation matrix $P$ such that
+be \emph{reducible} if there exists a permutation matrix $P$ such that
 \[
@@
 \]
-Otherwise it is \emph{irreducble}. We have the following:
+Otherwise it is \emph{irreducible}. We have the following:
 
@@
 
-\paragraph{Two Valid SOCP Constraint}
+\paragraph{Two Valid SOCP Constraints}
 

### 2020-11-10 RIP-MISDP-Paper/scripts/RandomInstancesMISDP.tex
correct and update random instances and RIP instances document

@@
   \end{tabular}
+  \caption{Parameters used by Kobayashi and Takano~\cite{Kob20}}.
+  \label{tab:KTinstanceSizes}
 \end{table}
@@
 script~\textsf{SCIPSDP/tools/InstanceGeneration/generateRandomMISDP.prl}
-takes as input~$(n,m_b,m_c,\alpha)$ and generates the
-instance~\textsf{randomMISDP\_n\_mb\_mc\_alpha.cbf} of the
-form~\eqref{eq:DualMISDP} in the CBF-format,
-see~\url{http://cblib.zib.de/}.
+takes as input~$(n,m_b,m_c,\alpha,\ell)$ and generates~$\ell$ random
+instances~\textsf{randomMISDP\_n\_mb\_mc\_alpha\_k.cbf}
+for~$k = 1,\dots,\ell$ of the form~\eqref{eq:DualMISDP} in the CBF-format,
+see~\url{http://cblib.zib.de/}. The
+script~\textsf{SCIPSDP/tools/InstanceGeneration/generateInstances.sh} can
+be used to generate such instances using the parameters from
+Table~\ref{tab:KTinstanceSizes}.
 
@@
 
-Li and Xie~\cite{LiX20} propose two valid inequalities for strengthening the MISDP:
+Li and Xie~\cite{LiX20} propose two valid inequalities for strengthening
+the MISDP formulation of SPCA, i.e., upper bound ($\beta_k$) of the RIP:
 \begin{align*}
@@
   \item ``rank''          : 1 (with rank-1 constraint), 0 (without)
-  \item ``socp''          : 1 (with valid inequalities from~\cite{LiX20}, see~\eqref{eq:ValidSOCPineqSDP}), 0 (without)
+  \item ``socp'' : variant of the valid inequalities from~\cite{LiX20},
+    see~\eqref{eq:ValidSOCPineqSDP}. Only for right side!
+    \begin{dense_itemize}
+      \item 0: no valid inequalities
+      \item 1: valid inequalities as SDP-constraints,
+        see~\eqref{eq:ValidSOCPineqSDP}
+      \item 2: valid inequalities directly as SOC-constraints
+    \end{dense_itemize}
   \item ``strgbnds''      : 1 (with stronger lower bounds~$X_{ij} \geq 0$
-    if~$A \geq 0$ entrywise), 0 (with)
+    if~$A \geq 0$ entrywise), 0 (without)! Only for right side!
   \item ``trineq''        : 1 (add the trace constraint as inequality
@@
 generated by
-\textsf{SCIPSDP/RIP-MISDP-Paper/scripts/RIPCBFdual.m}.\\
+\textsf{SCIPSDP/RIP-MISDP-Paper/scripts/RIPCBFdual.m}. Note that the valid
+SOCP inequality and the strengthening of the lower bounds if~$A$ is
+nonnegative is only valid for the right side of the RIP! \\
 
@@
 \textbf{CON} \\
-\# cons: $2n^2+n+2 - {\color{blue}\text{strgbnds}\cdot n^2}$, \qquad
-\# cones = $5 - {\color{blue} \text{strgbnds}}$\\
-{\color{violet} If ``boundver'' = 0 : \# cons: $3n+2 - \text{strgbnds}\cdot n$} \\
+\# cons: $2n^2+n+2 - {\color{blue}\text{strgbnds}\cdot n^2} +
+{\color{magenta} \ones_{\text{socp} = 2} \cdot n(n+2)}$, \qquad
+\# cones = $5 - {\color{blue} \text{strgbnds}} + {\color{magenta}
+  \ones_{\text{socp} = 2} \cdot n}$ \\
+{\color{violet} If ``boundver'' = 0 : \# cons: $3n+2 - \text{strgbnds}\cdot
+  n + \ones_{\text{socp} = 2} \cdot n(n+2)$} \\
 \[
@@
       (\text{if ``side'' = r})} \\
-    &\text{L-}  &1\; {\color{violet}(1)} \qquad & \sum z_i - k \leq 0
+    &\text{L-}  &1\; {\color{violet}(1)} \qquad & \sum z_i - k \leq 0 \\
+    &{\color{magenta} \text{Q}} & {\color{magenta} n+2} \qquad &
+    {\color{magenta} C_i \in L_{n+2}} \qquad &
+    {\color{magenta} \text{n times}}
   \end{aligned}
@@
 \textbf{ACOORD} (scalar cons $i$ | scalar var $j$ | val) \\
-\# entries: $4n^2+3n - {\color{blue} \text{strgbnds}\cdot 2n^2}$ \\
-{\color{violet} If ``boundver'' = 0 : \# entries: $7n -
-  \text{strgbnds}\cdot 2n$. In this case, all constraints are only added
+\# entries: $4n^2+3n - {\color{blue} \text{strgbnds}\cdot 2n^2} +
+{\color{magenta} \ones_{\text{socp} = 2} \cdot n(n+4)}$ \\
+{\color{violet} If ``boundver'' = 0 : \# entries:
+  $7n - \text{strgbnds}\cdot 2n + \ones_{\text{socp} = 2} \cdot n(n+4)$. In
+  this case, all constraints are only added
   for~$i=j$, so that the first column (scalar cons $i$) changes accordingly.} \\
 {\color{blue} If strgbnds = 1, then $n^2$ constraints are cancelled, so
-  that $n^2$ needs to be subtracted from the first column
-  (scalar cons $i$) in the fourth to seventh line.}
+  that $n^2$ needs to be subtracted from the first column (scalar cons $i$)
+  in the fourth to seventh line.}
 \[
@@
     0,\dots,n-1 \\
-    & n+in+j \quad & i \quad & -1.0\; (i = j),\,
-    {\color{orange} -\tfrac{1}{2}\; (i\neq j)} \quad & {\color{blue}-z_j -
-      X_{ij} \leq 0}
+    & n+in+j \quad & i \quad & -1.0\; (i = j),\, {\color{orange}
+      -\tfrac{1}{2}\; (i\neq j)} \quad & {\color{blue}-z_j - X_{ij} \leq 0}
     \quad & 0 \leq i,\, j \leq n-1\\
-    & n+in+j \quad & n+\tfrac{1}{2}i(i+1)+j \quad & -1.0
-    \quad & {\color{blue}-z_j - X_{ij} \leq 0} \quad & 0 \leq i,\, j \leq
+    & n+in+j \quad & n+\tfrac{1}{2}i(i+1)+j \quad & -1.0 \quad &
+    {\color{blue}-z_j - X_{ij} \leq 0} \quad & 0 \leq i,\, j \leq
     n-1, \; i \geq j \\
-    & n+in+j \quad & n+\tfrac{1}{2}j(j+1)+i \quad & -1.0
-    \quad & {\color{blue}-z_j - X_{ij} \leq 0} \quad & 0 \leq i,\, j \leq
+    & n+in+j \quad & n+\tfrac{1}{2}j(j+1)+i \quad & -1.0 \quad &
+    {\color{blue}-z_j - X_{ij} \leq 0} \quad & 0 \leq i,\, j \leq
     n-1, \; i < j \\
-    & n+n^2+in+j \quad & i \quad & 1.0\; (i
-    = j),\, {\color{orange} \tfrac{1}{2}\; (i\neq j)} \quad & z_j - X_{ij} \geq 0
-    \quad & 0
+    & n+n^2+in+j \quad & i \quad & 1.0\; (i = j),\, {\color{orange}
+      \tfrac{1}{2}\; (i\neq j)} \quad & z_j - X_{ij} \geq 0 \quad & 0
     \leq i,\, j \leq n-1 \\
-    & n+n^2+in+j \quad &
-    n+\tfrac{1}{2}i(i+1)+j \quad & -1.0 \quad & z_j - X_{ij} \geq 0 \quad &
+    & n+n^2+in+j \quad & n+\tfrac{1}{2}i(i+1)+j \quad & -1.0 \quad & z_j -
+    X_{ij} \geq 0 \quad &
     0 \leq i,\, j \leq n-1,\; i \geq j \\
-    & n+n^2+in+j \quad &
-    n+\tfrac{1}{2}j(j+1)+i \quad & -1.0 \quad & z_j - X_{ij} \geq 0 \quad &
+    & n+n^2+in+j \quad & n+\tfrac{1}{2}j(j+1)+i \quad & -1.0 \quad & z_j -
+    X_{ij} \geq 0 \quad &
     0 \leq i,\, j \leq n-1,\; i < j \\
@@
     & n+2n^2+1 \quad & i \quad & 1.0 \quad & \sum z_i - k \geq 0 \quad &
-    i = 0,\dots,n-1
+    i = 0,\dots,n-1 \\
+    & {\color{magenta} n+2n^2+2+...} \quad &
+    {\color{magenta}n+\tfrac{1}{2}i(i+1)+i} \quad & {\color{magenta}0.5}
+    \quad & {\color{magenta}C_i \in L_{n+2}} \quad & {\color{magenta}i =
+      0,\dots,n-1} \\
+    & {\color{magenta}n+2n^2+2+...} \quad & {\color{magenta} i} \quad &
+    {\color{magenta} 0.5} \quad & {\color{magenta} C_i \in L_{n+2}} \quad &
+    {\color{magenta} i = 0,\dots,n-1} \\
+    & {\color{magenta} n+2n^2+2+...}  \quad & {\color{magenta}
+      n+\tfrac{1}{2}i(i+1)+j} \quad & {\color{magenta} 0.5} \quad  &
+    {\color{magenta} C_i \in L_{n+2}} \quad & {\color{magenta} 0 \leq
+      i,j\leq n-1, i \geq j} \\
+    & {\color{magenta} n+2n^2+2+...}  \quad & {\color{magenta}
+      n+\tfrac{1}{2}j(j+1)+i} \quad & {\color{magenta} 0.5} \quad  &
+    {\color{magenta} C_i \in L_{n+2}} \quad & {\color{magenta} 0 \leq
+      i,j\leq n-1, i < j} \\
+    & {\color{magenta} n+2n^2+2+...}  \quad & {\color{magenta}
+      n+\tfrac{1}{2}i(i+1)+i} \quad & {\color{magenta} 0.5} \quad  &
+    {\color{magenta} C_i \in L_{n+2}} \quad & {\color{magenta} i =
+      0,\dots,n-1} \\
+    & {\color{magenta} n+2n^2+2+...}  \quad & {\color{magenta} i} \quad &
+    {\color{magenta} -0.5} \quad  & {\color{magenta} C_i \in L_{n+2}} \quad
+    & {\color{magenta} i = 0,\dots,n-1} \\
   \end{aligned}
 \]
+
 \textbf{BCOORD} (scalar cons $i$ | val) \\
@@
   lines in ACOORD with a \textcolor{blue}{blue constraint} are cancelled
-  as well.
+  as well. Only valid for the right side.
 \item If socp = 1: The valid SOCP inequalities
@@
   symmetric matrix of suitable dimension with ones at positions $(i,j)$ and
-  $(j,i)$, and zeros else.
+  $(j,i)$, and zeros else. Only valid for the right side.
+\item If socp = 2: The valid SOCP inequalities
+  $\sum_{j=0}^{n-1}X_{ij}^2 \leq X_{ii}z_i$ for $i = 0,\dots,n-1$
+  from~\cite{LiX20} are added directly as SOC-constraints, as those are
+  SOC-representable:
+  \begin{align*}
+    C_i \define \big(X_{i1}, \dots, X_{in}, \tfrac{1}{2}(X_{ii}-z_i),
+    \tfrac{1}{2}(X_{ii}+z_i \big) \in L_{n+2}.
+  \end{align*}
+  See the changes in CON and ACOORD in \textcolor{magenta}{magenta}.
 \end{itemize}
@@
 \textbf{CON} \\
-\# cons: $2n^2+n+2 + {\color{red}\text{socp}\cdot(\tfrac{1}{2}n(n+3)(n+2))} - {\color{blue}
-    \text{strgbnds} \cdot (\tfrac{1}{2}n(n-1))}$, \qquad
-\# cones = $5 \quad {\color{red} (6)}$\\
+\# cons:
+$2n^2+n+2 +
+{\color{red}\ones_{\text{socp}=1}\cdot(\tfrac{1}{2}n(n+3)(n+2))} +
+{\color{magenta}\ones_{\text{socp}=2}\cdot(n(n+2))} - {\color{blue}
+  \text{strgbnds} \cdot (\tfrac{1}{2}n(n-1))}$, \\
+\# cones = $5 + {\color{red} \ones_{\text{socp}=1}} + {\color{magenta}
+  \ones_{\text{socp}=2} \cdot n}$\\
 {\color{violet} If ``boundver'' = 0: \# cons: $3n+2 +
-  \text{socp}\cdot(\tfrac{1}{2}n(n+3)(n+2))$} \\
+  \ones_{\text{socp}=1}\cdot(\tfrac{1}{2}n(n+3)(n+2)) +
+  \ones_{\text{socp}=2}\cdot(n(n+2)) $} \\
 \[
@@
     {\color{cyan} \text{L+ (if ``side''} = \ell),\; \text{L- (if ``side'' = r})} \\
-    &\text{L-}  &1 \qquad & \sum z_i - k \leq 0
+    &\text{L-}  &1 \qquad & \sum z_i - k \leq 0 \\
+    &{\color{magenta} \text{Q}} & {\color{magenta} n+2} \qquad &
+    {\color{magenta} C_i \in L_{n+2}} \qquad &
+    {\color{magenta} \text{n times}}
  \end{aligned}
@@
 \textbf{FCOORD} (scalar cons $i$ | psd var $j$ | row | col | val) \\
-\# entries: $2n^2+n$ + {\color{red} $\text{socp} \cdot
-  n\cdot(3n+5+\tfrac{1}{2}n(n+1))$ } - {\color{blue} $\text{strgbnds} \cdot
+\# entries: $2n^2+n$ + {\color{red}
+  $\ones_{\text{socp}=1} \cdot n\cdot(3n+5+\tfrac{1}{2}n(n+1))$ } +
+{\color{magenta} $\ones_{\text{socp}=2} \cdot n(n+2)$ } - {\color{blue}
+  $\text{strgbnds} \cdot
   \tfrac{1}{2}n(n-1)$}\\
-{\color{violet} If ``boundver'' = 0: \# entries: $3n + \text{socp} \cdot
-  n\cdot(3n+5+\tfrac{1}{2}n(n+1)) - \text{strgbnds} \cdot
+{\color{violet} If ``boundver'' = 0: \# entries:
+  $3n + \ones_{\text{socp}=1} \cdot n\cdot(3n+5+\tfrac{1}{2}n(n+1)) +
+  \ones_{\text{socp}=2} \cdot n(n+2) - \text{strgbnds} \cdot
   \tfrac{1}{2}n(n-1)$.}\\
-{\color{red} The following entries need to be added for all~$i = 0,\dots,n-1$}:
+{\color{red} The following entries need to be added for
+  all~$i = 0,\dots,n-1$}:
 \[
@@
     &n+2n^2 \quad & 0 \quad & i \quad & i \quad & 1.0 \quad & \tr(X) - 1
-    \leq 0 \quad & i = 0,\dots,n-1
+    \leq 0 \quad & i = 0,\dots,n-1 \\
+    &{\color{magenta} n+2n^2+2+...} \quad & {\color{magenta} 0} \quad &
+    {\color{magenta} i} \quad & {\color{magenta} i} \quad &
+    {\color{magenta} 0.5} \quad & {\color{magenta} C_i \in L_{n+2}} \quad &
+    {\color{magenta} i = 0,\dots,n-1} \\
+    &{\color{magenta} n+2n^2+2+...} \quad & {\color{magenta} 0} \quad &
+    {\color{magenta} i} \quad & {\color{magenta} j} \quad &
+    {\color{magenta} 1.0} \quad & {\color{magenta} C_i \in L_{n+2}} \quad &
+    {\color{magenta} 0 \leq i,j\leq n-1, i \geq j} \\
+    &{\color{magenta} n+2n^2+2+...} \quad & {\color{magenta} 0} \quad &
+    {\color{magenta} j} \quad & {\color{magenta} i} \quad &
+    {\color{magenta} 1.0} \quad & {\color{magenta} C_i \in L_{n+2}} \quad &
+    {\color{magenta} 0 \leq i,j\leq n-1, i < j} \\
+    &{\color{magenta} n+2n^2+2+...} \quad & {\color{magenta} 0} \quad &
+    {\color{magenta} i} \quad & {\color{magenta} i} \quad &
+    {\color{magenta} 0.5} \quad & {\color{magenta} C_i \in L_{n+2}} \quad &
+    {\color{magenta} i = 0,\dots,n-1} \\
   \end{aligned}
@@
 \textbf{ACOORD} (scalar cons $i$ | scalar var $j$ val) \\
-\# entries: $2n^2+2n$ + {\color{red} $\text{socp}\cdot 2n$} - {\color{blue}
+\# entries: $2n^2+2n$ + {\color{red} $\ones_{\text{socp}> 0}\cdot 2n$} -
+{\color{blue}
   $\text{strgbnds} \cdot n^2$} \\
-{\color{violet} If ``boundver'' = 0: \# entries: $4n +\text{socp}\cdot 2n
-  - \text{strgbnds} \cdot n$} \\
+{\color{violet} If ``boundver'' = 0: \# entries: $4n +\ones_{\text{socp}>
+    0} \cdot 2n - \text{strgbnds} \cdot n$} \\
 \[
@@
     i = 0,\dots,n-1 \\
+    & {\color{magenta}n+2n^2+2+...} \quad & {\color{magenta} i} \quad &
+    {\color{magenta} 0.5} \quad & {\color{magenta} C_i \in L_{n+2}} \quad &
+    {\color{magenta} i = 0,\dots,n-1} \\
+    & {\color{magenta}n+2n^2+2+...} \quad & {\color{magenta} i} \quad &
+    {\color{magenta} -0.5} \quad & {\color{magenta} C_i \in L_{n+2}} \quad &
+    {\color{magenta} i = 0,\dots,n-1} \\
   \end{aligned}
@@
   bounds~$X_{ij} \geq 0$ for all $i,j =0,\dots,n-1$ are used. See the
-  changes in \textcolor{blue}{blue} in ACOORD.
+  changes in \textcolor{blue}{blue} in ACOORD. Only valid for the right
+  side.
 \item If socp = 1: The valid SOCP inequalities
@@
   symmetric matrix of suitable dimension with ones at positions $(i,j)$ and
-  $(j,i)$, and zeros else.
+  $(j,i)$, and zeros else. Only valid for the right side.
+\item If socp = 2: The valid SOCP inequalities
+  $\sum_{j=0}^{n-1}X_{ij}^2 \leq X_{ii}z_i$ for $i = 0,\dots,n-1$
+  from~\cite{LiX20} are added directly as SOC-constraints, as those are
+  SOC-representable:
+  \begin{align*}
+    C_i \define \big(X_{i1}, \dots, X_{in}, \tfrac{1}{2}(X_{ii}-z_i),
+    \tfrac{1}{2}(X_{ii}+z_i \big) \in L_{n+2}.
+  \end{align*}
+  See the changes in CON, FCOORD and ACOORD in
+  \textcolor{magenta}{magenta}.
 \end{itemize}

### 2020-12-04 tex/MISDP-Presolving/MISDPpresolving.tex
minor extension of argument

@@
 % -------------------------------------------------------------------------
-\subsection{Implied Linear Inequalities}
+\subsection{Implied Linear Inequalities}\label{sec:ImpliedLinEqs}
 
@@
 
-\begin{lemma}
+\begin{lemma}\label{lem:Propagation}
   For any solution $y \in \R^m$ of~\eqref{MISDP}, we have
@@
 could never be strengthened across the large testset.
+
+\begin{remark}
+  Note that in the pure form, the propagated inequalities from
+  Lemma~\ref{lem:Propagation} dominate the inequalities given in
+  Section~\ref{sec:ImpliedLinEqs}:
+  \begin{align*}
+    X_{st} \leq \sqrt{U_{ss} \cdot U_{tt}} \leq \frac{1}{2}\big( U_{ss} + U_{tt}\big),
+  \end{align*}
+  because of the inequality between the arithmetic and geometric mean. Does
+  this also hold in the general case of a (dual) LMI?
+\end{remark}
 
@@
 \begin{lemma}
-  $f(\mu)$ is concave.
+  For any two symmetric matrices $A^k$, $A^0 \in \R^{n \times n}$, the function
+  $f(\mu) = \lambda_{\min}(\mu\, A^k - A^0)$ is concave.
 \end{lemma}
@@
     & = \alpha\, \lambda_{\min}\big(\mu_1\, A^k - A^0\big) + (1 - \alpha)\, \lambda_{\min}\big(\mu_2\, A^k - A^0\big)\\
-    & = \alpha\, f(\mu_1) + (1 - \alpha)\, f(\mu_2),
+    & = \alpha\, f(\mu_1) + (1 - \alpha)\, f(\mu_2).
   \end{align*}
-  which shows the claim.
+  The inequality follows because for a symmetric matrix~$C \in \R^{n \times n}$:
+  \[
+    \lambda_{\min}(C) = \min_{\norm{x} = 1} x\T C x.
+  \]
+  Thus, for a symmetric matrix $D \in \R^{n \times n}$
+  \begin{align*}
+    \lambda_{\min}(C + D) & = \min_{\norm{x} = 1} x\T (C + D) x\\
+    & \geq \min_{\norm{x} = 1} x\T C  x + \min_{\norm{x} = 1} x\T D x =
+    \lambda_{\min}(C) + \lambda_{\min}(D),
+  \end{align*}
+  which shows the inequality.
 \end{proof}

### 2020-12-04 src/sdpi/sdpi.c
fixed return value

@@ -2948,7 +2948,7 @@ SCIP_Bool SCIPsdpiWasSolved(
 {
    assert( sdpi != NULL );
 
-   return ( sdpi->solved && SCIPsdpiSolverWasSolved(sdpi->sdpisolver) );
+   return sdpi->solved;
 }
 
 /** returns whether the original problem was solved, if SCIPsdpiWasSolved = true and SCIPsdpiSolvedOrig = false, then a penalty formulation was solved */

### 2020-12-05 src/sdpi/sdpi.c
add asserts and fix debug messages

@@ -858,7 +858,7 @@ SCIP_RETCODE computeLpLhsRhsAfterFixings(
             if ( lb > sdpilb[lpcol] + sdpi->epsilon )
             {
                /* this bound is stronger than the original one */
-               SCIPdebugMessage("Empty LP-row %d has been removed from SDP %d, lower bound of variable %d has been strenghened to %g "
+               SCIPdebugMessage("LP-row %d with one nonzero has been removed from SDP %d, lower bound of variable %d has been strenghened to %g "
                   "(originally %g)\n", currentrow, sdpi->sdpid, lpcol, lb, sdpilb[lpcol]);
                sdpilb[lpcol] = lb;
             }
@@ -867,7 +867,7 @@ SCIP_RETCODE computeLpLhsRhsAfterFixings(
             if ( ub < sdpiub[lpcol] - sdpi->epsilon )
             {
                /* this bound is stronger than the original one */
-               SCIPdebugMessage("Empty LP-row %d has been removed from SDP %d, upper bound of variable %d has been strenghened to %g "
+               SCIPdebugMessage("LP-row %d with one nonzero has been removed from SDP %d, upper bound of variable %d has been strenghened to %g "
                   "(originally %g)\n", currentrow, sdpi->sdpid, lpcol, ub, sdpiub[lpcol]);
                sdpiub[lpcol] = ub;
             }
@@ -1871,6 +1871,11 @@ SCIP_RETCODE SCIPsdpiLoadSDP(
          }
       }
    }
+   for (i = 0; i < nvars; ++i)
+   {
+      assert( lb[i] < SCIPsdpiInfinity(sdpi) );   /* lower bound should not be infinity */
+      assert( ub[i] > -SCIPsdpiInfinity(sdpi) );  /* upper bound should not be - infinity */
+   }
 #endif
 
    assert( nlpcons == 0 || lplhs != NULL );
@@ -1903,10 +1908,14 @@ SCIP_RETCODE SCIPsdpiLoadSDP(
       BMScopyMemoryArray(sdpi->sdpnblockvarnonz[b], sdpnblockvarnonz[b], sdpnblockvars[b]);
[…]

### 2020-12-05 src/scipsdp/cons_sdp.c
copy lock information

@@ -4560,7 +4560,7 @@ SCIP_DECL_CONSTRANS(consTransSdp)
       SCIP_CALL( SCIPduplicateBlockMemoryArray(scip, &(targetdata->val[i]), sourcedata->val[i], sourcedata->nvarnonz[i]) );
    }
    SCIP_CALL( SCIPallocBlockMemoryArray(scip, &(targetdata->vars), sourcedata->nvars) );
-   if ( sourcedata->locks )
+   if ( sourcedata->locks != NULL )
    {
       SCIP_CALL( SCIPduplicateBlockMemoryArray(scip, &(targetdata->locks), sourcedata->locks, sourcedata->nvars) );
    }
@@ -5516,6 +5516,15 @@ SCIP_DECL_CONSCOPY(consCopySdp)
       SCIP_CALL( SCIPcreateConsSdpRank1(scip, cons, copyname, sourcedata->nvars, sourcedata->nnonz, sourcedata->blocksize, sourcedata->nvarnonz,
             sourcedata->col, sourcedata->row, sourcedata->val, targetvars, sourcedata->constnnonz,
             sourcedata->constcol, sourcedata->constrow, sourcedata->constval, FALSE) );
+   }
+
+   /* copy lock information if available */
+   if ( sourcedata->locks != NULL )
+   {
+      SCIP_CONSDATA* targetdata;
+
+      targetdata = SCIPconsGetData(*cons);
+      SCIP_CALL( SCIPduplicateBlockMemoryArray(scip, &(targetdata->locks), sourcedata->locks, sourcedata->nvars) );
    }
 
    SCIPfreeBufferArray(scip, &targetvars);

### 2020-12-10 tex/MISDP-Presolving/MISDPpresolving.tex
fixed description of SOC formulation as SDP
add twominorprodconss and twominorsocconss to presolving document

@@
   $i \in [n]$:
-  \[
+  \begin{align}\label{eq:diaggezero}
     \sum_{k = 1}^m (A^k)_{ii}\, y_k - (A^0)_{ii} \geq 0.
-  \]
+  \end{align}
 \item If for some $i$, $j \in [n]$, $A^0_{ij} \neq 0$, and $A^0_{ii} = 0$,
@@
   the following inequality is valid:
-  \[
-    \sum_{\substack{k \in \mathcal{I}:\\ A^k_{ii} > 0}} y_k \geq 1.
-  \]
+  \begin{align}\label{eq:diagzeroimpl}
+    \sum_{\substack{k \in \mathcal{I}:\\ A^k_{ii} > 0}} y_k \geq1.
+  \end{align}
   Note that the conditions imply that $A^0$ is not psd.
 \item For all $X \succeq 0$, we have
-  \[
+  \begin{align}\label{eq:2minorlinP}
     X_{ii} + X_{jj} - 2\, X_{ij} \geq 0,\qquad
     X_{ii} + X_{jj} + 2\, X_{ij} \geq 0.
-  \]
+  \end{align}
   This follows by using a 2 by 2 minor and multiplying from left
@@
 \item The previous inequalities translated to the matrix pencil are:
+  \begin{equation}\label{eq:2minorlinD}
+    \begin{aligned}
+      & \sum_{k=1}^m A^k_{ii}\, y_k - A^0_{ii} + \sum_{k=1}^m A^k_{jj}\, y_k - A^0_{jj} - 2 \Big(\sum_{k=1}^m A^k_{ij}\, y_k - A^0_{ij}\Big) \geq 0\\
+      \quad\Leftrightarrow\quad& \sum_{k=1}^m \Big(A ^k_{ii} + A^k_{jj} -
+      2\, A^k_{ij}\Big)\, y_k \geq A^0_{ii} + A^0_{jj} - 2 A^0_{ij}.
+    \end{aligned}
+  \end{equation}
+\item If $A^k_{ii} = A^k_{jj} = 0$ for all~$k \in [m]$
+  and~$A^0_{ii}A^0_{jj} \geq 0$ for some $i \neq j \in [n]$, then the
+  following inequality is valid:
+  \begin{align}\label{eq:2minorprodD}
+    \sum_{k=1}^m A^k_{ij} y_k \geq A^0 - \sqrt{A^0_{ii}A^0_{jj}}.
+  \end{align}
+  Furhtermore, if exactly one~$A^k_{ij} \neq 0$, then this yields upper or
+  lower bounds for the corresponding variable~$y_k$, depending on the sign
+  of~$A^k_{ij}$.
+\item For a positive semidefinite matrix $X$ the following quadratic
+  inequality holds by using the 2 by 2 minor:
+  $X_{st}^2 \leq X_{ss} X_{tt}$. This constraint is SOC representable:
   \begin{align*}
-    & \sum_{k=1}^m A^k_{ii}\, y_k - A^0_{ii} + \sum_{k=1}^m A^k_{jj}\, y_k - A^0_{jj} - 2 \Big(\sum_{k=1}^m A^k_{ij}\, y_k - A^0_{ij}\Big) \geq 0\\
-    \quad\Leftrightarrow\quad&
-    \sum_{k=1}^m \Big(A ^k_{ii} + A^k_{jj} - 2\, A^k_{ij}\Big)\, y_k \geq A^0_{ii} + A^0_{jj} - 2 A^0_{ij}.
+    \left\|\binom{2\,X_{st}}{X_{ss} - X_{tt}}\right\| &\leq X_{ss} +
+                                                        X_{tt}. \\
+    \quad\Leftrightarrow\quad 4\, X_{st}^2 + (X_{ss}^2 - 2\,X_{ss}\, X_{tt}
+    + X_{tt}^2) & \leq X_{ss}^2 + 2\, X_{ss}\,X_{tt} + X_{tt}^2.
   \end{align*}
+  Translated to the matrix pencil notation the cut looks as follows:
+  \begin{equation}\label{eq:2minorSOC}
+    \begin{aligned}
+      &\left\|\binom{2\, \sum_{i=1}^m (A_i)_{st}\, y_i - 2\, (A_0)_{st}}{ q
+          \sum_{i=1}^m (A_i)_{ss}\, y_i - (A_0)_{ss} - \sum_{i=1}^m
+          (A_i)_{tt}\, y_i + (A_0)_{tt}}\right\| \\ &\leq \sum_{i=1}^m
+      (A_i)_{ss}\, y_i - (A_0)_{ss} + \sum_{i=1}^m (A_i)_{tt}\, y_i -
+      (A_0)_{tt}.
+    \end{aligned}
+  \end{equation}
+  These constraints can be added directly as SOC-constraints.
 \end{itemize}
@@
   Note that in the pure form, the propagated inequalities from
-  Lemma~\ref{lem:Propagation} dominate the inequalities given in
-  Section~\ref{sec:ImpliedLinEqs}:
+  Lemma~\ref{lem:Propagation} dominate the inequalities
+  in~\eqref{eq:2minorlinP}:
   \begin{align*}
@@
 \[
-  (x,t) \in L_{n+1} \quad\Leftrightarrow\quad
+  (x,\sqrt{t}) \in L_{n+1} \quad\Leftrightarrow\quad
   \tilde{X} \define
@@
 get: $\tilde{X} \succeq 0$ if and only if $t \geq 0$, $t - x\T x \geq
-0$. The latter is equivalent to $(x,t) \in L_{n+1}$.
+0$. The latter is equivalent to $(x,\sqrt{t}) \in L_{n+1}$.
 

### 2020-12-10 tex/MISDP-Presolving/MISDPpresolving.tex
add some todos to presolving document

@@
 \maketitle
+
+\todo[inline]{FM: We should argue that alle the inequalities and bounds are
+  as strong as possible.}
 
@@
 
+\todo[inline]{FM: If we tighten the coefficients, we need to scale the
+  linear constraints (for example, a trace constraint on~$A(y)$)
+  accordingly.}
+
 \begin{proof}
@@
 
+\todo[inline]{FM: If we tighten the bounds of a variable~$y_k$, we need to
+  scale the linear constraints accordingly.}
+
 \begin{proof}

### 2020-12-10 tex/MISDP-Presolving/MISDPpresolving.tex
add some remarks to presolving document

@@
   follows from~\eqref{eq:TwoByTwoMinorInequality}.
+  Note that an additional rank-1 constraint on the matrix~$X$ does not lead
+  to stronger inequalities.
 \item The previous inequalities translated to the matrix pencil are:
@@
 
-All the above methods are implemented and seem to help for some instances.
+All the above methods are implemented and seem to help for some
+instances.
+
+\begin{remark}
+  The assumptions under which inequalities~\eqref{eq:diagzeroimpl} are
+  applicable can be weakened. These inequalities are also valid
+  if~$A^0_{ii} > 0$, since, under the remaining assumptions,
+  \begin{align*}
+    0 \leq A(y)_{ii} \cdot A(y)_{jj} - (A^0_{ij})^2 \quad &\implies \quad
+    0 < A(y)_{ii} = \sum_{k} A^k_{ii}y_k - A^0_{ii}, \\
+    &\implies \quad 0 < \sum_{k} A^k_{ii}y_k
+  \end{align*}
+  and thus~$0 < \sum_{k} A^k_{ii}y_k$ has to hold. This implies the
+  validity of~\eqref{eq:diagzeroimpl}.
+\end{remark}
 
@@
 \end{proof}
+Note that these bounds are tight, even for a rank-1 matrix~$X$: Consider
+the all-ones matrix, which clearly has rank~1.
 
@@
 inequalities have to hold, for example McCormick inequalities.
+
+\begin{remark}
+  Most likely this can be generalized to the more interesting case of a
+  constraint of the form
+  \[
+  \begin{pmatrix}
+    t & x\T \\
+    x & X
+  \end{pmatrix}
+  \succeq 0,
+  \quad
+  \ell \leq (x,t) \leq u,
+\]
+where~$t$ is a scalar variable. Since constraints of this particular form
+appear in truss topology design and cardinality least-squares problems, we
+should check whether this helps for such instances.
+\end{remark}
 

### 2020-12-11 tex/MISDP-Presolving/MISDPpresolving.tex
minor

@@
   x_k^2 \leq t^2 - \sum_{\substack{i=1\\ i \neq k}}^n x_i^2 \leq t^2 -
-  \sum_{\substack{i=1\\ i \neq k}} \gamma_i \leq t^2,\qquad
+  \sum_{\substack{i=1\\ i \neq k}}^n \gamma_i \leq t^2,\qquad
   \gamma_i \define \min_{\gamma \in [\ell_i, u_i]} \gamma^2.
@@
 Thus, one question is whether one can derive strengthened bounds. This is
-currently unclear. As one idea, one can consider generalized Schur
+currently unclear, but it does not seem to be possible.
+
+\paragraph*{Some Ideas}
+As one idea, one can consider generalized Schur
 polynomials (since $X$ is not necessarily positive definite, one needs the

### 2020-12-11 tex/MISDP-Presolving/MISDPpresolving.tex
add part on convex hulls for RIP

@@
 
+\begin{question}
+  One idea is to perform spatial branching on the diagonal entries
+  $X_ii$. Does this help to solve the problem?
+\end{question}
+
+\subsubsection{Complete Description of Feasible Set}
+
+One interesting question is whether one can describe the convex hull of the
+feasible set. For this define
+\begin{align*}
+  S = \Big\{(X,z) \in S^n_+ \times [0,1]^n \suchthat & \tr(X) = 1,\; \sum_{i=1}^n z_i \leq k,\;     0 \leq X_{ii} \leq z_j \;\forall\, i \in [n],\\
+                                                     & -\tfrac{1}{2} z_j \leq X_{ij} \leq \tfrac{1}{2} z_j \;\forall\, i \neq j \in [n] \Big\},
+\end{align*}
+which is the strongest relaxation that we currently have. Its integer hull is
+\begin{align*}
+  S_I \define \conv \Big\{& (X,z) \in S^n_+ \times \{0,1\}^n \suchthat \tr(X) =
+                1,\; \sum_{i=1}^n z_i \leq k,\; 0 \leq X_{ii} \leq z_j \;\forall\, i \in [n] \Big\}.
+\end{align*}
+In general, we have $S \neq S_I$: If $k = 1$, then all off-diagonals have
+to be zero in $S_I$, but not necessarily in~$S$. Moreover, consider
+$2 \leq k \leq n$. Then for
+\[
+  z = \begin{pmatrix}
+    \tfrac{1}{k} \\
+    \vdots\\
+    \tfrac{1}{k}
+  \end{pmatrix},\quad
+  X \define
+  \begin{pmatrix}
+    \tfrac{1}{k} & \dots, & \tfrac{1}{k}\\
+    \vdots & \ddots & \vdots\\
+    \tfrac{1}{k} & \dots & \tfrac{1}{k}
+  \end{pmatrix},
+\]
+we have $(X,z) \in S$, but $(X,z) \notin S_I$.
+
+\paragraph{Linear Case}
+
+It might be instrumential to look at the linear analogue:
+\[
+  S \define \{(x,z) \in \R_+^n \times [0,1]^n \suchthat \sum_{i=1}^n x_i =
+  1,\; \sum_{i=1}^n z_i \leq k,\; 0 \leq x_i \leq z_i \;\forall\, i \in [n]\}.
+\]
+This is related to the \emph{hypersimplex}
+\[
+  \Delta_{n,k} \define \conv \{z \in \{0,1\}^n \suchthat \sum_{i=1}^n z_i = k\},
+\]
+for which complete linear descriptions are available, see
+Ziegler~\cite{Zie95}.
+
+\begin{question}
+  Does this give a linear description of $S$ and can this be extended to
+  the SDP case?
+\end{question}
+
+\subsubsection{Nonnegativity by the Perron-Frobenius Theorem}
+
 We apply the Perron-Frobenius theorem. To this end, we define a matrix $A$ to
@@
 
-\paragraph{Two Valid SOCP Constraints}
+\subsubsection{Two Valid SOCP Constraints}
 

### 2020-12-14 tex/MISDP-Presolving/MISDPpresolving.tex
extend section about bounds for rank-1 motivated structures

@@
 \begin{remark}
-  Most likely this can be generalized to the more interesting case of a
+  This can be generalized to the more interesting case of a
   constraint of the form
@@
   \quad
-  \ell \leq (x,t) \leq u,
-\]
-where~$t$ is a scalar variable. Since constraints of this particular form
-appear in truss topology design and cardinality least-squares problems, we
-should check whether this helps for such instances.
+  \ell \leq x \leq u,\quad \ell_t \leq t \leq u_t,
+\]
+where~$t$ is a scalar variable. Constraints of this particular form
+appear, e.g., in truss topology design and cardinality least-squares
+problems. The same argument as above yields the two inequalities
+\begin{align*}
+  X_{ii} &\geq \frac{2 \ell_i x_i - \ell_i^2}{u_t} \\
+  X_{ii} &\geq \frac{2 u_i x_i - u_i^2}{u_t}
+\end{align*}
+Assume that~$\ell_i < 0$ and ~$u_i > 0$. Then, these inequalities can be
+non-trivial, that is, the right-hand-side is nonnegative,
+if~$x_i \leq \ell_i/2$ and~$x_i \geq u_i/2$, respectively. Additionally, by
+replacing~$x_i$ by its lower bound~$\ell_i$, the first inequality can be
+used for propagation:
+\begin{align*}
+  X_{ii} &\geq \frac{2 \ell_i u_i - \ell_i^2}{u_t},
+\end{align*}
+where the right-hand-side is strictly positive, if $\ell_i > u_i/2$, $u_t >
+0$ and either~$u_i > 0$ or~$\ell_i$, $u_i < 0$.
+
+These inequalities are currently not implemented, but it may be worth
+testing them on different types of problems with LP- or SDP-solving to see
+if they have any impact.
 \end{remark}

### 2020-12-14 tex/MISDP-Presolving/MISDPpresolving.tex
add remark about Gershgorin for SDPrank1-constraints

@@
 
+\begin{remark}
+  This idea can be used to detect a cut off for an SDP-constraint that
+  should have rank 1. Recall that a positive semidefinite matrix has rank 1
+  if and only if all of its 2 by 2 minors are zero. This implies that the
+  smallest eigenvalue of each 2 by 2 submatrix needs to be 0. Thus, if
+  there exist two indices~$i$, $k$, so that
+  \begin{align*}
+    0 \notin D_i,\quad 0 \notin D_k, \quad D_i \cap D_k = \varnothing,
+  \end{align*}
+  then the current node can be cut off, since a minimal eigenvalue of 0 is
+  not obtainable for the 2 by 2 submatrix indexed by~$(i,k)$.
+
+  In the case of a dual SDP-constraint~$A(y) \geq 0$, the Gershgorin
+  discs~$D_i$, $D_k$ do not contain 0, if the lower bounds~$\ell_i$,
+  $\ell_k$ of the entries~$A(y)_{ii}$ and~$A(y)_{kk}$ are strictly
+  positive, and if the radius of the discs is strictly smaller
+  than~$\ell_i$, $\ell_k$, i.e.,
+  \begin{align*}
+    \ell_i > 0,\quad \ell_k > 0, \qquad \sum_{j\neq i} \abs{A(y)_{jj}} <
+    \ell_i, \quad \sum_{j\neq k} \abs{A(y)_{jj}} < \ell_k.
+  \end{align*}
+  Currently, this is not implemented for SDP-constraints with an additional
+  rank-1 constraint.
+\end{remark}
+
 % -------------------------------------------------------------------------

### 2020-12-14 tex/MISDP-Presolving/MISDPpresolving.tex
add remark about RIP in dimension two

@@
   One idea is to perform spatial branching on the diagonal entries
-  $X_ii$. Does this help to solve the problem?
+  $X_{ii}$. Does this help to solve the problem?
 \end{question}
+
+\begin{remark}
+  For the case of~$k=2$, an optimal solution~$X$ consists of a single 2 by
+  2 nonzero submatrix, and all other entries are 0. Let
+  \begin{align*}
+    \tilde{X} =  \begin{pmatrix}
+      X_{11} & X_{21} \\
+      X_{21} & X_{22}
+    \end{pmatrix}
+  \end{align*}
+  be a 2 by 2 submatrix of~$X$. The trace constraint then
+  reads~$X_{11} + X_{22} = 1$, and the psd constraint~$X \succeq 0$
+  simplifies to~$X_{11},\, X_{22} \geq 0$, and $X_{11}X_{22}\geq
+  X_{21}^2$. If~$X_{21} = 0$, then either~$X_{11} = 1$ or~$X_{22} = 1$,
+  depending on their coefficients in the objective function.
+  If~$X_{21} \neq 0$, aggregate~$X_{22} = 1 - X_{11}$. In this case, the
+  solution set is given by a disc centered at $(1/2,\,0)$ with
+  radius~$1/2$, where~$X_{11}$ is the first coordinate and~$X_{21}$ the
+  second. Optimizing in direction of the objective function yields an
+  optimal solution.
+\end{remark}
 

### 2020-12-14 tex/MISDP-Presolving/MISDPpresolving.tex
refine description a bit

@@
 
+Indeed, it is NP-hard to optimize over $S_I$ -- this is goal of the
+RIP-problem.
+
 \paragraph{Linear Case}
@@
 \]
-This is related to the \emph{hypersimplex}
-\[
-  \Delta_{n,k} \define \conv \{z \in \{0,1\}^n \suchthat \sum_{i=1}^n z_i = k\},
-\]
-for which complete linear descriptions are available, see
-Ziegler~\cite{Zie95}.
+Both the polytopes
+$\{x \in \R^n \suchthat \ones\T x = 1,\; 0 \leq x \leq z\}$ for fixed
+$z \in \{0,1\}^n$ and
+$\{z \in \R^n \suchthat \ones\T z \leq k,\; x \leq z\}$ for fixed
+$x \in \{0,1\}^n$ are integral.
 
 \begin{question}
-  Does this give a linear description of $S$ and can this be extended to
-  the SDP case?
+  Is the above description integral?
 \end{question}

### 2020-12-19 tex/MISDP-Presolving/MISDPpresolving.tex
handle linear case of feasibility set

@@
 
+% -------------------------------------------------------------------------
 \subsubsection{Complete Description of Feasible Set}
@@
 
+% -------------------------------------------------------------------------
 \paragraph{Linear Case}
 
-It might be instrumential to look at the linear analogue:
-\[
-  S \define \{(x,z) \in \R_+^n \times [0,1]^n \suchthat \sum_{i=1}^n x_i =
-  1,\; \sum_{i=1}^n z_i \leq k,\; 0 \leq x_i \leq z_i \;\forall\, i \in [n]\}.
-\]
-Both the polytopes
-$\{x \in \R^n \suchthat \ones\T x = 1,\; 0 \leq x \leq z\}$ for fixed
-$z \in \{0,1\}^n$ and
-$\{z \in \R^n \suchthat \ones\T z \leq k,\; x \leq z\}$ for fixed
-$x \in \{0,1\}^n$ are integral.
-
-\begin{question}
-  Is the above description integral?
-\end{question}
+It might be instrumential to look at the following linear analogue:
+\[
+S^k \define \{(x,z) \in \R_+^n \times [0,1]^n \suchthat \sum_{i=1}^n x_i =
+1,\; \sum_{i=1}^n z_i = k,\; 0 \leq x_i \leq z_i \;\forall\, i \in [n]\}.
+\]
+As before, $S^k_I$ denotes the integer hull
+$S^k_I \define \conv\big(S^k \cap (\R^n \times \{0,1\}^n)\big)$.
+
+\begin{theorem}
+  $S^k = S^k_I$, i.e., $S^k$ is integral.
+\end{theorem}
+
+\begin{proof}
+  The constraint matrix of $S^k$ with respect to the equations and the
+  inequalities $x - z \leq 0$ can be written as follows:
+  \[
+    M \define
+    \begin{pmatrix}
+      1 & \dots & 1 & 0 & \dots & 0 \\
+      0 & \dots & 0 & 1 & \dots & 1 \\
+      1 &       &   & -1 &      &   \\
+        & \ddots&   &    & \ddots & \\
+      0 &       & 1 & 0 &      & -1
+    \end{pmatrix}.
+  \]
+  This matrix is totally unimodular. This can be seen by multiplying the
+  first row with $-1$. Then we have exactly one $+1$ and $-1$ in each
+  column. Thus, $M$ is the node-arc indicence matrix of a directed graph
+  and hence TU. To take care of the variables bounds, one would add
+  identity matrices to $M$, which do not change the property of being TU.
+\end{proof}
+
+As a consequence, when solving
+\[
+  \min\, \{ c\T x + d\T z \suchthat (x,z) \in S^k \}
+\]
+one can even chose the $x$-variables to be integral, i.e., in 0 or 1, in an
+optimal solution.
+
+\begin{remark}
+  Aghezzaf and Wolsey~\cite{AghW92} show the following: Let
+  $T \subseteq \R^p_+ \times \{0,1\}^n$, $P = \conv(T)$, and
+  \[
+  P_k \define \{ (x,z) \in P \suchthat \sum_{i=1}^n z_i = k\}.
+  \]
+  For $k = 0, \dots, n$ define
+  \[
+  g_k \define \min\,\{c\T x + d\T z \suchthat (x,z) \in P_k,\; z \in
+  \{0,1\}^n\}.
+  \]
+  Then they show that
+  \[
+  \min\,\{c\T x + d\T z \suchthat (x,z) \in P_k\}
+  \]
+  has an optimal solution in $T$ if and only if $g_{k+1} - g_k$ is
+  non-decreasing for $k = 1, \dots, n - 1$.
+  % Alternatively, if for some $\mu$
+  % \[
+  % \min\, \{c\T x + d\T z - \mu \sum_{i=1}^n z_i \suchthat (x,z) \in P\}
+  % \]
+  % has integral optimal solutions $(x^1,z^1)$ and $(x^2,z^2)$ with
+  % $\ones\T z^1 = k_1$ and $\ones\T z^2 = k_2$ with $k_1 \leq k_2 - 2$, then
+  % there exists an alternative integral optimal solution with $\ones\T z = k$ and
+  % $k_1 < k < k_2$. Note that one can integrate the $\mu$-term into $d$.
+
+  In our situation we have
+  \[
+  T = \{(x,z) \in \R_+^n \times \{0,1\}^n \suchthat \sum_{i=1}^n x_i = 1,\;
+  0 \leq x_i \leq z_i \;\forall\, i \in [n]\}.
+  \]
+  The convex hull of $T$ is simply obtained by relaxing $z$ to be in $[0,1]^n$:
+  \[
+  P = \{(x,z) \in \R_+^n \times [0,1]^n \suchthat \sum_{i=1}^n x_i = 1,\;
+  0 \leq x_i \leq z_i \;\forall\, i \in [n]\}.
+  \]
+  Thus, $P_k = S^k$.
+
+  Then Aghezzaf and Wolsey consider lot sizing problems for given demands $D_1,
+  \dots, D_n$:
+  \begin{align*}
+    \min\quad & c\T x + d\T z\\
+    & \sum_{i=1}^t x_i \geq \sum_{i=1}^t D_i\quad \forall i = 1, \dots, n - 1,\\
+    & \sum_{i=1}^n x_i = \sum_{i=1}^n D_i,\\
+    & 0 \leq x \leq M z,\\
+    & z \in \{0,1\}^n.
+  \end{align*}
+  If $D_i = 0$, $i = 1, \dots, n-1$, $D_i = 1$, and $M = 1$, we obtain our problem.
+
+  Aghezzaf and Wolsey show that the optimum is attained in an integral
+  point for each $k$ if $c_1 \geq c_2 \geq \dots \geq c_n$. But the proof
+  is not completely clear.
+\end{remark}
 

### 2021-03-23 tex/MISDP-Presolving/MISDPpresolving.tex
add right hand side
correct bound tightening in presolving document

@@
 
-One can also use the methods in Section~\ref{sec:CoefficientTightening} to
-compute bounds on each $y_k$ variable. For this we only need that all $A^k$
-are psd, but $y_k$ does not need to be binary. We need the following:
+The methods in Section~\ref{sec:CoefficientTightening} cannot be used to
+tighten bounds on each $y_k$ variable. The following lemma is valid, bu it
+may cut off feasible solutions as shown in an example afterwards.
 
@@
 \end{lemma}
-
-\todo[inline]{FM: If we tighten the bounds of a variable~$y_k$, we need to
-  scale the linear constraints accordingly.}
 
@@
 
-From the lemma, we conclude that one can tighten the upper bounds of all
-variables $y_{\ell}$ to $\min\, \{u_{\ell}, \hat{\mu}_{\ell}\}$. This is
-implemented, but has not tightened any bound in presolving.
+The following example shows that the lemma may indeed cut off feasible
+solutions.
+
+\begin{example}
+  Consider the problem
+  \begin{align*}
+    \min\; &-y_1 \\
+    \text{s.t.}\; &
+    \begin{pmatrix} 2 & 0 \\ 0 & 2 \end{pmatrix} y_1
+    + \begin{pmatrix} 3 & 0 \\ 0 & 3 \end{pmatrix} y_2
+    + \begin{pmatrix} 1 & 0 \\ 0 & 1 \end{pmatrix} y_3
+    - \begin{pmatrix} 1 & 0 \\ 0 & 1 \end{pmatrix} \succeq 0, \\
+    & y_1 \in \{0,1\},\quad y_2 \in \{0,1\}, \quad y_3 \in [0,3]_\Z,
+  \end{align*}
+  which has optimal value $-1$. The scaling factor for the first variable
+  is~$\mu_1 = \frac{1}{2}$, but changing the upper bound of~$y_1$
+  to~$\mu_1$, i.e., fixing~$y_1 = 0$ cuts off all optimal
+  solutions. However, the above lemma is not violated, since after
+  fixing~$y_1 = \mu_1$, there still exists a feasible solution.
+\end{example}
+
+In order to correctly tighten variable bounds, assume that no coefficient
+matrix~$A^k$ is indefinite, and that the variable
+bounds~$\ell_k \leq y_k \leq u_k$ satisfy~$\ell_k >-\infty$
+if~$A^k \preceq 0$ and~$u_k < \infty$ if~$A^k \succeq 0$ for
+all~$k \in [m]$ . For an index~$k \in [m]$, define
+\begin{align*}
+  P_k \define \{i \neq k \suchthat A^i \succeq 0\},\qquad N_k \define \{i
+  \neq k \suchthat A^i \preceq 0\},
+\end{align*}
+as well as
+\begin{align*}
+  \nu_k =
+  \begin{cases}
+    \min\, \{\nu \suchthat \nu A^k + \sum_{i\in P_k}A^i u_i + \sum_{j
+      \in N_k} A^j \ell_j - A^0 \succeq 0 \}, & \text{if } A^k \succeq 0 \\
+    \max\, \{\nu \suchthat \nu A^k + \sum_{i\in P_k}A^i u_i + \sum_{j
+      \in N_k} A^j \ell_j - A^0 \succeq 0 \}, & \text{if } A^k \preceq 0.
+  \end{cases}
+\end{align*}
+
+The following lemma shows that the lower or upper bounds of the variables
+can be tightened, according to the semidefiniteness of the coefficient
+matrices.
+\begin{lemma}
+  Let all~$A^k$, $k \in [m]$ be not indefinite, and let~$\ell_k >-\infty$
+  if~$A^k \preceq 0$ as well as~$u_k < \infty$ if~$A^k \succeq 0$ for
+  all~$k \in [m]$. Then, $\sum_{k=1}^m A^k y_k - A^0 \succeq 0$ implies
+  \begin{align*}
+    y_k
+    \begin{cases}
+      \geq \max \, \{\ell_k,\, \nu_k\}, & \text{if } A^k \succeq 0, \\
+      \leq \min \, \{u_k,\, \nu_k\}, & \text{if } A^k \preceq 0.
+    \end{cases}
+  \end{align*}
+\end{lemma}
+
+\begin{proof}
+  Let all~$A^k$, $k \in [m]$ be not indefinite, and let~$\ell_k >-\infty$
+  if~$A^k \preceq 0$ as well as~$u_k < \infty$ if~$A^k \succeq 0$ for
+  all~$k \in [m]$. Furthermore, let $\sum_{k=1}^m A^k y_k - A^0 \succeq 0$,
+  and let~$k \in [m]$ with (w.l.o.g.)~$A^k \succeq 0$ and~$\nu_k \geq \ell_k$. Suppose
+  that~$y_k < \nu_k$. Then, by definition of~$\nu_k$, there
+  exists~$x \in \R^n$ with
+  \begin{align*}
+    0 &> x\T \Big(y_kA^k + \sum_{i \in P_k}A^i u_i + \sum_{i\in N_k}A^i
+        \ell_i - A^0 \Big) x \\
+      &= y_k x\T A^k x + \sum_{i \in P_k} u_i \underbrace{x\T A^i x}_{\geq 0} + \sum_{i\in N_k}
+        \ell_i \underbrace{x\T A^i x}_{\leq 0} - x\T A^0 x \\
+      &\geq y_k x\T A^k x + \sum_{i \in P_k} y_i x\T A^i x + \sum_{i\in N_k}
+        y_i x\T A^i x - x\T A^0 x \\
+      &= x\T \Big(\sum_{i=1}^m A^i y_i - A^0\Big) x,
+  \end{align*}
+  which is a contradiction to~$\sum_{k=1}^m A^k y_k - A^0 \succeq
+  0$. Thus,~$y_k \geq \nu_k$.
+\end{proof}
 
@@
   and hence TU. To take care of the variables bounds, one would add
-  identity matrices to $M$, which do not change the property of being TU.
+  identity matrices to $M$, which do not change the property of being
+  TU. Moreover, the right hand sides are integral and thus the polytope is
+  integral.
 \end{proof}
@@
     \min\quad & c\T x + d\T z\\
-    & \sum_{i=1}^t x_i \geq \sum_{i=1}^t D_i\quad \forall i = 1, \dots, n - 1,\\
+    & \sum_{i=1}^t x_i \geq \sum_{i=1}^t D_i\quad \forall \, t = 1, \dots, n - 1,\\
     & \sum_{i=1}^n x_i = \sum_{i=1}^n D_i,\\
@@
   \end{align*}
-  If $D_i = 0$, $i = 1, \dots, n-1$, $D_i = 1$, and $M = 1$, we obtain our problem.
+  If $D_i = 0$, $i = 1, \dots, n-1$, $D_n = 1$, and $M = 1$, we obtain our problem.
 

### 2022-04-08 src/sdpi/sdpisolver_mosek.c
minor cleanup

@@ -1429,21 +1429,7 @@ SCIP_RETCODE SCIPsdpiSolverLoadAndSolveWithPenalty(
          MOSEK_CALL( MSK_putintparam(sdpisolver->msktask, MSK_IPAR_INTPNT_SCALING, MSK_SCALING_NONE) );
       }
 
-      /* print whole problem (only for MOSEK < 9) and parameters if asked to */
-#ifdef SCIP_MORE_DEBUG
-#if MSK_VERSION_MAJOR < 9
-      {
-         int nmosekconss;
-         int nmosekvars;
-         int nmosekcones;
-
-         MOSEK_CALL( MSK_getnumcon(sdpisolver->msktask, &nmosekconss) );
-         MOSEK_CALL( MSK_getnumvar(sdpisolver->msktask, &nmosekvars) );
-         MOSEK_CALL( MSK_getnumcone(sdpisolver->msktask, &nmosekcones) );
-
-         MOSEK_CALL( MSK_printdata(sdpisolver->msktask, MSK_STREAM_LOG, 0, nmosekconss, 0, nmosekvars, 0, nmosekcones, 1, 1, 1, 1, 1, 1, 1, 1) );
-      }
-#endif
+      /* print parameters if asked to */
 #ifdef SCIP_PRINT_PARAMETERS
       MOSEK_CALL( MSK_printparam(sdpisolver->msktask) );
 #endif
@@ -1455,7 +1441,7 @@ SCIP_RETCODE SCIPsdpiSolverLoadAndSolveWithPenalty(
 #endif
 
       /* solve the problem */
-      MOSEK_CALL( MSK_optimizetrm(sdpisolver->msktask, &(sdpisolver->terminationcode)) );/*lint !e641*/
+      MOSEK_CALL( MSK_optimizetrm(sdpisolver->msktask, &sdpisolver->terminationcode) );/*lint !e641*/
       MOSEK_CALL( MSK_getdouinf(sdpisolver->msktask, MSK_DINF_OPTIMIZER_TIME, &sdpisolver->opttime) );
       MOSEK_CALL( MSK_getsolsta(sdpisolver->msktask, MSK_SOL_ITR, &sdpisolver->solstat) );/*lint !e641*/
 

### 2022-04-08 src/sdpi/sdpisolver_mosek.c
minor fixes

@@ -934,7 +934,7 @@ SCIP_RETCODE SCIPsdpiSolverLoadAndSolveWithPenalty(
    }
    assert( sdpisolver->nactivevars + nfixedvars == sdpisolver->nvars );
 
-   /* adjust maxabsrhscoef in penalty formulation */
+   /* adjust maxabsobjcoef in penalty formulation */
    if ( penaltyparam >= sdpisolver->epsilon )
    {
       if ( penaltyparam > maxabsobjcoef )
@@ -1095,7 +1095,7 @@ SCIP_RETCODE SCIPsdpiSolverLoadAndSolveWithPenalty(
       /* We currently do not include variable bounds in maxrhscoef, because it does not seem to be beneficial
        * overall. The bounds are very relevant for cardinality least square instances in which all variables are binary,
        * except for one continuous variable representing the objective value. The objective value can be
-       * large. Enlarging maxrhscoef will not particularly help in this context, since the objective values are measure
+       * large. Enlarging maxrhscoef will not particularly help in this context, since the objective values are measured
        * relatively and the bounds are filtered out later anyway. */
 #if CONVERT_ABSOLUTE_TOLERANCES
       if ( REALABS(mosekvarbounds[i]) > maxrhscoef )
@@ -1432,7 +1432,6 @@ SCIP_RETCODE SCIPsdpiSolverLoadAndSolveWithPenalty(
       /* print parameters if asked to */
 #ifdef SCIP_PRINT_PARAMETERS
       MOSEK_CALL( MSK_printparam(sdpisolver->msktask) );
-#endif
 #endif
 
       /* write to file if asked to */

### 2022-06-10 tex/ConflictAnalysis.tex
1st version of conflict analysis description

\documentclass[10pt, a4paper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[american]{babel}
\usepackage[dvipsnames]{xcolor}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{dsfont}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{booktabs}
\usepackage{siunitx}
\usepackage[lined,ruled,noend,linesnumbered]{algorithm2e}
\usepackage[color=red!50!blue!50,textsize=footnotesize,textwidth=4cm]{todonotes}
\usepackage[breaklinks,colorlinks,citecolor=blue,linkcolor=blue]{hyperref}
% \usepackage[tmargin=2.5cm,bmargin=2.5cm,lmargin=2.8cm,rmargin=2.8cm]{geometry}
\usepackage{cleveref}
\usepackage{xspace}

% operators
\DeclareMathOperator{\conv}{conv}
\DeclareMathOperator{\cone}{cone}
\DeclareMathOperator{\aff}{aff}
\DeclareMathOperator{\lin}{lin}
\DeclareMathOperator*{\argmin}{argmin}
\DeclareMathOperator*{\argmax}{argmax}
\DeclareMathOperator{\supp}{supp}
\DeclareMathOperator{\relint}{relint}
\DeclareMathOperator{\diag}{diag}
\DeclareMathOperator{\Diag}{Diag}
\DeclareMathOperator{\sgn}{sgn}
\DeclareMathOperator{\tr}{tr}
\renewcommand{\Re}{\text{Re}}

% macros
\newcommand{\suchthat}{\,:\,}
\newcommand{\abs}[1]{\lvert{#1}\rvert}
\newcommand{\card}[1]{\lvert{#1}\rvert}
\newcommand{\ceil}[1]{\lceil{#1}\rceil}
\newcommand{\floor}[1]{\lfloor{#1}\rfloor}
\newcommand{\define}{\coloneqq}
\newcommand{\enifed}{\eqqcolon}
\newcommand{\norm}[1]{\lVert{#1}\rVert}
\newcommand{\Norm}[2]{\lVert{#1}\rVert_{#2}}
\newcommand{\skal}[2]{\langle{#1},{#2}\rangle}
\newcommand{\order}[1]{O\left({#1}\right)}
\newcommand{\T}{^{\top}}

% sets
\newcommand{\R}{\mathds{R}}
\newcommand{\Z}{\mathds{Z}}
\newcommand{\Q}{\mathds{Q}}
\newcommand{\N}{\mathds{N}}
\newcommand{\C}{\mathds{C}}
\newcommand{\E}{\mathds{E}}
\newcommand{\ones}{\mathds{1}}
\renewcommand{\P}{\mathds{P}}
\renewcommand{\E}{\mathds{E}}

% environments
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{observation}[theorem]{Observation}
\newtheorem{example}[theorem]{Example}
\newtheorem{question}[theorem]{Question}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{assumption}[theorem]{Assumption}

% adjust itemize
\setlist[itemize]{topsep=0.5ex,partopsep=0ex,parsep=0ex,itemsep=0.5ex}

% other defines
\newcommand{\MP}[1]{\todo{#1}}
\newcommand{\MPin}[1]{\todo[inline]{#1}}

% title
\title{Conflict Analysis for Mixed-integer SDPs}
\author{Marc E. Pfetsch\thanks{Department of Mathematics, TU Darmstadt, Germany}}
\date{June 2022}

% -------------------------------------------------------------------------
% -------------------------------------------------------------------------
\begin{document}

\maketitle

% -------------------------------------------------------------------------
\section{Introduction}

We consider mixed-integer SDPs of the following form:
\begin{equation}\label{MISDP}
  \begin{aligned}
    \inf \quad & b\T y \\
    \text{s.t.} \quad & \sum_{j=1}^m A^j\, y_j - A^0 \succeq 0, \\
    & D y \geq d,\\
    & \ell \leq y \leq u,\\
    & y_i \in \Z,\quad i \in I.
  \end{aligned}
\end{equation}
In each node, we solve the following relaxations with local bounds $\ell$
and $u$:
\begin{equation}\label{SDP-D}
  \begin{aligned}
    \inf \quad & b\T y \\
    \text{s.t.} \quad & \sum_{j=1}^m A^j\, y_j - A^0 \succeq 0, \\
    & D y \geq d,\\
    & \ell \leq y \leq u,
  \end{aligned}
\end{equation}
where for simplicity of exposition, we assume that all entries of $\ell$
and $u$ are finite.

The general idea of conflict analysis is to generate globally valid
constraints as follows. Let any $\hat{X} \succeq 0$ and $\hat{z} \geq 0$ be
given. Then because $A(y) \define \sum_{j=1}^m A^j\, y_j - A^0$ is positive
semidefinite for all feasible $y \in \R^m$, we have
$\skal{A(y)}{\hat{X}} \geq 0$.  Similarly, the aggregated inequalities
$\hat{z}\T D \geq \hat{z}\T d$ hold. Together this yields:
\begin{align}
  & \skal{A(y)}{\hat{X}} + \hat{z}\T D \geq \hat{z}\T d\notag\\
  \Leftrightarrow\quad
  & \sum_{j=1}^m \skal{A^j}{\hat{X}} y_j + \hat{z}\T D + \geq
    \skal{A^0}{\hat{X}} + \hat{z}\T d,\label{eq:ConflictCut}
\end{align}
which is valid for all feasible $y$. Clearly,
Inequality~\eqref{eq:ConflictCut} is redundant for continuous $y$, but it
might allow to cut off certain mixed-integer solutions~$y$. In the implementation,
we only propagate bounds using~\eqref{eq:ConflictCut}.

More generally, we can also use the variable bounds $\ell \leq y$ and
$y \leq u$. If $\hat{r}^\ell$, $\hat{r}^u \geq 0$ are the corresponding
multiliers, this yields:
\begin{align}
  & \skal{A(y)}{\hat{X}} + \hat{z}\T D + (\hat{r}^\ell)\T y - (\hat{r}^u)\T y
    \geq \hat{z}\T d + (\hat{r}^\ell)\T \ell + (\hat{r}^u)\T u\notag\\
  \Leftrightarrow
  & \sum_{j=1}^m \skal{A^j}{\hat{X}} y_j + \hat{z}\T D +
    (\hat{r}^\ell)\T y - (\hat{r}^u)\T y \geq \skal{A^0}{\hat{X}} +
    \hat{z}\T d + (\hat{r}^\ell)\T \ell + (\hat{r}^u)\T u,\label{eq:ConflictCut2}
\end{align}
which is again valid for all feasible $y$.

% -------------------------------------------------------------------------
\section{Generation of Conflict Constraints}

Let $X$ be the dual variables for the SDP-constraint $A(y) \succeq 0$, $z$
be the dual variables for $Dy \geq d$, and $r^\ell$ and $r^u$ be the dual
variables for the bound constraints $y \geq \ell$ and $-y \geq -u$,
respectively. Then the corresponding primal problem for~\eqref{SDP-D} is:
\begin{equation}\label{SDP-P}
  \begin{aligned}
     \sup \quad & \skal{A^0}{X} + z\T d + \ell\T r^\ell - u\T r^u\\
     \text{s.t.} \quad & \skal{A^j}{X} + D\T z + r^\ell_j - r^u_j = b_j && \forall \, j \in [m],\\
     & X \succeq 0,\; z, \; r^\ell,\; r^u \geq 0.
  \end{aligned}
\end{equation}
Any feasible solution can be used to construct~\eqref{eq:ConflictCut}
or~\eqref{eq:ConflictCut2}.

Similarly, a primal ray $(\hat{X}, \hat{z}, \hat{r}^\ell, \hat{r}^u)$,
i.e.,
\begin{equation}\label{eq:ray}
\begin{aligned}
  & \skal{A^j}{\hat{X}} + D\T \hat{z} + \hat{r}^\ell_j - \hat{r}^u_j = 0 && \forall\; j \in [m],\\
  & \skal{A^0}{\hat{X}} + d\T \hat{z} + \ell\T \hat{r}^\ell - u\T \hat{r}^u > 0,
\end{aligned}
\end{equation}
yields an inequality as in~\eqref{eq:ConflictCut}
or~\eqref{eq:ConflictCut2}.

If $\ell \leq u$ and~\eqref{eq:ray} has a feasible solution
$(\tilde{X}, \tilde{r}^\ell, \tilde{r}^u)$, then
$(\hat{X}, \hat{r}^\ell, \hat{r}^u)$ with $\hat{X} = \tilde{X}$ and
$\hat{r}^\ell_j = - \min\,\{\skal{A^j}{\tilde{X}}, 0\} \geq 0$ and
$\hat{r}^u_j = \max\,\{\skal{A^j}{\tilde{X}}, 0\} \geq 0$ is feasible
for~\eqref{eq:ray}. Indeed, it is clear that the equations
of~\eqref{eq:ray} are satisfied. Therefore consider the contribution to
$\ell\T \hat{r}^\ell - u\T \hat{r}^u$ in component $j$:
\begin{align*}
  \ell_j \hat{r}^\ell_j - u_j \hat{r}^u_j & = - \ell_j\, \min\,\{\skal{A^j}{\tilde{X}}, 0\} - u_j\, \max\,\{\skal{A^j}{\tilde{X}}, 0\}\\
  & = - \ell_j\, \min\,\{\tilde{r}^u_j - \tilde{r}^\ell_j, 0\} - u_j\, \max\,\{\tilde{r}^u_j - \tilde{r}^\ell_j, 0\}\\
  & = \begin{cases}
    - \ell_j\, (\tilde{r}^u_j - \tilde{r}^\ell_j) & \text{if } \tilde{r}^u_j - \tilde{r}^\ell_j < 0,\\
    - u_j\, (\tilde{r}^u_j - \tilde{r}^\ell_j) & \text{if } \tilde{r}^u_j - \tilde{r}^\ell_j \geq 0,
  \end{cases}\\
  & = \begin{cases}
    - \ell_j\, \tilde{r}^u_j + \ell_j \tilde{r}^\ell_j & \text{if } \tilde{r}^u_j - \tilde{r}^\ell_j < 0,\\
    - u_j\, \tilde{r}^u_j + u_j \tilde{r}^\ell_j & \text{if } \tilde{r}^u_j - \tilde{r}^\ell_j \geq 0,
  \end{cases}\\
  & \geq \begin{cases}
    - u_j\, \tilde{r}^u_j + \ell_j \tilde{r}^\ell_j & \text{if }  \tilde{r}^u_j - \tilde{r}^\ell_j < 0,\\
    - u_j\, \tilde{r}^u_j + \ell_j \tilde{r}^\ell_j & \text{if } \tilde{r}^u_j - \tilde{r}^\ell_j \geq 0,
  \end{cases}\\
  & = \ell_j \tilde{r}^\ell_j - u_j \tilde{r}^u_j.
\end{align*}
Therefore the inequality in~\eqref{eq:ray} holds as well.

This implies that
\begin{align*}
  0 & < \skal{A^0}{\hat{X}} + \ell\T \hat{r}^\ell - u\T \hat{r}^u\\
  & = \skal{A^0}{\hat{X}} + \sum_{j=1}^m \ell_j \hat{r}^\ell_j -
    \sum_{j=1}^m u_j \hat{r}^u_j\\
  & = \skal{A^0}{\hat{X}} - \sum_{j: \skal{A^j}{\hat{X}} < 0} \ell_j \skal{A^j}{\hat{X}} -
    \sum_{j: \skal{A^j}{\hat{X}} > 0} u_j \skal{A^j}{\hat{X}}.
\end{align*}
This implies
\begin{align*}
  \sum_{j: \skal{A^j}{\hat{X}} < 0} \ell_j \skal{A^j}{\hat{X}} +
    \sum_{j: \skal{A^j}{\hat{X}} > 0} u_j \skal{A^j}{\hat{X}} < \skal{A^0}{\hat{X}}.
\end{align*}
However, any feasible solution $\hat{y}$ satisfies $A(\hat{y}) \succeq 0$ and $\ell
\leq \hat{y} \leq u$ and therefore
\begin{align*}
  0 & \leq \skal{A(\hat{y})}{\hat{X}} =
      \sum_{j=1}^m \skal{A^j}{\hat{X}}\, \hat{y}_j + \sum_{j=1}^m
      \skal{A^j}{\hat{X}}\, \hat{y}_j - \skal{A^0}{\hat{X}}\\
    & \leq \sum_{j: \skal{A^j}{\hat{X}} < 0} \skal{A^j}{\hat{X}}\, \ell_j +
    \sum_{j: \skal{A^j}{\hat{X}} > 0} \skal{A^j}{\hat{X}} u_j -
      \skal{A^0}{\hat{X}},
\end{align*}
a contradition. This implies that the valid linear inequality
$\skal{A(\hat{y})}{\hat{X}} \geq 0$ shows infeasibility with respect to the
local bounds.

% -------------------------------------------------------------------------
\section{Infeasibility Proof Analysis}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2022-06-11 tex/ConflictAnalysis.tex
extend description

@@
 where for simplicity of exposition, we assume that all entries of $\ell$
-and $u$ are finite.
+and $u$ are finite. Moreover, we always assume that $\ell \leq u$ holds.
 
@@
 $\skal{A(y)}{\hat{X}} \geq 0$.  Similarly, the aggregated inequalities
-$\hat{z}\T D \geq \hat{z}\T d$ hold. Together this yields:
+$\hat{z}\T D y \geq \hat{z}\T d$ hold. Together this yields:
 \begin{align}
-  & \skal{A(y)}{\hat{X}} + \hat{z}\T D \geq \hat{z}\T d\notag\\
+  & \skal{A(y)}{\hat{X}} + \hat{z}\T D y \geq \hat{z}\T d\notag\\
   \Leftrightarrow\quad
-  & \sum_{j=1}^m \skal{A^j}{\hat{X}} y_j + \hat{z}\T D + \geq
+  & \sum_{j=1}^m \skal{A^j}{\hat{X}} y_j + \hat{z}\T D y \geq
     \skal{A^0}{\hat{X}} + \hat{z}\T d,\label{eq:ConflictCut}
@@
 \begin{align}
-  & \skal{A(y)}{\hat{X}} + \hat{z}\T D + (\hat{r}^\ell)\T y - (\hat{r}^u)\T y
+  & \skal{A(y)}{\hat{X}} + \hat{z}\T D y + (\hat{r}^\ell)\T y - (\hat{r}^u)\T y
     \geq \hat{z}\T d + (\hat{r}^\ell)\T \ell + (\hat{r}^u)\T u\notag\\
-  \Leftrightarrow
-  & \sum_{j=1}^m \skal{A^j}{\hat{X}} y_j + \hat{z}\T D +
-    (\hat{r}^\ell)\T y - (\hat{r}^u)\T y \geq \skal{A^0}{\hat{X}} +
+  \Leftrightarrow\quad
+  & \sum_{j=1}^m \skal{A^j}{\hat{X}} y_j + \hat{z}\T D y +
+    (\hat{r}^\ell)\T y - (\hat{r}^u)\T y \geq \\
+  & \qquad \skal{A^0}{\hat{X}} +
     \hat{z}\T d + (\hat{r}^\ell)\T \ell + (\hat{r}^u)\T u,\label{eq:ConflictCut2}
@@
 
+It seems that~\eqref{eq:ConflictCut2} are less helpful
+than~\eqref{eq:ConflictCut}. However, one can cancel some coefficients
+in~\eqref{eq:ConflictCut} using variable bounds -- see
+Section~\ref{sec:Canceling}.
+
+% -------------------------------------------------------------------------
 % -------------------------------------------------------------------------
 \section{Generation of Conflict Constraints}
+
+% -------------------------------------------------------------------------
+\subsection{Feasible Problems}
 
@@
      \sup \quad & \skal{A^0}{X} + z\T d + \ell\T r^\ell - u\T r^u\\
-     \text{s.t.} \quad & \skal{A^j}{X} + D\T z + r^\ell_j - r^u_j = b_j && \forall \, j \in [m],\\
+     \text{s.t.} \quad & \skal{A^j}{X} + (D\T z)_j + r^\ell_j - r^u_j = b_j && \forall \, j \in [m],\\
      & X \succeq 0,\; z, \; r^\ell,\; r^u \geq 0.
@@
 \end{equation}
-Any feasible solution can be used to construct~\eqref{eq:ConflictCut}
-or~\eqref{eq:ConflictCut2}.
+Any feasible solution $(\hat{X}, \hat{z}, \hat{r}^\ell, \hat{r}^u)$ can be
+used to construct~\eqref{eq:ConflictCut} or~\eqref{eq:ConflictCut2}.
+
+Moreover, one can take an objective cutoff into account. Namely, assume
+that $y^\star$ is a feasible solution to~\eqref{MISDP}. We are then only
+interested in solutions that are at least as good, i.e.,
+$b\T y \leq b\T y^\star$ or $- b\T y \geq - b\T y^\star$. Adding this
+inequality to~\eqref{eq:ConflictCut} yields:
+\begin{align*}
+  & \sum_{j=1}^m \skal{A^j}{\hat{X}} y_j + \hat{z}\T D y - b\T y \geq
+    \skal{A^0}{\hat{X}} + \hat{z}\T d - b\T y^\star\\
+  \Leftrightarrow\quad
+  & \sum_{j=1}^m \big(\skal{A^j}{\hat{X}} + (\hat{z}\T D)_j - b_j\big) y_j \geq
+    \skal{A^0}{\hat{X}} + \hat{z}\T d - b\T y^\star.
+\end{align*}
+
+% -------------------------------------------------------------------------
+\subsection{Infeasible Problems}
 
 Similarly, a primal ray $(\hat{X}, \hat{z}, \hat{r}^\ell, \hat{r}^u)$,
-i.e.,
+i.e., it satsifies the system
 \begin{equation}\label{eq:ray}
 \begin{aligned}
-  & \skal{A^j}{\hat{X}} + D\T \hat{z} + \hat{r}^\ell_j - \hat{r}^u_j = 0 && \forall\; j \in [m],\\
-  & \skal{A^0}{\hat{X}} + d\T \hat{z} + \ell\T \hat{r}^\ell - u\T \hat{r}^u > 0,
+  & \skal{A^j}{X} + (D\T z)_j + r^\ell_j - r^u_j = 0 && \forall\; j \in [m],\\
+  & \skal{A^0}{X} + d\T z + \ell\T r^\ell - u\T r^u > 0,
 \end{aligned}
@@
 
-If $\ell \leq u$ and~\eqref{eq:ray} has a feasible solution
-$(\tilde{X}, \tilde{r}^\ell, \tilde{r}^u)$, then
-$(\hat{X}, \hat{r}^\ell, \hat{r}^u)$ with $\hat{X} = \tilde{X}$ and
-$\hat{r}^\ell_j = - \min\,\{\skal{A^j}{\tilde{X}}, 0\} \geq 0$ and
-$\hat{r}^u_j = \max\,\{\skal{A^j}{\tilde{X}}, 0\} \geq 0$ is feasible
-for~\eqref{eq:ray}. Indeed, it is clear that the equations
-of~\eqref{eq:ray} are satisfied. Therefore consider the contribution to
+In this case, one can prove that~\eqref{eq:ConflictCut} actually proves
+infeasibility for the local bounds~$\ell$ and $u$. To this end, we first
+show the following:
+
+\begin{lemma}\label{lemma:ComplemenatrySolution}
+  If~\eqref{eq:ray} has a feasible solution, then there exists a feasible
+  solution in which $r_j^\ell \cdot r_j^u = 0$ for all $j \in [m]$, i.e.,
+  at most one of $r_j^\ell$ and $r_j^u$ is nonzero.
+\end{lemma}
+
+\begin{proof}
+  Let $(\hat{X}, \hat{z}, \tilde{r}^\ell, \tilde{r}^u)$ be a feasible
+  solution of~\eqref{eq:ray}. Then define
+  \[
+    \hat{r}^\ell_j \define - \min\,\{\skal{A^j}{\hat{X}} + (D\T \hat{z})_j,\; 0\} \geq 0,
+    \quad
+    \hat{r}^u_j = \max\,\{\skal{A^j}{\hat{X}} + (D\T \hat{z})_j,\; 0\} \geq 0.
+  \]
+  We claim that $(\hat{X}, \hat{z}, \hat{r}^\ell, \hat{r}^u)$ is feasible
+  for~\eqref{eq:ray}. Note that by construction we have
+  $\hat{r}^\ell_j \cdot \hat{r}^u_j = 0$ for all $j \in [m]$.
+
+  First observere that the equations of~\eqref{eq:ray} are
+  satisfied. Therefore consider the contribution to
   $\ell\T \hat{r}^\ell - u\T \hat{r}^u$ in component $j$:
   \begin{align*}
-  \ell_j \hat{r}^\ell_j - u_j \hat{r}^u_j & = - \ell_j\, \min\,\{\skal{A^j}{\tilde{X}}, 0\} - u_j\, \max\,\{\skal{A^j}{\tilde{X}}, 0\}\\
+    \ell_j \hat{r}^\ell_j - u_j \hat{r}^u_j & = - \ell_j\, \min\,\{\skal{A^j}{\hat{X}} + (D\T \hat{z})_j, 0\} - u_j\, \max\,\{\skal{A^j}{\hat{X}} + (D\T \hat{z})_j, 0\}\\
     & = - \ell_j\, \min\,\{\tilde{r}^u_j - \tilde{r}^\ell_j, 0\} - u_j\, \max\,\{\tilde{r}^u_j - \tilde{r}^\ell_j, 0\}\\
@@
   Therefore the inequality in~\eqref{eq:ray} holds as well.
+\end{proof}
 
-This implies that
+\begin{proposition}
+  Let $(\hat{X}, \hat{z})$ be part of a solution of~\eqref{eq:ray}. Then~\eqref{eq:ConflictCut}, i.e.,
+  \[
+    \skal{A(y)}{\hat{X}} + \hat{z}\T D \geq \hat{z}\T d
+  \]
+  is infeasible with respect to the local bounds $\ell$ and $u$.
+\end{proposition}
+
+\begin{proof}
+  One the one hand, any feasible solution $\hat{y}$ of~\eqref{MISDP} satisfies
+  $A(\hat{y}) \succeq 0$, $D \hat{y} \geq d$, and
+  $\ell \leq \hat{y} \leq u$. Defining
+  $\hat{s}_j \define \skal{A^j}{\hat{X}} + (\hat{z}\T D)_j$ for
+  $j \in [m]$, we get from~\eqref{eq:ConflictCut}:
   \begin{align*}
-  0 & < \skal{A^0}{\hat{X}} + \ell\T \hat{r}^\ell - u\T \hat{r}^u\\
-  & = \skal{A^0}{\hat{X}} + \sum_{j=1}^m \ell_j \hat{r}^\ell_j -
+    \skal{A^0}{\hat{X}} + \hat{z}\T d & \leq \sum_{j=1}^m \big(\skal{A^j}{\hat{X}} + (\hat{z}\T D)_j\big)\, \hat{y}_j
+    = \sum_{j=1}^m \hat{s}_j\, \hat{y}_j\\
+    & \leq \sum_{j: \hat{s}_j < 0} \hat{s}_j\, \ell_j +
+        \sum_{j: \hat{s}_j > 0} \hat{s}_j\, u_j.
+  \end{align*}
+
+  On the other hand let $(\hat{X}, \hat{z}, \hat{r}^\ell, \hat{r}^u)$ be a
+  feasible solution of~\eqref{eq:ray} that satisfies the conditions of
+  Lemma~\ref{lemma:ComplemenatrySolution}. Then
+  $\hat{r}^u - \hat{r}^\ell = \hat{s}$.  We thus obtain
+  \begin{align*}
+    0 & < \skal{A^0}{\hat{X}} + d\T \hat{z} + \ell\T \hat{r}^\ell - u\T \hat{r}^u\\
+      & = \skal{A^0}{\hat{X}} + d\T \hat{z} + \sum_{j=1}^m \ell_j \hat{r}^\ell_j -
         \sum_{j=1}^m u_j \hat{r}^u_j\\
-  & = \skal{A^0}{\hat{X}} - \sum_{j: \skal{A^j}{\hat{X}} < 0} \ell_j \skal{A^j}{\hat{X}} -
-    \sum_{j: \skal{A^j}{\hat{X}} > 0} u_j \skal{A^j}{\hat{X}}.
+      & = \skal{A^0}{\hat{X}}  + d\T \hat{z} - \sum_{j: \hat{s}_j < 0} \ell_j\, \hat{s}_j -
+        \sum_{j: \hat{s}_j > 0} u_j\, \hat{s}_j,
   \end{align*}
-This implies
-\begin{align*}
-  \sum_{j: \skal{A^j}{\hat{X}} < 0} \ell_j \skal{A^j}{\hat{X}} +
-    \sum_{j: \skal{A^j}{\hat{X}} > 0} u_j \skal{A^j}{\hat{X}} < \skal{A^0}{\hat{X}}.
-\end{align*}
-However, any feasible solution $\hat{y}$ satisfies $A(\hat{y}) \succeq 0$ and $\ell
-\leq \hat{y} \leq u$ and therefore
-\begin{align*}
-  0 & \leq \skal{A(\hat{y})}{\hat{X}} =
-      \sum_{j=1}^m \skal{A^j}{\hat{X}}\, \hat{y}_j + \sum_{j=1}^m
-      \skal{A^j}{\hat{X}}\, \hat{y}_j - \skal{A^0}{\hat{X}}\\
-    & \leq \sum_{j: \skal{A^j}{\hat{X}} < 0} \skal{A^j}{\hat{X}}\, \ell_j +
-    \sum_{j: \skal{A^j}{\hat{X}} > 0} \skal{A^j}{\hat{X}} u_j -
-      \skal{A^0}{\hat{X}},
-\end{align*}
-a contradition. This implies that the valid linear inequality
-$\skal{A(\hat{y})}{\hat{X}} \geq 0$ shows infeasibility with respect to the
-local bounds.
+  a contradition.
+\end{proof}
 
 % -------------------------------------------------------------------------
-\section{Infeasibility Proof Analysis}
+\subsection{Canceling}
+\label{sec:Canceling}
+
+Starting from~\eqref{eq:ConflictCut}, one can possibly cancel coefficients
+as follows. Again using
+$s_j \define \skal{A^j}{\hat{X}} + (\hat{z}\T D)_j$, we get
+from~\eqref{eq:ConflictCut}:
+\begin{equation}\label{eq:ConflictCut3}
+  \sum_{j=1}^m s_j\, y_j \geq \skal{A^0}{\hat{X}} + \hat{z}\T d.
+\end{equation}
+
+Assume that $s_j < 0$ for some $j \in [m]$ and the global lower bound of
+$y_j$ is at least 0. Then we can add the valid inequality $-s_j\, y_j \geq 0$
+to~\eqref{eq:ConflictCut3}, which eliminates variable~$y_j$ from the
+inequality. Similarly, if $s_j > 0$ for some $j \in [m]$ and the global
+upper bound of $y_j$ is at most 0, we can add $s_j\, y_j \leq 0$ to
+eliminate variable~$y_j$.
+
+Witzig et al.~\cite{WitBH17} (see also Witzig~\cite{Wit22}) applied
+cancelation to continuous variables, but restricts attention to variables
+that are at their global bounds at the node at which the constraint is
+generated.
+
+\begin{small}
+   \bibliographystyle{plain}
+   \bibliography{ConflictAnalysis.bib}
+\end{small}
 

### 2022-06-11 src/sdpi/sdpi.c
also run for infeasible problems (return ray)

@@ -4062,15 +4062,17 @@ SCIP_RETCODE SCIPsdpiGetPrimalBoundVars(
       }
       *success = TRUE;
    }
-   /* If the dual is infeasible, there is no feasible solution; If the primal is infeasible, the dual is unbounded or
-    * infeasible. In both cases we should not return the solution (rather a ray). */
-   else if ( SCIPsdpiSolverIsDualInfeasible(sdpi->sdpisolver) || SCIPsdpiSolverIsPrimalInfeasible(sdpi->sdpisolver) )
+   /* If the primal is infeasible, we usually do not have a dual solution nor a primal ray. */
+   else if ( SCIPsdpiSolverIsPrimalInfeasible(sdpi->sdpisolver) )
    {
-      SCIPdebugMessage("Problem is infeasible, no primal solution available.\n");
+      SCIPdebugMessage("Primal problem is infeasible, no primal solution available.\n");
    }
    else
    {
       SCIP_RETCODE retcode;
+
+      /* We assume that there exists a primal ray if the dual is infeasible and it is returned by the same function. */
+      assert ( SCIPsdpiSolverIsDualInfeasible(sdpi->sdpisolver) || SCIPsdpiSolverIsDualFeasible(sdpi->sdpisolver) );
 
       retcode = SCIPsdpiSolverGetPrimalBoundVars(sdpi->sdpisolver, lbvals, ubvals);
       if ( retcode == SCIP_OKAY )
@@ -4126,11 +4128,10 @@ SCIP_RETCODE SCIPsdpiGetPrimalLPSides(
       }
       *success = TRUE;
    }
-   /* If the dual is infeasible, there is no feasible solution; If the primal is infeasible, the dual is unbounded or
-    * infeasible. In both cases we should not return the solution (rather a ray). */
-   else if ( SCIPsdpiSolverIsDualInfeasible(sdpi->sdpisolver) || SCIPsdpiSolverIsPrimalInfeasible(sdpi->sdpisolver) )
+   /* If the primal is infeasible, we usually do not have a dual solution nor a primal ray. */
+   else if ( SCIPsdpiSolverIsPrimalInfeasible(sdpi->sdpisolver) )
    {
-      SCIPdebugMessage("Problem is infeasible, no primal solution available.\n");
+      SCIPdebugMessage("Primal problem is infeasible, no primal solution available.\n");
    }
    else
[…]

### 2022-06-11 src/scipsdp/relax_sdp.c
parameters for deciding application for (in)feasible instances

@@ -100,6 +100,9 @@
 #define DEFAULT_USESCALING          TRUE     /**< whether the SDP-solver should use scaling */
 #define DEFAULT_SCALEOBJ            FALSE    /**< whether the objective should be scaled in order to get a more stable behavior */
 #define DEFAULT_CONFLICTCONSS       FALSE    /**< whether conflict constraints should be generated */
+#define DEFAULT_CONFLICTFEAS        FALSE    /**< whether conflict constraints should be generated for feasible subproblems */
+#define DEFAULT_CONFLICTOBJCUT      FALSE    /**< whether an objective cut should be used to generate conflict constraints for feasible subproblems */
+#define DEFAULT_CONFLICTINFEAS      TRUE     /**< whether conflict constraints should be generated for infeasible subproblems */
 
 #define WARMSTART_MINVAL            0.01     /**< if we get a value less than this when warmstarting (currently only for the linear part when combining with analytic center), the value is set to this */
 #define WARMSTART_PROJ_MINRHSOBJ    1        /**< minimum value for rhs/obj when computing minimum eigenvalue for warmstart-projection */
@@ -141,6 +144,9 @@ struct SCIP_RelaxData
    SCIP_Bool             usescaling;         /**< whether the SDP-solver should use scaling */
    SCIP_Bool             scaleobj;           /**< whether the objective should be scaled in order to get a more stable behavior */
    SCIP_Bool             conflictconss;      /**< whether conflict constraints should be generated */
+   SCIP_Bool             conflictfeas;       /**< whether conflict constraints should be generated for feasible subproblems */
+   SCIP_Bool             conflictobjcut;     /**< whether an objective cut should be used to generate conflict constraints for feasible subproblems */
+   SCIP_Bool             conflictinfeas;     /**< whether conflict constraints should be generated for infeasible subproblems */
    int                   slatercheck;        /**< Should the Slater condition for the dual problem be checked ahead of solving every SDP ? */
    SCIP_Bool             sdpinfo;            /**< Should the SDP solver output information to the screen? */
[…]

### 2022-06-15 tex/ConflictAnalysis.tex
add first overview of literature

@@
 \maketitle
+
+The original intention of conflict analysis was to derive useful
+information from infeasible subproblems in a branch-and-bound
+procedure. There are two types of conflict analysis: 
+\begin{itemize}
+\item If infeasibility was determined using propagation of variable bounds,
+  one can use a graph-based conflict analysis. Here, a directed graph
+  encodes the dependency of bound changes on each other, i.e., whether one
+  bound was used to strengthen the other bound. One can then try to find a
+  small subset of bound changes that arise from branching decisions that
+  still lead to infeasibility. One can then add constraints that enforce
+  that at least one branching decision is taken in a different
+  way. See~\cite{Ach07b,Ach07} for more information in the context of MIPs.
+
+  This kind of conflict analysis can also be used in the context of MISDPs
+  and has already been investigated in~\cite{MatP22}. Overall, this kind of
+  conflict analysis was not beneficial.
+\item If the relaxation is infeasible, there are two ways to derive
+  conflict information: The first is a greedy approach, which relaxes some
+  of the branching decisions leading to the current relaxation as long as
+  the relaxation remains infeasible. Then a constraint as in graph-based
+  conflict analysis can be derived. See~\cite{Ach07b,Ach07} for more
+  information in the context of MIPs.
+
+  The other method is to use dual information to derive valid inequalities,
+  which can then be propagated during the tree. This is called dual ray
+  analysis by Witzig~\cite{Wit22} (see also \cite{WitBH17}). We follow this
+  approach in the following.
+\end{itemize}
 

### 2022-06-16 tex/ConflictAnalysis.tex
one more reference
extend description at bit

@@
 
+Very similar ideas are needed for an outer approximation algorithm, in
+which one adds aggregated inequalities that will cut off nodes that are
+infeasible or bound exceeding. Coey et al.~\cite{CoeLV20} describe such an
+algorithm for mixed-integer conic problems (including MISDP).
+
 % -------------------------------------------------------------------------
@@
 interested in solutions that are at least as good, i.e.,
-$b\T y \leq b\T y^\star$ or $- b\T y \geq - b\T y^\star$. Adding this
-inequality to~\eqref{eq:ConflictCut} yields:
+$b\T y \leq b\T y^\star$ or $- b\T y \geq - \overline{b}$, where
+$\overline{b} = b\T y^\star - \varepsilon$ for $\varepsilon > 0$. Adding
+this inequality to~\eqref{eq:ConflictCut} yields:
 \begin{align*}
   & \sum_{j=1}^m \skal{A^j}{\hat{X}} y_j + \hat{z}\T D y - b\T y \geq
-    \skal{A^0}{\hat{X}} + \hat{z}\T d - b\T y^\star\\
+    \skal{A^0}{\hat{X}} + \hat{z}\T d - \overline{b}\\
   \Leftrightarrow\quad
   & \sum_{j=1}^m \big(\skal{A^j}{\hat{X}} + (\hat{z}\T D)_j - b_j\big) y_j \geq
-    \skal{A^0}{\hat{X}} + \hat{z}\T d - b\T y^\star.
+    \skal{A^0}{\hat{X}} + \hat{z}\T d - \overline{b}.
 \end{align*}
+
+If the mixed-integer feasible point $y^\star$ was produced by the current
+node relaxation, then this cut will cut off~$y^\star$: First, by
+complementary slackness, the aggregated inequalities are satisfied with
+equality. Moreover, $b\T y \leq \overline{b} < b\T y^\star$ will be
+violated by construction. Thus, $y^\star$ is cut off. It remains to see
+whether propagation of the local bounds at this node will detect
+infeasibility.
 

### 2022-06-16 tex/ConflictAnalysis.tex
restructure, prove cut off for feasible problems

--- 2022-06-15 tex/ConflictAnalysis.tex
@@
 and $u$ are finite. Moreover, we always assume that $\ell \leq u$ holds.
+
+Let $X$ be the dual variables for the SDP-constraint $A(y) \succeq 0$, $z$
+be the dual variables for $Dy \geq d$, and $r^\ell$ and $r^u$ be the dual
+variables for the bound constraints $y \geq \ell$ and $-y \geq -u$,
+respectively. Then the corresponding primal problem for~\eqref{SDP-D} is:
+\begin{equation}\label{SDP-P}
+  \begin{aligned}
+     \sup \quad & \skal{A^0}{X} + z\T d + \ell\T r^\ell - u\T r^u\\
+     \text{s.t.} \quad & \skal{A^j}{X} + (D\T z)_j + r^\ell_j - r^u_j = b_j && \forall \, j \in [m],\\
+     & X \succeq 0,\; z, \; r^\ell,\; r^u \geq 0.
+  \end{aligned}
+\end{equation}
+
+% -------------------------------------------------------------------------
+% -------------------------------------------------------------------------
+\section{Generation of Conflict Constraints}
 
@@
 % -------------------------------------------------------------------------
-% -------------------------------------------------------------------------
-\section{Generation of Conflict Constraints}
-
-% -------------------------------------------------------------------------
-\subsection{Feasible Problems}
-
-Let $X$ be the dual variables for the SDP-constraint $A(y) \succeq 0$, $z$
-be the dual variables for $Dy \geq d$, and $r^\ell$ and $r^u$ be the dual
-variables for the bound constraints $y \geq \ell$ and $-y \geq -u$,
-respectively. Then the corresponding primal problem for~\eqref{SDP-D} is:
-\begin{equation}\label{SDP-P}
-  \begin{aligned}
-     \sup \quad & \skal{A^0}{X} + z\T d + \ell\T r^\ell - u\T r^u\\
-     \text{s.t.} \quad & \skal{A^j}{X} + (D\T z)_j + r^\ell_j - r^u_j = b_j && \forall \, j \in [m],\\
-     & X \succeq 0,\; z, \; r^\ell,\; r^u \geq 0.
-  \end{aligned}
-\end{equation}
-Any feasible solution $(\hat{X}, \hat{z}, \hat{r}^\ell, \hat{r}^u)$ can be
-used to construct~\eqref{eq:ConflictCut} or~\eqref{eq:ConflictCut2}.
-
-Moreover, one can take an objective cutoff into account. Namely, assume
-that $y^\star$ is a feasible solution to~\eqref{MISDP}. We are then only
-interested in solutions that are at least as good, i.e.,
-$b\T y \leq b\T y^\star$ or $- b\T y \geq - b\T y^\star$. Adding this
-inequality to~\eqref{eq:ConflictCut} yields:
-\begin{align*}
-  & \sum_{j=1}^m \skal{A^j}{\hat{X}} y_j + \hat{z}\T D y - b\T y \geq
-    \skal{A^0}{\hat{X}} + \hat{z}\T d - b\T y^\star\\
-  \Leftrightarrow\quad
-  & \sum_{j=1}^m \big(\skal{A^j}{\hat{X}} + (\hat{z}\T D)_j - b_j\big) y_j \geq
-    \skal{A^0}{\hat{X}} + \hat{z}\T d - b\T y^\star.
-\end{align*}
-
-% -------------------------------------------------------------------------
 \subsection{Infeasible Problems}
 
-Similarly, a primal ray $(\hat{X}, \hat{z}, \hat{r}^\ell, \hat{r}^u)$,
-i.e., it satsifies the system
+A primal ray, i.e., a point $(\hat{X}, \hat{z}, \hat{r}^\ell, \hat{r}^u)$
+satsifying the system
 \begin{equation}\label{eq:ray}
@@
 % -------------------------------------------------------------------------
+\subsection{Feasible Problems}
+
+Any feasible solution $(\hat{X}, \hat{z}, \hat{r}^\ell, \hat{r}^u)$
+of~\eqref{SDP-D} can be used to construct~\eqref{eq:ConflictCut}
+or~\eqref{eq:ConflictCut2}.
+
+Moreover, one can take an objective cutoff into account. Namely, assume
+that $y^\star$ is a feasible solution to~\eqref{MISDP}. We are then only
+interested in solutions that are at least as good, i.e.,
+$b\T y \leq \overline{b}$ or equivalently $- b\T y \geq - \overline{b}$,
+where $\overline{b} = b\T y^\star - \varepsilon$ for $\varepsilon >
+0$. Adding this inequality to~\eqref{eq:ConflictCut} yields:
+\begin{align}
+  & \sum_{j=1}^m \skal{A^j}{\hat{X}} y_j + \hat{z}\T D y - b\T y \geq
+    \skal{A^0}{\hat{X}} + \hat{z}\T d - \overline{b}\notag\\
+  \Leftrightarrow\quad
+  & \sum_{j=1}^m \big(\skal{A^j}{\hat{X}} + (\hat{z}\T D)_j - b_j\big) y_j \geq
+    \skal{A^0}{\hat{X}} + \hat{z}\T d - \overline{b}.\label{eq:ConflictCutObj}
+\end{align}
+
+\begin{proposition}
+  Let $(\hat{X}, \hat{z})$ be part of a solution of~\eqref{SDP-P} and
+  $\hat{y}$ be a mixed-integer local solution of the corresponding dual
+  problem. Assume that strong duality holds. Define
+  $\overline{b} = b\T \hat{y} -
+  \varepsilon$. Then~\eqref{eq:ConflictCutObj}, i.e.,
+  \[
+    \sum_{j=1}^m \big(\skal{A^j}{\hat{X}} + (\hat{z}\T D)_j - b_j\big) y_j \geq
+    \skal{A^0}{\hat{X}} + \hat{z}\T d - \overline{b}.
+  \]
+  is infeasible with respect to the local bounds $\ell$ and $u$.
+\end{proposition}
+
+\begin{proof}
+  Define $\hat{s}_j \define \skal{A^j}{\hat{X}} + (\hat{z}\T D)_j - b_j$
+  for $j \in [m]$. Then we get from~\eqref{eq:ConflictCutObj} for any
+  feasible solution $y$ with $b\T y \leq \overline{b}$:
+  \begin{align*}
+    \skal{A^0}{\hat{X}} + \hat{z}\T d - \overline{b} & \leq \sum_{j=1}^m \big(\skal{A^j}{\hat{X}} + (\hat{z}\T D)_j - b_j\big)\, y_j
+    = \sum_{j=1}^m \hat{s}_j\, y_j\\
+    & \leq \sum_{j: \hat{s}_j < 0} \hat{s}_j\, \ell_j +
+        \sum_{j: \hat{s}_j > 0} \hat{s}_j\, u_j.
+  \end{align*}
+
+  On the other hand let $(\hat{X}, \hat{z}, \hat{r}^\ell, \hat{r}^u)$ be a
+  feasible solution of~\eqref{SDP-P} that satisfies the conditions of
+  Lemma~\ref{lemma:ComplemenatrySolution}. Then
+  $\hat{r}^u - \hat{r}^\ell = \hat{s}$.  We thus obtain by strong duality:
+  \begin{align*}
+    b\T \hat{y} & = \skal{A^0}{\hat{X}} + d\T \hat{z} + \ell\T \hat{r}^\ell - u\T \hat{r}^u\\
+      & = \skal{A^0}{\hat{X}} + d\T \hat{z} + \sum_{j=1}^m \ell_j \hat{r}^\ell_j -
+        \sum_{j=1}^m u_j \hat{r}^u_j\\
+      & = \skal{A^0}{\hat{X}}  + d\T \hat{z} - \sum_{j: \hat{s}_j < 0} \ell_j\, \hat{s}_j -
+        \sum_{j: \hat{s}_j > 0} u_j\, \hat{s}_j.
+  \end{align*}
+  This means that
+  \begin{align*}
+    \sum_{j: \hat{s}_j < 0} \ell_j\, \hat{s}_j +
+        \sum_{j: \hat{s}_j > 0} u_j\, \hat{s}_j = \skal{A^0}{\hat{X}}  +
+    d\T \hat{z} - \overline{b} - \varepsilon
+  \end{align*}
+  and therefore
+  \begin{align*}
+    \sum_{j: \hat{s}_j < 0} \ell_j\, \hat{s}_j +
+        \sum_{j: \hat{s}_j > 0} u_j\, \hat{s}_j < \skal{A^0}{\hat{X}}  +
+    d\T \hat{z} - \overline{b},
+  \end{align*}
+  a contradiction.
+\end{proof}
+
+% -------------------------------------------------------------------------
 \subsection{Canceling}

### 2022-06-16 tex/ConflictAnalysis.tex
rework

@@
 
-The original intention of conflict analysis was to derive useful
-information from infeasible subproblems in a branch-and-bound
-procedure. There are two types of conflict analysis:
-\begin{itemize}
-\item If infeasibility was determined using propagation of variable bounds,
-  one can use a graph-based conflict analysis. Here, a directed graph
-  encodes the dependency of bound changes on each other, i.e., whether one
-  bound was used to strengthen the other bound. One can then try to find a
-  small subset of bound changes that arise from branching decisions that
-  still lead to infeasibility. One can then add constraints that enforce
-  that at least one branching decision is taken in a different
-  way. See~\cite{Ach07b,Ach07} for more information in the context of MIPs.
-
-  This kind of conflict analysis can also be used in the context of MISDPs
-  and has already been investigated in~\cite{MatP22}. Overall, this kind of
-  conflict analysis was not beneficial.
-\item If the relaxation is infeasible, there are two ways to derive
-  conflict information: The first is a greedy approach, which relaxes some
-  of the branching decisions leading to the current relaxation as long as
-  the relaxation remains infeasible. Then a constraint as in graph-based
-  conflict analysis can be derived. See~\cite{Ach07b,Ach07} for more
-  information in the context of MIPs.
-
-  The other method is to use dual information to derive valid inequalities,
-  which can then be propagated during the tree. This is called dual ray
-  analysis by Witzig~\cite{Wit22} (see also \cite{WitBH17}). We follow this
-  approach in the following.
-\end{itemize}
-
-Very similar ideas are needed for an outer approximation algorithm, in
-which one adds aggregated inequalities that will cut off nodes that are
-infeasible or bound exceeding. Coey et al.~\cite{CoeLV20} describe such an
-algorithm for mixed-integer conic problems (including MISDP).
-
 % -------------------------------------------------------------------------
@@
 
+We will need the following result.
+
+\begin{lemma}\label{lemma:ComplemenatrySolution}
+  Let $\hat{r}^\ell$, $\hat{r}^u \in \R^m_+$ with
+  $\hat{r}^\ell - \hat{r}^u = \hat{s}$ for some $\hat{s} \in \R^m$. Then
+  there exists $r^\ell$, $r^u \in \R^m_+$ with $r^\ell - r^u = s$, such
+  that $r_j^\ell \cdot r_j^u = 0$ for all $j \in [m]$ and for
+  $\ell \leq u$:
+  \begin{equation}\label{eq:objpart}
+    \ell\T r^\ell - u\T r^u \geq \ell\T \hat{r}^\ell - u\T \hat{r}^u.
+  \end{equation}
+\end{lemma}
+
+\begin{proof}
+  Define
+  \[
+    r^\ell_j \define - \min\,\{\hat{s}_j, 0\} \geq 0,
+    \quad
+    r^u_j = \max\,\{\hat{s}_j, 0\} \geq 0.
+  \]
+  By construction $r^\ell_j \cdot r^u_j = 0$ for all
+  $j \in [m]$ and $r^\ell - r^u = s$. Then consider:
+  \begin{align*}
+    \ell_j r^\ell_j - u_j r^u_j & = - \ell_j\, \min\,\{\hat{s}_j, 0\} - u_j\, \max\,\{\hat{s}_j, 0\}\\
+    & = - \ell_j\, \min\,\{\hat{r}^u_j - \hat{r}^\ell_j, 0\} - u_j\, \max\,\{\hat{r}^u_j - \hat{r}^\ell_j, 0\}\\
+    & = \begin{cases}
+      - \ell_j\, (\hat{r}^u_j - \hat{r}^\ell_j) & \text{if } \hat{r}^u_j - \hat{r}^\ell_j < 0,\\
+      - u_j\, (\hat{r}^u_j - \hat{r}^\ell_j) & \text{if } \hat{r}^u_j - \hat{r}^\ell_j \geq 0,
+    \end{cases}\\
+    & = \begin{cases}
+      - \ell_j\, \hat{r}^u_j + \ell_j \hat{r}^\ell_j & \text{if } \hat{r}^u_j - \hat{r}^\ell_j < 0,\\
+      - u_j\, \hat{r}^u_j + u_j \hat{r}^\ell_j & \text{if } \hat{r}^u_j - \hat{r}^\ell_j \geq 0,
+    \end{cases}\\
+    & \geq \begin{cases}
+      - u_j\, \hat{r}^u_j + \ell_j \hat{r}^\ell_j & \text{if }  \hat{r}^u_j - \hat{r}^\ell_j < 0,\\
+      - u_j\, \hat{r}^u_j + \ell_j \hat{r}^\ell_j & \text{if } \hat{r}^u_j - \hat{r}^\ell_j \geq 0,
+    \end{cases}\\
+    & = \ell_j \hat{r}^\ell_j - u_j \hat{r}^u_j,
+  \end{align*}
+  which shows~\eqref{eq:objpart}.
+\end{proof}
+
+% -------------------------------------------------------------------------
+% -------------------------------------------------------------------------
+\section{Brief Literature Review for Conflict Analysis}
+
+The original intention of conflict analysis was to derive useful
+information from infeasible subproblems in a branch-and-bound
+procedure. There are three types of conflict analysis:
+\begin{itemize}
+\item If infeasibility was determined using propagation of variable bounds,
+  one can use a \emph{graph-based conflict analysis}. Here, a directed graph
+  encodes the dependency of bound changes on each other, i.e., whether one
+  bound was used to strengthen the other bound. One can then try to find a
+  small subset of bound changes that arise from branching decisions that
+  still lead to infeasibility. One can then add constraints enforcing
+  that at least one branching decision is taken in a different
+  way. See~\cite{Ach07b,Ach07} for more information in the context of MIPs.
+
+  This kind of conflict analysis can also be used in the context of MISDPs
+  and has already been investigated in~\cite{MatP22}. Overall, this kind of
+  conflict analysis was not beneficial.
+\item If the relaxation is infeasible, there is a \emph{greedy approach},
+  which relaxes some of the branching decisions defining the current node
+  as long as the relaxation remains infeasible. Then a constraint as in
+  graph-based conflict analysis can be derived. See~\cite{Ach07b,Ach07} for
+  more information in the context of MIPs. This method does not seem very
+  promising for MISDPs, since we do not have warm start in this case.
+\item If the relaxation is infeasible, one can also use dual information
+  to derive valid inequalities, which can then be propagated during the
+  tree. This is called dual ray analysis by Witzig~\cite{Wit22} (see also
+  \cite{WitBH17}).
+\end{itemize}
+
+In the following, we will generalize the last approach to MISDPs and call
+it \emph{dual conflict analysis}.
+
+Very similar ideas are needed for an outer approximation algorithm, in
+which one adds aggregated inequalities that will cut off nodes that are
+infeasible or bound exceeding. Coey et al.~\cite{CoeLV20} describe such an
+algorithm for mixed-integer conic problems (including MISDP).
+
 % -------------------------------------------------------------------------
@@
 
-The general idea of conflict analysis is to generate globally valid
+The general idea of dual conflict analysis is to generate globally valid
 constraints as follows. Let any $\hat{X} \succeq 0$ and $\hat{z} \geq 0$ be
@@
 $y \leq u$. If $\hat{r}^\ell$, $\hat{r}^u \geq 0$ are the corresponding
-multiliers, this yields:
+multipliers, this yields:
 \begin{align}
@@
 
-It seems that~\eqref{eq:ConflictCut2} are less helpful
+It seems that Inequalities~\eqref{eq:ConflictCut2} are less helpful
 than~\eqref{eq:ConflictCut}. However, one can cancel some coefficients
@@
 A primal ray, i.e., a point $(\hat{X}, \hat{z}, \hat{r}^\ell, \hat{r}^u)$
-satsifying the system
+satisfying the system
 \begin{equation}\label{eq:ray}
@@
 In this case, one can prove that~\eqref{eq:ConflictCut} actually proves
-infeasibility for the local bounds~$\ell$ and $u$. To this end, we first
-show the following:
-
-\begin{lemma}\label{lemma:ComplemenatrySolution}
-  If~\eqref{eq:ray} has a feasible solution, then there exists a feasible
-  solution in which $r_j^\ell \cdot r_j^u = 0$ for all $j \in [m]$, i.e.,
-  at most one of $r_j^\ell$ and $r_j^u$ is nonzero.
-\end{lemma}
-
-\begin{proof}
-  Let $(\hat{X}, \hat{z}, \tilde{r}^\ell, \tilde{r}^u)$ be a feasible
-  solution of~\eqref{eq:ray}. Then define
-  \[
-    \hat{r}^\ell_j \define - \min\,\{\skal{A^j}{\hat{X}} + (D\T \hat{z})_j,\; 0\} \geq 0,
-    \quad
-    \hat{r}^u_j = \max\,\{\skal{A^j}{\hat{X}} + (D\T \hat{z})_j,\; 0\} \geq 0.
-  \]
-  We claim that $(\hat{X}, \hat{z}, \hat{r}^\ell, \hat{r}^u)$ is feasible
-  for~\eqref{eq:ray}. Note that by construction we have
-  $\hat{r}^\ell_j \cdot \hat{r}^u_j = 0$ for all $j \in [m]$.
-
-  First observere that the equations of~\eqref{eq:ray} are
-  satisfied. Therefore consider the contribution to
-  $\ell\T \hat{r}^\ell - u\T \hat{r}^u$ in component $j$:
-  \begin{align*}
-    \ell_j \hat{r}^\ell_j - u_j \hat{r}^u_j & = - \ell_j\, \min\,\{\skal{A^j}{\hat{X}} + (D\T \hat{z})_j, 0\} - u_j\, \max\,\{\skal{A^j}{\hat{X}} + (D\T \hat{z})_j, 0\}\\
-    & = - \ell_j\, \min\,\{\tilde{r}^u_j - \tilde{r}^\ell_j, 0\} - u_j\, \max\,\{\tilde{r}^u_j - \tilde{r}^\ell_j, 0\}\\
-    & = \begin{cases}
-      - \ell_j\, (\tilde{r}^u_j - \tilde{r}^\ell_j) & \text{if } \tilde{r}^u_j - \tilde{r}^\ell_j < 0,\\
-      - u_j\, (\tilde{r}^u_j - \tilde{r}^\ell_j) & \text{if } \tilde{r}^u_j - \tilde{r}^\ell_j \geq 0,
-    \end{cases}\\
-    & = \begin{cases}
-      - \ell_j\, \tilde{r}^u_j + \ell_j \tilde{r}^\ell_j & \text{if } \tilde{r}^u_j - \tilde{r}^\ell_j < 0,\\
-      - u_j\, \tilde{r}^u_j + u_j \tilde{r}^\ell_j & \text{if } \tilde{r}^u_j - \tilde{r}^\ell_j \geq 0,
-    \end{cases}\\
-    & \geq \begin{cases}
-      - u_j\, \tilde{r}^u_j + \ell_j \tilde{r}^\ell_j & \text{if }  \tilde{r}^u_j - \tilde{r}^\ell_j < 0,\\
-      - u_j\, \tilde{r}^u_j + \ell_j \tilde{r}^\ell_j & \text{if } \tilde{r}^u_j - \tilde{r}^\ell_j \geq 0,
-    \end{cases}\\
-    & = \ell_j \tilde{r}^\ell_j - u_j \tilde{r}^u_j.
-  \end{align*}
-  Therefore the inequality in~\eqref{eq:ray} holds as well.
-\end{proof}
+infeasibility for the local bounds~$\ell$ and $u$ via propagation:
 
 \begin{proposition}
-  Let $(\hat{X}, \hat{z})$ be part of a solution of~\eqref{eq:ray}. Then~\eqref{eq:ConflictCut}, i.e.,
+  Let $(\hat{X}, \hat{z}, \hat{r}^\ell, \hat{r}^u)$ be a solution
+  of~\eqref{eq:ray}. Then~\eqref{eq:ConflictCut}, i.e.,
   \[
@@
 
-  On the other hand let $(\hat{X}, \hat{z}, \hat{r}^\ell, \hat{r}^u)$ be a
-  feasible solution of~\eqref{eq:ray} that satisfies the conditions of
-  Lemma~\ref{lemma:ComplemenatrySolution}. Then
-  $\hat{r}^u - \hat{r}^\ell = \hat{s}$.  We thus obtain
+  On the other hand, by applying Lemma~\ref{lemma:ComplemenatrySolution} to
+  $\hat{r}^\ell - \hat{r}^u = \hat{s}$, we can assume that at most one of
+  $\hat{r}^\ell_j$ and $\hat{r}^u_j$ is nonzero and the inequality
+  in~\eqref{eq:ray} is still statisfied. We thus obtain
   \begin{align*}
@@
   \end{align*}
-  a contradition.
+  a contradiction.
 \end{proof}
+
+The hope is that~\eqref{eq:ConflictCut} is still violated for local bounds
+not far from $\ell \leq u$.
 
@@
 $b\T y \leq \overline{b}$ or equivalently $- b\T y \geq - \overline{b}$,
-where $\overline{b} = b\T y^\star - \varepsilon$ for $\varepsilon >
-0$. Adding this inequality to~\eqref{eq:ConflictCut} yields:
+where $\overline{b} \define b\T y^\star - \varepsilon$ for
+$\varepsilon > 0$. Adding this inequality to~\eqref{eq:ConflictCut} yields:
 \begin{align}
@@
 \begin{proposition}
-  Let $(\hat{X}, \hat{z})$ be part of a solution of~\eqref{SDP-P} and
-  $\hat{y}$ be a mixed-integer local solution of the corresponding dual
-  problem. Assume that strong duality holds. Define
-  $\overline{b} = b\T \hat{y} -
-  \varepsilon$. Then~\eqref{eq:ConflictCutObj}, i.e.,
+  Let $(\hat{X}, \hat{z}, \hat{r}^\ell, \hat{r}^u)$ be a solution
+  of~\eqref{SDP-P} and $\hat{y}$ be a mixed-integer local solution of the
+  corresponding dual problem. Assume that strong duality holds. Define
+  $\overline{b} \define b\T \hat{y} - \varepsilon$ for $\varepsilon >
+  0$. Then~\eqref{eq:ConflictCutObj}, i.e.,
   \[
@@
 
-  On the other hand let $(\hat{X}, \hat{z}, \hat{r}^\ell, \hat{r}^u)$ be a
-  feasible solution of~\eqref{SDP-P} that satisfies the conditions of
-  Lemma~\ref{lemma:ComplemenatrySolution}. Then
-  $\hat{r}^u - \hat{r}^\ell = \hat{s}$.  We thus obtain by strong duality:
+  On the other hand, by applying Lemma~\ref{lemma:ComplemenatrySolution} to
+  $\hat{r}^\ell - \hat{r}^u = \hat{s}$, we can assume that at most one of
+  $\hat{r}^\ell_j$ and $\hat{r}^u_j$ is nonzero and the corresponding
+  solution is still optimal. We thus obtain by strong duality:
   \begin{align*}
@@
 
+The hope is that~\eqref{eq:ConflictCutObj} is still violated for local
+bounds not far from $\ell \leq u$.
+
 % -------------------------------------------------------------------------
@@
 as follows. Again using
-$s_j \define \skal{A^j}{\hat{X}} + (\hat{z}\T D)_j$, we get
+$\hat{s}_j \define \skal{A^j}{\hat{X}} + (\hat{z}\T D)_j$, we get
 from~\eqref{eq:ConflictCut}:
 \begin{equation}\label{eq:ConflictCut3}
-  \sum_{j=1}^m s_j\, y_j \geq \skal{A^0}{\hat{X}} + \hat{z}\T d.
+  \sum_{j=1}^m \hat{s}_j\, y_j \geq \skal{A^0}{\hat{X}} + \hat{z}\T d.
 \end{equation}
 
-Assume that $s_j < 0$ for some $j \in [m]$ and the global lower bound of
-$y_j$ is at least 0. Then we can add the valid inequality $-s_j\, y_j \geq 0$
-to~\eqref{eq:ConflictCut3}, which eliminates variable~$y_j$ from the
-inequality. Similarly, if $s_j > 0$ for some $j \in [m]$ and the global
-upper bound of $y_j$ is at most 0, we can add $s_j\, y_j \leq 0$ to
-eliminate variable~$y_j$.
+Assume that $\hat{s}_j < 0$ for some $j \in [m]$ and the global lower bound
+of $y_j$ is at least 0. Then we can add the valid inequality
+$-\hat{s}_j\, y_j \geq 0$ to~\eqref{eq:ConflictCut3}, which eliminates
+variable~$y_j$ from the inequality. Similarly, if $\hat{s}_j > 0$ for some
+$j \in [m]$ and the global upper bound of $y_j$ is at most 0, we can add
+$\hat{s}_j\, y_j \leq 0$ to eliminate variable~$y_j$.
 
 Witzig et al.~\cite{WitBH17} (see also Witzig~\cite{Wit22}) applied
-cancelation to continuous variables, but restricts attention to variables
+cancellation to continuous variables, but restricts attention to variables
 that are at their global bounds at the node at which the constraint is

### 2022-07-08 tex/ConflictAnalysis.tex
minor
fixes to tex document

@@
   $\hat{r}^\ell - \hat{r}^u = \hat{s}$ for some $\hat{s} \in \R^m$. Then
-  there exists $r^\ell$, $r^u \in \R^m_+$ with $r^\ell - r^u = s$, such
+  there exists $r^\ell$, $r^u \in \R^m_+$ with $r^\ell - r^u = \hat{s}$, such
   that $r_j^\ell \cdot r_j^u = 0$ for all $j \in [m]$ and for
@@
   By construction $r^\ell_j \cdot r^u_j = 0$ for all
-  $j \in [m]$ and $r^\ell - r^u = s$. Then consider:
+  $j \in [m]$ and $r^\ell - r^u = \hat{s}$. Then consider:
   \begin{align*}
@@
   & \skal{A(y)}{\hat{X}} + \hat{z}\T D y + (\hat{r}^\ell)\T y - (\hat{r}^u)\T y
-    \geq \hat{z}\T d + (\hat{r}^\ell)\T \ell + (\hat{r}^u)\T u\notag\\
+    \geq \hat{z}\T d + (\hat{r}^\ell)\T \ell - (\hat{r}^u)\T u\notag\\
   \Leftrightarrow\quad
@@
   & \qquad \skal{A^0}{\hat{X}} +
-    \hat{z}\T d + (\hat{r}^\ell)\T \ell + (\hat{r}^u)\T u,\label{eq:ConflictCut2}
+    \hat{z}\T d + (\hat{r}^\ell)\T \ell - (\hat{r}^u)\T u,\label{eq:ConflictCut2}
 \end{align}
@@
   On the other hand, by applying Lemma~\ref{lemma:ComplemenatrySolution} to
-  $\hat{r}^\ell - \hat{r}^u = \hat{s}$, we can assume that at most one of
+  $\hat{r}^\ell - \hat{r}^u = -\hat{s}$, we can assume that at most one of
   $\hat{r}^\ell_j$ and $\hat{r}^u_j$ is nonzero and the inequality
-  in~\eqref{eq:ray} is still statisfied. We thus obtain
+  in~\eqref{eq:ray} is still satisfied. We thus obtain
   \begin{align*}
@@
 Any feasible solution $(\hat{X}, \hat{z}, \hat{r}^\ell, \hat{r}^u)$
-of~\eqref{SDP-D} can be used to construct~\eqref{eq:ConflictCut}
+of~\eqref{SDP-P} can be used to construct~\eqref{eq:ConflictCut}
 or~\eqref{eq:ConflictCut2}.
@@
   for $j \in [m]$. Then we get from~\eqref{eq:ConflictCutObj} for any
-  feasible solution $y$ with $b\T y \leq \overline{b}$:
+  feasible solution $y$ with $b\T y \leq \overline{b}$ with $\ell \leq y
+  \leq u$:
   \begin{align*}
@@
   On the other hand, by applying Lemma~\ref{lemma:ComplemenatrySolution} to
-  $\hat{r}^\ell - \hat{r}^u = \hat{s}$, we can assume that at most one of
+  $\hat{r}^\ell - \hat{r}^u = -\hat{s}$, we can assume that at most one of
   $\hat{r}^\ell_j$ and $\hat{r}^u_j$ is nonzero and the corresponding
@@
 
-Witzig et al.~\cite{WitBH17} (see also Witzig~\cite{Wit22}) applied
-cancellation to continuous variables, but restricts attention to variables
+Witzig et al.~\cite{WitBH17} (see also Witzig~\cite{Wit22}) apply
+cancellation to continuous variables, but restrict attention to variables
 that are at their global bounds at the node at which the constraint is

### 2022-08-01 tex/ConflictAnalysis.tex
add question to tex document
add explanation about primal solutions

@@
 \end{equation}
+\todo[inline]{Frederic: I think we also need~$X \succeq 0$ and~$z,\, r^{\ell},\, u
+  \geq 0$ in~\eqref{eq:ray}.}
 yields an inequality as in~\eqref{eq:ConflictCut}
@@
 
+% -------------------------------------------------------------------------
+\subsection{Primal Solutions for Special Cases}
+\label{sec:Primal}
+
+In order to produce conflict constraints, one needs solutions for the
+primal of the LP-relaxation~\eqref{SDP-P}. Because SCIP-SDP preprocesses
+certain SDP problems, we briefly discuss how primal solutions can be
+obtained in these cases.
+
+% -------------------------------------------------------------------------
+\subsubsection{SDPs with One Variable}
+
+For the special case of SDP-relaxations in which only one variable is
+unfixed, SCIP-SDP uses specialized methods to solve them. In order to
+derive a primal solution, consider
+\begin{equation}\label{eq:basic}
+  \inf\; \{\gamma\, \mu \suchthat \mu\, A - B \succeq 0,\; \ell \leq \mu \leq u\},
+\end{equation}
+where $A$, $B \in \R^{n \times n}$ are symmetric matrices, $\gamma \in \R$,
+and $\ell \in \R \cup \{-\infty\}$, $u \in \R \cup \{\infty\}$.
+
+The dual SDP of~\eqref{eq:basic} is
+\begin{equation}\label{eq:dual}
+  \sup\; \{\skal{B}{X} + \ell\, \alpha - u\, \beta \suchthat \skal{A}{X} +
+  \alpha - \beta = \gamma,\; X \succeq 0,\; \alpha,\; \beta \geq 0\}.
+\end{equation}
+
+If $\mu^*$ is an optimal solution for~\eqref{eq:basic}, we can construct an
+optimal primal solution $(X^*, \alpha^*, \beta^*)$ as follows.
+
+\paragraph{Case 0, $\boldsymbol{\gamma = 0}$:}
+
+Here we can take $(X^*, \alpha^*, \beta^*) = (0,0,0)$. Thus, in the
+following, we assume that $\gamma \neq 0$.
+
+\paragraph{Case 1, $\boldsymbol{\ell < \mu^* < u}$:}
+
+Then by complementary slackness necessarily $\alpha^* = \beta^* = 0$, and
+thus $\skal{A}{X^*} = \gamma$. Optimality implies that
+\[
+\gamma\, \mu^* = \skal{B}{X^*}
+\quad\Leftrightarrow\quad
+\mu^* \skal{A}{X^*} = \skal{B}{X^*}
+\quad\Leftrightarrow\quad
+\skal{\mu^* A - B}{X^*} = 0.
+\]
+
+Note that in~\eqref{eq:dual} there is only one affine constraint, which
+implies that there exists a rank-1 solution. Thus, we are aiming to
+construct a rank-1 solution $X^* = x^* (x^*)\T \succeq 0$.  This solution
+has to satisfy:
+\[
+0 = \skal{\mu^* A - B}{x^* (x^*)\T} = (x^*)\T (\mu^* A - B) x^*.
+\]
+Therefore, $x^*$ needs to be an eigenvector for the eigenvalue~$0$ of
+$\mu^* A - B$ and thus $x^* \in \ker(\mu^* A - B)$. Moreover,
+$\skal{A}{x^*(x^*)\T} = (x^*)\T A x^* = \gamma$ needs to hold.  Consider a
+generator matrix $C \in \R^{n \times r}$ of the kernel of $\mu^* A - B$,
+i.e., $\ker(\mu^* A - B) = \{C y \suchthat y \in \R^r\}$.  Then by scaling,
+the problem reduces to the question whether there exists $y \in \R^r$ such
+that
+\[
+y\T C\T A C y
+\begin{cases}
+  > 0 & \text{if } \gamma > 0,\\
+  < 0 & \text{if } \gamma < 0.
+\end{cases}
+\]
+Computing a maximal/minimal eigenvalue/eigenvector of $C\T A C$ will decide
+this.
+
+\paragraph{Case 2, $\boldsymbol{\mu^* = \ell}$:}
+
+In this case, $\beta^* = 0$ has to hold.
+
+If $\gamma \geq 0$, we can take
+$(X^*, \alpha^*, \beta^*) = (0, \gamma, 0)$, which is feasible with
+objective value $\gamma \ell = \gamma \mu^*$ as needed. Therefore, this
+point is optimal.
+
+If $\gamma < 0$, the conditions are
+\[
+\skal{B}{X^*} + \ell\, \alpha^* = \gamma\, \ell,\quad
+\skal{A}{X^*} + \alpha^* = \gamma,\quad X^* \succeq 0,\; \alpha^* \geq 0.
+\]
+This yields
+\[
+\skal{B}{X^*} + \ell\, \alpha^* = \skal{A}{X^*}\, \ell + \alpha^*\, \ell
+\quad\Leftrightarrow\quad
+\skal{\ell\, A - B}{X^*} = 0.
+\]
+As above, we are looking for a rank-1 solution $(x^*)\T x^*$. Let
+$C \in \R^{n \times r}$ be a generator matrix of $\ker(\ell\, A - B)$. Then
+we need $y^* \in \R^r$ such that
+\[
+(y^*)\T C\T A C y^* < 0.
+\]
+If this is impossible, the problem is infeasible (e.g., if $A \succeq
+0$). Otherwise, we can scale $x^* = C y^*$ and set $\alpha^* = 0$.
+
+\paragraph{Case 3, $\boldsymbol{\mu^* = u}$:}
+
+This case is symmetric to Case 2. If $\gamma < 0$, we can set
+$(X^*, \alpha^*, \beta^*) = (0, 0, -\gamma)$.
+
+If $\gamma \geq 0$, we can construct a rank-1 solution as above.
+
 \begin{small}

### 2022-08-01 tex/ConflictAnalysis.tex
add case with all fixed variables
add something on the infeasible case
some fixes

@@
 \]
-Therefore, $x^*$ needs to be an eigenvector for the eigenvalue~$0$ of
-$\mu^* A - B$ and thus $x^* \in \ker(\mu^* A - B)$. Moreover,
-$\skal{A}{x^*(x^*)\T} = (x^*)\T A x^* = \gamma$ needs to hold.  Consider a
-generator matrix $C \in \R^{n \times r}$ of the kernel of $\mu^* A - B$,
-i.e., $\ker(\mu^* A - B) = \{C y \suchthat y \in \R^r\}$.  Then by scaling,
-the problem reduces to the question whether there exists $y \in \R^r$ such
-that
+Since $\mu^* A - B \succeq 0$, there exists a matrix
+$D \in \R^{k \times n}$ such that $D\T D = \mu^* A - B$. Thus, we are
+searching for $x^*$ with $(D x^*)\T D x^* = 0$. Therefore,
+$x^* \in \ker(D)$. In fact,
+$\ker(D) = \ker(\mu^* A - B)$.\footnote{Clearly,
+  $\ker(D) \subseteq \ker(\mu^* A - B)$. Conversely, let
+  $x \in \ker(\mu^* A - B)$. Then
+  $(Dx)\T Dx = \norm{Dx}_2^2 = (\mu^* A - B)x = 0$. But then $Dx = 0$.}
+Moreover, $\skal{A}{x^*(x^*)\T} = (x^*)\T A x^* = \gamma$ needs to hold.
+Consider a generator matrix $C \in \R^{n \times r}$ of the kernel of
+$\mu^* A - B$, i.e., $\ker(\mu^* A - B) = \{C y \suchthat y \in \R^r\}$.
+Then by scaling, the problem reduces to the question whether there exists
+$y^* \in \R^r$ such that
 \[
-y\T C\T A C y
+(y^*)\T C\T A C y^*
 \begin{cases}
@@
 
+\paragraph{Case 4, infeasible:}
+
+If~\eqref{eq:basic} is infeasible, we are searching for a rank-1 solution
+$X^* = (x^*)\T x^*$ such that
+\[
+\skal{B}{X^*}
+\begin{cases}
+ > 0 & \text{if } \gamma > 0,\\
+ < 0 & \text{if } \gamma < 0
+\end{cases}
+\]
+and $\skal{A}{X^*} = 0$ (and $\alpha^* = \beta^* = 0$). Let again $C^{n
+  \times r}$ be a generator matrix of $\ker(A)$ and compute
+\[
+y\T C\T B C y
+\begin{cases}
+  > 0 & \text{if } \gamma > 0,\\
+  < 0 & \text{if } \gamma < 0.
+\end{cases}
+\]
+By scaling, this produces the wanted solution.
+
+% -------------------------------------------------------------------------
+\subsubsection{SDPs with All Fixed Variables}
+
+Consider the general problem with fixed variables:
+\[
+  \inf\; \{ b\T y \suchthat \sum_{i=1}^m A^i\, y_i - A^0 \succeq 0,\; \ell \leq y \leq \ell\},
+\]
+and its dual:
+\[
+  \sup\; \{\skal{A^0}{X} + \ell\T \alpha - \ell\T \beta \suchthat \skal{A}{X} +
+  \alpha_i - \beta_i = b_i\; \forall i \in [m],\; X \succeq 0,\; \alpha,\; \beta \geq 0\}.
+\]
+We can take
+\[
+X^* = 0,\quad
+\alpha^*_i = \max\,\{0, b_i\}, \quad
+\beta^*_i = -\min\,\{0, b_i\}.
+\]
+This solution is feasible for the dual and has objective value $\ell\T b$.
+
 \begin{small}

### 2022-08-01 src/scipsdp/relax_sdp.c
add function SCIPsdpiHavePrimalSol()

@@ -1296,6 +1296,12 @@ SCIP_RETCODE computeConflictCut(
    assert( conflictcutlhs != NULL );
    assert( success != NULL );
 
+   /* only run if we can get a primal solution */
+   if ( ! SCIPsdpiHavePrimalSol(sdpi) )
+   {
+      *success = FALSE;
+      return SCIP_OKAY;
+   }
    *success = TRUE;
 
    nvars = SCIPgetNVars(scip);

### 2022-08-01 src/sdpi/sdpi.c

@@ -3268,6 +3268,18 @@ SCIP_Bool SCIPsdpiSolvedOrig(
    return ( SCIPsdpiWasSolved(sdpi) && (! sdpi->penalty) );
 }
 
+/** returns whether a primal solution or ray is available */
+SCIP_Bool SCIPsdpiHavePrimalSol(
+   SCIP_SDPI*            sdpi                /**< SDP-interface structure */
+   )
+{
+   if ( ! sdpi->solved || sdpi->infeasible || sdpi->allfixed || sdpi->solvedonevarsdp == SCIP_ONEVAR_INFEASIBLE || SCIPsdpiSolverIsPrimalInfeasible(sdpi->sdpisolver) )
+   {
+      return FALSE;
+   }
+   return TRUE;
+}
+
 /** returns true if the solver could determine whether the problem is feasible, so it returns true if the
  *  solver knows that the problem is feasible/infeasible/unbounded, it returns false if the solver does not know
  *  anything about the feasibility status and thus the functions IsPrimalFeasible etc. should not be used

### 2022-08-01 src/sdpi/sdpi.h

@@ -449,6 +449,12 @@ SCIP_Bool SCIPsdpiSolvedOrig(
    SCIP_SDPI*            sdpi                /**< SDP-interface structure */
    );
 
+/** returns whether a primal solution or ray is available */
+SCIP_EXPORT
+SCIP_Bool SCIPsdpiHavePrimalSol(
+   SCIP_SDPI*            sdpi                /**< SDP-interface structure */
+   );
+
 /** returns true if the solver could determine whether the problem is feasible, so it returns true if the
  *  solver knows that the problem is feasible/infeasible/unbounded, it returns false if the solver does not know
  *  anything about the feasibility status and thus the functions IsPrimalFeasible etc. should not be used

### 2022-08-02 tex/ConflictAnalysis.tex
Extensions

@@
 certain SDP problems, we briefly discuss how primal solutions can be
-obtained in these cases.
+obtained in these cases. We start with a general discussion and then
+discuss how such solutions can be obtained from the semi-smooth Newton
+algorithm for solving SDPs with one variable.
 
@@
 % -------------------------------------------------------------------------
+\subsubsection{Obtaining Results from SDP with One Variable}
+
+The semi-smooth Newton algorithm for solving SDPs with one variable can
+also produce the corresponding solutions, see~\cite{MatP22}.
+
+\paragraph{Infeasibility I}
+
+The implementation can determine one kind of infeasibility
+of~\eqref{eq:basic} by computing a unit eigenvector $w$ for a minimal
+eigenvalue $\lambda$ of $A u - B$. If $\lambda < 0$ and $w\T A w > 0$, then
+the problem is infeasible.
+
+In this case, we can use $(X^*, \alpha^*, \beta^*) = (w w\T, 0, w\T A w)$
+as a ray. To see this, note that
+\[
+\skal{A}{X^*} + \alpha^* - \beta^* = w\T A w - w\T A w = 0.
+\]
+Moreover, the objective function satisfies
+\[
+\skal{B}{X^*} + \ell \alpha^* - u \beta^* = w\T B w - u\, w\T A w = w\T (B -
+u A) w = -\lambda\, w\T w > 0,
+\]
+and this holds for each positive multiple of $(X^*, \alpha^*, \beta^*)$.
+
+\paragraph{Infeasibility II}
+
+A second case of infeasibility of~\eqref{eq:basic} arises if a unit
+eigenvector $v$ for a minimal eigenvalue $\lambda$ of $A \ell - B$
+satisfies $\lambda < 0$ and $v\T A v \leq 0$.
+
+In this case, we can use $(X^*, \alpha^*, \beta^*) = (v v\T, -v\T A v, 0)$
+as a ray. To see this, note that
+\[
+\skal{A}{X^*} + \alpha^* - \beta^* = v\T A v - v\T A v = 0.
+\]
+Moreover, the objective function satisfies
+\[
+\skal{B}{X^*} + \ell \alpha^* - u \beta^* = v\T B v - \ell\, v\T A v = v\T (B -
+\ell A) v = -\lambda\, v\T v > 0,
+\]
+and this holds for each positive multiple of $(X^*, \alpha^*, \beta^*)$.
+
+\paragraph{Infeasibility III}
+
+The third case of infeasibility arises if the iteration terminates with
+$\mu \define \mu_k$ and $(v^k)\T A v^k = 0$, where $v \define v^k$ is a
+unit eigenvector of $A \mu - B$ corresponding to eigenvalue
+$\lambda \define \lambda_k < 0$. In this case, we can take
+$(X^*, \alpha^*, \beta^*) = (v v\T, 0, 0)$. To see this, first note that
+$(A \mu - B)v = \lambda v$. This implies that
+$\mu\, v\T A v - v\T B v = \lambda$, i.e., $- v\T B v = \lambda < 0$. Then
+\[
+\skal{A}{X^*} + \alpha^* - \beta^* = v\T A v = 0.
+\]
+Moreover, the objective function satisfies
+\[
+\skal{B}{X^*} + \ell \alpha^* - u \beta^* = v\T B v > 0,
+\]
+and this holds for each positive multiple of $(X^*, \alpha^*, \beta^*)$.
+
+\paragraph{Optimality}
+
+Let $\mu$ and $v$ be the final values of the algorithm in the case that no
+infeasibility was determined. Then $v$ is a unit eigenvector of $A \mu - B$
+corresponding to eigenvalue $\lambda$. Thus, the algorithm terminated
+because the condition in the while loop is not satisfied anymore, i.e.,
+$\lambda \geq 0$ or $v\T A v \leq 0$. Since the problem is feasible, we
+always have $\lambda \geq 0$ and in exact arithmetic the algorithm would
+stop with $\lambda = 0$.
+
+First assume that $\delta \define v\T A v > 0$. Note that $\lambda = 0$
+implies $(A \mu - B)v = 0$ and thus $v\T B v = \mu\, v\T A v$. Then we can
+use $(X^*, \alpha^*, \beta^*) = (\frac{\gamma}{\delta} v v\T, 0, 0)$.  To
+see this, note that
+\[
+\skal{A}{X^*} + \alpha^* - \beta^* = \frac{\gamma}{\delta}\, v\T A v = \gamma.
+\]
+Moreover, the objective function satisfies
+\[
+\skal{B}{X^*} + \ell \alpha^* - u \beta^* = \frac{\gamma}{\delta}\, v\T B v
+= \frac{\gamma}{\delta}\, \mu\, v\T A v = \gamma\, \mu,
+\]
+as it should be.
+
+Now assume that $v\T A v = 0$. In this case, strong duality does not need
+to hold and it is harder to construct any feasible solution, because we
+would need $\skal{A}{X^*} = \gamma$. For instance, if $A$ is negative
+semidefinite and $\gamma > 0$, this is impossible to achieve.
+
+% -------------------------------------------------------------------------
 \subsubsection{SDPs with All Fixed Variables}

### 2022-08-02 src/sdpi/sdpi.c
allow to return certificate of 1d SDP

@@ -311,6 +311,9 @@ struct SCIP_SDPi
    SCIP_Real             onevarsdpobjval;    /**< objective value of one variable SDP */
    SCIP_Real             onevarsdpoptval;    /**< optimal value of one variable SDP */
    int                   onevarsdpidx;       /**< index of variable for one variable SDP */
+   SCIP_Real*            onevarsdpcertvec;   /**< one variable SDP certificate vector (eigenvector) */
+   int                   onevarsdpcertsize;  /**< block size for one variable SDP certificate vector */
+   SCIP_Real             onevarsdpcertval;   /**< one variable SDP certificate value (supergradient) */
 };
 
 
@@ -1637,6 +1640,9 @@ SCIP_RETCODE SCIPsdpiCreate(
    (*sdpi)->onevarsdpobjval = SCIP_INVALID;
    (*sdpi)->onevarsdpoptval = SCIP_INVALID;
    (*sdpi)->onevarsdpidx = -1;
+   (*sdpi)->onevarsdpcertvec = NULL;
+   (*sdpi)->onevarsdpcertsize = -1;
+   (*sdpi)->onevarsdpcertval = SCIP_INVALID;
 
    (*sdpi)->nallfixed = 0;
    (*sdpi)->ninfeasible = 0;
@@ -1726,6 +1732,8 @@ SCIP_RETCODE SCIPsdpiFree(
    BMSfreeBlockMemoryArrayNull((*sdpi)->blkmem, &((*sdpi)->ub), (*sdpi)->maxnvars);/*lint !e737*/
    BMSfreeBlockMemoryArrayNull((*sdpi)->blkmem, &((*sdpi)->lb), (*sdpi)->maxnvars);/*lint !e737*/
    BMSfreeBlockMemoryArrayNull((*sdpi)->blkmem, &((*sdpi)->obj), (*sdpi)->maxnvars);/*lint !e737*/
+
+   BMSfreeBlockMemoryArrayNull((*sdpi)->blkmem, &((*sdpi)->onevarsdpcertvec), (*sdpi)->onevarsdpcertsize);
 
    /* free the solver */
    SCIP_CALL( SCIPsdpiSolverFree(&((*sdpi)->sdpisolver)) );
@@ -1884,6 +1892,9 @@ SCIP_RETCODE SCIPsdpiClone(
    newsdpi->onevarsdpobjval = SCIP_INVALID;
[…]

### 2022-08-02 src/sdpi/solveonevarsdp.c

@@ -172,6 +172,8 @@ SCIP_RETCODE SCIPsolveOneVarSDP(
    SCIP_Real*            sdpval,             /**< array of nonzero values */
    SCIP_Real             infinity,           /**< infinity value */
    SCIP_Real             feastol,            /**< feasibility tolerance */
+   SCIP_Real*            certificatevector,  /**< array to store a certificate (eigen)vector (or NULL if not required) */
+   SCIP_Real*            certificatevalue,   /**< array to store a certificate value (or NULL if not required) */
    SCIP_Real*            objval,             /**< pointer to store optimal objective value */
    SCIP_Real*            optval              /**< pointer to store optimal value of variable */
    )
@@ -200,6 +202,8 @@ SCIP_RETCODE SCIPsolveOneVarSDP(
 
    *objval = SCIP_INVALID;
    *optval = SCIP_INVALID;
+   if ( certificatevalue != NULL )
+      *certificatevalue = SCIP_INVALID;
 
    /* can currently only treat the case with finite lower and upper bounds */
    if ( lb <= -infinity )
@@ -265,6 +269,11 @@ SCIP_RETCODE SCIPsolveOneVarSDP(
          *objval = infinity;
          *optval = ub;
 
+         if ( certificatevalue != NULL )
+            *certificatevalue = supergradient;
+         if ( certificatevector != NULL )
+            BMScopyMemoryArray(certificatevector, eigenvector, blocksize);
+
          goto TERMINATE;
       }
    }
@@ -282,6 +291,7 @@ SCIP_RETCODE SCIPsolveOneVarSDP(
       SCIPdebugMessage("Lower bound is optimal.\n");
       *objval = obj * lb;
       *optval = lb;
+      supergradient = SCIP_INVALID;
 
       goto TERMINATE;
    }
@@ -350,6 +360,11 @@ SCIP_RETCODE SCIPsolveOneVarSDP(
    }
 
  TERMINATE:
+   if ( certificatevalue != NULL )
+      *certificatevalue = supergradient;
+   if ( certificatevector != NULL )
+      BMScopyMemoryArray(certificatevector, eigenvector, blocksize);
+
    BMSfreeBufferMemoryArrayNull(bufmem, &tmpmatrix);
    BMSfreeBufferMemoryArrayNull(bufmem, &fullmatrix);
    BMSfreeBufferMemoryArrayNull(bufmem, &fullconstmatrix);
